### 2020-01-21 dissertation/l4proj.tex
Write initial draft of background/related work

\documentclass{l4proj}

%
% put any additional packages here
%

\begin{document}

%==============================================================================
%% METADATA
\title{Boolean types and semantic subtyping for Featherweight Java}
\author{Artem Usov}
\date{January 18, 2020}

\maketitle

%==============================================================================
%% ABSTRACT
\begin{abstract}
    Every abstract follows a similar pattern. Motivate; set aims; describe work; explain results.
    \vskip 0.5em
    ``XYZ is bad. This project investigated ABC to determine if it was better. 
    ABC used XXX and YYY to implement ZZZ. This is particularly interesting as XXX and YYY have
    never been used together. It was found that  
    ABC was 20\% better than XYZ, though it caused rabies in half of subjects.''
\end{abstract}

%==============================================================================

% EDUCATION REUSE CONSENT FORM
% If you consent to your project being shown to future students for educational purposes
% then insert your name and the date below to  sign the education use form that appears in the front of the document. 
% You must explicitly give consent if you wish to do so.
% If you sign, your project may be included in the Hall of Fame if it scores particularly highly.
%
% Please note that you are under no obligation to sign 
% this declaration, but doing so would help future students.
%
\def\consentname {Artem Usov} % your full name
\def\consentdate {19 January 2020} % the date you agree
\educationalconsent

%==============================================================================
\tableofcontents

%==============================================================================
%% Notes on formatting
%==============================================================================
% The first page, abstract and table of contents are numbered using Roman numerals and are not
% included in the page count. 
%
% From now on pages are numbered
% using Arabic numerals. Therefore, immediately after the first call to \chapter we need the call
% \pagenumbering{arabic} and this should be called once only in the document. 
%
% Do not alter the bibliography style.
%
% The first Chapter should then be on page 1. You are allowed 40 pages for a 40 credit project and 30 pages for a 
% 20 credit report. This includes everything numbered in Arabic numerals (excluding front matter) up
% to but excluding the appendices and bibliography.
%
% You must not alter text size (it is currently 10pt) or alter margins or spacing.
%
%
%==================================================================================================================================
%
% IMPORTANT
% The chapter headings here are **suggestions**. You don't have to follow this model if
% it doesn't fit your project. Every project should have an introduction and conclusion,
% however. 
%
%==================================================================================================================================
\chapter{Introduction}

% reset page numbering. Don't remove this!
\pagenumbering{arabic}

Why should the reader care about what are you doing and what are you actually doing?
\section{Guidance}

\textbf{Motivate} first, then state the general problem clearly.

\section{Writing guidance}
\subsection{Who is the reader?}

This is the key question for any writing. Your reader:

\begin{itemize}
    \item
          is a trained computer scientist: \emph{don't explain basics}.
    \item
          has limited time: \emph{keep on topic}.
    \item
          has no idea why anyone would want to do this: \emph{motivate clearly}
    \item
          might not know \emph{anything} about your project in particular:
          \emph{explain your project}.
    \item
          but might know precise details and check them: \emph{be precise and
              strive for accuracy.}
    \item
          doesn't know or care about you: \emph{personal discussions are
              irrelevant}.
\end{itemize}

Remember, you will be marked by your supervisor and one or more members
of staff. You might also have your project read by a prize-awarding
committee or possibly a future employer. Bear that in mind.

\subsection{References and style guides}
There are many style guides on good English writing. You don't need to
read these, but they will improve how you write.

\begin{itemize}
    \item
          \emph{How to write a great research paper} \cite{Pey17} (\textbf{recommended}, even
          though you aren't writing a research paper)
    \item
          \emph{How to Write with Style} \cite{Von80}. Short and easy to read. Available online.
    \item
          \emph{Style: The Basics of Clarity and Grace} \cite{Wil09} A very popular modern English 
          style guide.
    \item
          \emph{Politics and the English Language} \cite{Orw68}  A famous essay on effective, clear
          writing in English.
    \item
          \emph{The Elements of Style} \cite{StrWhi07} Outdated, and American, but a classic.
    \item
          \emph{The Sense of Style} \cite{Pin15} Excellent, though quite in-depth.
\end{itemize}

\subsubsection{Citation styles}

\begin{itemize}
    \item If you are referring to a reference as a noun, then cite it as: ``\citet{Orw68} discusses
          the role of language in political thought.''
    \item If you are referring implicitly to references, use: ``There are many good books on writing
          \citep{Orw68, Wil09, Pin15}.''
\end{itemize}

There is a complete guide on good citation practice by Peter Coxhead available here: 
\url{http://www.cs.bham.ac.uk/~pxc/refs/index.html}. 
If you are unsure about how to cite online sources, please see this guide: 
\url{https://student.unsw.edu.au/how-do-i-cite-electronic-sources}.

\subsection{Plagiarism warning}

\begin{highlight_title}{WARNING}

    If you include material from other sources without full and correct attribution, you are 
    commiting plagiarism. The penalties for plagiarism are severe. Quote any included text and cite
    it correctly. Cite all images, figures, etc. clearly in the caption of the figure.
\end{highlight_title}

%==================================================================================================================================
\chapter{Background}
What did other people do, and how is it relevant to what you want to do?

\section{Type Systems}

Type systems are by far the most popular and best developed methods in modern software engineering 
to reduce the number of run-time errors in a program. A possible definition of a type system is 
given by \citet{Pierce2002} as:

\textit{A type system is a tractable syntactic method for proving the absence of certain program
    behaviors by classifying phrases according to the kinds of values they compute.}

In particular, the most important detail in the above definition is the emphasis on classifying 
phrases in our language into a particular type based on the kinds of values they hold or compute. 
For example this would allow us to check that the arguments given to an arithmetic operation are 
always numbers. This therefore allows us to validate that a program is absent of certain bad 
run-time behaviour.

However what the above definition is missing that is it defines a type system as solely a syntactic 
method, where it is defined by a formal system of deduction rules, which is by far the most common 
approach for a type system. There does exist an alternative method in the semantic approach. 
\citet{Frisch2002} describe the semantic approach as instead starting with a model of the language 
and all the possible values in the language and define an interpretation of types as subsets of 
values the model.

Such as model comes with several advantages, such as when $t \leq s$ does not hold, then it is
possible to exhibit an element of the model which is in the interpretation of type $t$ and not $s$, 
even in the presence of arrow \textit{(function)} types. This can then be used to show more 
informative error messages to the programmer \citep{Castagna2005}. However due to being a more 
technical, such as it being not trivial to actually define the interpretation of types as subsets of
a model, it has received less attention than syntactic methods.

The first problem we encounter in the semantic approach is that in the semantic approach, $t_{1}$
is a subtype of $t_{2}$ if all the $t_{1}$-values are also $t_{2}$-values. However, in this way, 
subtyping is defined by relying on the notion of well-typed values; hence, we need the typing 
relation to determine typing judgements for values; but the typing rules use the subtyping relation
which we are going to define \citep{Dardha2017}. So, there is a circularity as demonstrated in 
Figure \ref{fig:circ}.

To solve this problem, we follow the framework defined in \citep{Frisch2008}. The general idea of
the framework is that we first extend the types in the language with \textit{Boolean Combinators}: 
union $\lor$, intersection $\land$ and negation $\neg$. We then define an abstract model $\beta$ 
where our types are interpreted according to the set-theoretic meaning of the combinators and can 
define subtyping as set-inclusion and therefore define the typing rules. Separately, if we have a 
notion of well-typed values, we can define another interpretation of types as sets of values. We can
again define a subtyping relation $\nu$. While these may be different relation, if the models are 
chosen carefully such that 
\begin{equation}
    s \leq_{\beta} t \iff s \leq_{\nu} t,
\end{equation}
then these models coincide and this closes the circularity.

\begin{figure}
    \centering
    \includegraphics[width=0.4\linewidth]{images/circularity.PNG}
    \caption{TODO caption, \citep{Castagna2005}}
    \label{fig:circ}
\end{figure}

\section{Featherweight Java}

The syntax of the language is exactly that of Featherweight Java, or FJ \citep{Igarashi1999}, with 
the associated type system only being different. This was chosen on purpose as the full features of 
Java such as concurrency and reflection are orthogonal (\textit{TODO reword}) to the purpose of the 
project while the type system of FJ still exactly corresponds to that of Java. The main key 
simplification in FJ is the omission of assignment. All fields of an object are initialised in the
constructor and not changed afterwards. While this restricts FJ to what is essentially a functional 
fragment of Java, it is still fully computationally complete. The syntax of FJ can be seen in Figure
\ref{fig:syntax} and a sample code fragment can be seen in Figure \ref{fig:class}.

\begin{figure}
    \centering
    \includegraphics[width=0.4\linewidth]{images/syntax.PNG}
    \caption{TODO caption, \citep{Igarashi1999}}
    \label{fig:syntax}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=0.4\linewidth]{images/class.PNG}
    \caption{TODO caption, \citep{Igarashi1999}}
    \label{fig:class}
\end{figure}

\section{Related Work}
The closest similar area of research would likely be the work on $\mathbb{C}$Duce 
\citep{Benzaken2003}, which is also a functional language with a semantic type system designed for 
working with XML documents and a continuation of the work on XDuce \citep{Hosoya2003}. The language
extended XDuce by introducing less XML specific types such as records, boolean connectives and arrow
types. This therefore makes it similar to (\textit{TODO reword}) our language in that a class-based 
semantic type system is a combination of the $\mathbb{C}$Duce record types with arrow types 
\textit{(fields and methods)}

Our work and the work on $\mathbb{C}$Duce follow the functional style of $\lambda$-calculus,
whereas the work by \citet{Castagna2008} extends $\pi$-calculus with semantic subtyping.
(\textit{TODO reword}) Similar work to ours creating an implementation for this would result in a 
Golang-like language \citep{Meyerson2014}, creating a concurrency-focused language with more 
intuitive types. The paper found that it was required to be able to decide and resolve the atomicy, 
that is whether the only proper subtype is the empty type, of types in order to decide the subtyping
relation, and observes that this same problem appears in $\lambda$-calculus and any other 
semantic-based system, for which we will show where we encountered this problem further in this 
paper.

%==================================================================================================================================
\chapter{Analysis/Requirements}
What is the problem that you want to solve, and how did you arrive at it?
\section{Guidance}
Make it clear how you derived the constrained form of your problem via a clear and logical process.

%==================================================================================================================================
\chapter{Design}
How is this problem to be approached, without reference to specific implementation details? 
\section{Guidance}
Design should cover the abstract design in such a way that someone else might be able to do what you did, but with a different language or library or tool.

%==================================================================================================================================
\chapter{Implementation}
What did you do to implement this idea, and what technical achievements did you make?
\section{Guidance}
You can't talk about everything. Cover the high level first, then cover important, relevant or impressive details.

\section{General points}

These points apply to the whole dissertation, not just this chapter.

\subsection{Figures}
\emph{Always} refer to figures included, like Figure \ref{fig:relu}, in the body of the text. Include full, explanatory captions and make sure the figures look good on the page.
You may include multiple figures in one float, as in Figure \ref{fig:synthetic}, using \texttt{subcaption}, which is enabled in the template.

% Figures are important. Use them well.
\begin{figure}
    \centering
    \includegraphics[width=0.5\linewidth]{images/relu.pdf}

    \caption{In figure captions, explain what the reader is looking at: ``A schematic of the rectifying linear unit, where $a$ is the output amplitude,
        $d$ is a configurable dead-zone, and $Z_j$ is the input signal'', as well as why the reader is looking at this: 
        ``It is notable that there is no activation \emph{at all} below 0, which explains our initial results.'' 
        \textbf{Use vector image formats (.pdf) where possible}. Size figures appropriately, and do not make them over-large or too small to read.
    }

    % use the notation fig:name to cross reference a figure
    \label{fig:relu}
\end{figure}

\begin{figure}
    \centering
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{images/synthetic.png}
        \caption{Synthetic image, black on white.}
        \label{fig:syn1}
    \end{subfigure}
    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
    %(or a blank line to force the subfigure onto a new line)
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{images/synthetic_2.png}
        \caption{Synthetic image, white on black.}
        \label{fig:syn2}
    \end{subfigure}
    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
    %(or a blank line to force the subfigure onto a new line)    
    \caption{Synthetic test images for edge detection algorithms. \subref{fig:syn1} shows various gray levels that require an adaptive algorithm. \subref{fig:syn2}
        shows more challenging edge detection tests that have crossing lines. Fusing these into full segments typically requires algorithms like the Hough transform.
        This is an example of using subfigures, with \texttt{subref}s in the caption.
    }\label{fig:synthetic}
\end{figure}

\clearpage

\subsection{Equations}

Equations should be typeset correctly and precisely. Make sure you get parenthesis sizing correct, and punctuate equations correctly 
(the comma is important and goes \textit{inside} the equation block). Explain any symbols used clearly if not defined earlier.

For example, we might define:
\begin{equation}
    \hat{f}(\xi) = \frac{1}{2}\left[ \int_{-\infty}^{\infty} f(x) e^{2\pi i x \xi} \right],
\end{equation}
where $\hat{f}(\xi)$ is the Fourier transform of the time domain signal $f(x)$.

\subsection{Algorithms}
Algorithms can be set using \texttt{algorithm2e}, as in Algorithm \ref{alg:metropolis}.

% NOTE: line ends are denoted by \; in algorithm2e
\begin{algorithm}
    \DontPrintSemicolon
    \KwData{$f_X(x)$, a probability density function returing the density at $x$.\; $\sigma$ a standard deviation specifying the spread of the proposal distribution.\;
        $x_0$, an initial starting condition.}
    \KwResult{$s=[x_1, x_2, \dots, x_n]$, $n$ samples approximately drawn from a distribution with PDF $f_X(x)$.}
    \Begin{
        $s \longleftarrow []$\;
        $p \longleftarrow f_X(x)$\;
        $i \longleftarrow 0$\;
        \While{$i < n$}
        {
            $x^\prime \longleftarrow \mathcal{N}(x, \sigma^2)$\;
            $p^\prime \longleftarrow f_X(x^\prime)$\;
            $a \longleftarrow \frac{p^\prime}{p}$\;
            $r \longleftarrow U(0,1)$\;
            \If{$r<a$}
            {
                $x \longleftarrow x^\prime$\;
                $p \longleftarrow f_X(x)$\;
                $i \longleftarrow i+1$\;
                append $x$ to $s$\;
            }
        }
    }

    \caption{The Metropolis-Hastings MCMC algorithm for drawing samples from arbitrary probability distributions,
        specialised for normal proposal distributions $q(x^\prime|x) = \mathcal{N}(x, \sigma^2)$. The symmetry of the normal distribution means the acceptance rule takes the simplified form.}\label{alg:metropolis}
\end{algorithm}

\subsection{Tables}

If you need to include tables, like Table \ref{tab:operators}, use a tool like https://www.tablesgenerator.com/ to generate the table as it is
extremely tedious otherwise.

\begin{table}[]
    \caption{The standard table of operators in Python, along with their functional equivalents from the \texttt{operator} package. Note that table
        captions go above the table, not below. Do not add additional rules/lines to tables. }\label{tab:operators}
    %\tt 
    \rowcolors{2}{}{gray!3}
    \begin{tabular}{@{}lll@{}}
        %\toprule
        \textbf{Operation}    & \textbf{Syntax}                         & \textbf{Function}                          \\ %\midrule % optional rule for header
        Addition              & \texttt{a + b}                          & \texttt{add(a, b)}                         \\
        Concatenation         & \texttt{seq1 + seq2}                    & \texttt{concat(seq1, seq2)}                \\
        Containment Test      & \texttt{obj in seq}                     & \texttt{contains(seq, obj)}                \\
        Division              & \texttt{a / b}                          & \texttt{div(a, b) }                        \\
        Division              & \texttt{a / b}                          & \texttt{truediv(a, b) }                    \\
        Division              & \texttt{a // b}                         & \texttt{floordiv(a, b)}                    \\
        Bitwise And           & \texttt{a \& b}                         & \texttt{and\_(a, b)}                       \\
        Bitwise Exclusive Or  & \texttt{a \textasciicircum b}           & \texttt{xor(a, b)}                         \\
        Bitwise Inversion     & \texttt{$\sim$a}                        & \texttt{invert(a)}                         \\
        Bitwise Or            & \texttt{a | b}                          & \texttt{or\_(a, b)}                        \\
        Exponentiation        & \texttt{a ** b}                         & \texttt{pow(a, b)}                         \\
        Identity              & \texttt{a is b}                         & \texttt{is\_(a, b)}                        \\
        Identity              & \texttt{a is not b}                     & \texttt{is\_not(a, b)}                     \\
        Indexed Assignment    & \texttt{obj{[}k{]} = v}                 & \texttt{setitem(obj, k, v)}                \\
        Indexed Deletion      & \texttt{del obj{[}k{]}}                 & \texttt{delitem(obj, k)}                   \\
        Indexing              & \texttt{obj{[}k{]}}                     & \texttt{getitem(obj, k)}                   \\
        Left Shift            & \texttt{a \textless{}\textless b}       & \texttt{lshift(a, b)}                      \\
        Modulo                & \texttt{a \% b}                         & \texttt{mod(a, b)}                         \\
        Multiplication        & \texttt{a * b}                          & \texttt{mul(a, b)}                         \\
        Negation (Arithmetic) & \texttt{- a}                            & \texttt{neg(a)}                            \\
        Negation (Logical)    & \texttt{not a}                          & \texttt{not\_(a)}                          \\
        Positive              & \texttt{+ a}                            & \texttt{pos(a)}                            \\
        Right Shift           & \texttt{a \textgreater{}\textgreater b} & \texttt{rshift(a, b)}                      \\
        Sequence Repetition   & \texttt{seq * i}                        & \texttt{repeat(seq, i)}                    \\
        Slice Assignment      & \texttt{seq{[}i:j{]} = values}          & \texttt{setitem(seq, slice(i, j), values)} \\
        Slice Deletion        & \texttt{del seq{[}i:j{]}}               & \texttt{delitem(seq, slice(i, j))}         \\
        Slicing               & \texttt{seq{[}i:j{]}}                   & \texttt{getitem(seq, slice(i, j))}         \\
        String Formatting     & \texttt{s \% obj}                       & \texttt{mod(s, obj)}                       \\
        Subtraction           & \texttt{a - b}                          & \texttt{sub(a, b)}                         \\
        Truth Test            & \texttt{obj}                            & \texttt{truth(obj)}                        \\
        Ordering              & \texttt{a \textless b}                  & \texttt{lt(a, b)}                          \\
        Ordering              & \texttt{a \textless{}= b}               & \texttt{le(a, b)}                          \\
        % \bottomrule
    \end{tabular}
\end{table}
\subsection{Code}

Avoid putting large blocks of code in the report (more than a page in one block, for example). Use syntax highlighting if possible, as in Listing \ref{lst:callahan}.

\begin{lstlisting}[language=python, float, caption={The algorithm for packing the $3\times 3$ outer-totalistic binary CA successor rule into a 
    $16\times 16\times 16\times 16$ 4 bit lookup table, running an equivalent, notionally 16-state $2\times 2$ CA.}, label=lst:callahan]
    def create_callahan_table(rule="b3s23"):
        """Generate the lookup table for the cells."""        
        s_table = np.zeros((16, 16, 16, 16), dtype=np.uint8)
        birth, survive = parse_rule(rule)

        # generate all 16 bit strings
        for iv in range(65536):
            bv = [(iv >> z) & 1 for z in range(16)]
            a, b, c, d, e, f, g, h, i, j, k, l, m, n, o, p = bv

            # compute next state of the inner 2x2
            nw = apply_rule(f, a, b, c, e, g, i, j, k)
            ne = apply_rule(g, b, c, d, f, h, j, k, l)
            sw = apply_rule(j, e, f, g, i, k, m, n, o)
            se = apply_rule(k, f, g, h, j, l, n, o, p)

            # compute the index of this 4x4
            nw_code = a | (b << 1) | (e << 2) | (f << 3)
            ne_code = c | (d << 1) | (g << 2) | (h << 3)
            sw_code = i | (j << 1) | (m << 2) | (n << 3)
            se_code = k | (l << 1) | (o << 2) | (p << 3)

            # compute the state for the 2x2
            next_code = nw | (ne << 1) | (sw << 2) | (se << 3)

            # get the 4x4 index, and write into the table
            s_table[nw_code, ne_code, sw_code, se_code] = next_code

        return s_table

\end{lstlisting}

%==================================================================================================================================
\chapter{Evaluation}
How good is your solution? How well did you solve the general problem, and what evidence do you have to support that?

\section{Guidance}
\begin{itemize}
    \item
          Ask specific questions that address the general problem.
    \item
          Answer them with precise evidence (graphs, numbers, statistical
          analysis, qualitative analysis).
    \item
          Be fair and be scientific.
    \item
          The key thing is to show that you know how to evaluate your work, not
          that your work is the most amazing product ever.
\end{itemize}

\section{Evidence}
Make sure you present your evidence well. Use appropriate visualisations, reporting techniques and statistical analysis, as appropriate.

If you visualise, follow the basic rules, as illustrated in Figure \ref{fig:boxplot}:
\begin{itemize}
    \item Label everything correctly (axis, title, units).
    \item Caption thoroughly.
    \item Reference in text.
    \item \textbf{Include appropriate display of uncertainty (e.g. error bars, Box plot)}
    \item Minimize clutter.
\end{itemize}

See the file \texttt{guide\_to\_visualising.pdf} for further information and guidance.

\begin{figure}
    \centering
    \includegraphics[width=1.0\linewidth]{images/boxplot_finger_distance.pdf}

    \caption{Average number of fingers detected by the touch sensor at different heights above the surface, averaged over all gestures. Dashed lines indicate
        the true number of fingers present. The Box plots include bootstrapped uncertainty notches for the median. It is clear that the device is biased toward 
        undercounting fingers, particularly at higher $z$ distances.
    }

    % use the notation fig:name to cross reference a figure
    \label{fig:boxplot}
\end{figure}

%==================================================================================================================================
\chapter{Conclusion}
Summarise the whole project for a lazy reader who didn't read the rest (e.g. a prize-awarding committee).
\section{Guidance}
\begin{itemize}
    \item
          Summarise briefly and fairly.
    \item
          You should be addressing the general problem you introduced in the
          Introduction.        
    \item
          Include summary of concrete results (``the new compiler ran 2x
          faster'')
    \item
          Indicate what future work could be done, but remember: \textbf{you
              won't get credit for things you haven't done}.
\end{itemize}

%==================================================================================================================================
%
% 
%==================================================================================================================================
%  APPENDICES

\begin{appendices}

    \chapter{Appendices}

    Typical inclusions in the appendices are:

    \begin{itemize}
        \item
              Copies of ethics approvals (required if obtained)
        \item
              Copies of questionnaires etc. used to gather data from subjects.
        \item
              Extensive tables or figures that are too bulky to fit in the main body of
              the report, particularly ones that are repetitive and summarised in the body.

        \item Outline of the source code (e.g. directory structure), or other architecture documentation like class diagrams.

        \item User manuals, and any guides to starting/running the software.

    \end{itemize}

    \textbf{Don't include your source code in the appendices}. It will be
    submitted separately.

\end{appendices}

%==================================================================================================================================
%   BIBLIOGRAPHY

% The bibliography style is abbrvnat
% The bibliography always appears last, after the appendices.

\bibliographystyle{abbrvnat}

\bibliography{l4proj}

\end{document}

### 2020-02-17 src/main/java/domain/Class.java
initial implementation of method types

@@ -56,6 +56,10 @@ public class Class {
     return methods;
   }
 
+  public void addMethod(final Method method) {
+    this.getMethods().put(method.getName(), method);
+  }
+
   public String getSuperName() {
     return superName;
   }
@@ -68,8 +72,8 @@ public class Class {
     return token;
   }
 
-  public boolean isResolved() {
-    return resolved;
+  public boolean isNotResolved() {
+    return !resolved;
   }
 
   public void setResolved(final boolean resolved) {

### 2020-02-17 src/main/java/domain/Method.java

@@ -9,7 +9,7 @@ public class Method {
   // TODO - Encapsulate return and parameter type
   private BooleanType returnType;
   private final String name;
-  private final Parameter parameter;
+  private final MethodParameter parameter;
   private final ExpressionContext expression;
   private final Token token;
 
@@ -21,7 +21,7 @@ public class Method {
    * @param parameter  parameter that method takes
    * @param ctx        the expression context within method that we store to type check later
    */
-  public Method(final BooleanType returnType, final String name, final Parameter parameter,
+  public Method(final BooleanType returnType, final String name, final MethodParameter parameter,
                 final ExpressionContext ctx, final Token token) {
     this.returnType = returnType;
     this.name = name;
@@ -34,7 +34,7 @@ public class Method {
     return name;
   }
 
-  public Parameter getParameter() {
+  public MethodParameter getParameter() {
     return parameter;
   }
 

### 2020-02-17 src/main/java/domain/MethodParameter.java

@@ -1,13 +1,15 @@
 package domain;
 
 import domain.type.BooleanType;
+import domain.type.MethodType;
 import org.antlr.v4.runtime.Token;
+import org.antlr.v4.runtime.misc.Pair;
 
 import java.util.Objects;
 
-public class Parameter {
+public class MethodParameter {
   private final String name;
-  private final BooleanType type;
+  private final Pair<BooleanType, MethodType> type;
   private final Token token;
 
   /**
@@ -16,14 +18,14 @@ public class Parameter {
    * @param name name of parameter
    * @param type type of parameter, currently as a string
    */
-  public Parameter(final String name, final BooleanType type, final Token token) {
+  public MethodParameter(final String name, final Pair<BooleanType, MethodType> type, final Token token) {
     super();
     this.name = name;
     this.type = type;
     this.token = token;
   }
 
-  public BooleanType getType() {
+  public Pair<BooleanType, MethodType> getType() {
     return type;
   }
 
@@ -39,7 +41,7 @@ public class Parameter {
   public boolean equals(Object o) {
     if (this == o) return true;
     if (o == null || getClass() != o.getClass()) return false;
-    Parameter parameter = (Parameter) o;
+    MethodParameter parameter = (MethodParameter) o;
     return Objects.equals(type, parameter.type);
   }
 

### 2020-02-17 src/main/java/domain/Program.java

@@ -1,18 +1,25 @@
 package domain;
 
 import java.util.Map;
+import jsf.jsfParser.ExpressionContext;
 
 public class Program {
 
   private final Map<String, Class> classes;
+  private final ExpressionContext expression;
 
-  public Program(final Map<String, Class> classes) {
+  public Program(final Map<String, Class> classes, ExpressionContext expression) {
     super();
     this.classes = classes;
+    this.expression = expression;
   }
 
   public Map<String, Class> getClasses() {
     return classes;
+  }
+
+  public ExpressionContext getExpression() {
+    return expression;
   }
 
   @Override

### 2020-02-17 src/main/java/domain/type/BooleanType.java

@@ -1,6 +1,7 @@
 package domain.type;
 
 import java.util.HashSet;
+import java.util.Objects;
 import java.util.Set;
 
 import util.TypeHelper;
@@ -58,7 +59,7 @@ public class BooleanType {
 
     if (this.isLeafNode()) {
       if (!TypeHelper.SUB_CLASSES.containsKey(this.getType())) {
-        return new HashSet<>();
+        throw new RuntimeException("Trying to use class that does not exist: " + this.getType().getName());
       }
       result = new HashSet<>(TypeHelper.SUB_CLASSES.get(this.getType()));
 
@@ -93,7 +94,7 @@ public class BooleanType {
   }
 
   /**
-   * Gets all the unique concrete types in the boolean expressions, this ignores NOT's and connectives.
+   * Gets all the unique concrete types in the boolean expressions, this ignores connectives.
    *
    * @return set of all types included in the expression
    */
@@ -107,7 +108,9 @@ public class BooleanType {
 
   private void typeGetHelper(final Set<Type> set) {
     if (this.isLeafNode()) {
-      set.add(this.getType());
+      if (!this.isNot()) {
+        set.add(this.getType());
+      }
     } else {
       this.getLeft().typeGetHelper(set);
       this.getRight().typeGetHelper(set);
@@ -145,6 +148,24 @@ public class BooleanType {
     } else {
       return connective + ": " + left + ", " + right;
     }
+  }
+
+  @Override
+  public boolean equals(Object o) {
+    if (this == o) return true;
+    if (o == null || getClass() != o.getClass()) return false;
+    BooleanType that = (BooleanType) o;
+    return leafNode == that.leafNode &&
+        connective == that.connective &&
+        Objects.equals(left, that.left) &&
+        Objects.equals(right, that.right) &&
+        Objects.equals(type, that.type) &&
+        Objects.equals(not, that.not);
+  }
+
+  @Override
+  public int hashCode() {
+    return Objects.hash(leafNode, connective, left, right, type, not);
   }
 
   public enum BooleanConnective {

### 2020-02-17 src/main/java/domain/type/MethodType.java

@@ -1,0 +1,66 @@
+package domain.type;
+
+import domain.Program;
+import util.TypeHelper;
+
+import java.util.HashSet;
+import java.util.Set;
+
+public class MethodType {
+  private final String name;
+  private final BooleanType parameter;
+  private final BooleanType returnType;
+  private final boolean isNot;
+
+  public MethodType(String name, BooleanType parameter, BooleanType returnType, boolean isNot) {
+    this.name = name;
+    this.parameter = parameter;
+    this.returnType = returnType;
+    this.isNot = isNot;
+  }
+
+  public String getName() {
+    return name;
+  }
+
+  public BooleanType getParameter() {
+    return parameter;
+  }
+
+  public BooleanType getReturnType() {
+    return returnType;
+  }
+
+  public boolean isNot() {
+    return isNot;
+  }
+
+  public Set<Type> getTypes(Program program) {
+    // TODO make this static or something
+    var set = new HashSet<Type>();
+
+    for (var c : program.getClasses().values()) {
+      if (c.getMethods().containsKey(this.getName())) {
+        var method = c.getMethods().get(this.getName());
+
+        if (method.getReturnType().equals(this.getReturnType())) {
+          if (method.getParameter().getType().a != null && method.getParameter().getType().a.equals(this.getParameter())){
+            set.add(c.getType());
+          }
+          // TODO FIX THIS
+//          if (method.getParameter().getType().a == null && method.getParameter().getType().b.equals(this)) {
+//            set.add(c.getType());
+//          }
+        }
+      }
+    }
+
+    if (isNot) {
+      final var universe = new HashSet<>(TypeHelper.UNIVERSE_SET);
+      universe.removeAll(set);
+      return universe;
+    } else {
+      return set;
+    }
+  }
+}

### 2020-02-17 src/main/java/jsf/App.java

@@ -35,12 +35,12 @@ public final class App {
       final var visitor = new ProgramVisitor();
       final var program = visitor.visit(tree);
 
-      System.out.println(program);
+      //System.out.println(program);
 
       visitor.visit(program);
-    } catch (IOException e) {
+    } catch (IOException | RuntimeException e) {
       if (LOGGER.isLoggable(Level.SEVERE)) {
-        LOGGER.severe("Exception occurred: " + e);
+        LOGGER.severe("Exception occurred: " + e.getMessage());
       }
     }
   }

### 2020-02-17 src/main/java/parsing/ClassVisitor.java

@@ -7,8 +7,10 @@ import domain.Field;
 import domain.Method;
 import domain.Program;
 import domain.type.ClassType;
+
 import java.util.HashMap;
 import java.util.concurrent.atomic.AtomicBoolean;
+
 import jsf.jsfBaseVisitor;
 import jsf.jsfParser.ClassDeclContext;
 
@@ -39,8 +41,7 @@ public class ClassVisitor extends jsfBaseVisitor<Class> {
       }
     });
 
-    final var constructorVisitor = new ConstructorVisitor(name);
-    final var constructor = ctx.constructorDecl().accept(constructorVisitor);
+    final var constructor = new ConstructorVisitor(name).visit(ctx.constructorDecl());
 
     final var methodVisitor = new MethodVisitor();
     ctx.methodDecl().forEach(methodCtx -> {
@@ -48,7 +49,6 @@ public class ClassVisitor extends jsfBaseVisitor<Class> {
       if (methods.containsKey(m.getName())) {
         reportError("repeated method name: " + m.getName(), m.getToken());
       } else {
-        // TODO CHECK FOR CLASS TYPE TO RESOLVE
         methods.put(m.getName(), m);
       }
     });
@@ -63,8 +63,7 @@ public class ClassVisitor extends jsfBaseVisitor<Class> {
    * @param program    the whole program structure
    */
   public void visit(final Class visitClass, final Program program) {
-    final var constructorVisitor = new ConstructorVisitor(visitClass.getName());
-    constructorVisitor.visit(visitClass.getConstructor(), program);
+    new ConstructorVisitor(visitClass.getName()).visit(visitClass.getConstructor(), program);
 
     final var methodVisitor = new MethodVisitor();
     visitClass.getMethods().values().forEach(m -> methodVisitor.visit(m, program));

### 2020-02-17 src/main/java/parsing/ConstructorVisitor.java

@@ -4,6 +4,7 @@ import static util.TypeResolverUtils.isNotValidSubtype;
 import static util.TypeResolverUtils.reportError;
 
 import domain.Constructor;
+import domain.Field;
 import domain.Parameter;
 import domain.Program;
 
@@ -12,6 +13,7 @@ import java.util.HashMap;
 import java.util.LinkedHashMap;
 import java.util.stream.IntStream;
 
+import domain.type.ClassType;
 import jsf.jsfBaseVisitor;
 import jsf.jsfParser.ConstructorDeclContext;
 import org.antlr.v4.runtime.Token;
@@ -34,6 +36,8 @@ public class ConstructorVisitor extends jsfBaseVisitor<Constructor> {
 
     final var typeVisitor = new TypeVisitor();
 
+    // Constructor Parameters --------------------------------------------------------
+
     final var parameters = new LinkedHashMap<String, Parameter>();
 
     if (ctx.type().size() > 0) {
@@ -44,8 +48,7 @@ public class ConstructorVisitor extends jsfBaseVisitor<Constructor> {
 
         for (var t : parameter.getType().getTypes()) {
           if (t.getName().equals(owner)) {
-            reportError("type of parameter in constructor cant be itself: " + parameter.getName()
-                , parameter.getToken());
+            reportError("type of parameter in constructor cant be itself", parameter.getToken());
           }
         }
 
@@ -57,18 +60,20 @@ public class ConstructorVisitor extends jsfBaseVisitor<Constructor> {
       });
     }
 
-    final var superVisitor = new SuperVisitor();
-    final var superArgumentNames = superVisitor.visit(ctx.superDecl());
+    // Super call arguments ---------------------------------------------
+
+    final var superArgumentNames = new SuperVisitor().visit(ctx.superDecl());
 
     final var superArguments = new ArrayList<Parameter>();
     superArgumentNames.forEach(a -> {
       if (!parameters.containsKey(a)) {
-        reportError("Parameter " + a + "doesnt exist in " + "constructor parameters of " + owner,
-            ctx.superDecl().start);
+        reportError(a + "doesnt exist in " + "constructor parameters of " + owner, ctx.superDecl().start);
       } else {
         superArguments.add(parameters.get(a));
       }
     });
+
+    // Field assignments -----------------------------------------------
 
     final var fieldAssignments = new ArrayList<Pair<Token, Parameter>>();
     final var fieldAssignmentVisitor = new FieldAssignmentVisitor();
[…]

### 2020-02-17 src/main/java/parsing/ExpressionVisitor.java

@@ -3,6 +3,8 @@ package parsing;
 import static util.TypeResolverUtils.getFromValue;
 import static util.TypeResolverUtils.isNotValidSubtype;
 
+import com.google.common.collect.Sets;
+import domain.MethodParameter;
 import domain.Parameter;
 import domain.Program;
 import domain.type.BasicType;
@@ -15,17 +17,19 @@ import java.util.concurrent.atomic.AtomicBoolean;
 
 import jsf.jsfBaseVisitor;
 import jsf.jsfParser.DecimalContext;
+import jsf.jsfParser.FalseContext;
 import jsf.jsfParser.ExpressionContext;
 import jsf.jsfParser.FieldContext;
 import jsf.jsfParser.MethodContext;
 import jsf.jsfParser.NumContext;
 import jsf.jsfParser.ObjectContext;
+import jsf.jsfParser.TrueContext;
 import jsf.jsfParser.VarContext;
 
 public class ExpressionVisitor extends jsfBaseVisitor<Set<Type>> {
 
   private final Program program;
-  private final Parameter parameter;
+  private final MethodParameter parameter;
 
   /**
    * Create an expression visitor, done on second round analysis.
@@ -33,7 +37,7 @@ public class ExpressionVisitor extends jsfBaseVisitor<Set<Type>> {
    * @param program   whole program context
    * @param parameter the parameter of the method the expression is in
    */
-  public ExpressionVisitor(final Program program, final Parameter parameter) {
+  public ExpressionVisitor(final Program program, final MethodParameter parameter) {
     super();
     this.program = program;
     this.parameter = parameter;
@@ -46,6 +50,7 @@ public class ExpressionVisitor extends jsfBaseVisitor<Set<Type>> {
     if (ctx.e2 != null) {
       final var types2 = ctx.e2.accept(this);
 
+      // TODO most likely needs fixed
       types.retainAll(types2);
 
       final var hasNumber = new AtomicBoolean(false);
@@ -67,16 +72,22 @@ public class ExpressionVisitor extends jsfBaseVisitor<Set<Type>> {
 
   @Override
   public Set<Type> visitNum(final NumContext ctx) {
-    final var set = new HashSet<Type>();
-    set.add(getFromValue(ctx.NUMBER().getText()));
-    return set;
+    return Sets.newHashSet(getFromValue(ctx.NUMBER().getText()));
   }
 
   @Override
   public Set<Type> visitDecimal(final DecimalContext ctx) {
-    final var set = new HashSet<Type>();
[…]

### 2020-02-17 src/main/java/parsing/MethodTypeVisitor.java

@@ -1,0 +1,29 @@
+package parsing;
+
+import domain.type.BooleanType;
+import domain.type.MethodType;
+import jsf.jsfBaseVisitor;
+import jsf.jsfParser.MethodtypeContext;
+import jsf.jsfParser.NormalContext;
+import org.antlr.v4.runtime.misc.Pair;
+
+public class MethodTypeVisitor extends jsfBaseVisitor<Pair<BooleanType, MethodType>> {
+
+  @Override
+  public Pair<BooleanType, MethodType> visitNormal(NormalContext ctx) {
+    var type = new TypeVisitor().visit(ctx.type());
+    return new Pair<>(type, null);
+  }
+
+  @Override
+  public Pair<BooleanType, MethodType> visitMethodtype(MethodtypeContext ctx) {
+    var typeVisitor = new TypeVisitor();
+    var param = typeVisitor.visit(ctx.param);
+    var returnType = typeVisitor.visit(ctx.returnType);
+
+    boolean isNot = ctx.NOT() != null;
+
+    var methodType = new MethodType(ctx.ID().getText(), param, returnType, isNot);
+    return new Pair<>(null, methodType);
+  }
+}

### 2020-02-17 src/main/java/parsing/MethodVisitor.java

@@ -4,12 +4,13 @@ import static util.TypeResolverUtils.isNotValidSubtype;
 import static util.TypeResolverUtils.reportError;
 
 import domain.Method;
-import domain.Parameter;
+import domain.MethodParameter;
 import domain.Program;
 import domain.type.BasicType;
 import domain.type.BooleanType;
 import jsf.jsfBaseVisitor;
 import jsf.jsfParser.MethodDeclContext;
+import org.antlr.v4.runtime.misc.Pair;
 
 public class MethodVisitor extends jsfBaseVisitor<Method> {
 
@@ -19,12 +20,14 @@ public class MethodVisitor extends jsfBaseVisitor<Method> {
 
     final var returnType = typeVisitor.visit(ctx.returntype);
 
-    Parameter parameter;
+    MethodParameter parameter;
     if (ctx.paramname != null && ctx.paramtype != null) {
       final var name = ctx.paramname.getText();
-      parameter = new Parameter(name, typeVisitor.visit(ctx.paramtype), ctx.paramname);
+
+      var type = new MethodTypeVisitor().visit(ctx.paramtype);
+      parameter = new MethodParameter(name, type, ctx.paramname);
     } else {
-      parameter = new Parameter("", new BooleanType(BasicType.VOID, false), ctx.paramname);
+      parameter = new MethodParameter("", new Pair<>(new BooleanType(BasicType.VOID, false), null), ctx.paramname);
     }
 
     final var name = ctx.name.getText();
@@ -43,8 +46,9 @@ public class MethodVisitor extends jsfBaseVisitor<Method> {
     final var expressionTypes = method.getExpression().accept(expressionVisitor);
 
     if (isNotValidSubtype(method.getReturnType().getSet(), expressionTypes)) {
+      // TODO Improve message
       reportError("Return type of expression of method " + method.getName() + " does not match: "
-              + expressionTypes + " != " + method.getReturnType(), method.getToken());
+              + expressionTypes + " != " + method.getReturnType().getSet(), method.getToken());
     }
   }
 }

### 2020-02-17 src/main/java/parsing/ProgramVisitor.java

@@ -28,7 +28,7 @@ public class ProgramVisitor extends jsfBaseVisitor<Program> {
       }
     });
 
-    return new Program(classes);
+    return new Program(classes, ctx.expression());
   }
 
   /**
@@ -43,7 +43,7 @@ public class ProgramVisitor extends jsfBaseVisitor<Program> {
     } while (hasChanged);
 
     for (final var c : program.getClasses().values()) {
-      if (!c.isResolved()) {
+      if (c.isNotResolved()) {
         // TODO make error message nicer
         throw new RuntimeException("Cannot resolve the types");
       }
@@ -56,6 +56,7 @@ public class ProgramVisitor extends jsfBaseVisitor<Program> {
               + " does not exist", c.getToken());
         } else {
           final var superClass = program.getClasses().get(c.getSuperName());
+
           for (final var field : superClass.getFields().values()) {
             if (c.getFields().get(field.getName()) != null) {
               reportError("Overwriting field of super class", c.getFields().get(field.getName()).getToken());
@@ -63,8 +64,15 @@ public class ProgramVisitor extends jsfBaseVisitor<Program> {
               c.addField(field);
             }
           }
-        }
 
+          for (final var method : superClass.getMethods().values()) {
+            if (c.getMethods().get(method.getName()) != null) {
+              reportError("Overwriting field of super class", c.getMethods().get(method.getName()).getToken());
+            } else {
+              c.addMethod(method);
+            }
+          }
+        }
       }
     }
 
@@ -74,6 +82,10 @@ public class ProgramVisitor extends jsfBaseVisitor<Program> {
 
     final var classVisitor = new ClassVisitor();
     program.getClasses().values().forEach(c -> classVisitor.visit(c, program));
+
+    // TODO check null handled correctly
+    var expressionVisitor = new ExpressionVisitor(program, null);
+    expressionVisitor.visit(program.getExpression());
   }
 
   private void generateRelation(final Collection<Class> classes, final Program program) {
@@ -92,16 +104,23 @@ public class ProgramVisitor extends jsfBaseVisitor<Program> {
 
   private boolean resolve(final Program program) {
     var hasChanged = false;
-    var allResolved = true;
 
     for (final var c : program.getClasses().values()) {
-      if (!c.isResolved()) {
+      if (c.isNotResolved()) {
+        var allResolved = true;
+
         for (final var field : c.getFields().values()) {
           for (final var type : field.getType().getTypes()) {
             if (type instanceof ClassType) {
[…]

### 2020-02-17 src/main/java/parsing/TypeVisitor.java

@@ -19,7 +19,7 @@ public class TypeVisitor extends jsfBaseVisitor<BooleanType> {
 
   @Override
   public BooleanType visitNotClass(final jsfParser.NotClassContext ctx) {
-    return new BooleanType(getFromTypeName(ctx.getText()), true);
+    return new BooleanType(getFromTypeName(ctx.classlbl.getText()), true);
   }
 
   @Override

### 2020-02-17 src/main/java/util/TypeHelper.java

@@ -2,6 +2,7 @@ package util;
 
 import domain.Class;
 import domain.Field;
+import domain.Method;
 import domain.Program;
 import domain.type.BasicType;
 import domain.type.ClassType;
@@ -68,16 +69,31 @@ public final class TypeHelper {
             }
           }
 
-          isValid = canAdd(existingClass, field);
+          isValid = canAddField(existingClass, field);
         }
+
+        if (isValid) {
+          for (final var method : newClass.getMethods().values()) {
+            isValid = canAddMethod(existingClass, method);
+          }
+        }
+
         if (isValid) {
           subclasses.add(existing);
         }
 
+
         isValid = true;
         for (final var field : existingClass.getFields().values()) {
-          isValid = canAdd(newClass, field);
+          isValid = canAddField(newClass, field);
         }
+
+        if (isValid) {
+          for (final var method : existingClass.getMethods().values()) {
+            isValid = canAddMethod(newClass, method);
+          }
+        }
+
         if (isValid) {
           SUB_CLASSES.get(existing).add(newClass.getType());
         }
@@ -89,7 +105,7 @@ public final class TypeHelper {
     return true;
   }
 
-  private static boolean canAdd(final Class c, final Field field) {
+  private static boolean canAddField(final Class c, final Field field) {
     if (c.getFields().containsKey(field.getName())) {
       final var newSet = new HashSet<>(c.getFields().get(field.getName()).getType().getSet());
 
@@ -102,5 +118,19 @@ public final class TypeHelper {
       return false;
     }
     return true;
+  }
+
+  private static boolean canAddMethod(final Class c, final Method method) {
+    if (c.getMethods().containsKey(method.getName())) {
+      var existingMethod = c.getMethods().get(method.getName());
+
+      if (! existingMethod.getReturnType().equals(method.getReturnType())) {
+       return false;
+      }
+
+      return existingMethod.getParameter().equals(method.getParameter());
+    } else {
+      return false;
+    }
   }
 }

### 2020-02-17 src/main/java/util/TypeResolverUtils.java

@@ -3,13 +3,14 @@ package util;
 import com.google.common.primitives.Doubles;
 import com.google.common.primitives.Floats;
 import com.google.common.primitives.Ints;
+import com.google.common.primitives.Longs;
 import domain.type.BasicType;
 import domain.type.BooleanType;
 import domain.type.ClassType;
 import domain.type.Type;
 
+import java.lang.Byte;
 import java.util.Arrays;
-import java.util.HashSet;
 import java.util.Optional;
 import java.util.Set;
 
@@ -47,9 +48,9 @@ public final class TypeResolverUtils {
   /**
    * Given Boolean type t1, check if Boolean type t2 is a subtype by checking if set intersection is empty.
    *
-   * @param   t1 Boolean type of super type
-   * @param   t2 Boolean type which is checked if subtype of t1
-   * @return  true if t2 is not a valid subtype of t1
+   * @param t1 Boolean type of super type
+   * @param t2 Boolean type which is checked if subtype of t1
+   * @return true if t2 is not a valid subtype of t1
    */
   public static boolean isNotValidSubtype(final BooleanType t1, final BooleanType t2) {
     return !t1.getSet().containsAll(t2.getSet());
@@ -59,9 +60,9 @@ public final class TypeResolverUtils {
    * Given a Boolean type t, check if the type represented by the set t1 is a subtype by checking if set intersection
    * is empty. The set passed is not modified but copied.
    *
-   * @param t1   Boolean type of super type
-   * @param t2  Set of types representing the type which is being checked
-   * @return    true if type represented by t1 is not a subtype of t
+   * @param t1 Boolean type of super type
+   * @param t2 Set of types representing the type which is being checked
+   * @return true if type represented by t1 is not a subtype of t
    */
   public static boolean isNotValidSubtype(final BooleanType t1, final Set<Type> t2) {
     return !t1.getSet().containsAll(t2);
@@ -70,9 +71,10 @@ public final class TypeResolverUtils {
   /**
    * Given a set representing type t, and a set representing type t1, check if t1 is a subtype of t by checking if set
    * intersection is empty. The sets passed are not modified by copied.
-   * @param t1   Set of types representing super type
-   * @param t2  Set of types representing the type being checked
-   * @return    true if type represented by t1 is not a subtype of t2
+   *
+   * @param t1 Set of types representing super type
[…]

### 2020-02-17 conference/paper.tex
add template latex file for conference paper

\documentclass[10pt,a4paper,twocolumn]{article}

\usepackage[cm]{fullpage}
\usepackage{newtxtext}
\usepackage{newtxmath}
\usepackage[hyphens]{url}
\usepackage{minted}
\usepackage[english]{babel}
\usepackage{cite}

\title{JSF}
\author{Artem Usov}
\date{February 2020}

\begin{document}

\maketitle

\section{Introduction}

\end{document}

### 2020-02-18 conference/paper.tex
use lncs template, get started on paper

@@
-\documentclass[10pt,a4paper,twocolumn]{article}
-
-\usepackage[cm]{fullpage}
-\usepackage{newtxtext}
-\usepackage{newtxmath}
-\usepackage[hyphens]{url}
-\usepackage{minted}
-\usepackage[english]{babel}
-\usepackage{cite}
-
-\title{JSF}
-\author{Artem Usov}
-\date{February 2020}
+\documentclass[runningheads]{llncs}
+\usepackage{graphicx}
+% \renewcommand\UrlFont{\color{blue}\rmfamily}
 
@@
 
+\title{SFJ: An implementation of Semantic Featherweight Java}
+%\titlerunning{Abbreviated paper title}
+
+\author{Artem Usov \and Ornela Dardha\orcidID{0000-0001-9927-7875}}
+
+%\authorrunning{F. Author et al.}
+
+\institute{School of Computing Science, University of Glasgow, United Kingdom \\
+    \email{\{artem.usov, ornela.dardha\}@@glasgow.ac.uk}}
+
 \maketitle
+
+\begin{abstract}
+    The abstract should briefly summarize the contents of the paper in 15--250 words.
+
+    \keywords{nominal and structural subtyping  \and Featherweight Java 
+    \and object-oriented languages \and semantic subtyping \and type theory.}
+\end{abstract}
 
@@
 
+A typing system for a programming language is a set of deduction rules that allow
+
+\begin{equation}
+    \tau ::= \alpha \mid \mu
+\end{equation}
+
+Since we want our types to represent sets of values, we restrict our types to finite trees whose 
+leaves are constants with no cycles. For example a recursive type $\alpha = [a : \alpha]$ would
+be an infinite tree \textbf{new} $C$(\textbf{new} $C$($\cdots{}$)). Similarly the types $\alpha = 
+[b: \beta]$, $\beta = [a = \alpha]$ would also be impossible to instantiate.
+
+Therefore, to not allow this, when processing the AST of a program, we do not allow the type of 
+the class we are defining to be a field and we mark any classes with only basic types in it's fields
+as \textit{resolved} and otherwise if it contains a class type, we mark it as \textit{unresolved}. 
+After processing the whole AST, we perform the following algorithm to decide whether the type 
+definitions in the program are valid.
+
+\begin{verbatim}
+boolean resolutionOccured = false
+do 
+    for each class that is unresolved:
+        boolean resolved = true
+
+        for each field that is a class type:
+            if class type is not resolved 
+                resolved = false
+
+        if resolved == true
+            set class to be resolved
+            resolutionOccured = true
+while resolutionOccured == true
+
+if not all classes are resolved
+    program contains invalid type definition
+\end{verbatim}
+
+\subsubsection{Sample Heading (Third Level)}
+Only two levels of headings should be numbered. Lower level headings remain unnumbered; they are 
+formatted as run-in headings.
+
+\paragraph{Sample Heading (Fourth Level)}
+The contribution should contain no more than four levels of headings. Table~\ref{tab1} gives a 
+summary of all heading levels.
+
+\begin{table}
+    \caption{Table captions should be placed above the tables.}
+    \label{tab1}
+    \begin{tabular}{|l|l|l|}
+        \hline
+        Heading level     & Example                                          & Font size and style \\
+        \hline
+        Title (centered)  & {\Large\bfseries Lecture Notes}                  & 14 point, bold      \\
+        1st-level heading & {\large\bfseries 1 Introduction}                 & 12 point, bold      \\
+        2nd-level heading & {\bfseries 2.1 Printing Area}                    & 10 point, bold      \\
+        3rd-level heading & {\bfseries Run-in Heading in Bold.} Text follows & 10 point, bold      \\
+        4th-level heading & {\itshape Lowest Level Heading.} Text follows    & 10 point, italic    \\
+        \hline
+    \end{tabular}
+\end{table}
+
+\noindent Displayed equations are centered and set on a separateline.
+
+\begin{equation}
+    x + y = z
+\end{equation}
+Please try to avoid rasterized images for line-art diagrams and schemas. Whenever possible, use 
+vector graphics instead.
+
+%\begin{figure}
+%   \includegraphics[width=\textwidth]{fig1.eps}
+%  \caption{A figure caption is always placed below the illustration.
+%     Please note that short captions are centered, while long ones are
+%     justified by the macro package automatically.} \label{fig1}
+%\end{figure}
+
+\begin{theorem}
+    This is a sample theorem. The run-in heading is set in bold, while
+    the following text appears in italics. Definitions, lemmas,
+    propositions, and corollaries are styled the same way.
+\end{theorem}
+%
+% the environments 'definition', 'lemma', 'proposition', 'corollary',
+% 'remark', and 'example' are defined in the LLNCS documentclass as well.
+%
+\begin{proof}
+    Proofs, examples, and remarks have the initial word in italics,
+    while the following text appears in normal font.
+\end{proof}
+For citations of references, we prefer the use of square brackets
+and consecutive numbers. Citations using labels or the author/year
+convention are also acceptable. The following bibliography provides
+a sample reference list with entries for journal
+articles~\cite{Dardha2017}.
+
+\bibliographystyle{splncs04}
+\bibliography{paper}
+
 \end{document}

### 2020-02-19 conference/paper.tex
add algorithms of checking type trees and creating relation

@@
 \usepackage{graphicx}
+\usepackage{amsmath}
 % \renewcommand\UrlFont{\color{blue}\rmfamily}
@@
 do 
-    for each class that is unresolved:
+    for class that is unresolved:
         boolean resolved = true
 
-        for each field that is a class type:
+        for field that contains a class type:
             if class type is not resolved 
@@
 
-        if resolved == true
+        if resolved is true
             set class to be resolved
             resolutionOccured = true
-while resolutionOccured == true
+while resolutionOccured is true
 
-if not all classes are resolved
+if all classes are not resolved
     program contains invalid type definition
+\end{verbatim}
+
+Given that we now know that all the types in the program are valid, we can create the subtyping 
+relation which will be a map of types to a set of its subtypes. For every program, we assume the 
+initial subtyping relation for all of our basic types. We would like to highlight that Int is not
+a subtype of Float as it cannot represent the whole domain of Int accurately therefore the domain is 
+not fully contained in Float. Similarly so for Double and Long.
+
+\begin{flalign*}
+    Boolean &= \left\{Boolean \right\} &\\
+    Double  &= \left\{Double, Float, Int, Short, Byte\right\} &\\
+    Float &= \left\{Float, Short, Byte\right\} &\\
+    Long &= \left\{Long, Int, Short, Byte\right\} &\\
+    Int &= \left\{Int, Short, Byte\right\} &\\
+    Short &= \left\{Short, Byte\right\} &\\
+    Byte &= \left\{Byte\right\} & 
+\end{flalign*}
+
+Knowing that all our types are finite trees with leaves as constants and given this initial
+relation, we can now confidently create a subtyping relation for all class types using the following
+algorithm.
+
+\begin{verbatim}
+function generateRelation(classes):
+    List<class> untyped = []
+
+    for class in classes:
+        if addClass(class) is false:
+            untyped.add(class)
+
+    if untyped is not []:
+        generateRelation(untyped)
+
+function addClass(class):
+    for existing class type in relation:
+        if tryAddSubtype(class, existingClass) is false:
+            return false
+
+        tryAddSubtype(existingClass, class)
+
+        add class to its own subtype relation
+
+function tryAddSubtype(class, other):
+    boolean flag = true
+
+    for field in class:
+        if field contains type not in relation:
+            return false
+
+        if other does not have field:
+            flag = false
+        else:
+            if other.field.types not fully contains field.types:
+                flag = false
+
+    for method in class:
+        if method contains type not in relation:
+            return false
+
+        if other does not have method:
+            flag = false
+        else:
+            if other.method.types not fully contains method.types:
+                flag = false
+
+    if flag == true:
+        add class to other subtype relation
 \end{verbatim}

### 2020-02-19 conference/paper.tex
packages
running authors and title
restructuring the paper

@@
 \documentclass[runningheads]{llncs}
+%%%%%%%%%%%%%%%%%%%%%%%%
+%PACKAGES
+%%%%%%%%%%%%%%%%%%%%%%%%
+\usepackage[latin1]{inputenc}
+\usepackage[english]{babel}
+
+\usepackage{amsmath}
+\usepackage{amsfonts}
+\usepackage{amssymb}
+%\usepackage{marvosym}
+%\usepackage{txfonts}
+%\usepackage{multicol}
+\usepackage{proof-dashed}
 \usepackage{graphicx}
-\usepackage{amsmath}
+\usepackage{listings}
+
+% Non standard
+\usepackage{varacca}
+\input semmacros
 % \renewcommand\UrlFont{\color{blue}\rmfamily}
@@
 \title{SFJ: An implementation of Semantic Featherweight Java}
-%\titlerunning{Abbreviated paper title}
+\titlerunning{SFJ}
 
 \author{Artem Usov \and Ornela Dardha\orcidID{0000-0001-9927-7875}}
-
-%\authorrunning{F. Author et al.}
+\authorrunning{A. Usov and O. Dardha}
 
@@
 \end{equation}
+
+\section{Semantic Subtyping for Featherweight Java}
+
+\section{SFJ: Design and Implementation}
 
@@
 
-\subsubsection{Sample Heading (Third Level)}
-Only two levels of headings should be numbered. Lower level headings remain unnumbered; they are 
-formatted as run-in headings.
-
-\paragraph{Sample Heading (Fourth Level)}
-The contribution should contain no more than four levels of headings. Table~\ref{tab1} gives a 
-summary of all heading levels.
-
-\begin{table}
-    \caption{Table captions should be placed above the tables.}
-    \label{tab1}
-    \begin{tabular}{|l|l|l|}
-        \hline
-        Heading level     & Example                                          & Font size and style \\
-        \hline
-        Title (centered)  & {\Large\bfseries Lecture Notes}                  & 14 point, bold      \\
-        1st-level heading & {\large\bfseries 1 Introduction}                 & 12 point, bold      \\
-        2nd-level heading & {\bfseries 2.1 Printing Area}                    & 10 point, bold      \\
-        3rd-level heading & {\bfseries Run-in Heading in Bold.} Text follows & 10 point, bold      \\
-        4th-level heading & {\itshape Lowest Level Heading.} Text follows    & 10 point, italic    \\
-        \hline
-    \end{tabular}
-\end{table}
-
-\noindent Displayed equations are centered and set on a separateline.
-
-\begin{equation}
-    x + y = z
-\end{equation}
-Please try to avoid rasterized images for line-art diagrams and schemas. Whenever possible, use 
-vector graphics instead.
-
-%\begin{figure}
-%   \includegraphics[width=\textwidth]{fig1.eps}
-%  \caption{A figure caption is always placed below the illustration.
-%     Please note that short captions are centered, while long ones are
-%     justified by the macro package automatically.} \label{fig1}
-%\end{figure}
-
-\begin{theorem}
-    This is a sample theorem. The run-in heading is set in bold, while
-    the following text appears in italics. Definitions, lemmas,
-    propositions, and corollaries are styled the same way.
-\end{theorem}
-%
-% the environments 'definition', 'lemma', 'proposition', 'corollary',
-% 'remark', and 'example' are defined in the LLNCS documentclass as well.
-%
-\begin{proof}
-    Proofs, examples, and remarks have the initial word in italics,
-    while the following text appears in normal font.
-\end{proof}
-For citations of references, we prefer the use of square brackets
-and consecutive numbers. Citations using labels or the author/year
-convention are also acceptable. The following bibliography provides
+The following bibliography provides
 a sample reference list with entries for journal
@@
 
+\section{Related Work and Conclusion}
+
+%% References
 \bibliographystyle{splncs04}

### 2020-02-19 conference/paper.tex
getting started on intro

@@
 %%%%%%%%%%%%%%%%%%%%%%%%
-\usepackage[latin1]{inputenc}
+\usepackage[utf8]{inputenc}
 \usepackage[english]{babel}
@@
 \usepackage{listings}
+\usepackage{hyperref}
+\usepackage{stmaryrd}
 
 % Non standard
-\usepackage{varacca}
+%\usepackage{varacca}
 \input semmacros
-% \renewcommand\UrlFont{\color{blue}\rmfamily}
 
@@
 
-\title{SFJ: An implementation of Semantic Featherweight Java}
+\title{SFJ: An implementation of Semantic Featherweight Java\thanks{Research supported by the UK EPSRC grant EP/K034413/1, ``From Data Types to Session Types: A Basis for Concurrency and Distribution (ABCD)'', and by the EU HORIZON 2020 MSCA RISE project 778233 ``BehAPI: Behavioural Application Program Interfaces''.}}
 \titlerunning{SFJ}
@@
 \institute{School of Computing Science, University of Glasgow, United Kingdom \\
-    \email{\{artem.usov, ornela.dardha\}@@glasgow.ac.uk}}
+    \email{\{artem.usov, ornela.dardha\}@glasgow.ac.uk}}
 
@@
 \begin{abstract}
-    The abstract should briefly summarize the contents of the paper in 15--250 words.
+Subtyping is a key notion in programming languages, as it allows more flexibility in coding.
+There are two approaches to defining subtyping relations: the \emph{syntactic} and the \emph{semantic} approach.
+In the latter, one starts by defining a model of the programming language at hand and an interpretation of its types as subsets of such model.
+Then, subtyping is defined as \emph{set inclusion of types} denoting sets.
+An orthogonal issue, typical of object-oriented languages, is the \emph{nominal} and \emph{structural} subtyping.
+Dardha \emph{et al.} \cite{Dardha2013,Dardha2017} combined structural subtyping and boolean types, on one hand, with nominal subtyping, on the other, in Featherweight Java (FJ), by thus exploiting the benefits of both approaches.
 
-    \keywords{nominal and structural subtyping  \and Featherweight Java 
-    \and object-oriented languages \and semantic subtyping \and type theory.}
+However, these benefits were only presented and exemplified at a theoretical level and not exploited practically. 
+At last, in this work we have implemented SFJ---Semantic Featherweight Java, which features structural subtyping and boolean types as well as nominal subtyping, typical of FJ itself.
+The benefits of SFJ, also illustrated in the accompanied video, show how static typechecking of boolean types and semantic subtyping gives higher guarantees of program correctness, more flexibility and compactness of program writing.
+\keywords{Nominal subtyping \and Structural subtyping  \and Semantic Featherweight Java \and Object-oriented languages \and Boolean types \and Type theory.}
 \end{abstract}
@@
 \section{Introduction}
+There are two approaches to defining subtyping relations: the \emph{syntactic} and the \emph{semantic} approach.
+In the latter, one starts by defining a model of the programming language at hand and an interpretation of its types as subsets of such model.
+The syntactic subtyping is more mainstream in programming languages design and implementation, and the relation is given by means of a set of formal deductive subtyping rules.
+The semantic subtyping is less know and the relation is given by inclusion of sets denoting types.
+One starts from a formal model of the programming language at hand and defines an interpretation, or semantics, of types as subsets of this model; subtyping naturally is defined as inclusion of these sets, hence the semantic adjective attached to it.
+Orthogonally, especially for object-oriented languages, there are two approaches to defining subtyping relation: the \emph{nominal} and the \emph{structural} approach.
+The nominal subtyping relation is based on \emph{declarations} by the developer and is \emph{name}-based, hance the adjective nominal: $A$ is a subtype of $B$ if and only if it is declared to be so, that is if the class $A$ extends (or implements) the class (or interface) $B$.
+The structural subtyping relation instead is based on the \emph{structure} of a class, its fields and methods: a class $A$ is a subtype of a class $B$ if and only if the fields and methods of $A$ are a superset of the fields and methods of $B$, and their types in $A$ are subtypes of the types in $B$.
+At this point it is not surprising that the nominal subtyping aligns well with the syntactic approach and hence is more well-known than the structural subtyping which aligns well with the semantic approach.
+
+Dardha \emph{et al.} \cite{Dardha2013,Dardha2017} define semantic subtyping for \emph{Featherweight Java} (FJ) \cite{featherweight}, which is the functional fragment of the Java language.
+This approach to defining subtyping enriches FJ with boolean connectives and naturally allows for structural subtyping in the language.
+However, the nominal subtyping, which one would  commonly expect to be part of FJ, can be achieved though use of semantic subtyping and structural subtyping as shown in the case studies in Dardha \emph{et al.} \cite[\S 8.4]{Dardha2017}.
+
+\begin{example}[The Polygons]
+We will use the \emph{Polygons} as our running example, both in the paper and in the tool video.
+Consider the set of polygons, such as triangles, rectangles, rumbles given by a class hierarchy.
+Our task is to define a method \emph{diagonal} that takes a polygon and returns the length of its longest diagonal.
+This method makes sense only if the polygon passed to it has at least four sides, hence triangles are excluded.
+In Java this could be implemented in the following ways:
+$$
+\begin{array}{l}
+\synclass Polygon \ \{ \cdots\}
+\\
+\syndecl{Triangle}{Polygon} {\cdots}
+\\
+\syndecl{Other\_Polygons}{Polygon} {
+\\
+	\qquad \mathbf{real}\ diagonal(Other\_Polygons \ p)\ \{\ldots \}
+	\\
+	\hspace{3.5cm} \ldots
+	\\
+}
+\\
+\syndecl{Rectangle}{Other\_Polygons}{\cdots}
+\\
+\syndecl{Rumble}{Other\_Polygons} {\cdots}
+\end{array}
+$$
+Or by means of an interface:
+$$
+\begin{array}{l}
+public \ \syninterface Diagonal\ \{
+	\\
+	\qquad \mathbf{real}\ diagonal(Polygon \ p);
+	\\
+\}
+\\
+\synclass Polygon \ \{\cdots\}
+\\
+\syndecl{Triangle}{Polygon} {\cdots}
+\\
+\synidecl{Rectangle}{Polygon} \synimpl Diagonal \ \{\cdots\}
+\\
+\synidecl{Rumble}{Polygon} \synimpl Diagonal \ \{\cdots\}
+\\
+\hspace{4.5cm} \ldots
+\end{array}
+$$
+Now, suppose that our class hierarchy is such that \emph{Polygon} is the parent class and all other geometric figure classes that extend \emph{Polygon}, which is exactly how one would naturally define the set of polygons.
+If we think of the class hierarchy as given and part of legacy code, then again the natural way to solve this in Java is by defining the method \emph{diagonal} in the class \emph{Polygon} and using an \textbf{instanceof}, for example inside a \textbf{try-catch} that throws an exception at run time if the argument passed to the method is a triangle.
+
+In our framework we propose a different approach which allows the class hierarchy to be defined naturally as above and does not require changes to it, where this maybe is not viable, and yet it does not require exceptions at run time, but mere static typechecking.
+
+Our framework combines boolean types with semantic subtyping for a more elegant solution: it is enough to define a method \emph{diagonal} that has an argument of type \emph{Polygon} and not \emph{Triangle} thus allowing the typechecker to check at compile time the restrictions on types:
+$$
+\begin{array}{l}
+\syndecl{Polygon}{Object} {\cdots}
+\\
+\syndecl{Triangle}{Polygon} {\cdots}
+\\
+\syndecl{Rectangle}{Polygon} {\cdots}
+\\
+\syndecl{Rumble}{Polygon} {\cdots}
+\\
+\hspace{2.5cm} \ldots
+\\
+\syndecl{Diagonal}{Object} {
+	\\
+	\qquad \mathbf{real}\ diagonal(\textit{Polygon} \text{ and not}  \textit{Triangle} \ p)\{\ldots\}
+	\\
+}
+\end{array}
+$$
+Finally,  we can call the method \emph{diagonal} on an argument of type \emph{Polygon}: if the polygon is not a \emph{Triangle}, then the method returns the length of its longest diagonal; otherwise, if the polygon is a triangle, then there will be a type error at compile time.
+\end{example}
 
@@
 
-\section{Semantic Subtyping for Featherweight Java}
+\section{Background}
 
@@
 \section{Related Work and Conclusion}
+The first use of the semantic approach goes back to 15 years ago, as in \cite{Aiken,Damm}.
+Hosoya and Pierce have also adopted this approach in \cite{XML1,XML2,XML3} to define XDuce, an XML-oriented language designed specifically to transform XML documents in other XML documents satisfying certain properties.
+The values of this language are fragments of XML documents; types are interpreted as sets of documents, more precisely as sets of values.
+The subtyping relation is established as inclusion of these sets.
+The type-system contains boolean types, product types and recursive types.
+There are no function types and no functions in the language.
+
+Castagna et al.
+in \cite{gentle,Cas05,FCB08} extend the XDuce language with first-class functions and arrow types and define and implement a higher-order language named $\mathbb{C}$Duce adopting the semantic approach to subtyping.
+The starting point of their framework is a higher-order $\lambda-$calculus with pairs and projections.
+The set of types is extended with intersection, union and negation types interpreted in a set-theoretic way.
+
+The semantic approach can also be applied to the $\pi-$calculus \cite{milner,sangiorgi}.
+Castagna, Varacca and De Nicola in \cite{Cpi} have used this technique to define the $\mathbb{C}\pi$ language, a variant of the asynchronous $\pi-$calculus, where channel types are augmented with boolean connectives interpreted in an obvious way.
 
 %% References
+\newpage
 \bibliographystyle{splncs04}

### 2020-02-20 conference/paper.tex
a pass on the intro

@@
 \usepackage{stmaryrd}
+
+\addto\extrasenglish{%
+  \renewcommand{\sectionautorefname}{\S}%
+}
 
@@
 This approach to defining subtyping enriches FJ with boolean connectives and naturally allows for structural subtyping in the language.
-However, the nominal subtyping, which one would  commonly expect to be part of FJ, can be achieved though use of semantic subtyping and structural subtyping as shown in the case studies in Dardha \emph{et al.} \cite[\S 8.4]{Dardha2017}.
+However, nominal subtyping, which one would  commonly expect to be part of FJ, can be achieved though use of semantic subtyping and structural subtyping as shown in the extensive case studies in Dardha \emph{et al.} \cite[\S 8.4]{Dardha2017}.
 
 \begin{example}[The Polygons]
-We will use the \emph{Polygons} as our running example, both in the paper and in the tool video.
+We will use the \emph{Polygons} as our running example, to illustrate the benefits of the approach by Dardha \emph{et al.} \cite{Dardha2013,Dardha2017}, both in the paper and in the tool video.
 Consider the set of polygons, such as triangles, rectangles, rumbles given by a class hierarchy.
-Our task is to define a method \emph{diagonal} that takes a polygon and returns the length of its longest diagonal.
+We want to define a method \emph{diagonal} that takes a polygon and returns the length of its longest diagonal.
 This method makes sense only if the polygon passed to it has at least four sides, hence triangles are excluded.
@@
 $$
-Now, suppose that our class hierarchy is such that \emph{Polygon} is the parent class and all other geometric figure classes that extend \emph{Polygon}, which is exactly how one would naturally define the set of polygons.
-If we think of the class hierarchy as given and part of legacy code, then again the natural way to solve this in Java is by defining the method \emph{diagonal} in the class \emph{Polygon} and using an \textbf{instanceof}, for example inside a \textbf{try-catch} that throws an exception at run time if the argument passed to the method is a triangle.
 
-In our framework we propose a different approach which allows the class hierarchy to be defined naturally as above and does not require changes to it, where this maybe is not viable, and yet it does not require exceptions at run time, but mere static typechecking.
+Now, suppose that our class hierarchy is such that \emph{Polygon} is the parent class and all other geometric figure classes extend \emph{Polygon}, which is exactly how one would naturally define the set of polygons.
+If we think of the class hierarchy as given and part of legacy code, which cannot be changed, then again the natural way to solve this in Java is by defining the method \emph{diagonal} in the class \emph{Polygon} and using an \textbf{instanceof}, for example inside a \textbf{try-catch} that throws an exception at run time, if the argument passed to the method is a triangle.
 
-Our framework combines boolean types with semantic subtyping for a more elegant solution: it is enough to define a method \emph{diagonal} that has an argument of type \emph{Polygon} and not \emph{Triangle} thus allowing the typechecker to check at compile time the restrictions on types:
+The approach proposed by Dardha \emph{et al.} \cite{Dardha2013,Dardha2017} allows for the class hierarchy to be defined in such a natural way without needing to change it, but on the other hand it does not require handling exceptions at run time,  rather just static typechecking.
+To achieve this, the framework combines boolean types with semantic subtyping: it is enough to define a method \emph{diagonal} that has an argument of type \emph{Polygon} and not \emph{Triangle} thus allowing the typechecker to check at compile time the restrictions on types:
 $$
@@
 	\\
-	\qquad \mathbf{real}\ diagonal(\textit{Polygon} \text{ and not}  \textit{Triangle} \ p)\{\ldots\}
+	\qquad \mathbf{real}\ diagonal(\textit{Polygon} \textbf{ and not }  \textit{Triangle} \ p)\{\ldots\}
 	\\
@@
 $$
-Finally,  we can call the method \emph{diagonal} on an argument of type \emph{Polygon}: if the polygon is not a \emph{Triangle}, then the method returns the length of its longest diagonal; otherwise, if the polygon is a triangle, then there will be a type error at compile time.
+We can now call \emph{diagonal} on an argument of type \emph{Polygon}: if the polygon is \textbf{not} a \emph{Triangle}, then the method returns the length of its longest diagonal; otherwise, if the polygon is a triangle, then there will be a type error at compile time.
 \end{example}
+The combination of boolean types and semantic subtyping gives a more elegant solution, as illustrate by the \emph{Polygons}---of course a toy example used only as a proof of concept of the approach---however, this approach is not yet practical as there are no boolean types or semantic subtyping in Java, hence its benefits unfortunately remained theoretical.
+
+After a few years, finally we can now exploit the benefits of this approach at a practical level: in this paper we present SFJ---Semantic Featherweight Java \autoref{sec:design}, an implementation of Featherweight Java with integrated boolean types and semantic subtyping as well as nominal subtyping, the latter more natural in FJ.
+In SFJ the programmer has a larger, yet simple, and more expressive set of types, by using \textbf{and}, \textbf{not}, \textbf{or} boolean connectives.
+This added expressivity does not however introduce added complexity, but rather the opposite is true as the programmer has an easier, more compact and more elegant way of programming.
+Moreover, as demonstrated in Dardha \emph{et al.} \cite[\S 8.4]{Dardha2017}, the added expressivity allows for standard programming constructs and features in Java, for e.g., lists, to be easily encoded in our framework, and the programmer can choose between semantic or nominal subtyping, making SFJ a conservative extension of FJ.
+
+The technical developments behind semantic subtyping and its properties \autoref{sec:background} are not easy, however, they are completely transparent to the programmer who only needs to know boolean types and the difference between semantic and nominal subtyping and which to choose based on their goals.
+
+\section{Background}
+\label{sec:background}
 
@@
 
-\section{Background}
-
 \section{SFJ: Design and Implementation}
-
+\label{sec:design}
 Since we want our types to represent sets of values, we restrict our types to finite trees whose 
@@
 \section{Related Work and Conclusion}
+\label{sec:conclusion}
 The first use of the semantic approach goes back to 15 years ago, as in \cite{Aiken,Damm}.

### 2020-02-20 conference/paper.tex
finalising intro

@@
 
-Dardha \emph{et al.} \cite{Dardha2013,Dardha2017} define semantic subtyping for \emph{Featherweight Java} (FJ) \cite{featherweight}, which is the functional fragment of the Java language.
-This approach to defining subtyping enriches FJ with boolean connectives and naturally allows for structural subtyping in the language.
-However, nominal subtyping, which one would  commonly expect to be part of FJ, can be achieved though use of semantic subtyping and structural subtyping as shown in the extensive case studies in Dardha \emph{et al.} \cite[\S 8.4]{Dardha2017}.
+Dardha \emph{et al.} \cite{Dardha2013,Dardha2017} define boolean types and semantic subtyping for \emph{Featherweight Java} (FJ) \cite{featherweight}, which is the functional fragment of the Java language.
+This approach to defining subtyping naturally allows for structural subtyping in FJ.
+On the other hand, nominal subtyping, typical of FJ, can be obtained in the framework exploiting boolean types and semantic subtyping, as shown in the extensive case studies in Dardha \emph{et al.} \cite[\S 8.4]{Dardha2017}.
 
@@
 
-After a few years, finally we can now exploit the benefits of this approach at a practical level: in this paper we present SFJ---Semantic Featherweight Java \autoref{sec:design}, an implementation of Featherweight Java with integrated boolean types and semantic subtyping as well as nominal subtyping, the latter more natural in FJ.
-In SFJ the programmer has a larger, yet simple, and more expressive set of types, by using \textbf{and}, \textbf{not}, \textbf{or} boolean connectives.
+After a few years, finally we can now exploit the benefits of this approach at a practical level: in this paper we present SFJ---Semantic Featherweight Java \autoref{sec:design}, an implementation of Featherweight Java with integrated boolean types and semantic subtyping.
+In SFJ the programmer has a larger and more expressive set of types, by using \textbf{and}, \textbf{not}, \textbf{or} boolean connectives.
 This added expressivity does not however introduce added complexity, but rather the opposite is true as the programmer has an easier, more compact and more elegant way of programming.
-Moreover, as demonstrated in Dardha \emph{et al.} \cite[\S 8.4]{Dardha2017}, the added expressivity allows for standard programming constructs and features in Java, for e.g., lists, to be easily encoded in our framework, and the programmer can choose between semantic or nominal subtyping, making SFJ a conservative extension of FJ.
+Moreover, as demonstrated in Dardha \emph{et al.} \cite[\S 8.4]{Dardha2017}, the added expressivity allows for standard programming constructs and features in Java, for e.g., lists, or overloading classes (via multimethods below), to be easily encoded in our framework.
+Another benefit of semantic subtyping is that it can allow a natural integration of structural and nominal subtyping which the programmer has the option to choose, making SFJ a conservative extension of FJ.
 
-The technical developments behind semantic subtyping and its properties \autoref{sec:background} are not easy, however, they are completely transparent to the programmer who only needs to know boolean types and the difference between semantic and nominal subtyping and which to choose based on their goals.
+\paragraph{On multimethods}
+Since FJ is a core language, some features of the full Java are removed: overloading methods is one of them.
+In our framework, leveraging the expressivity of boolean connectives and semantic subtyping, many of these features, and language constructs, e.g., lists, are restored: overloading methods is one of them.
+We can thus model \emph{multimethods}, \cite{BC97}, which according to the authors is ``\emph{very clean and easy to understand [...] it would be the best solution for a brand new language}".
+As an example Dardha \emph{et al.}  \cite{Dardha2013,Dardha2017}, consider the following class declarations:
+$$
+\begin{array}{ll}
+\syndecl{A}{\mathit{Object}} {
+&
+\qquad
+\syndecl{B}{A} {
+\\
+	%\qquad\quad \ldots
+%&
+	%\qquad\qquad\quad \ldots
+%\\
+	\qquad\quad \mathbf{int}\ \mathit{length}\ (\mathbf{string}\ s)\{\ \ldots\ \}\qquad
+&
+	\qquad\qquad\quad \mathbf{int}\ \mathit{length}\ (\mathbf{int}\ n)\{\ \ldots\ \}
+\\
+}
+&
+\qquad }
+\end{array}
+$$
+Method {\it length} has type $\mathbf{string} \rightarrow \mathbf{int}$ in $A$.
+However, in $B$ it has type $(\mathbf{string} \rightarrow \mathbf{int}) \synwedge (\mathbf{int} \rightarrow \mathbf{int})$,
+which can be simplified as $(\mathbf{string} \synvee \mathbf{int}) \rightarrow \mathbf{int}$.
 
-\section{Background}
-\label{sec:background}
-
-A typing system for a programming language is a set of deduction rules that allow
-
-\begin{equation}
-    \tau ::= \alpha \mid \mu
-\end{equation}
+The technical developments behind semantic subtyping and its properties are not easy, however, they are completely transparent to the programmer.
+Due to space limit we will remove theoretical details from this submission,as they are not necessary to understand the implementation of SFJ which is the focus of this paper.
+We refer the reader to the relevant work \cite{Dardha2013,Dardha2017} and will include the background on semantic subtyping in a possible full paper.
 

### 2020-02-20 conference/paper.tex
related work and conclusion done

@@
 The technical developments behind semantic subtyping and its properties are not easy, however, they are completely transparent to the programmer.
-Due to space limit we will remove theoretical details from this submission,as they are not necessary to understand the implementation of SFJ which is the focus of this paper.
-We refer the reader to the relevant work \cite{Dardha2013,Dardha2017} and will include the background on semantic subtyping in a possible full paper.
+Due to space limit we will remove theoretical details from this submission, as they are not necessary to understand the design and implementation of SFJ, which is the focus of this paper.
+We refer the reader to the relevant work \cite{Dardha2013,Dardha2017} and we will include the background on semantic subtyping in a possible full paper.
 
@@
 \label{sec:conclusion}
-The first use of the semantic approach goes back to 15 years ago, as in \cite{Aiken,Damm}.
-Hosoya and Pierce have also adopted this approach in \cite{XML1,XML2,XML3} to define XDuce, an XML-oriented language designed specifically to transform XML documents in other XML documents satisfying certain properties.
-The values of this language are fragments of XML documents; types are interpreted as sets of documents, more precisely as sets of values.
-The subtyping relation is established as inclusion of these sets.
-The type-system contains boolean types, product types and recursive types.
-There are no function types and no functions in the language.
+Semantic subtyping approach goes back to more than two decades ago \cite{Aiken,Damm}.
+Notable lines of work include:
+Hosoya and Pierce \cite{XML1,XML2,XML3} who define XDuce, an XML-oriented language designed specifically to transform XML documents in other XML documents satisfying certain properties.
+Castagna \emph{et al.} \cite{gentle,Cas05,FCB08} extend XDuce with first-class functions and arrow types and implement it as $\mathbb{C}$Duce.
+The starting point of their framework is a higher-order $\lambda-$calculus with pairs and projections.
+Castagna \emph{et al.} \cite{Cpi} define $\mathbb{C}\pi$, a variant of the asynchronous $\pi-$calculus \cite{sangiorgi}, where channel types are augmented with boolean connectives.
+Ancona and Lagorio~\cite{AL10} define subtyping for infinite types coinductively by using union and object type constructors, where types are interpreted as sets of value of the language.
+Bonsangue \emph{et al.}~\cite{BRABR14} study a coalgebraic approach to coinductive types and define a set-theoretic interpretation of coinductive types with union types.
+Pearce \cite{Pearce19} defines semantic subtyping for rewriting rules in the Whiley Rewrite Language and for a flow-typing calculus  \cite{Pearce13}.
 
-Castagna et al.
-in \cite{gentle,Cas05,FCB08} extend the XDuce language with first-class functions and arrow types and define and implement a higher-order language named $\mathbb{C}$Duce adopting the semantic approach to subtyping.
-The starting point of their framework is a higher-order $\lambda-$calculus with pairs and projections.
-The set of types is extended with intersection, union and negation types interpreted in a set-theoretic way.
-
-The semantic approach can also be applied to the $\pi-$calculus \cite{milner,sangiorgi}.
-Castagna, Varacca and De Nicola in \cite{Cpi} have used this technique to define the $\mathbb{C}\pi$ language, a variant of the asynchronous $\pi-$calculus, where channel types are augmented with boolean connectives interpreted in an obvious way.
+To conclude, in this paper we presented the design and implementation of SFJ, an extension of Featherweight Java with boolean types and semantic subtyping, which combined allow for both structural and nominal subtyping in FJ as well as restore standard Java constructs for e.g., lists and features for e.g., overloading which were not present in FJ.
 

### 2020-02-25 conference/paper.tex
reduce spacing in algorithms, add further explanations and details

@@
 }
+
+\makeatletter
+\def\verbatim@font{\linespread{0.75}\normalfont\ttfamily}
+\makeatother
 
@@
 At last, in this work we have implemented SFJ---Semantic Featherweight Java, which features structural subtyping and boolean types as well as nominal subtyping, typical of FJ itself.
-The benefits of SFJ, also illustrated in the accompanied video, show how static typechecking of boolean types and semantic subtyping gives higher guarantees of program correctness, more flexibility and compactness of program writing.
+The benefits of SFJ, also illustrated in the accompanied video, show how static type checking of boolean types and semantic subtyping gives higher guarantees of program correctness, more flexibility and compactness of program writing.
 \keywords{Nominal subtyping \and Structural subtyping  \and Semantic Featherweight Java \and Object-oriented languages \and Boolean types \and Type theory.}
@@
 Orthogonally, especially for object-oriented languages, there are two approaches to defining subtyping relation: the \emph{nominal} and the \emph{structural} approach.
-The nominal subtyping relation is based on \emph{declarations} by the developer and is \emph{name}-based, hance the adjective nominal: $A$ is a subtype of $B$ if and only if it is declared to be so, that is if the class $A$ extends (or implements) the class (or interface) $B$.
+The nominal subtyping relation is based on \emph{declarations} by the developer and is \emph{name}-based, hence the adjective nominal: $A$ is a subtype of $B$ if and only if it is declared to be so, that is if the class $A$ extends (or implements) the class (or interface) $B$.
 The structural subtyping relation instead is based on the \emph{structure} of a class, its fields and methods: a class $A$ is a subtype of a class $B$ if and only if the fields and methods of $A$ are a superset of the fields and methods of $B$, and their types in $A$ are subtypes of the types in $B$.
@@
 
-The approach proposed by Dardha \emph{et al.} \cite{Dardha2013,Dardha2017} allows for the class hierarchy to be defined in such a natural way without needing to change it, but on the other hand it does not require handling exceptions at run time,  rather just static typechecking.
-To achieve this, the framework combines boolean types with semantic subtyping: it is enough to define a method \emph{diagonal} that has an argument of type \emph{Polygon} and not \emph{Triangle} thus allowing the typechecker to check at compile time the restrictions on types:
+The approach proposed by Dardha \emph{et al.} \cite{Dardha2013,Dardha2017} allows for the class hierarchy to be defined in such a natural way without needing to change it, but on the other hand it does not require handling exceptions at run time,  rather just static type checking.
+To achieve this, the framework combines boolean types with semantic subtyping: it is enough to define a method \emph{diagonal} that has an argument of type \emph{Polygon} and not \emph{Triangle} thus allowing the type checker to check at compile time the restrictions on types:
 $$
@@
 \label{sec:design}
-Since we want our types to represent sets of values, we restrict our types to finite trees whose 
+The key notion in semantic subtyping is denoting types as sets of values. Since we want these types
+to be able to be finitely representable, we restrict our types to finite trees whose
 leaves are constants with no cycles. For example a recursive type $\alpha = [a : \alpha]$ would
@@
 
-Therefore, to not allow this, when processing the AST of a program, we do not allow the type of 
-the class we are defining to be a field and we mark any classes with only basic types in it's fields
-as \textit{resolved} and otherwise if it contains a class type, we mark it as \textit{unresolved}. 
-After processing the whole AST, we perform the following algorithm to decide whether the type 
-definitions in the program are valid.
-
+Therefore, when processing the abstract syntax tree (AST) of a program, we do not allow the type of
+the class we are defining to be the type of its own field. Secondly we mark any classes with only
+basic types \textit{(bool, int, float ...)} for the types of its fields as \textit{resolved} and otherwise, if it contains a class type, we mark it as \textit{unresolved}.
+After processing the whole AST, we perform the following algorithm to decide whether the type definitions in the program are valid.
 \begin{verbatim}
+do
     boolean resolutionOccured = false
-do 
     for class that is unresolved:
         boolean resolved = true
-
         for field that contains a class type:
@@
                 resolved = false
-
-        if resolved is true
+        if resolved is true:
             set class to be resolved
@@
 
-if all classes are not resolved
+if all classes are not resolved:
     program contains invalid type definition
@@
 
+If the types in the program are finite trees finite trees whose leaves are constants with no cycles,
+then at each iteration of the algorithm, either all the types are going to be resolved, or we are
+going to be able to resolve at least one type. If we do not resolve at least one type and not all 
+types are resolved, we know we have encountered a cycle in the type definition.
+
 Given that we now know that all the types in the program are valid, we can create the subtyping
-relation which will be a map of types to a set of its subtypes. For every program, we assume the 
-initial subtyping relation for all of our basic types. We would like to highlight that Int is not
-a subtype of Float as it cannot represent the whole domain of Int accurately therefore the domain is 
-not fully contained in Float. Similarly so for Double and Long.
-
+relation which will be a map of types to a set of its subtypes. For every program, we assume an
+initial subtyping relation for all of our basic types. We would like to highlight that \textit{Int}
+is not a subtype of \textit{Float} as it cannot represent the whole domain of \textit{Int} accurately
+and therefore the domain is not fully contained, however this is not the case for \textit{Double} 
+and \textit{Int}. Similarly so \textit{Long} is not a subtype of \textit{Double}.
 \begin{flalign*}
@@
 Knowing that all our types are finite trees with leaves as constants and given this initial
-relation, we can now confidently create a subtyping relation for all class types using the following
+relation, we can create a subtyping relation for all class types using the following
 algorithm.
@@
     List<class> untyped = []
-
     for class in classes:
@@
             untyped.add(class)
-
     if untyped is not []:
@@
     for existing class type in relation:
-        if tryAddSubtype(class, existingClass) is false:
+        if checkSuperSet(class, existingClass) is false:
             return false
-
-        tryAddSubtype(existingClass, class)
-
+        checkSuperSet(existingClass, class)
         add class to its own subtype relation
 
-function tryAddSubtype(class, other):
+function checkSuperSet(class, other):
     boolean flag = true
-
     for field in class:
@@
             return false
-
         if other does not have field:
@@
                 flag = false
-
     for method in class:
@@
             return false
-
         if other does not have method:
@@
                 flag = false
-
     if flag == true:
@@
 
-The following bibliography provides
-a sample reference list with entries for journal
-articles~\cite{Dardha2017}.
+Continuing with the Polygons example, the algorithm would generate the following relation, including all base relations we defined before.
+\begin{flalign*}
+    Polygon &= \left\{Polygon, Triangle, Rectangle, Rumble \right\} &&\\
+    Triangle  &= \left\{Triangle\right\}  
+    Rectangle = \left\{Rectangle\right\} 
+    Diagonal = \left\{Diagonal\right\} 
+\end{flalign*}
+
+Given method \textit{diagonal} from the example, we can see that the result of the set operation on its parameter type would be 
+\begin{flalign*}
+    Polygon \textbf{ and not } Triangle = \left\{Polygon, Rectangle, Rumble \right\} 
+\end{flalign*}
+Therefore if we try to pass an argument of type $Triangle$ to this method, the type
+if not fully contained within the computed set, so it does not type check.
+
+The above subtyping algorithm will find all nominal subtypes, as nominal subtypes inherit all fields
+and methods of their super class, so are always guaranteed to be a superset. Similarly, we can check
+every pair of types if one is a superset of the other, i.e.\ they share common fields and
+methods, and if so, we know it is a structural subtype of the other. For example, the Empty type,
+$empty = []$, will have all other classes as structural subtypes. The above algorithm will 
+without differentiating, find all nominal and structural subtypes as we always consider all other
+classes in the program to be potential subtype candidates.
+
+The argument against structural subtyping comes up here, as a type $coord = [x:int, y:int, z:int]$ will be structurally similar to a type $colour = [x:int, y:int, z:int]$. While they can be used 
+interchangeable in our type system, the meaning of the types are completely different so it means
+we ignore possible wrong behaviour in our program. We however argue that it is no more unsafe than
+nominal subtyping in Java where an overridden method can be defined to perform the opposite
+logic to what the super class expects. Therefore both methods leave an expectation on the programmer
+to check what they are doing is correct.
+
+Furthermore, we have extended FJ with method types, demonstrated below to show another possible
+implementation of the Polygons example, where we only implement the \textit{diagonal} method on 
+polygons that support it, and not in the Triangle class. 
+$$
+\syndecl{Diagonal}{Object} {
+	\mathbf{int}\ diagonal(\textit{(getDiagonal: void$\rightarrow$int)}  \ p)\{\ldots\}
+}
+$$
+
+While we do not generate a relation for each method in all classes as this is computationally 
+unnecessary, we can on encountering a method type generate the set of classes which implement this
+method and combine their subtype relations using the \textit{or} operator. We can then use the same set inclusion to see if the argument passed to this function type checks.
+
+In the implementation of SFJ, we are liberal in the use of fields and method which might not be 
+implemented for all types in the subtyping relation rather than restricting only it to those which
+we know are shared by all types. Improper use of this allows for run-time exceptions however it
+more closely follows the idea of giving the programmer tools to create more expressive programs
+at the cost of checking program correctness such as with the use of structural subtypes.
 

### 2020-02-26 conference/paper.tex
update formatting of line breaks
start of final passes

@@
 One starts from a formal model of the programming language at hand and defines an interpretation, or semantics, of types as subsets of this model; subtyping naturally is defined as inclusion of these sets, hence the semantic adjective attached to it.
-Orthogonally, especially for object-oriented languages, there are two approaches to defining subtyping relation: the \emph{nominal} and the \emph{structural} approach.
+Orthogonally, especially for object-oriented languages, there are two approaches to defining the subtyping relation: the \emph{nominal} and the \emph{structural} approach.
 The nominal subtyping relation is based on \emph{declarations} by the developer and is \emph{name}-based, hence the adjective nominal: $A$ is a subtype of $B$ if and only if it is declared to be so, that is if the class $A$ extends (or implements) the class (or interface) $B$.
@@
 \\
-\syndecl{Rumble}{Other\_Polygons} {\cdots}
+\syndecl{Rhombus}{Other\_Polygons} {\cdots}
 \end{array}
@@
 \\
-\synidecl{Rumble}{Polygon} \synimpl Diagonal \ \{\cdots\}
+\synidecl{Rhombus}{Polygon} \synimpl Diagonal \ \{\cdots\}
 \\
@@
 \\
-\syndecl{Rumble}{Polygon} {\cdots}
+\syndecl{Rhombus}{Polygon} {\cdots}
 \\
@@
 This added expressivity does not however introduce added complexity, but rather the opposite is true as the programmer has an easier, more compact and more elegant way of programming.
-Moreover, as demonstrated in Dardha \emph{et al.} \cite[\S 8.4]{Dardha2017}, the added expressivity allows for standard programming constructs and features in Java, for e.g., lists, or overloading classes (via multimethods below), to be easily encoded in our framework.
+Moreover, as demonstrated in Dardha \emph{et al.} \cite[\S 8.4]{Dardha2017}, the added expressivity allows for standard programming constructs and for features in Java e.g., lists, or overloading classes (via multimethods below), to be easily encoded in our framework.
 Another benefit of semantic subtyping is that it can allow a natural integration of structural and nominal subtyping which the programmer has the option to choose, making SFJ a conservative extension of FJ.
@@
 \label{sec:design}
-The key notion in semantic subtyping is denoting types as sets of values. Since we want these types
-to be able to be finitely representable, we restrict our types to finite trees whose
-leaves are constants with no cycles. For example a recursive type $\alpha = [a : \alpha]$ would
-be an infinite tree \textbf{new} $C$(\textbf{new} $C$($\cdots{}$)). Similarly the types $\alpha =
-[b: \beta]$, $\beta = [a = \alpha]$ would also be impossible to instantiate.
+\paragraph{Implementation}
+The key notion in semantic subtyping is denoting types as sets of values. 
+Since we want these types to be able to be finitely representable, we restrict our types to finite trees whose leaves are constants with no cycles. 
+For example a recursive type $\alpha = [a : \alpha]$ would be an infinite tree \textbf{new} $C$(\textbf{new} $C$($\cdots{}$)). 
+Similarly the types $\alpha = [b: \beta]$, $\beta = [a = \alpha]$ would also be impossible to instantiate.
 
-Therefore, when processing the abstract syntax tree (AST) of a program, we do not allow the type of
-the class we are defining to be the type of its own field. Secondly we mark any classes with only
-basic types \textit{(bool, int, float ...)} for the types of its fields as \textit{resolved} and otherwise, if it contains a class type, we mark it as \textit{unresolved}.
+Therefore, when processing the abstract syntax tree (AST) of a program in SFJ, we do not allow the type of the class we are defining to be a type of its own field. 
+Secondly we mark any classes with only basic types \emph{(bool, int, float ...)} for the types of its fields as \emph{resolved} and otherwise, if it contains a class type, we mark it as \emph{unresolved}.
 After processing the whole AST, we perform the following algorithm to decide whether the type definitions in the program are valid.
@@
 
-If the types in the program are finite trees finite trees whose leaves are constants with no cycles,
-then at each iteration of the algorithm, either all the types are going to be resolved, or we are
-going to be able to resolve at least one type. If we do not resolve at least one type and not all 
-types are resolved, we know we have encountered a cycle in the type definition.
+If the types in the program are finite trees finite trees whose leaves are constants with no cycles, then at each iteration of the algorithm, either all the types are going to be resolved, or we are going to be able to resolve at least one type.
+If we do not resolve at least one type and not all types are resolved, we know we have encountered a cycle in the type definition.
 
-Given that we now know that all the types in the program are valid, we can create the subtyping
-relation which will be a map of types to a set of its subtypes. For every program, we assume an
-initial subtyping relation for all of our basic types. We would like to highlight that \textit{Int}
-is not a subtype of \textit{Float} as it cannot represent the whole domain of \textit{Int} accurately
-and therefore the domain is not fully contained, however this is not the case for \textit{Double} 
-and \textit{Int}. Similarly so \textit{Long} is not a subtype of \textit{Double}.
+Given that we now know that all the types in the program are valid, we can create the subtyping relation which will be a map of types to a set of its subtypes. 
+For every program, we assume an initial subtyping relation for all of our basic types. 
+We would like to highlight that \emph{Int} is not a subtype of \emph{Float} as it cannot represent the whole domain of \emph{Int} accurately and therefore the domain is not fully contained, however this is not the case for \emph{Double} and \emph{Int}.
+Similarly so \emph{Long} is not a subtype of \emph{Double}.
 \begin{flalign*}
@@
 
-Knowing that all our types are finite trees with leaves as constants and given this initial
-relation, we can create a subtyping relation for all class types using the following
-algorithm.
+Knowing that all our types are finite trees with leaves as constants and given this initial relation, we can create a subtyping relation for all class types using the following algorithm.
 
@@
 
-Continuing with the Polygons example, the algorithm would generate the following relation, including all base relations we defined before.
+Continuing with the \emph{Polygons} example, the algorithm would generate the following relation, including all base relations we defined before.
 \begin{flalign*}
@@
 
-Given method \textit{diagonal} from the example, we can see that the result of the set operation on its parameter type would be 
+Given method \emph{diagonal} from the example, we can see that the result of the set operation on its parameter type would be 
 \begin{flalign*}
-    Polygon \textbf{ and not } Triangle = \left\{Polygon, Rectangle, Rumble \right\} 
+    Polygon \textbf{ and not } Triangle = \left\{Polygon, Rectangle, Rhombus \right\} 
 \end{flalign*}
-Therefore if we try to pass an argument of type $Triangle$ to this method, the type
-if not fully contained within the computed set, so it does not type check.
+Therefore if we try to pass an argument of type \emph{Triangle} to this method, the type if not fully contained within the computed set, so it does not type check.
 
-The above subtyping algorithm will find all nominal subtypes, as nominal subtypes inherit all fields
-and methods of their super class, so are always guaranteed to be a superset. Similarly, we can check
-every pair of types if one is a superset of the other, i.e.\ they share common fields and
-methods, and if so, we know it is a structural subtype of the other. For example, the Empty type,
-$empty = []$, will have all other classes as structural subtypes. The above algorithm will 
-without differentiating, find all nominal and structural subtypes as we always consider all other
-classes in the program to be potential subtype candidates.
+The above subtyping algorithm will find all nominal subtypes, as nominal subtypes inherit all fields and methods of their super class, so are always guaranteed to be a superset. 
+Similarly, we can check every pair of types if one is a superset of the other, i.e.\ they share common fields and methods, and if so, we know it is a structural subtype of the other. 
+For example, the Empty type, $empty = []$, will have all classes as structural subtypes. 
+The above algorithm will without differentiating, find all nominal and structural subtypes as we always consider all other classes in the program to be potential subtype candidates.
 
-The argument against structural subtyping comes up here, as a type $coord = [x:int, y:int, z:int]$ will be structurally similar to a type $colour = [x:int, y:int, z:int]$. While they can be used 
-interchangeable in our type system, the meaning of the types are completely different so it means
-we ignore possible wrong behaviour in our program. We however argue that it is no more unsafe than
-nominal subtyping in Java where an overridden method can be defined to perform the opposite
-logic to what the super class expects. Therefore both methods leave an expectation on the programmer
-to check what they are doing is correct.
+The argument against structural subtyping comes up here, as a type $coord = [x:int, y:int, z:int]$ will be structurally similar to a type $colour = [x:int, y:int, z:int]$. 
+While they can be used interchangeable in our type system, the meaning of the types are completely different so it means we ignore possible wrong behaviour in our program. 
+We however argue that it is no more unsafe than nominal subtyping in Java where an overridden method can be defined to perform the opposite logic to what the super class expects. 
+Therefore both methods leave an expectation on the programmer to check what they are doing is correct.
 
-Furthermore, we have extended FJ with method types, demonstrated below to show another possible
-implementation of the Polygons example, where we only implement the \textit{diagonal} method on 
-polygons that support it, and not in the Triangle class. 
+Furthermore, we have extended FJ with method types, demonstrated below to show another possible implementation of the Polygons example, where we only implement the \textit{diagonal} method on polygons that support it, and not in the Triangle class.
 $$
@@
 
-While we do not generate a relation for each method in all classes as this is computationally 
-unnecessary, we can on encountering a method type generate the set of classes which implement this
-method and combine their subtype relations using the \textit{or} operator. We can then use the same set inclusion to see if the argument passed to this function type checks.
+While we do not generate a relation for each method in all classes as this is computationally unnecessary, we can on encountering a method type generate the set of classes which implement this method and combine their subtype relations using the \textit{or} operator.
+We can then use the same set inclusion to see if the argument passed to this function type checks.
 
-In the implementation of SFJ, we are liberal in the use of fields and method which might not be 
-implemented for all types in the subtyping relation rather than restricting only it to those which
-we know are shared by all types. Improper use of this allows for run-time exceptions however it
-more closely follows the idea of giving the programmer tools to create more expressive programs
-at the cost of checking program correctness such as with the use of structural subtypes.
+In the implementation of SFJ, we are liberal in the use of fields and method which might not be implemented for all types in the subtyping relation rather than restricting only it to those which we know are shared by all types.
+Improper use of this allows for run-time exceptions however it more closely follows the idea of giving the programmer tools to create more expressive programs at the cost of checking program correctness such as with the use of structural subtypes.
 

### 2020-02-26 conference/paper.tex
cutting on intro

--- 2020-02-25 conference/paper.tex
@@
 
-\title{SFJ: An implementation of Semantic Featherweight Java\thanks{Research supported by the UK EPSRC grant EP/K034413/1, ``From Data Types to Session Types: A Basis for Concurrency and Distribution (ABCD)'', and by the EU HORIZON 2020 MSCA RISE project 778233 ``BehAPI: Behavioural Application Program Interfaces''.}}
+\title{SFJ: An implementation of Semantic Featherweight Java
+%\thanks{Research supported by the UK EPSRC grant EP/K034413/1, ``From Data Types to Session Types: A Basis for Concurrency and Distribution (ABCD)'', and by the EU HORIZON 2020 MSCA RISE project 778233 ``BehAPI: Behavioural Application Program Interfaces''.}
+}
 \titlerunning{SFJ}
@@
 \begin{abstract}
-Subtyping is a key notion in programming languages, as it allows more flexibility in coding.
+\emph{Subtyping} is a key notion in programming languages, as it allows more flexibility in coding.
 There are two approaches to defining subtyping relations: the \emph{syntactic} and the \emph{semantic} approach.
-In the latter, one starts by defining a model of the programming language at hand and an interpretation of its types as subsets of such model.
-Then, subtyping is defined as \emph{set inclusion of types} denoting sets.
-An orthogonal issue, typical of object-oriented languages, is the \emph{nominal} and \emph{structural} subtyping.
+In the latter, one starts by defining a model of the language and an interpretation of types as subsets of this model.
+Subtyping is defined as \emph{set inclusion of types} denoting sets.
+An orthogonal issue, typical of object-oriented languages, is the \emph{nominal} vs. \emph{structural} subtyping.
 Dardha \emph{et al.} \cite{Dardha2013,Dardha2017} combined structural subtyping and boolean types, on one hand, with nominal subtyping, on the other, in Featherweight Java (FJ), by thus exploiting the benefits of both approaches.
 
-However, these benefits were only presented and exemplified at a theoretical level and not exploited practically.
-At last, in this work we have implemented SFJ---Semantic Featherweight Java, which features structural subtyping and boolean types as well as nominal subtyping, typical of FJ itself.
+However, these benefits were illustrated only at a theoretical level but not exploited practically.
+In this paper, we present SFJ---Semantic Featherweight Java, an implementation of FJ with features structural subtyping and boolean types as well as nominal subtyping.
 The benefits of SFJ, also illustrated in the accompanied video, show how static type checking of boolean types and semantic subtyping gives higher guarantees of program correctness, more flexibility and compactness of program writing.
-\keywords{Nominal subtyping \and Structural subtyping  \and Semantic Featherweight Java \and Object-oriented languages \and Boolean types \and Type theory.}
+%\keywords{Nominal subtyping \and Structural subtyping  \and Semantic Featherweight Java \and Object-oriented languages \and Boolean types \and Type theory.}
 \end{abstract}
@@
 There are two approaches to defining subtyping relations: the \emph{syntactic} and the \emph{semantic} approach.
-In the latter, one starts by defining a model of the programming language at hand and an interpretation of its types as subsets of such model.
-The syntactic subtyping is more mainstream in programming languages design and implementation, and the relation is given by means of a set of formal deductive subtyping rules.
-The semantic subtyping is less know and the relation is given by inclusion of sets denoting types.
-One starts from a formal model of the programming language at hand and defines an interpretation, or semantics, of types as subsets of this model; subtyping naturally is defined as inclusion of these sets, hence the semantic adjective attached to it.
-Orthogonally, especially for object-oriented languages, there are two approaches to defining subtyping relation: the \emph{nominal} and the \emph{structural} approach.
-The nominal subtyping relation is based on \emph{declarations} by the developer and is \emph{name}-based, hence the adjective nominal: $A$ is a subtype of $B$ if and only if it is declared to be so, that is if the class $A$ extends (or implements) the class (or interface) $B$.
-The structural subtyping relation instead is based on the \emph{structure} of a class, its fields and methods: a class $A$ is a subtype of a class $B$ if and only if the fields and methods of $A$ are a superset of the fields and methods of $B$, and their types in $A$ are subtypes of the types in $B$.
-At this point it is not surprising that the nominal subtyping aligns well with the syntactic approach and hence is more well-known than the structural subtyping which aligns well with the semantic approach.
+The syntactic subtyping is more mainstream in programming languages and is defined by means of a set of formal deductive subtyping rules.
+The semantic subtyping is less know and is defined as inclusion of sets denoting types.
+Starting from a formal model of the  language, we define an interpretation of types as subsets of this model, typically, subsets of values in the language.
+Then, subtyping is defined as inclusion of these subsets.
 
-Dardha \emph{et al.} \cite{Dardha2013,Dardha2017} define boolean types and semantic subtyping for \emph{Featherweight Java} (FJ) \cite{featherweight}, which is the functional fragment of the Java language.
-This approach to defining subtyping naturally allows for structural subtyping in FJ.
-On the other hand, nominal subtyping, typical of FJ, can be obtained in the framework exploiting boolean types and semantic subtyping, as shown in the extensive case studies in Dardha \emph{et al.} \cite[\S 8.4]{Dardha2017}.
+Orthogonally, for object-oriented languages there are two approaches to defining subtyping relations: the \emph{nominal} and the \emph{structural} approach.
+The nominal subtyping is based on {declarations} by the developer and is {name}-based: $A$ is a subtype of $B$ if and only if it is declared to be so, that is if the class $A$ extends (or implements) the class (or interface) $B$.
+The structural subtyping instead is based on the \emph{structure} of a class, its fields and methods: a class $A$ is a subtype of a class $B$ if and only if the fields and methods of $A$ are a superset of the fields and methods of $B$, and their types in $A$ are subtypes of the types in $B$.
+At this point it is not surprising that the nominal subtyping aligns well with the syntactic approach and in fact is more well-known than the structural subtyping which in turn aligns well with the semantic approach.
 
-\begin{example}[The Polygons]
-We will use the \emph{Polygons} as our running example, to illustrate the benefits of the approach by Dardha \emph{et al.} \cite{Dardha2013,Dardha2017}, both in the paper and in the tool video.
+Dardha \emph{et al.} \cite{Dardha2013,Dardha2017} define semantic subtyping for \emph{Featherweight Java} (FJ) \cite{featherweight} with \emph{boolean types}.
+This approach allows the co-existence of nominal and structural subtyping in FJ, bringing in higher guarantees of program correctness, flexibility and compactness in programming.
+Unfortunately, these benefits were only presented at a theoretical level and not exploited practically, due to the lack of an implementation of the language and its types and type system.
+
+In this paper, we present SFJ---\emph{Semantic Featherweight Java} \autoref{sec:design}, an implementation of FJ with integrated boolean types and semantic subtyping.
+In SFJ the developer has a larger and more expressive set of types, by using boolean connectives \textbf{and}, \textbf{not}, \textbf{or}.
+On the other hand, this added expressivity does not introduce added complexity, but rather the opposite is true, as the developer has an easier, more compact and more elegant way of programming.
+SFJ integrates both structural and nominal subtyping, and the developer can choose which one to use.
+Finally, as discussed in Dardha \emph{et al.} \cite[\S 8.4]{Dardha2017}, we can easily encode in SFJ standard programming constructs and features of Java, such as lists, or overloading classes (via multimethods, see below), which are missing in FJ, making SFJ a more complete and closer to Java programming language.
+
+\begin{example}[Polygons]
+This will be our running example both in the paper and in the tool video \cite{UD20} to illustrate the benefits of boolean types and semantic subtyping developed by Dardha \emph{et al.} \cite{Dardha2013,Dardha2017} and implemented as SFJ.
+
 Consider the set of polygons, such as triangles, rectangles, rumbles given by a class hierarchy.
@@
 \begin{array}{l}
-\synclass Polygon \ \{ \cdots\}
+\synclass Polygon \ \{ \ldots\}
 \\
-\syndecl{Triangle}{Polygon} {\cdots}
+\syndecl{Triangle}{Polygon} {\ldots}
 \\
@@
 \\
-\syndecl{Rectangle}{Other\_Polygons}{\cdots}
+\syndecl{Rectangle}{Other\_Polygons}{\ldots}
 \\
-\syndecl{Rumble}{Other\_Polygons} {\cdots}
+\syndecl{Rumble}{Other\_Polygons} {\ldots}
 \end{array}
 $$
-Or by means of an interface:
+Or by means of an interface \emph{Diagonal}:
 $$
@@
 \\
-\synclass Polygon \ \{\cdots\}
+\synclass Polygon \ \{\ldots\}
 \\
-\syndecl{Triangle}{Polygon} {\cdots}
+\syndecl{Triangle}{Polygon} {\ldots}
 \\
-\synidecl{Rectangle}{Polygon} \synimpl Diagonal \ \{\cdots\}
+\synidecl{Rectangle}{Polygon} \synimpl Diagonal \ \{\ldots\}
 \\
-\synidecl{Rumble}{Polygon} \synimpl Diagonal \ \{\cdots\}
+\synidecl{Rumble}{Polygon} \synimpl Diagonal \ \{\ldots\}
 \\
@@
 
-Now, suppose that our class hierarchy is such that \emph{Polygon} is the parent class and all other geometric figure classes extend \emph{Polygon}, which is exactly how one would naturally define the set of polygons.
-If we think of the class hierarchy as given and part of legacy code, which cannot be changed, then again the natural way to solve this in Java is by defining the method \emph{diagonal} in the class \emph{Polygon} and using an \textbf{instanceof}, for example inside a \textbf{try-catch} that throws an exception at run time, if the argument passed to the method is a triangle.
+Now, suppose that our class hierarchy is such that \emph{Polygon} is the parent class and all other geometric figure classes extend \emph{Polygon}, which is how one would naturally define the set of polygons.
+If we think of the class hierarchy as given and part of legacy code, which cannot be changed, then again a natural way to implement this in Java is by defining the method \emph{diagonal} in the class \emph{Polygon} and using an \textbf{instanceof}, for e.g., inside a \textbf{try-catch} that throws an exception at run time, if the argument passed to the method is a triangle.
 
-The approach proposed by Dardha \emph{et al.} \cite{Dardha2013,Dardha2017} allows for the class hierarchy to be defined in such a natural way without needing to change it, but on the other hand it does not require handling exceptions at run time,  rather just static type checking.
-To achieve this, the framework combines boolean types with semantic subtyping: it is enough to define a method \emph{diagonal} that has an argument of type \emph{Polygon} and not \emph{Triangle} thus allowing the type checker to check at compile time the restrictions on types:
+We propose a more elegant solution, by combining boolean types and semantic subtyping, where only static typechecking is required and implement this in SFJ \cite{UD20}: it is enough to define a method \emph{diagonal} that has an argument of type \emph{Polygon} \textbf{and not} \emph{Triangle} thus allowing the typechecker to check at compile time the restrictions on types:
 $$
 \begin{array}{l}
-\syndecl{Polygon}{Object} {\cdots}
+\syndecl{Polygon}{Object} {\ldots}
 \\
-\syndecl{Triangle}{Polygon} {\cdots}
+\syndecl{Triangle}{Polygon} {\ldots}
 \\
-\syndecl{Rectangle}{Polygon} {\cdots}
+\syndecl{Rectangle}{Polygon} {\ldots}
 \\
-\syndecl{Rumble}{Polygon} {\cdots}
+\syndecl{Rumble}{Polygon} {\ldots}
 \\
@@
 \end{example}
-The combination of boolean types and semantic subtyping gives a more elegant solution, as illustrate by the \emph{Polygons}---of course a toy example used only as a proof of concept of the approach---however, this approach is not yet practical as there are no boolean types or semantic subtyping in Java, hence its benefits unfortunately remained theoretical.
-
-After a few years, finally we can now exploit the benefits of this approach at a practical level: in this paper we present SFJ---Semantic Featherweight Java \autoref{sec:design}, an implementation of Featherweight Java with integrated boolean types and semantic subtyping.
-In SFJ the programmer has a larger and more expressive set of types, by using \textbf{and}, \textbf{not}, \textbf{or} boolean connectives.
-This added expressivity does not however introduce added complexity, but rather the opposite is true as the programmer has an easier, more compact and more elegant way of programming.
-Moreover, as demonstrated in Dardha \emph{et al.} \cite[\S 8.4]{Dardha2017}, the added expressivity allows for standard programming constructs and features in Java, for e.g., lists, or overloading classes (via multimethods below), to be easily encoded in our framework.
-Another benefit of semantic subtyping is that it can allow a natural integration of structural and nominal subtyping which the programmer has the option to choose, making SFJ a conservative extension of FJ.
-
-\paragraph{On multimethods}
+\emph{On multimethods}
 Since FJ is a core language, some features of the full Java are removed: overloading methods is one of them.
@@
 
-The technical developments behind semantic subtyping and its properties are not easy, however, they are completely transparent to the programmer.
+The technical developments behind semantic subtyping and its properties are not easy, however, they are completely transparent to the developer.
 Due to space limit we will remove theoretical details from this submission, as they are not necessary to understand the design and implementation of SFJ, which is the focus of this paper.
@@
 \label{sec:design}
-The key notion in semantic subtyping is denoting types as sets of values. Since we want these types
+As already discussed in the introduction, the key idea behind semantic subtyping is interpreting types as sets, in particular,  as sets of values of the language at hand. Since we want these types
 to be able to be finitely representable, we restrict our types to finite trees whose
 leaves are constants with no cycles. For example a recursive type $\alpha = [a : \alpha]$ would
-be an infinite tree \textbf{new} $C$(\textbf{new} $C$($\cdots{}$)). Similarly the types $\alpha =
+be an infinite tree \textbf{new} $C$(\textbf{new} $C$($\ldots{}$)). Similarly the types $\alpha =
 [b: \beta]$, $\beta = [a = \alpha]$ would also be impossible to instantiate.

### 2020-02-26 conference/paper.tex
now?

@@
 \\
-\syndecl{Rumble}{Other\_Polygons} {\ldots}
+\syndecl{Rhombus}{Other\_Polygons} {\ldots}
 \end{array}
@@
 \\
-\synidecl{Rumble}{Polygon} \synimpl Diagonal \ \{\ldots\}
+\synidecl{Rhombus}{Polygon} \synimpl Diagonal \ \{\ldots\}
 \\
@@
 \\
-\syndecl{Rumble}{Polygon} {\ldots}
+\syndecl{Rhombus}{Polygon} {\ldots}
 \\
@@
 \label{sec:design}
-As already discussed in the introduction, the key idea behind semantic subtyping is interpreting types as sets, in particular,  as sets of values of the language at hand. Since we want these types
-to be able to be finitely representable, we restrict our types to finite trees whose
-leaves are constants with no cycles. For example a recursive type $\alpha = [a : \alpha]$ would
-be an infinite tree \textbf{new} $C$(\textbf{new} $C$($\ldots{}$)). Similarly the types $\alpha =
-[b: \beta]$, $\beta = [a = \alpha]$ would also be impossible to instantiate.
+As already discussed in the introduction, the key idea behind semantic subtyping is interpreting types as sets, in particular,  as sets of values of the language at hand.
+Since we want these types to be able to be finitely representable, we restrict our types to finite trees whose leaves are constants with no cycles.
+For example a recursive type $\alpha = [a : \alpha]$ would be an infinite tree \textbf{new} $C$(\textbf{new} $C$($\cdots{}$)). 
+Similarly the types $\alpha = [b: \beta]$, $\beta = [a = \alpha]$ would also be impossible to instantiate.
 
-Therefore, when processing the abstract syntax tree (AST) of a program, we do not allow the type of
-the class we are defining to be the type of its own field. Secondly we mark any classes with only
-basic types \textit{(bool, int, float ...)} for the types of its fields as \textit{resolved} and otherwise, if it contains a class type, we mark it as \textit{unresolved}.
+Therefore, when processing the abstract syntax tree (AST) of a program in SFJ, we do not allow the type of the class we are defining to be a type of its own field. 
+Secondly we mark any classes with only basic types \emph{(bool, int, float ...)} for the types of its fields as \emph{resolved} and otherwise, if it contains a class type, we mark it as \emph{unresolved}.
 After processing the whole AST, we perform the following algorithm to decide whether the type definitions in the program are valid.

### 2020-02-27 conference/paper.tex
types and terms

@@
 \label{sec:design}
-As already discussed in the introduction, the key idea behind semantic subtyping is interpreting types as sets, in particular,  as sets of values of the language at hand.
-Since we want these types to be able to be finitely representable, we restrict our types to finite trees whose leaves are constants with no cycles.
-For example a recursive type $\alpha = [a : \alpha]$ would be an infinite tree \textbf{new} $C$(\textbf{new} $C$($\cdots{}$)). 
-Similarly the types $\alpha = [b: \beta]$, $\beta = [a = \alpha]$ would also be impossible to instantiate.
+The syntax of types is given by the following grammar \cite{}:
+$$
+\begin{array}{llll}
+\tau & ::= & \alpha \ |\ \mu
+& \mbox{{\em Type term}}
+\\
+\alpha & ::=  & \zero \ |\ \btypes \ |\ [\wt{l:\tau}] \ |\ \alpha \synwedge \alpha \ |\ \synneg\alpha
+\qquad
+& \mbox{{\em Object type} ($\alpha $-type)}
+\\
+\mu & ::=  & \alpha \to \alpha \ |\ \mu \synwedge \mu \ |\ \synneg\mu
+& \mbox{{\em Method type} ($\mu $-type)}
+\end{array}
+$$
+$\alpha$-types are used to type fields and $\mu$-types are used to type methods.
+Type $\zero$ is the empty type.
+Type $\btypes$ denotes the \emph{basic} types, such as integers, booleans, etc.
+{Record} types $[\wt{l:\tau}]$, where $\wt l$ is a sequence of disjoint labels are used to type objects.
+Arrow types $\alpha \to \alpha$ are used to type methods.
+The boolean types using $\synwedge$ and $\synneg$ have their expected meanings and $\synvee$ is obtained by their combination.\\
+The syntax of terms is given by the following grammar and is based on the standard syntax of terms in FJ \cite{}:
+\begin{align*}
+& \mbox{\textit{Class declaration}} & L \; ::= \; & \syndecl{C}{C}{\wt{\alpha \ a};\ K; \ \wt{M}\ }\\
+& \mbox{\textit{Constructor}} & K \;::=\; & C\ (\wt{\alpha\ x})\ \{\ \synsuper(\wt{x});\ \wt{\synthis.a}=\wt{x}; \}\\
+& \mbox{\textit{Method declaration}} & M \; ::= \; & \alpha \ m\ (\alpha \ x)\ \{\ \synret e; \} \\
+& \mbox{\textit{Expressions}} & e \; ::=\; & x\ |\  c\ |\ e.a\ |\ e.m(e) \ | \ \synnew C(\wt{e})
+\end{align*}
+Since we want to use types $\tau$ in practice in SFJ, we restrict them to finite trees whose leaves are constants with no cycles.
+For e.g., a recursive type $\alpha = [a : \alpha]$ would be an infinite tree \textbf{new} $C$(\textbf{new} $C$($\cdots{}$)), hence we cannot instate it.
+%Similarly the types $\alpha = [b: \beta]$, $\beta = [a = \alpha]$ would also be impossible to instantiate.
 
 Therefore, when processing the abstract syntax tree (AST) of a program in SFJ, we do not allow the type of the class we are defining to be a type of its own field. 
-Secondly we mark any classes with only basic types \emph{(bool, int, float ...)} for the types of its fields as \emph{resolved} and otherwise, if it contains a class type, we mark it as \emph{unresolved}.
+Secondly, we mark any classes with fields only of basic types as \emph{resolved} and otherwise, if it contains a record type, we mark it as \emph{unresolved}. 
 After processing the whole AST, we perform the following algorithm to decide whether the type definitions in the program are valid.

### 2020-02-27 conference/paper.tex
explaining mapping of types and subtyping algorithm

@@
 \begin{example}[Polygons]
+\label{polygons}
 This will be our running example both in the paper and in the tool video \cite{UD20} to illustrate the benefits of boolean types and semantic subtyping developed by Dardha \emph{et al.} \cite{Dardha2013,Dardha2017} and implemented as SFJ.
@@
 \label{sec:design}
-The syntax of types is given by the following grammar \cite{}:
+The syntax of types is given by the following grammar \cite{Dardha2013,Dardha2017}:
 $$
@@
 The boolean types using $\synwedge$ and $\synneg$ have their expected meanings and $\synvee$ is obtained by their combination.\\
-The syntax of terms is given by the following grammar and is based on the standard syntax of terms in FJ \cite{}:
+The syntax of terms is given by the following grammar and is based on the standard syntax of terms in FJ \cite{featherweight,Dardha2013,Dardha2017}:
 \begin{align*}
@@
 Since we want to use types $\tau$ in practice in SFJ, we restrict them to finite trees whose leaves are constants with no cycles.
-For e.g., a recursive type $\alpha = [a : \alpha]$ would be an infinite tree \textbf{new} $C$(\textbf{new} $C$($\cdots{}$)), hence we cannot instate it.
+For e.g., a recursive type $\alpha = [a : \alpha]$ denotes an infinite program tree \textbf{new} $C$(\textbf{new} $C$($\cdots{}$)), hence we avoid it as it is inhabitable.
 %Similarly the types $\alpha = [b: \beta]$, $\beta = [a = \alpha]$ would also be impossible to instantiate.
-
-Therefore, when processing the abstract syntax tree (AST) of a program in SFJ, we do not allow the type of the class we are defining to be a type of its own field. 
-Secondly, we mark any classes with fields only of basic types as \emph{resolved} and otherwise, if it contains a record type, we mark it as \emph{unresolved}. 
-After processing the whole AST, we perform the following algorithm to decide whether the type definitions in the program are valid.
+When processing the abstract syntax tree (AST) of a program in SFJ, we forbid fields to have the same type as the type of the class they are defined in. 
+Secondly, we mark any classes containing fields typed with only basic types as \emph{resolved} otherwise, as \emph{unresolved}. 
+This is captured by the following algorithm, which checks if the type definitions in the program are valid.
 \begin{verbatim}
@@
 \end{verbatim}
-
-If the types in the program are finite trees finite trees whose leaves are constants with no cycles, then at each iteration of the algorithm, either all the types are going to be resolved, or we are going to be able to resolve at least one type.
-If we do not resolve at least one type and not all types are resolved, we know we have encountered a cycle in the type definition.
-
-Given that we now know that all the types in the program are valid, we can create the subtyping relation which will be a map of types to a set of its subtypes. 
-For every program, we assume an initial subtyping relation for all of our basic types. 
-We would like to highlight that \emph{Int} is not a subtype of \emph{Float} as it cannot represent the whole domain of \emph{Int} accurately and therefore the domain is not fully contained, however this is not the case for \emph{Double} and \emph{Int}.
-Similarly so \emph{Long} is not a subtype of \emph{Double}.
+%If the types in the program are finite trees whose leaves are constants with no cycles, then at each iteration of the algorithm, either all the types are going to be resolved, or we are going to be able to resolve at least one type.
+%If we do not resolve at least one type and not all types are resolved, we know we have encountered a cycle in the type definition. 
+%
+Now, we can define the subtyping relation.
+Building upon the types-as-sets-of-values approach of the semantic subtyping for FJ \cite{Dardha2013,Dardha2017}, in our implementation of SFJ we keep track of this relation by defining a map from a type to the set of its subtypes, with the property that the set of values of a subtype is included in the set of values of the type.
+As a first step, we start with basic types.
+\setlength{\abovedisplayskip}{5pt}
+\setlength{\belowdisplayskip}{5pt}
 \begin{flalign*}
-    Boolean &= \left\{Boolean \right\} &\\
-    Double  &= \left\{Double, Float, Int, Short, Byte\right\} &\\
-    Float &= \left\{Float, Short, Byte\right\} &\\
-    Long &= \left\{Long, Int, Short, Byte\right\} &\\
-    Int &= \left\{Int, Short, Byte\right\} &\\
-    Short &= \left\{Short, Byte\right\} &\\
-    Byte &= \left\{Byte\right\} &
+    &Double  = \left\{Double, Float, Int, Short, Byte\right\} &&Float = \left\{Float, Short, Byte\right\}\\
+    &Long = \left\{Long, Int, Short, Byte\right\}  &&Int = \left\{Int, Short, Byte\right\}\\
+    &Short = \left\{Short, Byte\right\} && Byte = \left\{Byte\right\}\\
+    &Boolean = \left\{Boolean \right\} 
 \end{flalign*}
-
-Knowing that all our types are finite trees with leaves as constants and given this initial relation, we can create a subtyping relation for all class types using the following algorithm.
-
+%
+Note that \emph{Int} is not a subtype of \emph{Float} as the latter cannot represent the whole set of \emph{Int} values accurately and therefore \emph{Int} is not fully set-contained, however this is not the case for \emph{Int} and \emph{Double}.
+Similarly, \emph{Long} is not a subtype of \emph{Double}.
+Finally, the following algorithm defines the subtyping relation for all class types.
 \begin{verbatim}
@@
 \end{verbatim}
-
-Continuing with the \emph{Polygons} example, the algorithm would generate the following relation, including all base relations we defined before.
+Let us illustrate the subtyping relation algorithm with our \emph{Polygons} example \autoref{polygons}.
+The algorithm generates the following subtyping relation, which also includes the subtyping relation for basic types defined in the first step.
 \begin{flalign*}
-    Polygon &= \left\{Polygon, Triangle, Rectangle, Rumble \right\} &&\\
-    Triangle  &= \left\{Triangle\right\}  
-    Rectangle = \left\{Rectangle\right\} 
+    &Polygon = \left\{Polygon, Triangle, Rectangle, Rumble \right\} && Triangle  = \left\{Triangle\right\}\\ 
+    &Rectangle = \left\{Rectangle\right\} &&
     Diagonal = \left\{Diagonal\right\} 
 \end{flalign*}
-
-Given method \emph{diagonal} from the example, we can see that the result of the set operation on its parameter type would be 
+Recall method \emph{diagonal} in class \emph{Diagonal}, we can see that the result of the set operation on its parameter type gives the following set of polygons: 
 \begin{flalign*}
@@
 \end{flalign*}
-Therefore if we try to pass an argument of type \emph{Triangle} to this method, the type if not fully contained within the computed set, so it does not type check.
+Therefore, if we pass to this method an argument of type \emph{Triangle}, as it is not contained in the computed set of subtypes, we get a typechecking error at compile type \cite{UD20}.
 

### 2020-02-27 conference/paper.tex
nominal vs structural

@@
 \end{flalign*}
-Therefore, if we pass to this method an argument of type \emph{Triangle}, as it is not contained in the computed set of subtypes, we get a typechecking error at compile type \cite{UD20}.
+If we pass to \emph{diagonal} an argument of type \emph{Triangle}, as it is not contained in the computed set of subtypes, we get a typechecking error at compile type \cite{UD20}.
 
-The above subtyping algorithm will find all nominal subtypes, as nominal subtypes inherit all fields and methods of their super class, so are always guaranteed to be a superset. 
-Similarly, we can check every pair of types if one is a superset of the other, i.e.\ they share common fields and methods, and if so, we know it is a structural subtype of the other. 
-For example, the Empty type, $empty = []$, will have all classes as structural subtypes. 
-The above algorithm will without differentiating, find all nominal and structural subtypes as we always consider all other classes in the program to be potential subtype candidates.
-
-The argument against structural subtyping comes up here, as a type $coord = [x:int, y:int, z:int]$ will be structurally similar to a type $colour = [x:int, y:int, z:int]$. 
-While they can be used interchangeable in our type system, the meaning of the types are completely different so it means we ignore possible wrong behaviour in our program. 
+The subtyping algorithm finds all nominal and structural subtypes.
+It will find all nominal subtypes, as nominal subtypes inherit all fields and methods of their super class, so are always guaranteed to be a superset. 
+Similarly, given two types, the algorithm can check if one is a superset of the other, i.e.,\ they share common fields and methods, and if so, they are related by structural subtyping. 
+%For e.g., the Empty type, $empty = []$, will have all classes as structural subtypes. 
+%The above algorithm will without differentiating, find all nominal and structural subtypes as we always consider all other classes in the program to be potential subtype candidates.
+Consider two structurally equivalent types $coord = [x:int, y:int, z:int]$ and $colour = [x:int, y:int, z:int]$. 
+While they can be used interchangeably in our type system, the meaning of the types is completely different, meaning we ignore possible wrong behaviour in our program. 
 We however argue that it is no more unsafe than nominal subtyping in Java where an overridden method can be defined to perform the opposite logic to what the super class expects. 

### 2020-02-27 conference/paper.tex
pass over, clarify explanations

@@
 However, these benefits were illustrated only at a theoretical level but not exploited practically.
-In this paper, we present SFJ---Semantic Featherweight Java, an implementation of FJ with features structural subtyping and boolean types as well as nominal subtyping.
+In this paper, we present SFJ---Semantic Featherweight Java, an implementation of FJ which features structural subtyping and boolean types as well as nominal subtyping.
 The benefits of SFJ, also illustrated in the accompanied video, show how static type checking of boolean types and semantic subtyping gives higher guarantees of program correctness, more flexibility and compactness of program writing.
@@
 There are two approaches to defining subtyping relations: the \emph{syntactic} and the \emph{semantic} approach.
-The syntactic subtyping is more mainstream in programming languages and is defined by means of a set of formal deductive subtyping rules.
-The semantic subtyping is less know and is defined as inclusion of sets denoting types.
+Syntactic subtyping is more mainstream in programming languages and is defined by means of a set of formal deductive subtyping rules.
+Semantic subtyping is less know and is defined as inclusion of sets denoting types.
 Starting from a formal model of the  language, we define an interpretation of types as subsets of this model, typically, subsets of values in the language.
@@
 Orthogonally, for object-oriented languages there are two approaches to defining subtyping relations: the \emph{nominal} and the \emph{structural} approach.
-The nominal subtyping is based on {declarations} by the developer and is {name}-based: $A$ is a subtype of $B$ if and only if it is declared to be so, that is if the class $A$ extends (or implements) the class (or interface) $B$.
-The structural subtyping instead is based on the \emph{structure} of a class, its fields and methods: a class $A$ is a subtype of a class $B$ if and only if the fields and methods of $A$ are a superset of the fields and methods of $B$, and their types in $A$ are subtypes of the types in $B$.
+Nominal subtyping is based on {declarations} by the developer and is {name}-based: $A$ is a subtype of $B$ if and only if it is declared to be so, that is if the class $A$ extends (or implements) the class (or interface) $B$.
+Structural subtyping instead is based on the \emph{structure} of a class, its fields and methods: a class $A$ is a subtype of a class $B$ if and only if the fields and methods of $A$ are a superset of the fields and methods of $B$, and their types in $A$ are subtypes of the types in $B$.
 At this point it is not surprising that the nominal subtyping aligns well with the syntactic approach and in fact is more well-known than the structural subtyping which in turn aligns well with the semantic approach.
@@
 SFJ integrates both structural and nominal subtyping, and the developer can choose which one to use.
-Finally, as discussed in Dardha \emph{et al.} \cite[\S 8.4]{Dardha2017}, we can easily encode in SFJ standard programming constructs and features of Java, such as lists, or overloading classes (via multimethods, see below), which are missing in FJ, making SFJ a more complete and closer to Java programming language.
+Finally, as discussed in Dardha \emph{et al.} \cite[\S 8.4]{Dardha2017}, we can easily encode in SFJ standard programming constructs and features of Java, such as lists, or overloading classes (via multimethods, see below), which are missing in FJ, making SFJ more complete and closer to Java.
 
@@
 
-Consider the set of polygons, such as triangles, rectangles, rumbles given by a class hierarchy.
+Consider the set of polygons, such as triangles, rectangles and rhombuses given by a class hierarchy.
 We want to define a method \emph{diagonal} that takes a polygon and returns the length of its longest diagonal.
@@
 
-We propose a more elegant solution, by combining boolean types and semantic subtyping, where only static typechecking is required and implement this in SFJ \cite{UD20}: it is enough to define a method \emph{diagonal} that has an argument of type \emph{Polygon} \textbf{and not} \emph{Triangle} thus allowing the typechecker to check at compile time the restrictions on types:
+We propose a more elegant solution, by combining boolean types and semantic subtyping, where only static typechecking is required and implement this in SFJ \cite{UD20}: it is enough to define a method \emph{diagonal} that has an argument of type \emph{Polygon} \textbf{and not} \emph{Triangle}, thus allowing the typechecker to check at compile time the restrictions on types:
 $$
@@
         boolean resolved = true
-        for field that contains a class type:
+        for field in class that contains a class type:
             if class type is not resolved
@@
 %
-Note that \emph{Int} is not a subtype of \emph{Float} as the latter cannot represent the whole set of \emph{Int} values accurately and therefore \emph{Int} is not fully set-contained, however this is not the case for \emph{Int} and \emph{Double}.
+Note that \emph{Int} is not a subtype of \emph{Float} as a 32-bit \emph{float} cannot represent the whole set of 32-bit \emph{integer} values accurately and therefore \emph{Int} is not fully set-contained, however this is not the case for \emph{Int} and \emph{Double}.
 Similarly, \emph{Long} is not a subtype of \emph{Double}.
@@
 \begin{flalign*}
-    &Polygon = \left\{Polygon, Triangle, Rectangle, Rumble \right\} && Triangle  = \left\{Triangle\right\}\\ 
+    &Polygon = \left\{Polygon, Triangle, Rectangle, Rhombus \right\} && Triangle  = \left\{Triangle\right\}\\ 
     &Rectangle = \left\{Rectangle\right\} &&
@@
 \end{flalign*}
-If we pass to \emph{diagonal} an argument of type \emph{Triangle}, as it is not contained in the computed set of subtypes, we get a typechecking error at compile type \cite{UD20}.
+If we pass to \emph{diagonal} an argument of type \emph{Triangle}, as it is not contained in the computed set of subtypes, we get a type checking error at compile time \cite{UD20}.
 
@@
 It will find all nominal subtypes, as nominal subtypes inherit all fields and methods of their super class, so are always guaranteed to be a superset. 
-Similarly, given two types, the algorithm can check if one is a superset of the other, i.e.,\ they share common fields and methods, and if so, they are related by structural subtyping. 
+Similarly, given all pairs of types, the algorithm checks if one is a superset of the other, i.e.,\ they share common fields and methods, and if so, they are related by structural subtyping. 
 %For e.g., the Empty type, $empty = []$, will have all classes as structural subtypes. 
 %The above algorithm will without differentiating, find all nominal and structural subtypes as we always consider all other classes in the program to be potential subtype candidates.
-Consider two structurally equivalent types $coord = [x:int, y:int, z:int]$ and $colour = [x:int, y:int, z:int]$. 
-While they can be used interchangeably in our type system, the meaning of the types is completely different, meaning we ignore possible wrong behaviour in our program. 
+
+However, it can be argued that structural subtyping is counterproductive in writing logically correct programs.
+Consider two structurally equivalent types $coordinate = [x:int, y:int, z:int]$ and $colour = [x:int, y:int, z:int]$. 
+While they can be used interchangeably in our type system, the meaning of the types is completely different, meaning we ignore what should be a type check error in our program. 
 We however argue that it is no more unsafe than nominal subtyping in Java where an overridden method can be defined to perform the opposite logic to what the super class expects. 
@@
 
-Furthermore, we have extended FJ with method types, demonstrated below to show another possible implementation of the Polygons example, where we only implement the \textit{diagonal} method on polygons that support it, and not in the Triangle class.
+Furthermore, we have extended FJ with method types, demonstrated below to show another possible implementation of the Polygons example \autoref{polygons}.
 $$
@@
 
-While we do not generate a relation for each method in all classes as this is computationally unnecessary, we can on encountering a method type generate the set of classes which implement this method and combine their subtype relations using the \textit{or} operator.
-We can then use the same set inclusion to see if the argument passed to this function type checks.
+We instead only implement the \emph{diagonal} method on polygons that support it.
+We can then specify to only accept classes that implement the method \emph{diagonal}.
+If we pass an argument of type \emph{Triangle}, it is not in the computed set of subtypes, so we get a type checking error at compile time.
+We do not have to generate a relation for all methods in all classes as this is computationally unnecessary and instead
+only generate the set of subtypes on encountering a method type.
 
-In the implementation of SFJ, we are liberal in the use of fields and method which might not be implemented for all types in the subtyping relation rather than restricting only it to those which we know are shared by all types.
-Improper use of this allows for run-time exceptions however it more closely follows the idea of giving the programmer tools to create more expressive programs at the cost of checking program correctness such as with the use of structural subtypes.
+%In the implementation of SFJ, we are liberal in the use of fields and method which might not be implemented for all types in the subtyping relation rather than restricting only it to those which we know are shared by all types.
+%Improper use of this allows for run-time exceptions however it more closely follows the idea of giving the programmer tools to create more expressive programs at the cost of checking program correctness such as with the use of structural subtypes.
 

### 2020-02-27 conference/paper.tex
final pass, small amount of cutting still needed

@@
 Consider two structurally equivalent types $coordinate = [x:int, y:int, z:int]$ and $colour = [x:int, y:int, z:int]$. 
-While they can be used interchangeably in our type system, the meaning of the types is completely different, meaning we ignore what should be a type check error in our program. 
+While they can be used interchangeably in our type system, the meaning of the types is completely different, meaning we ignore what should likely be a type check error in our program. 
 We however argue that it is no more unsafe than nominal subtyping in Java where an overridden method can be defined to perform the opposite logic to what the super class expects. 
@@
 The starting point of their framework is a higher-order $\lambda-$calculus with pairs and projections.
+Muehlboeck and Tate \cite{Muehlboeck2018} define a syntactic framework with boolean connectives which has been implemented in the Ceylon programming language \cite{Ceylon2016}.
 Castagna \emph{et al.} \cite{Cpi} define $\mathbb{C}\pi$, a variant of the asynchronous $\pi-$calculus \cite{sangiorgi}, where channel types are augmented with boolean connectives.

### 2020-02-27 conference/paper.tex
cut multimethods explanations to fit page limits

@@
 SFJ integrates both structural and nominal subtyping, and the developer can choose which one to use.
-Finally, as discussed in Dardha \emph{et al.} \cite[\S 8.4]{Dardha2017}, we can easily encode in SFJ standard programming constructs and features of Java, such as lists, or overloading classes (via multimethods, see below), which are missing in FJ, making SFJ more complete and closer to Java.
+Finally, as discussed in Dardha \emph{et al.} \cite[\S 8.4]{Dardha2017}, we can easily encode in SFJ standard programming constructs and features of Java, such as lists, or overloading classes via multimethods \cite{BC97}, which are missing in FJ, making SFJ more complete and closer to Java.
 
@@
 \end{example}
-\emph{On multimethods}
-Since FJ is a core language, some features of the full Java are removed: overloading methods is one of them.
-In our framework, leveraging the expressivity of boolean connectives and semantic subtyping, many of these features, and language constructs, e.g., lists, are restored: overloading methods is one of them.
-We can thus model \emph{multimethods}, \cite{BC97}, which according to the authors is ``\emph{very clean and easy to understand [...] it would be the best solution for a brand new language}".
-As an example Dardha \emph{et al.}  \cite{Dardha2013,Dardha2017}, consider the following class declarations:
-$$
-\begin{array}{ll}
-\syndecl{A}{\mathit{Object}} {
-&
-\qquad
-\syndecl{B}{A} {
-\\
-	%\qquad\quad \ldots
-%&
-	%\qquad\qquad\quad \ldots
-%\\
-	\qquad\quad \mathbf{int}\ \mathit{length}\ (\mathbf{string}\ s)\{\ \ldots\ \}\qquad
-&
-	\qquad\qquad\quad \mathbf{int}\ \mathit{length}\ (\mathbf{int}\ n)\{\ \ldots\ \}
-\\
-}
-&
-\qquad }
-\end{array}
-$$
-Method {\it length} has type $\mathbf{string} \rightarrow \mathbf{int}$ in $A$.
-However, in $B$ it has type $(\mathbf{string} \rightarrow \mathbf{int}) \synwedge (\mathbf{int} \rightarrow \mathbf{int})$,
-which can be simplified as $(\mathbf{string} \synvee \mathbf{int}) \rightarrow \mathbf{int}$.
 

### 2020-02-27 conference/paper.tex
fix typo

@@
 
-We instead only implement the \emph{diagonal} method on polygons that support it.
-We can then specify to only accept classes that implement the method \emph{diagonal}.
+We instead only implement the \emph{getDiagonal} method on polygons that support it.
+We can then specify to only accept classes that implement the method \emph{getDiagonal}.
 If we pass an argument of type \emph{Triangle}, it is not in the computed set of subtypes, so we get a type checking error at compile time.

### 2020-02-28 conference/paper.tex
final pass on paper

@@
 \begin{abstract}
-\emph{Subtyping} is a key notion in programming languages, as it allows more flexibility in coding.
+{Subtyping} is a key notion in programming languages, as it allows more flexibility in coding.
 There are two approaches to defining subtyping relations: the \emph{syntactic} and the \emph{semantic} approach.
-In the latter, one starts by defining a model of the language and an interpretation of types as subsets of this model.
-Subtyping is defined as \emph{set inclusion of types} denoting sets.
-An orthogonal issue, typical of object-oriented languages, is the \emph{nominal} vs. \emph{structural} subtyping.
-Dardha \emph{et al.} \cite{Dardha2013,Dardha2017} combined structural subtyping and boolean types, on one hand, with nominal subtyping, on the other, in Featherweight Java (FJ), by thus exploiting the benefits of both approaches.
+%
+In semantic subtyping, one defines a model of the language and an interpretation of types as subsets of this model.
+Subtyping is defined as inclusion of subsets denoting types.
+%
+An orthogonal subtyping question, typical of object-oriented languages, is the \emph{nominal} vs. \emph{structural} subtyping.
+Dardha \emph{et al.} \cite{Dardha2013,Dardha2017} defined boolean types and semantic subtyping for Featherweight Java (FJ) and integrated both structural and nominal subtyping, thus exploiting the benefits of both approaches.
+However, these benefits were illustrated only at a theoretical level, but not exploited practically.
 
-However, these benefits were illustrated only at a theoretical level but not exploited practically.
-In this paper, we present SFJ---Semantic Featherweight Java, an implementation of FJ which features structural subtyping and boolean types as well as nominal subtyping.
-The benefits of SFJ, also illustrated in the accompanied video, show how static type checking of boolean types and semantic subtyping gives higher guarantees of program correctness, more flexibility and compactness of program writing.
+In this paper, we present SFJ---Semantic Featherweight Java, an implementation of FJ which features boolean types and structural subtyping as well as nominal subtyping.
+The benefits of SFJ, also illustrated in the accompanying video \cite{UD20}, show how static typechecking of boolean types and semantic subtyping gives higher guarantees of program correctness, more flexibility and compactness of program writing.
 %\keywords{Nominal subtyping \and Structural subtyping  \and Semantic Featherweight Java \and Object-oriented languages \and Boolean types \and Type theory.}
@@
 Syntactic subtyping is more mainstream in programming languages and is defined by means of a set of formal deductive subtyping rules.
-Semantic subtyping is less know and is defined as inclusion of sets denoting types.
-Starting from a formal model of the  language, we define an interpretation of types as subsets of this model, typically, subsets of values in the language.
-Then, subtyping is defined as inclusion of these subsets.
+Semantic subtyping is less know:  one defines a formal model of the  language and an interpretation of types as subsets of this model, ultimately, these will be subsets of values in the language.
+Then, subtyping is defined as set inclusion of these subsets denoting types.
+%
+Orthogonally, for object-oriented languages there are two approaches to defining subtyping relations: the \emph{nominal} and the \emph{structural} approach.
+Nominal subtyping is based on {declarations} by the developer and is \emph{name}-based: $A$ is a subtype of $B$ if and only if it is declared to be so, that is if the class $A$ extends (or implements) the class (or interface) $B$.
+Structural subtyping instead is based on the \emph{structure} of a class, its fields and methods: a class $A$ is a subtype of a class $B$ if and only if the fields and methods of $A$ are a superset of the fields and methods of $B$, and their types in $A$ are subtypes of the types in $B$.
+It is not surprising that the nominal subtyping aligns well with the syntactic approach, and in fact it is more well-known than the structural subtyping, which in turn aligns well with the semantic approach.
 
-Orthogonally, for object-oriented languages there are two approaches to defining subtyping relations: the \emph{nominal} and the \emph{structural} approach.
-Nominal subtyping is based on {declarations} by the developer and is {name}-based: $A$ is a subtype of $B$ if and only if it is declared to be so, that is if the class $A$ extends (or implements) the class (or interface) $B$.
-Structural subtyping instead is based on the \emph{structure} of a class, its fields and methods: a class $A$ is a subtype of a class $B$ if and only if the fields and methods of $A$ are a superset of the fields and methods of $B$, and their types in $A$ are subtypes of the types in $B$.
-At this point it is not surprising that the nominal subtyping aligns well with the syntactic approach and in fact is more well-known than the structural subtyping which in turn aligns well with the semantic approach.
-
-Dardha \emph{et al.} \cite{Dardha2013,Dardha2017} define semantic subtyping for \emph{Featherweight Java} (FJ) \cite{featherweight} with \emph{boolean types}.
-This approach allows the co-existence of nominal and structural subtyping in FJ, bringing in higher guarantees of program correctness, flexibility and compactness in programming.
+Dardha \emph{et al.} \cite{Dardha2013,Dardha2017} define {boolean types} and semantic subtyping for \emph{Featherweight Java} (FJ) \cite{featherweight}.
+This approach allows for the integration of both nominal and structural subtyping in FJ, bringing in higher guarantees of program correctness, flexibility and compactness in programming.
 Unfortunately, these benefits were only presented at a theoretical level and not exploited practically, due to the lack of an implementation of the language and its types and type system.
 
-In this paper, we present SFJ---\emph{Semantic Featherweight Java} \autoref{sec:design}, an implementation of FJ with integrated boolean types and semantic subtyping.
-In SFJ the developer has a larger and more expressive set of types, by using boolean connectives \textbf{and}, \textbf{not}, \textbf{or}.
+In this paper, we present SFJ---\emph{Semantic Featherweight Java} \autoref{sec:design}, an implementation of FJ with boolean types and semantic subtyping.
+In SFJ the developer has a larger and more expressive set of types, by using boolean connectives \textbf{and}, \textbf{not}, \textbf{or}, with the expected set-theoretic interpretation.
 On the other hand, this added expressivity does not introduce added complexity, but rather the opposite is true, as the developer has an easier, more compact and more elegant way of programming.
 SFJ integrates both structural and nominal subtyping, and the developer can choose which one to use.
-Finally, as discussed in Dardha \emph{et al.} \cite[\S 8.4]{Dardha2017}, we can easily encode in SFJ standard programming constructs and features of Java, such as lists, or overloading classes via multimethods \cite{BC97}, which are missing in FJ, making SFJ more complete and closer to Java.
+Finally, as discussed in Dardha \emph{et al.} \cite[\S 8.4]{Dardha2017}, thanks to semantic subtyping, we can easily encode in SFJ standard programming constructs and features of Java, such as lists, or overloading classes via multimethods \cite{BC97}, which are missing in FJ, thus making SFJ a more complete language closer to Java.
 
@@
 Now, suppose that our class hierarchy is such that \emph{Polygon} is the parent class and all other geometric figure classes extend \emph{Polygon}, which is how one would naturally define the set of polygons.
-If we think of the class hierarchy as given and part of legacy code, which cannot be changed, then again a natural way to implement this in Java is by defining the method \emph{diagonal} in the class \emph{Polygon} and using an \textbf{instanceof}, for e.g., inside a \textbf{try-catch} that throws an exception at run time, if the argument passed to the method is a triangle.
+If we think of the class hierarchy as given and part of legacy code, which cannot be changed, then again a natural way to implement this in Java is by defining the method \emph{diagonal} in the class \emph{Polygon} and using an \textbf{instanceof}, for e.g., inside a \textbf{try-catch} which throws an exception at runtime, if the argument passed to the method is a triangle.
 
@@
 
-The technical developments behind semantic subtyping and its properties are not easy, however, they are completely transparent to the developer.
-Due to space limit we will remove theoretical details from this submission, as they are not necessary to understand the design and implementation of SFJ, which is the focus of this paper.
-We refer the reader to the relevant work \cite{Dardha2013,Dardha2017} and we will include the background on semantic subtyping in a possible full paper.
+The technical developments behind semantic subtyping and its properties are complex, however, they are completely transparent to the developer.
+Due to space limit we will remove most theoretical details from this submission, as they are not necessary to understand the design and implementation of SFJ, which is the focus of this paper.
+We refer the reader to the relevant work \cite{Dardha2013,Dardha2017} and we will include a more complete background on semantic subtyping in a possible full paper.
 
@@
 \\
-\alpha & ::=  & \zero \ |\ \btypes \ |\ [\wt{l:\tau}] \ |\ \alpha \synwedge \alpha \ |\ \synneg\alpha
+\alpha & ::=  & \zero \ |\ \btypes \ |\ [\wt{l:\tau}] \ |\ \alpha \ \bf{and}\ \alpha \ |\ \bf{not}\ \alpha
 \qquad
@@
 \\
-\mu & ::=  & \alpha \to \alpha \ |\ \mu \synwedge \mu \ |\ \synneg\mu
+\mu & ::=  & \alpha \to \alpha \ |\ \mu \ \bf{and}\ \mu \ |\ \bf{not}\ \mu
 & \mbox{{\em Method type} ($\mu $-type)}
@@
 Arrow types $\alpha \to \alpha$ are used to type methods.
-The boolean types using $\synwedge$ and $\synneg$ have their expected meanings and $\synvee$ is obtained by their combination.\\
+The boolean types using $\bf{and}$ and $\bf{not}$ have their expected set-theoretic meanings, and $\bf{or}$ is obtained by their combination.\\
 The syntax of terms is given by the following grammar and is based on the standard syntax of terms in FJ \cite{featherweight,Dardha2013,Dardha2017}:
@@
 Now, we can define the subtyping relation.
-Building upon the types-as-sets-of-values approach of the semantic subtyping for FJ \cite{Dardha2013,Dardha2017}, in our implementation of SFJ we keep track of this relation by defining a map from a type to the set of its subtypes, with the property that the set of values of a subtype is included in the set of values of the type.
+Building upon the interpretation of types as sets of values to define the semantic subtyping for FJ \cite{Dardha2013,Dardha2017}, in our implementation of SFJ we keep track of the subtyping relation by defining a map from a type to the set of its subtypes, with the property that the set of values of a subtype is included in the set of values of the type.
 As a first step, we start with basic types.
@@
 Similarly, \emph{Long} is not a subtype of \emph{Double}.
-Finally, the following algorithm defines the subtyping relation for all class types.
+
+Finally, the following algorithm defines the subtyping relation for all class types, which concludes the process.
 \begin{verbatim}
@@
 \end{verbatim}
-Let us illustrate the subtyping relation algorithm with our \emph{Polygons} example \autoref{polygons}.
+Let us illustrate the subtyping algorithm with our \emph{Polygons} example \autoref{polygons}.
 The algorithm generates the following subtyping relation, which also includes the subtyping relation for basic types defined in the first step.
@@
 \end{flalign*}
-If we pass to \emph{diagonal} an argument of type \emph{Triangle}, as it is not contained in the computed set of subtypes, we get a type checking error at compile time \cite{UD20}.
+If we pass to \emph{diagonal} an argument of type \emph{Triangle}, as it is not contained in the computed set of subtypes, we get a typechecking error at compile time \cite{UD20}.
 
@@
 Similarly, given all pairs of types, the algorithm checks if one is a superset of the other, i.e.,\ they share common fields and methods, and if so, they are related by structural subtyping. 
-%For e.g., the Empty type, $empty = []$, will have all classes as structural subtypes. 
+For e.g., type $empty = []$, will have all classes as structural subtypes. 
 %The above algorithm will without differentiating, find all nominal and structural subtypes as we always consider all other classes in the program to be potential subtype candidates.
 
-However, it can be argued that structural subtyping is counterproductive in writing logically correct programs.
-Consider two structurally equivalent types $coordinate = [x:int, y:int, z:int]$ and $colour = [x:int, y:int, z:int]$. 
-While they can be used interchangeably in our type system, the meaning of the types is completely different, meaning we ignore what should likely be a type check error in our program. 
-We however argue that it is no more unsafe than nominal subtyping in Java where an overridden method can be defined to perform the opposite logic to what the super class expects. 
-Therefore both methods leave an expectation on the programmer to check what they are doing is correct.
+Note that both approaches to subtyping have their drawbacks, hence their integration is the key to overcoming them, and in SFJ the developer can choose which approach to use for a given task.
+%However, it can be argued that structural subtyping is counterproductive in writing logically correct programs.
+Consider two structurally equivalent types $coordinate = [x:int, y:int]$ and $colour = [x:int, y:int]$. 
+While they can be used interchangeably in our type system, their ``meaning'' is completely different.
+On the other hand, with nominal subtyping in Java one can define an overridden method to perform the opposite logic to what the super class expects. 
+Therefore both approaches leave an expectation on the programmer to check what they are doing is correct.
 
@@
 We can then specify to only accept classes that implement the method \emph{getDiagonal}.
-If we pass an argument of type \emph{Triangle}, it is not in the computed set of subtypes, so we get a type checking error at compile time.
+If we pass an argument of type \emph{Triangle}, it is not in the computed set of subtypes, so we get a typechecking error at compile time.
 We do not have to generate a relation for all methods in all classes as this is computationally unnecessary and instead
@@
 
-To conclude, in this paper we presented the design and implementation of SFJ, an extension of Featherweight Java with boolean types and semantic subtyping, which combined allow for both structural and nominal subtyping in FJ as well as restore standard Java constructs for e.g., lists and features for e.g., overloading which were not present in FJ.
+To conclude, in this paper we presented the design and implementation of SFJ, an extension of Featherweight Java with boolean types and semantic subtyping, which allow for both structural and nominal subtyping in FJ as well as restore standard Java constructs for e.g., lists and features for e.g., overloading which were not present in FJ, making SFJ a more complete language.
 

### 2020-02-28 conference/paper.tex
more on nominal vs structural

@@
 
-Note that both approaches to subtyping have their drawbacks, hence their integration is the key to overcoming them, and in SFJ the developer can choose which approach to use for a given task.
+Note that both approaches to subtyping have their drawbacks.
 %However, it can be argued that structural subtyping is counterproductive in writing logically correct programs.
-Consider two structurally equivalent types $coordinate = [x:int, y:int]$ and $colour = [x:int, y:int]$. 
-While they can be used interchangeably in our type system, their ``meaning'' is completely different.
-On the other hand, with nominal subtyping in Java one can define an overridden method to perform the opposite logic to what the super class expects. 
-Therefore both approaches leave an expectation on the programmer to check what they are doing is correct.
+Consider two structurally equivalent record types $coordinate = [x:int, y:int, z:int]$ and $colour = [x:int, y:int, z:int]$. 
+While structurally they can be used interchangeably in our type system, their ``meaning'' is completely different.
+On the other hand, with nominal subtyping in Java one can define an overridden method to perform the opposite logic to what the super class is expecting. 
+Therefore, both approaches leave an expectation on the developer to check what they are doing is correct, hence the integration of both in SFJ is the key to overcoming these drawbacks, as one can choose which approach to use for a given task.
 

### 2020-02-28 conference/paper.tex
added Square

@@
 
-Consider the set of polygons, such as triangles, rectangles and rhombuses given by a class hierarchy.
+Consider the set of polygons, such as triangles, squares and rhombuses given by a class hierarchy.
 We want to define a method \emph{diagonal} that takes a polygon and returns the length of its longest diagonal.
@@
 \\
-\syndecl{Rectangle}{Other\_Polygons}{\ldots}
+\syndecl{Square}{Other\_Polygons}{\ldots}
 \\
@@
 \\
-\synidecl{Rectangle}{Polygon} \synimpl Diagonal \ \{\ldots\}
+\synidecl{Square}{Polygon} \synimpl Diagonal \ \{\ldots\}
 \\
@@
 \\
-\syndecl{Rectangle}{Polygon} {\ldots}
+\syndecl{Square}{Polygon} {\ldots}
 \\
@@
 \begin{flalign*}
-    &Polygon = \left\{Polygon, Triangle, Rectangle, Rhombus \right\} && Triangle  = \left\{Triangle\right\}\\ 
-    &Rectangle = \left\{Rectangle\right\} &&
+    &Polygon = \left\{Polygon, Triangle, Square, Rhombus \right\} && Triangle  = \left\{Triangle\right\}\\ 
+    &Square = \left\{Square\right\} &&
     Diagonal = \left\{Diagonal\right\} 
@@
 \begin{flalign*}
-    Polygon \textbf{ and not } Triangle = \left\{Polygon, Rectangle, Rhombus \right\} 
+    Polygon \textbf{ and not } Triangle = \left\{Polygon, Square, Rhombus \right\} 
 \end{flalign*}

### 2020-02-28 conference/paper.tex
links and final pass

@@
 \\
-	\qquad \mathbf{real}\ diagonal(Other\_Polygons \ p)\ \{\ldots \}
+	\qquad \mathbf{double}\ diagonal(Other\_Polygons \ shape)\ \{\ldots \}
 	\\
@@
 	\\
-	\qquad \mathbf{real}\ diagonal(Polygon \ p);
+	\qquad \mathbf{double}\ diagonal(Polygon \ shape);
 	\\
@@
 \begin{array}{l}
-\syndecl{Polygon}{Object} {\ldots}
+\synclass Polygon \ \{\ldots\}
 \\
@@
 \\
-\syndecl{Diagonal}{Object} {
+\synclass Diagonal \ \{
 	\\
-	\qquad \mathbf{real}\ diagonal(\textit{Polygon} \textbf{ and not }  \textit{Triangle} \ p)\{\ldots\}
+	%\qquad Diagonal ( )\   \{\ super();\ \}
+	\hspace{2.5cm} \ldots\\
+	\qquad \mathbf{double}\ diagonal((\textit{Polygon} \textbf{ and not }  \textit{Triangle}) \ shape)\{\ldots\}
 	\\
-}
+\}
 \end{array}
@@
 \begin{flalign*}
-    &Double  = \left\{Double, Float, Int, Short, Byte\right\} &&Float = \left\{Float, Short, Byte\right\}\\
-    &Long = \left\{Long, Int, Short, Byte\right\}  &&Int = \left\{Int, Short, Byte\right\}\\
-    &Short = \left\{Short, Byte\right\} && Byte = \left\{Byte\right\}\\
-    &Boolean = \left\{Boolean \right\} 
+    &Double  = \left\{Double, Float, Int, Short, Byte\right\} && Float = \left\{Float, Short, Byte\right\}\\
+    &Long = \left\{Long, Int, Short, Byte\right\}  && Int = \left\{Int, Short, Byte\right\}\\
+    &Short = \left\{Short, Byte\right\} \qquad Byte = \left\{Byte\right\} && Boolean = \left\{Boolean \right\} 
 \end{flalign*}
@@
 \end{verbatim}
-Let us illustrate the subtyping algorithm with our \emph{Polygons} example \autoref{polygons}.
+Let us illustrate the subtyping algorithm with our \emph{Polygons} \autoref{polygons}.
 The algorithm generates the following subtyping relation, which also includes the subtyping relation for basic types defined in the first step.
 \begin{flalign*}
-    &Polygon = \left\{Polygon, Triangle, Square, Rhombus \right\} && Triangle  = \left\{Triangle\right\}\\ 
-    &Square = \left\{Square\right\} &&
+    &Polygon = \left\{Polygon, Triangle, Square, Rhombus \right\} \qquad Triangle = \left\{Triangle\right\}\\ 
+    &Square = \left\{Square\right\} \qquad  Rhombus  = \left\{Rhombus\right\} \qquad
     Diagonal = \left\{Diagonal\right\} 
@@
 The subtyping algorithm finds all nominal and structural subtypes.
-It will find all nominal subtypes, as nominal subtypes inherit all fields and methods of their super class, so are always guaranteed to be a superset. 
+It finds all nominal subtypes as they inherit all fields and methods of their super class, so are always guaranteed to be a superset. 
 Similarly, given all pairs of types, the algorithm checks if one is a superset of the other, i.e.,\ they share common fields and methods, and if so, they are related by structural subtyping.
@@
 %However, it can be argued that structural subtyping is counterproductive in writing logically correct programs.
-Consider two structurally equivalent record types $coordinate = [x:int, y:int, z:int]$ and $colour = [x:int, y:int, z:int]$. 
+Consider two structurally equivalent class/record types $coordinate = [x:int, y:int, z:int]$ and $colour = [x:int, y:int, z:int]$. 
 While structurally they can be used interchangeably in our type system, their ``meaning'' is completely different.
@@
 
-Furthermore, we have extended FJ with method types, demonstrated below to show another possible implementation of the Polygons example \autoref{polygons}.
+Furthermore, we have extended FJ with method types, demonstrated below to show another possible implementation of the \emph{Polygons} \autoref{polygons}.
 $$
-\syndecl{Diagonal}{Object} {
-	\mathbf{int}\ diagonal(\textit{(getDiagonal: void$\rightarrow$int)}  \ p)\{\ldots\}
-}
+\synclass Diagonal \ \{
+	\mathbf{int}\ diagonal(\textit{(getDiagonal: void$\rightarrow$int)}  \ shape)\{\ldots\}
+\}
 $$

### 2020-02-28 conference/paper.tex
audio and subtittles

@@
 In this paper, we present SFJ---Semantic Featherweight Java, an implementation of FJ which features boolean types and structural subtyping as well as nominal subtyping.
-The benefits of SFJ, also illustrated in the accompanying video \cite{UD20}, show how static typechecking of boolean types and semantic subtyping gives higher guarantees of program correctness, more flexibility and compactness of program writing.
+The benefits of SFJ, illustrated in the video (with audio and subtitles) \cite{UD20}, show how static typechecking of boolean types and semantic subtyping gives higher guarantees of program correctness, more flexibility and compactness of program writing.
 %\keywords{Nominal subtyping \and Structural subtyping  \and Semantic Featherweight Java \and Object-oriented languages \and Boolean types \and Type theory.}

### 2020-02-28 conference/paper.tex
minor

--- 2020-02-28 conference/paper.tex
@@
 In this paper, we present SFJ---Semantic Featherweight Java, an implementation of FJ which features boolean types and structural subtyping as well as nominal subtyping.
-The benefits of SFJ, also illustrated in the accompanying video \cite{UD20}, show how static typechecking of boolean types and semantic subtyping gives higher guarantees of program correctness, more flexibility and compactness of program writing.
+The benefits of SFJ, illustrated in the paper and the video (with audio/subtitles) \cite{UD20}, show how static typechecking of boolean types and semantic subtyping gives higher guarantees of program correctness, more flexibility and compactness of program writing.
 %\keywords{Nominal subtyping \and Structural subtyping  \and Semantic Featherweight Java \and Object-oriented languages \and Boolean types \and Type theory.}

### 2020-03-22 conference/paper.tex
finish introduction

--- 2020-02-28 conference/paper.tex
@@
 \\
-	\qquad \mathbf{double}\ diagonal(Other\_Polygons \ shape)\ \{\ldots \}
+	\qquad \mathbf{real}\ diagonal(Other\_Polygons \ p)\ \{\ldots \}
 	\\
@@
 	\\
-	\qquad \mathbf{double}\ diagonal(Polygon \ shape);
+	\qquad \mathbf{real}\ diagonal(Polygon \ p);
 	\\
@@
 \begin{array}{l}
-\synclass Polygon \ \{\ldots\}
+\syndecl{Polygon}{Object} {\ldots}
 \\
@@
 \\
-\synclass Diagonal \ \{
+\syndecl{Diagonal}{Object} {
 	\\
-	%\qquad Diagonal ( )\   \{\ super();\ \}
-	\hspace{2.5cm} \ldots\\
-	\qquad \mathbf{double}\ diagonal((\textit{Polygon} \textbf{ and not }  \textit{Triangle}) \ shape)\{\ldots\}
+	\qquad \mathbf{real}\ diagonal(\textit{Polygon} \textbf{ and not }  \textit{Triangle} \ p)\{\ldots\}
 	\\
-\}
+}
 \end{array}
@@
 The syntax of types is given by the following grammar \cite{Dardha2013,Dardha2017}:
+
 $$
@@
 $$
+
 $\alpha$-types are used to type fields and $\mu$-types are used to type methods.
@@
 \begin{flalign*}
-    &Double  = \left\{Double, Float, Int, Short, Byte\right\} && Float = \left\{Float, Short, Byte\right\}\\
-    &Long = \left\{Long, Int, Short, Byte\right\}  && Int = \left\{Int, Short, Byte\right\}\\
-    &Short = \left\{Short, Byte\right\} \qquad Byte = \left\{Byte\right\} && Boolean = \left\{Boolean \right\} 
+    &Double  = \left\{Double, Float, Int, Short, Byte\right\} &&Float = \left\{Float, Short, Byte\right\}\\
+    &Long = \left\{Long, Int, Short, Byte\right\}  &&Int = \left\{Int, Short, Byte\right\}\\
+    &Short = \left\{Short, Byte\right\} && Byte = \left\{Byte\right\}\\
+    &Boolean = \left\{Boolean \right\}
 \end{flalign*}
@@
 \end{verbatim}
-Let us illustrate the subtyping algorithm with our \emph{Polygons} \autoref{polygons}.
+Let us illustrate the subtyping algorithm with our \emph{Polygons} example \autoref{polygons}.
 The algorithm generates the following subtyping relation, which also includes the subtyping relation for basic types defined in the first step.
 \begin{flalign*}
-    &Polygon = \left\{Polygon, Triangle, Square, Rhombus \right\} \qquad Triangle = \left\{Triangle\right\}\\ 
-    &Square = \left\{Square\right\} \qquad  Rhombus  = \left\{Rhombus\right\} \qquad
+    &Polygon = \left\{Polygon, Triangle, Square, Rhombus \right\} && Triangle  = \left\{Triangle\right\}\\
+    &Square = \left\{Square\right\} &&
     Diagonal = \left\{Diagonal\right\}
@@
 The subtyping algorithm finds all nominal and structural subtypes.
-It finds all nominal subtypes as they inherit all fields and methods of their super class, so are always guaranteed to be a superset. 
+It will find all nominal subtypes, as nominal subtypes inherit all fields and methods of their super class, so are always guaranteed to be a superset.
 Similarly, given all pairs of types, the algorithm checks if one is a superset of the other, i.e.,\ they share common fields and methods, and if so, they are related by structural subtyping.
@@
 %However, it can be argued that structural subtyping is counterproductive in writing logically correct programs.
-Consider two structurally equivalent class/record types $coordinate = [x:int, y:int, z:int]$ and $colour = [x:int, y:int, z:int]$. 
+Consider two structurally equivalent record types $coordinate = [x:int, y:int, z:int]$ and $colour = [x:int, y:int, z:int]$.
 While structurally they can be used interchangeably in our type system, their ``meaning'' is completely different.
@@
 
-Furthermore, we have extended FJ with method types, demonstrated below to show another possible implementation of the \emph{Polygons} \autoref{polygons}.
+Furthermore, we have extended FJ with method types, demonstrated below to show another possible implementation of the Polygons example \autoref{polygons}.
 $$
-\synclass Diagonal \ \{
-	\mathbf{int}\ diagonal(\textit{(getDiagonal: void$\rightarrow$int)}  \ shape)\{\ldots\}
-\}
+\syndecl{Diagonal}{Object} {
+	\mathbf{int}\ diagonal(\textit{(getDiagonal: void$\rightarrow$int)}  \ p)\{\ldots\}
+}
 $$

### 2020-03-22 dissertation/l4proj.tex
finish introduction

@@
 
-%
-% put any additional packages here
-%
+\input{semmacros}
 
@@
 
-%==============================================================================
-%% METADATA
+%===================================================================================================
+
 \title{Boolean types and semantic subtyping for Featherweight Java}
@@
 
-%==============================================================================
-%% ABSTRACT
+%===================================================================================================
+
 \begin{abstract}
@@
 
-%==============================================================================
+%===================================================================================================
 
-% EDUCATION REUSE CONSENT FORM
-% If you consent to your project being shown to future students for educational purposes
-% then insert your name and the date below to  sign the education use form that appears in the front of the document. 
-% You must explicitly give consent if you wish to do so.
-% If you sign, your project may be included in the Hall of Fame if it scores particularly highly.
-%
-% Please note that you are under no obligation to sign 
-% this declaration, but doing so would help future students.
-%
-\def\consentname {Artem Usov} % your full name
-\def\consentdate {19 January 2020} % the date you agree
+\def\consentname {Artem Usov}
+\def\consentdate {19 January 2020}
 \educationalconsent
 
-%==============================================================================
+%===================================================================================================
+
 \tableofcontents
 
-%==============================================================================
-%% Notes on formatting
-%==============================================================================
-% The first page, abstract and table of contents are numbered using Roman numerals and are not
-% included in the page count. 
-%
-% From now on pages are numbered
-% using Arabic numerals. Therefore, immediately after the first call to \chapter we need the call
-% \pagenumbering{arabic} and this should be called once only in the document. 
-%
-% Do not alter the bibliography style.
-%
-% The first Chapter should then be on page 1. You are allowed 40 pages for a 40 credit project and 30 pages for a 
-% 20 credit report. This includes everything numbered in Arabic numerals (excluding front matter) up
-% to but excluding the appendices and bibliography.
-%
-% You must not alter text size (it is currently 10pt) or alter margins or spacing.
-%
-%
-%==================================================================================================================================
-%
-% IMPORTANT
-% The chapter headings here are **suggestions**. You don't have to follow this model if
-% it doesn't fit your project. Every project should have an introduction and conclusion,
-% however. 
-%
-%==================================================================================================================================
+%===================================================================================================
+
 \chapter{Introduction}
 
-% reset page numbering. Don't remove this!
 \pagenumbering{arabic}
 
-Why should the reader care about what are you doing and what are you actually doing?
-\section{Guidance}
+\section{End goal of programming}
 
-\textbf{Motivate} first, then state the general problem clearly.
+The end goal in all programming projects, is to create a solution to the task at hand that performs exactly as expected by the programmer, and exactly solves the problem that was initially defined.
+However, this very rarely actually happens.
 
-\section{Writing guidance}
-\subsection{Who is the reader?}
+On the one hand, this is often because the original problem is not statically defined.
+If it is a solution for an external client, their requirements and needs will change over time, or the needs and wants of the users using this project can also change.
+This means the project needs to be adapted over time.
+On the other hand, and almost always much less evident, is that programs do not perform exactly as the programmer imagines them to work.
 
-This is the key question for any writing. Your reader:
+These are in general called \emph{bugs} in a program, and have become more and more common as the projects we create have become more complex compared to the early days of computing and the underlying hardware we work on has also become more complex.
+Therefore in both academia and industry, there have been great efforts over the years to create tools that allow us to decrease the amount of bugs in our programs and increase our productivity.
+
+\subsection{Static Code Analysis}
+
+One area of great effort has been in the development of static code analysis tools.
+Some of these exist integrated into the language itself such as with Spark \citep{Carre1990}.
+Spark is a formally defined language based on the Ada \citep{Ada1979} language intended for the development of high integrity software such as flight control systems.
+Spark uses contracts and flow analysis to logically prove that programs are correct.
+Others exist as separate, well-known tools that exist alongside the language such as the Clang analysis tools for C and C++ \citep{kremenek2008}.
+
+While they are not necessarily needed to write correct programs, these tools have quickly become industry standards for maintaining a level of correctness within codebases and avoiding typical programming errors.
+
+\subsection{Into Static Analysis}
+
+Underlying all of the previously mentioned static code analysis tools, is their use of the \textbf{type system} of the language.
+Type systems at their most basic work by assigning types to various constructs of a program so that incompatible types cannot be used.
+
+For example, one common area of error in C code that the Clang tools warn about is implicit conversions where the value of an expression is a different type from the one expected, such as the use of a floating point number where an integer is expected.
+The C language will implicitly convert the value to an integer, however this can cause unexpected behaviour for the programmer.
+It knows that this conversion needs to happen as the type system dictates that you cannot use any other type other than an integer where an integer is expected.
+
+\section{Varying Type Systems}
+
+The effectiveness of static analysis tools and more generally the \emph{safety} and \emph{expressivity} of a programming language therefore depends on the type system that is used.
+Type systems have evolved from the relatively simple systems in languages like C to more complex ones such as in Java where there are custom object types and subtyping relations where types can be substituted for one another.
+
+More modern advances in type systems include the use of linear types in the Cyclone \citep{grossman2002} language that allow us to more naturally and logically define finite state machines.
+There exist other new developments in type systems such as Boolean types which are explored in this paper.
+
+\section{Goals of this Project}
+
+In this section, the issue and aims of this paper are presented.
+
+\subsection{Problem Statement}
+
+The typical hierarchical subtyping definitions in Java quite often restrict and complicate the logic that the programmer wants to implement.
+These problems come up particularly often when working with legacy code that has defines incorrect abstractions in its classes.
+This therefore leads to less understandable and less maintainable code, that could cause problems in the future.
+
+\subsection{Aims}
+
+The aim of this project is to implement a new programming language that uses a proposed type system consisting of boolean types and a semantic subtyping algorithm.
+This project aims to transform the mathematical definition of such a type system into an actual solution in software, solving the issues encountered doing so.
+The language should be able to illustrate the improved solutions to the
+common example problems encounter in a language without these features such as Java.
+Furthermore, the language will be evaluated so see whether these additional language features are intuitive to use by programmers.
+
+\section{Dissertation Outline}
+
+The dissertation is structured into seven chapters as follow:
 
@@
     \item
-          is a trained computer scientist: \emph{don't explain basics}.
+          \textbf{Chapter 2} provides background about type systems and Featherweight Java, the language on top of which the new type system is implemented.
+We also discuss some existing work on semantic type systems and boolean types.
     \item
-          has limited time: \emph{keep on topic}.
+          \textbf{Chapter 3} TODO
     \item
-          has no idea why anyone would want to do this: \emph{motivate clearly}
+          \textbf{Chapter 4} TODO
     \item
-          might not know \emph{anything} about your project in particular:
-          \emph{explain your project}.
+          \textbf{Chapter 5} TODO
     \item
-          but might know precise details and check them: \emph{be precise and
-              strive for accuracy.}
+          \textbf{Chapter 6} contains details about the evaluation method, evaluation results and analysis of results.
     \item
-          doesn't know or care about you: \emph{personal discussions are
-              irrelevant}.
+          Finally \textbf{chapter 8} concludes the dissertation with a summary of the important results.
 \end{itemize}
 
-Remember, you will be marked by your supervisor and one or more members
-of staff. You might also have your project read by a prize-awarding
-committee or possibly a future employer. Bear that in mind.
+%===================================================================================================
 
-\subsection{References and style guides}
-There are many style guides on good English writing. You don't need to
-read these, but they will improve how you write.
-
-\begin{itemize}
-    \item
-          \emph{How to write a great research paper} \cite{Pey17} (\textbf{recommended}, even
-          though you aren't writing a research paper)
-    \item
-          \emph{How to Write with Style} \cite{Von80}. Short and easy to read. Available online.
-    \item
-          \emph{Style: The Basics of Clarity and Grace} \cite{Wil09} A very popular modern English 
-          style guide.
-    \item
-          \emph{Politics and the English Language} \cite{Orw68}  A famous essay on effective, clear
-          writing in English.
-    \item
-          \emph{The Elements of Style} \cite{StrWhi07} Outdated, and American, but a classic.
-    \item
-          \emph{The Sense of Style} \cite{Pin15} Excellent, though quite in-depth.
-\end{itemize}
-
-\subsubsection{Citation styles}
-
-\begin{itemize}
-    \item If you are referring to a reference as a noun, then cite it as: ``\citet{Orw68} discusses
-          the role of language in political thought.''
-    \item If you are referring implicitly to references, use: ``There are many good books on writing
-          \citep{Orw68, Wil09, Pin15}.''
-\end{itemize}
-
-There is a complete guide on good citation practice by Peter Coxhead available here: 
-\url{http://www.cs.bham.ac.uk/~pxc/refs/index.html}. 
-If you are unsure about how to cite online sources, please see this guide: 
-\url{https://student.unsw.edu.au/how-do-i-cite-electronic-sources}.
-
-\subsection{Plagiarism warning}
-
-\begin{highlight_title}{WARNING}
-
-    If you include material from other sources without full and correct attribution, you are 
-    commiting plagiarism. The penalties for plagiarism are severe. Quote any included text and cite
-    it correctly. Cite all images, figures, etc. clearly in the caption of the figure.
-\end{highlight_title}
-
-%==================================================================================================================================
 \chapter{Background}
-What did other people do, and how is it relevant to what you want to do?
 
@@
 
-Type systems are by far the most popular and best developed methods in modern software engineering 
-to reduce the number of run-time errors in a program. A possible definition of a type system is 
-given by \citet{Pierce2002} as:
+In all modern programming languages, a type systems is arguably the main way that incorrect behaviour in a program is reduced.
+A possible definition of a type system is given by \citet{Pierce2002} as:
 
-\textit{A type system is a tractable syntactic method for proving the absence of certain program
-    behaviors by classifying phrases according to the kinds of values they compute.}
+\emph{A type system is a tractable syntactic method for proving the absence of certain program behaviors by classifying phrases according to the kinds of values they compute.}
 
-In particular, the most important detail in the above definition is the emphasis on classifying 
-phrases in our language into a particular type based on the kinds of values they hold or compute. 
-For example this would allow us to check that the arguments given to an arithmetic operation are 
-always numbers. This therefore allows us to validate that a program is absent of certain bad 
-run-time behaviour.
+In particular, the most important detail in the above definition is the emphasis on classifying phrases in our language into a particular type based on the kinds of values they hold or compute.
+For example by using a type system, we can check that the arguments given to an arithmetic operation are always numbers.
+Therefore we can validate that a specific program is absent of certain bad run-time behaviour such as trying to add a boolean value \emph{true} to a number \emph{2}.
 
-However what the above definition is missing that is it defines a type system as solely a syntactic 
-method, where it is defined by a formal system of deduction rules, which is by far the most common 
-approach for a type system. There does exist an alternative method in the semantic approach. 
-\citet{Frisch2002} describe the semantic approach as instead starting with a model of the language 
-and all the possible values in the language and define an interpretation of types as subsets of 
-values the model.
+However what the above definition is inaccurate about is that is it defines a type system to be solely a syntactic method, in which the system is defined as a list of formal deduction rules.
+The syntactic approach is certainly by far the most common approach for a type system.
+However, there does exist an alternative method in the semantic approach.
+\citet{Frisch2002} describe the semantic approach as instead starting with a model of the language and all the possible values in the language and defines an interpretation of types as subsets of values the model.
 
-Such as model comes with several advantages, such as when $t \leq s$ does not hold, then it is
-possible to exhibit an element of the model which is in the interpretation of type $t$ and not $s$, 
-even in the presence of arrow \textit{(function)} types. This can then be used to show more 
-informative error messages to the programmer \citep{Castagna2005}. However due to being a more 
-technical, such as it being not trivial to actually define the interpretation of types as subsets of
-a model, it has received less attention than syntactic methods.
+Such as model comes with several advantages.
+For example, given two types $s$ and $t$, which represent a certain subset of values in the model, when $s \subseteq t$ does not hold, then it is possible to exhibit an element of the model which is in the interpretation of type $s$ and not $t$.
+This can then be used to show more informative error messages to the programmer \citep{Castagna2005}, such as showing which value in $s$ is causing the equality not to hold.
+However due to being a more technical approach, such as the non-triviality of defining the interpretation of types as subsets of a model, the semantic method has received less attention than syntactic methods.
 
-The first problem we encounter in the semantic approach is that in the semantic approach, $t_{1}$
-is a subtype of $t_{2}$ if all the $t_{1}$-values are also $t_{2}$-values. However, in this way, 
-subtyping is defined by relying on the notion of well-typed values; hence, we need the typing 
-relation to determine typing judgements for values; but the typing rules use the subtyping relation
-which we are going to define \citep{Dardha2017}. So, there is a circularity as demonstrated in 
-Figure \ref{fig:circ}.
+The first problem we encounter in the semantic approach is that in such a model, $t_{1}$ is a subtype of $t_{2}$ if all the $t_{1}$-values are also $t_{2}$-values, i.e.
+the equality $t_{1} \subseteq t_{2}$ holds.
+However, in this way, subtyping is defined by relying on the notion of well-typed values; hence, we need the typing relation to determine typing judgements for values; but the typing rules use the subtyping relation defined in \citep{Dardha2017}.
+So, there is a circularity in our definition, as demonstrated in Figure \ref{fig:circ}.
 
-To solve this problem, we follow the framework defined in \citep{Frisch2008}. The general idea of
-the framework is that we first extend the types in the language with \textit{Boolean Combinators}: 
-union $\lor$, intersection $\land$ and negation $\neg$. We then define an abstract model $\beta$ 
-where our types are interpreted according to the set-theoretic meaning of the combinators and can 
-define subtyping as set-inclusion and therefore define the typing rules. Separately, if we have a 
-notion of well-typed values, we can define another interpretation of types as sets of values. We can
-again define a subtyping relation $\nu$. While these may be different relation, if the models are 
-chosen carefully such that 
-\begin{equation}
+To solve this problem, we follow the framework defined by \citet{Frisch2008}.
+The general idea of the framework is that we first extend the types in the language with \emph{Boolean Combinators}: union $\lor$, intersection $\land$ and negation $\neg$.
+We can then define an abstract model $\beta$ where our types are interpreted according to the set-theoretic meaning of the combinators and can define subtyping as set-inclusion and therefore define the typing rules.
+Separately, if we have a notion of well-typed values, we can define another interpretation of types as sets of values.
+We can again define a subtyping relation $\nu$.
+While these may be different relation, if the models are chosen carefully such that:
+\begin{equation*}
     s \leq_{\beta} t \iff s \leq_{\nu} t,
-\end{equation}
+\end{equation*}
+
 then these models coincide and this closes the circularity.
@@
     \includegraphics[width=0.4\linewidth]{images/circularity.PNG}
-    \caption{TODO caption, \citep{Castagna2005}}
+    \caption{Circularity in the semantic subtyping definition \citep{Castagna2005}}
     \label{fig:circ}
@@
 
-The syntax of the language is exactly that of Featherweight Java, or FJ \citep{Igarashi1999}, with 
-the associated type system only being different. This was chosen on purpose as the full features of 
-Java such as concurrency and reflection are orthogonal (\textit{TODO reword}) to the purpose of the 
-project while the type system of FJ still exactly corresponds to that of Java. The main key 
-simplification in FJ is the omission of assignment. All fields of an object are initialised in the
-constructor and not changed afterwards. While this restricts FJ to what is essentially a functional 
-fragment of Java, it is still fully computationally complete. The syntax of FJ can be seen in Figure
-\ref{fig:syntax} and a sample code fragment can be seen in Figure \ref{fig:class}.
+The syntax of the JSF is exactly that of Featherweight Java (FJ) \citep{Igarashi1999}, with only the associated type system being different.
+This was chosen on purpose as the full features of Java such as concurrency and reflection are orthogonal to the purpose of demonstrating the use of the novel semantic type system.
+The main key simplification in FJ from Java is the omission of assignment.
+All fields of an object cannot be changed after initialisation and all methods are pure functions
+While this restricts FJ to what is essentially a functional fragment of Java, it is still fully computationally complete.
 
-\begin{figure}
-    \centering
-    \includegraphics[width=0.4\linewidth]{images/syntax.PNG}
-    \caption{TODO caption, \citep{Igarashi1999}}
-    \label{fig:syntax}
-\end{figure}
+\section{Tools}
 
-\begin{figure}
-    \centering
-    \includegraphics[width=0.4\linewidth]{images/class.PNG}
-    \caption{TODO caption, \citep{Igarashi1999}}
-    \label{fig:class}
-\end{figure}
+The type system of FSJ was built using Antlr to define the grammar and type checking of the language and the
 
@@
 How is this problem to be approached, without reference to specific implementation details?
+
+\label{sec:design}
+The syntax of types is given by the following grammar \citep{Dardha2013,Dardha2017}:
+$$
+    \begin{array}{llll}
+        \tau   & ::=                                       & \alpha \ |\ \mu
+               & \mbox{{\em Type term}}
+        \\
+        \alpha & ::=                                       & \zero \ |\ \btypes \ |\ [\wt{l:\tau}] \ |\ \alpha \ \bf{and}\ \alpha \ |\ \bf{not}\ \alpha
+        \qquad
+               & \mbox{{\em Object type} ($\alpha $-type)}
+        \\
+        \mu    & ::=                                       & \alpha \to \alpha \ |\ \mu \ \bf{and}\ \mu \ |\ \bf{not}\ \mu
+               & \mbox{{\em Method type} ($\mu $-type)}
+    \end{array}
+$$
+
+The syntax of terms is given by the following grammar and is based on the standard syntax of terms in FJ \cite{Igarashi1999,Dardha2013,Dardha2017}:
+\begin{align*}
+     & \mbox{\textit{Class declaration}}  & L \; ::= \; & \syndecl{C}{C}{\wt{\alpha \ a};\ K; \ \wt{M}\ }                         \\
+     & \mbox{\textit{Constructor}}        & K \;::=\;   & C\ (\wt{\alpha\ x})\ \{\ \synsuper(\wt{x});\ \wt{\synthis.a}=\wt{x}; \} \\
+     & \mbox{\textit{Method declaration}} & M \; ::= \; & \alpha \ m\ (\alpha \ x)\ \{\ \synret e; \}                             \\
+     & \mbox{\textit{Expressions}}        & e \; ::=\;  & x\ |\  c\ |\ e.a\ |\ e.m(e) \ | \ \synnew C(\wt{e})
+\end{align*}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.4\linewidth]{images/class.PNG}
+    \caption{TODO caption, \citep{Igarashi1999}}
+    \label{fig:class}
+\end{figure}
+
 \section{Guidance}

### 2020-03-22 conference/paper.tex

--- 2020-02-28 conference/paper.tex
@@
             %\qquad Diagonal ( )\   \{\ super();\ \}
-	\hspace{2.5cm} \ldots\\
+            \hspace{2.5cm} \ldots \\
             \qquad \mathbf{double}\ diagonal((\textit{Polygon} \textbf{ and not }  \textit{Triangle}) \ shape)\{\ldots\}
@@
 \begin{align*}
-& \mbox{\textit{Class declaration}} & L \; ::= \; & \syndecl{C}{C}{\wt{\alpha \ a};\ K; \ \wt{M}\ }\\
-& \mbox{\textit{Constructor}} & K \;::=\; & C\ (\wt{\alpha\ x})\ \{\ \synsuper(\wt{x});\ \wt{\synthis.a}=\wt{x}; \}\\
+     & \mbox{\textit{Class declaration}}  & L \; ::= \; & \syndecl{C}{C}{\wt{\alpha \ a};\ K; \ \wt{M}\ }                         \\
+     & \mbox{\textit{Constructor}}        & K \;::=\;   & C\ (\wt{\alpha\ x})\ \{\ \synsuper(\wt{x});\ \wt{\synthis.a}=\wt{x}; \} \\
      & \mbox{\textit{Method declaration}} & M \; ::= \; & \alpha \ m\ (\alpha \ x)\ \{\ \synret e; \}                             \\
@@
 \end{flalign*}
-%
+
 Note that \emph{Int} is not a subtype of \emph{Float} as a 32-bit \emph{float} cannot represent the whole set of 32-bit \emph{integer} values accurately and therefore \emph{Int} is not fully set-contained, however this is not the case for \emph{Int} and \emph{Double}.
@@
     Diagonal = \left\{Diagonal\right\}
+    >>>>>>> c2071d9f146f39da75304752181339c4518fa178
 \end{flalign*}

### 2020-03-24 dissertation/l4proj.tex
more work on background

@@
 \input{semmacros}
+\usepackage{amsmath, amssymb}
 
@@
 
-The end goal in all programming projects, is to create a solution to the task at hand that performs exactly as expected by the programmer, and exactly solves the problem that was initially defined.
+The end goal in all programming projects is to create a solution to the task at hand that exactly solves the initially defined problem and performs exactly as expected by the programmer.
 However, this very rarely actually happens.
@@
 On the one hand, this is often because the original problem is not statically defined.
-If it is a solution for an external client, their requirements and needs will change over time, or the needs and wants of the users using this project can also change.
+If it is a solution for an external client, their requirements and needs will change over time, or similarly the needs and wants of the users using this project can also change.
 This means the project needs to be adapted over time.
@@
 
-For example, one common area of error in C code that the Clang tools warn about is implicit conversions where the value of an expression is a different type from the one expected, such as the use of a floating point number where an integer is expected.
-The C language will implicitly convert the value to an integer, however this can cause unexpected behaviour for the programmer.
-It knows that this conversion needs to happen as the type system dictates that you cannot use any other type other than an integer where an integer is expected.
+For example, one common area of error in C code that the Clang tools warn about is implicit conversions, where the value of an expression is a different type from the one expected, such as the use of a floating point number where an integer is expected.
+The C language will implicitly convert the value to an integer, however this can cause unexpected behaviour, unbeknownst to the programmer.
+The language knows that this conversion needs to happen as the type system dictates that you cannot use any other type other than an integer where an integer is expected.
 
@@
 
-In this section, the issue and aims of this paper are presented.
+In this section, the issues and aims of this paper are presented.
 
@@
 The typical hierarchical subtyping definitions in Java quite often restrict and complicate the logic that the programmer wants to implement.
-These problems come up particularly often when working with legacy code that has defines incorrect abstractions in its classes.
-This therefore leads to less understandable and less maintainable code, that could cause problems in the future.
+These problems come up particularly often when working with legacy code that has incorrectly defined abstractions in its classes.
+This therefore leads to less understandable and less maintainable code that could cause problems in the future.
 
@@
 
-The aim of this project is to implement a new programming language that uses a proposed type system consisting of boolean types and a semantic subtyping algorithm.
+The aim of this project is to implement a new programming language, which we further call Semantic Featherweight Java (SFJ), that uses a proposed type system consisting of boolean types and a semantic subtyping algorithm.
 This project aims to transform the mathematical definition of such a type system into an actual solution in software, solving the issues encountered doing so.
-The language should be able to illustrate the improved solutions to the
-common example problems encounter in a language without these features such as Java.
-Furthermore, the language will be evaluated so see whether these additional language features are intuitive to use by programmers.
+The language should be able to illustrate the improved solutions to the common example problems encountered in a language without these features such as Java.
+Furthermore, the language will be evaluated so see whether these additional language features are intuitive to use.
 
@@
 
+In this chapter, we define and explore type systems and the different ways they can be defined.
+We especially explore the semantic method of defining a type system and explore the benefits and challenges that it brings.
+
 \section{Type Systems}
@@
 
-In particular, the most important detail in the above definition is the emphasis on classifying phrases in our language into a particular type based on the kinds of values they hold or compute.
-For example by using a type system, we can check that the arguments given to an arithmetic operation are always numbers.
-Therefore we can validate that a specific program is absent of certain bad run-time behaviour such as trying to add a boolean value \emph{true} to a number \emph{2}.
+In particular, the detail we want to focus on in the above definition is the emphasis on classifying phrases in our language into a particular type based on the kinds of values they hold or compute.
+For example by using a type system, we can check that the arguments given to an arithmetic operation are always correct.
+Given a number and another arithmetic expression, we know they both hold or compute a number type, so the operation is correct.
+Therefore we can validate that a specific program is absent of certain bad run-time behaviour such as possibly trying to add a boolean value \emph{true} to a number \emph{2}.
 
 However what the above definition is inaccurate about is that is it defines a type system to be solely a syntactic method, in which the system is defined as a list of formal deduction rules.
-The syntactic approach is certainly by far the most common approach for a type system.
-However, there does exist an alternative method in the semantic approach.
-\citet{Frisch2002} describe the semantic approach as instead starting with a model of the language and all the possible values in the language and defines an interpretation of types as subsets of values the model.
+The syntactic approach is certainly by far the most common approach for a type system, but there does exist an alternative method in the semantic approach.
+\citet{Frisch2002} describe the semantic approach as instead starting with a model of the language and all the possible values in the language and defining an interpretation of types as subsets of values the model.
 
@@
 the equality $t_{1} \subseteq t_{2}$ holds.
-However, in this way, subtyping is defined by relying on the notion of well-typed values; hence, we need the typing relation to determine typing judgements for values; but the typing rules use the subtyping relation defined in \citep{Dardha2017}.
+However, in this way, subtyping is defined by relying on the notion of well-typed values; hence, we need the typing relation to determine typing judgements for values; but the typing rules require a subtyping relation to be fully defined.
 So, there is a circularity in our definition, as demonstrated in Figure \ref{fig:circ}.
@@
 
-The syntax of the JSF is exactly that of Featherweight Java (FJ) \citep{Igarashi1999}, with only the associated type system being different.
-This was chosen on purpose as the full features of Java such as concurrency and reflection are orthogonal to the purpose of demonstrating the use of the novel semantic type system.
-The main key simplification in FJ from Java is the omission of assignment.
-All fields of an object cannot be changed after initialisation and all methods are pure functions
+The syntax of SFJ is exactly that of Featherweight Java (FJ) \citep{Igarashi1999}, with only the associated type system being different.
+FJ is closely related to Java, with a key simplification in the removal of assignment operators.
+All fields of an object cannot be changed after initialisation and all methods are pure functions.
 While this restricts FJ to what is essentially a functional fragment of Java, it is still fully computationally complete.
+SFJ was intentionally based on FJ instead of the full Java language because features such as concurrency and reflection are orthogonal to the purpose of demonstrating the implementation of our novel semantic type system.
 
@@
 
-The type system of FSJ was built using Antlr to define the grammar and type checking of the language and the
+The type system of FSJ was built using ANTLR \citep{parr2013} to define the grammar of the language and create a parser for this grammar rather than having to create these by hand, making the initial development process much quicker.
+It accepts a grammar using Extended Backus-Naur Form (EBNF) notation to create ANTLR rules which are a list of productions or alternatives. A general form of a rule is therefore as follow:
+
+$$
+    \begin{array}{llll}
+        rule & : & alternative_{1}
+        \\
+             & | & alternative_{2}
+        \\
+             &   & \vdots
+        \\
+             & | & alternative_{n}
+        \\
+             & ; & 
+    \end{array}
+$$
+
+Each alternative production in a rule can itself be a list of elements, where an element can be another rule or just a terminating token.
+Since we are using EBNF notation, we can also use \emph{*} and \emph{?} to respectively signify repeated and conditional elements in a production.
+We show an excerpt taken from the SFJ grammar as an example:
+
+$$
+    \begin{array}{llll}
+        expression     & : & \mbox{primExpression ( (PLUS | MINUS | DIV | MULT) primExpression)?}
+        \\
+                       & ;
+        \\
+        \\
+        primExpression & : & \mbox{NUMBER}
+        \\
+                       & | & \mbox{TRUE}
+        \\
+                       & ; & 
+        \\
+        \\
+        TRUE           & : & \mbox{'true'}
+        \\
+                       & ; & 
+        \\
+        \\
+        NUMBER         & : & \mbox{DIGIT (DIGIT)*}
+        \\
+                       & ; & 
+        \\
+        \\
+        DIGIT          & : & '0'..'9'
+        \\
+                       & ; & 
+    \end{array}
+$$
 
@@
 
-%==================================================================================================================================
+%===================================================================================================
 \chapter{Analysis/Requirements}

### 2020-03-24 conference/paper.tex

@@
     Diagonal = \left\{Diagonal\right\}
-    >>>>>>> c2071d9f146f39da75304752181339c4518fa178
 \end{flalign*}

### 2020-03-31 dissertation/l4proj.tex
add more to background

@@
 On the one hand, this is often because the original problem is not statically defined.
-If it is a solution for an external client, their requirements and needs will change over time, or similarly the needs and wants of the users using this project can also change.
+If it is a solution for an external client, then their requirements and needs will change over time, or similarly the needs and wants of the users using this project can also change.
 This means the project needs to be adapted over time.
-On the other hand, and almost always much less evident, is that programs do not perform exactly as the programmer imagines them to work.
+On the other hand, and almost always much less evident, is that programs do not work exactly as the programmer imagines them to work.
 
-These are in general called \emph{bugs} in a program, and have become more and more common as the projects we create have become more complex compared to the early days of computing and the underlying hardware we work on has also become more complex.
-Therefore in both academia and industry, there have been great efforts over the years to create tools that allow us to decrease the amount of bugs in our programs and increase our productivity.
+These are in general called \emph{bugs} in a program, and have become more and more common as projects have become more complex compared to the early days of computing and the underlying hardware we work on has also become more complex.
+Therefore in both academia and industry there have been great efforts over the years to create tools that allow us to decrease the amount of bugs in our programs and increase our productivity.
 
@@
 Spark is a formally defined language based on the Ada \citep{Ada1979} language intended for the development of high integrity software such as flight control systems.
-Spark uses contracts and flow analysis to logically prove that programs are correct.
 Others exist as separate, well-known tools that exist alongside the language such as the Clang analysis tools for C and C++ \citep{kremenek2008}.
 
-While they are not necessarily needed to write correct programs, these tools have quickly become industry standards for maintaining a level of correctness within codebases and avoiding typical programming errors.
+Whilst they are not necessarily needed to write correct programs, these tools have quickly become industry standards for maintaining a level of correctness within codebases and avoiding typical programming errors.
 
@@
 
-Underlying all of the previously mentioned static code analysis tools, is their use of the \textbf{type system} of the language.
+Underlying all of the previously mentioned static code analysis tools, is their use of the \emph{type system} of the language.
 Type systems at their most basic work by assigning types to various constructs of a program so that incompatible types cannot be used.
@@
 The C language will implicitly convert the value to an integer, however this can cause unexpected behaviour, unbeknownst to the programmer.
-The language knows that this conversion needs to happen as the type system dictates that you cannot use any other type other than an integer where an integer is expected.
+Whilst being inconvenient, this conversion needs to happen as the type system of the language dictates that you cannot use any other type other than an integer where an integer is expected.
 
@@
 More modern advances in type systems include the use of linear types in the Cyclone \citep{grossman2002} language that allow us to more naturally and logically define finite state machines.
-There exist other new developments in type systems such as Boolean types which are explored in this paper.
+There also exist other new developments in type systems such as Boolean types which are explored in this paper.
 
@@
 
-The aim of this project is to implement a new programming language, which we further call Semantic Featherweight Java (SFJ), that uses a proposed type system consisting of boolean types and a semantic subtyping algorithm.
-This project aims to transform the mathematical definition of such a type system into an actual solution in software, solving the issues encountered doing so.
+The aim of this project is to implement a new programming language, which we further call Semantic Featherweight Java (SFJ), which uses a new proposed type system consisting of boolean types and a semantic subtyping algorithm.
+This project aims to transform the mathematical definition of such a type system into an actual software implementation, solving the issues encountered doing so.
 The language should be able to illustrate the improved solutions to the common example problems encountered in a language without these features such as Java.
@@
     \item
-          Finally \textbf{chapter 8} concludes the dissertation with a summary of the important results.
+          Finally \textbf{chapter 7} concludes the dissertation with a summary of the important results.
 \end{itemize}
@@
 We especially explore the semantic method of defining a type system and explore the benefits and challenges that it brings.
+We also discuss the tools that were used in this project and the related work that has already been done with boolean types or semantic type systems.
 
@@
 
-In all modern programming languages, a type systems is arguably the main way that incorrect behaviour in a program is reduced.
+In all modern programming languages, a \emph{type system} is arguably the main way that incorrect behaviour in a program is reduced by preventing the occurrence of \emph{type errors}.
 A possible definition of a type system is given by \citet{Pierce2002} as:
@@
 
-In particular, the detail we want to focus on in the above definition is the emphasis on classifying phrases in our language into a particular type based on the kinds of values they hold or compute.
-For example by using a type system, we can check that the arguments given to an arithmetic operation are always correct.
-Given a number and another arithmetic expression, we know they both hold or compute a number type, so the operation is correct.
-Therefore we can validate that a specific program is absent of certain bad run-time behaviour such as possibly trying to add a boolean value \emph{true} to a number \emph{2}.
+A type error is defined by \citet{wright1994} as:
 
-However what the above definition is inaccurate about is that is it defines a type system to be solely a syntactic method, in which the system is defined as a list of formal deduction rules.
+\emph{The use of a function on arguments for which it is not defined, and the attempted application of a non-function.}
+
+Most importantly, we want to emphasize from the first definition that in a type system, we aim to classify phrases in our language into a specific \emph{type}.
+Then, using the type system, we can check that the types of the arguments given to any operation are always correct, thus avoiding a type error.
+
+For example, taking the addition operation which takes two arguments of the number type, then given the arguments $42$ and $2 + 2$, we know they both hold or compute a number type and so there is guaranteed to be no type error.
+Conversely, given the arguments $42$ and the boolean value $true$, they are not both a number type and so cause a type error.
+We therefore avoid unexpected bad run-time behaviour and instead find these problems before even running the program.
+
+However what the first definition is missing is that it defines a type system to be solely a syntactic method, in which the system is defined as a list of formal deduction rules.
 The syntactic approach is certainly by far the most common approach for a type system, but there does exist an alternative method in the semantic approach.
@@
 
+For example, given a language with a model $D$:
+
+\begin{equation*}
+    D = \left[ a + bi \mid a, b \in \mathbb{R}\right]
+\end{equation*}
+
+we can define the types \emph{int}, \emph{real} and \emph{complex} as subsets of the values in $D$ corresponding to their usual mathematical definitions.
+
 Such as model comes with several advantages.
-For example, given two types $s$ and $t$, which represent a certain subset of values in the model, when $s \subseteq t$ does not hold, then it is possible to exhibit an element of the model which is in the interpretation of type $s$ and not $t$.
-This can then be used to show more informative error messages to the programmer \citep{Castagna2005}, such as showing which value in $s$ is causing the equality not to hold.
-However due to being a more technical approach, such as the non-triviality of defining the interpretation of types as subsets of a model, the semantic method has received less attention than syntactic methods.
+For example, given two types $s$ and $t$, which represent subsets of values in the model, when $s \subseteq t$ does not hold, it is possible to exhibit an element of the model which is in the interpretation of type $s$ and not $t$.
+This can then be used to show more informative error messages to the programmer as shown by \citet{Castagna2005}, such as showing which value in $s$ is causing the equality not to hold.
+Secondly the set based approach to the type system leads itself to able to incorporate boolean operators into its types easily.
+However due to being a more technical approach, such as the non-triviality of defining the interpretation of types as subsets of a model as discussed below, the semantic method has received less attention than syntactic methods.
 
@@
 the equality $t_{1} \subseteq t_{2}$ holds.
-However, in this way, subtyping is defined by relying on the notion of well-typed values; hence, we need the typing relation to determine typing judgements for values; but the typing rules require a subtyping relation to be fully defined.
+However, in this way, subtyping is defined by relying on the notion of well-typed values; hence, we need the typing relation to able to type values; but the typing rules require a subtyping relation to be fully defined.
 So, there is a circularity in our definition, as demonstrated in Figure \ref{fig:circ}.
@@
 To solve this problem, we follow the framework defined by \citet{Frisch2008}.
-The general idea of the framework is that we first extend the types in the language with \emph{Boolean Combinators}: union $\lor$, intersection $\land$ and negation $\neg$.
-We can then define an abstract model $\beta$ where our types are interpreted according to the set-theoretic meaning of the combinators and can define subtyping as set-inclusion and therefore define the typing rules.
-Separately, if we have a notion of well-typed values, we can define another interpretation of types as sets of values.
-We can again define a subtyping relation $\nu$.
-While these may be different relation, if the models are chosen carefully such that:
+The general idea of the framework is that we first extend the types in the language with \emph{Boolean Combinators}: union $\lor$, intersection $\land$ and negation $\neg$ to give a type algebra $\tau$.
+We can then define an abstract model $\beta$ with a interpretation function $[\![ ]\!]_{\beta} : \tau \to \rho(\beta)$ (where $\rho(\beta)$ is the power-set of $\beta$).
+This model must capture the meaning of each type but also interprets the combinators in a set-theoretic way, such as $[\![s \land t]\!]_{\beta} = [\![s]\!]_{\beta} \cap [\![t]\!]_{\beta}$ etc.
+This model then induces the following subtyping relation:
+
+\begin{equation*}
+    s \leq_{\beta} t \iff [\![s]\!]_{\beta} \subseteq [\![t]\!]_{\beta}
+\end{equation*}
+
+and we can therefore define the typing rules.
+
+Separately, now that we have typing rules, we can define a new interpretation of types based on values $[\![t]\!]_{\mathrm{v}} = \left\{\mathrm{v} \in V \mid \: \vdash_{\beta} \mathrm{v} : t \right\}$ and then define a ``new''subtyping relation as above:
+
+\begin{equation*}
+    s \leq_{\mathrm{v}} t \iff [\![s]\!]_{\mathrm{v}} \subseteq [\![t]\!]_{\mathrm{v}}
+\end{equation*}
+
+While these may be different relations, if the models are chosen carefully such that:
 \begin{equation*}
@@
 
-then these models coincide and this closes the circularity.
+then these subtyping relations coincide and this closes the circularity.
 
@@
 
+\paragraph{On multimethods}
+Since FJ is a core language, some features of the full Java are removed: overloading methods is one of them.
+In our framework, leveraging the expressivity of boolean connectives and semantic subtyping, many of these features, and language constructs, e.g., lists, are restored: overloading methods is one of them.
+We can thus model \emph{multimethods} \citep{BC97}, which according to the authors is ``\emph{very clean and easy to understand [...] it would be the best solution for a brand new language}".
+As an example \citet{Dardha2013,Dardha2017}, consider the following class declarations:
+$$
+    \begin{array}{ll}
+        \syndecl{A}{\mathit{Object}} {
+         &
+            \qquad
+            \syndecl{B}{A} {
+                \\
+                \qquad\quad \mathbf{int}\ \mathit{length}\ (\mathbf{string}\ s)\{\ \ldots\ \}\qquad
+         &
+                \qquad\qquad\quad \mathbf{int}\ \mathit{length}\ (\mathbf{int}\ n)\{\ \ldots\ \}
+                \\
+            }
+         &
+            \qquad }
+    \end{array}
+$$
+Method {\it length} has type $\mathbf{string} \rightarrow \mathbf{int}$ in $A$.
+However, in $B$ it has type $(\mathbf{string} \rightarrow \mathbf{int}) \synwedge (\mathbf{int} \rightarrow \mathbf{int})$,
+which can be simplified as $(\mathbf{string} \synvee \mathbf{int}) \rightarrow \mathbf{int}$.
+
 \section{Tools}
 
-The type system of FSJ was built using ANTLR \citep{parr2013} to define the grammar of the language and create a parser for this grammar rather than having to create these by hand, making the initial development process much quicker.
-It accepts a grammar using Extended Backus-Naur Form (EBNF) notation to create ANTLR rules which are a list of productions or alternatives. A general form of a rule is therefore as follow:
+The type system of FSJ was built using ANTLR \citep{parr2013} to define the grammar of the language and automatically create a parser for this grammar rather than having to create one by hand, which made the initial development process much quicker.
+It accepts a grammar using Extended Backus-Naur Form (EBNF) \footnote{See ISO/IEC 14977 for reference, although it is not the standard used in the examples} notation to create ANTLR rules which are a list of productions or alternatives.
+A general form of a rule is as follow:
 
@@
 Since we are using EBNF notation, we can also use \emph{*} and \emph{?} to respectively signify repeated and conditional elements in a production.
+We can also use $|$ to give several alternatives for a single element.
 We show an excerpt taken from the SFJ grammar as an example:
@@
 The closest similar area of research would likely be the work on $\mathbb{C}$Duce
-\citep{Benzaken2003}, which is also a functional language with a semantic type system designed for
+\citep{Benzaken2003}, which is xalso a functional language with a semantic type system designed for
 working with XML documents and a continuation of the work on XDuce \citep{Hosoya2003}. The language
@@
 (\textit{TODO reword}) Similar work to ours creating an implementation for this would result in a
-Golang-like language \citep{Meyerson2014}, creating a concurrency-focused language with more
+Golang-like \footnote{https://golang.org/} like language, creating a concurrency-focused language with more
 intuitive types. The paper found that it was required to be able to decide and resolve the atomicy,
@@
 %===================================================================================================
-\chapter{Analysis/Requirements}
-What is the problem that you want to solve, and how did you arrive at it?
-\section{Guidance}
-Make it clear how you derived the constrained form of your problem via a clear and logical process.
 
-%==================================================================================================================================
-\chapter{Design}
-How is this problem to be approached, without reference to specific implementation details?
+% \chapter{Analysis/Requirements}
+% What is the problem that you want to solve, and how did you arrive at it?
+% \section{Guidance}
+% Make it clear how you derived the constrained form of your problem via a clear and logical process.
+
+%===================================================================================================
+\chapter{SFJ: Design and Implementation}
+
+\section{Syntax}
+
+\subsection{Syntax of Types}
 
@@
 
-The syntax of terms is given by the following grammar and is based on the standard syntax of terms in FJ \cite{Igarashi1999,Dardha2013,Dardha2017}:
+$\alpha$-types are used to type fields and $\mu$-types are used to type methods.
+Type $\zero$ is the empty type.
+Type $\btypes$ denotes the \emph{basic} types, such as integers, booleans, etc.
+    {Record} types $[\wt{l:\tau}]$, where $\wt l$ is a sequence of disjoint labels are used to type objects.
+Arrow types $\alpha \to \alpha$ are used to type methods.
+The boolean types using ${and}$ and ${not}$ have their expected set-theoretic meanings, and ${or}$ is obtained by their combination.\\
+
+\subsection{Syntax of Terms}
+
+The syntax of terms is given by the following grammar and is based on the standard syntax of terms in FJ \cite{Igarashi1999,Dardha2013,Dardha2017}.
+We assume an infinite countable set of names, with some special names: $\mathit{Object}$ indicates the root class, $\synthis$  indicates the current and $\synsuper$ indicates the parent object.
+We let  $A, B, C, \ldots$ range over classes; $a, b, \ldots$ over fields; $m, n, \ldots$ over methods and $x, y, z, \ldots$ range over variables.
+Constants $c$ range over an infinite countable set $\mathcal{K}$.
+
 \begin{align*}
@@
 
-\begin{figure}
-    \centering
-    \includegraphics[width=0.4\linewidth]{images/class.PNG}
-    \caption{TODO caption, \citep{Igarashi1999}}
-    \label{fig:class}
-\end{figure}
+A {program} $(\wt{L}, e)$ is a pair of a sequence of class declarations $\wt L$, giving rise to a class hierarchy as specified by the inheritance relation, and an expression $e$ to be evaluated.
+%
+A class declaration $L$ specifies the name of the class, the name of the parent class it extends, its typed fields , the constructor $K$ and its method declarations $M$.
+The constructor $K$ initializes the fields of the object by assigning values to the fields inherited by the super-class and to the fields declared in the current \textbf{this} class.
+%
+A method declaration $M$ specifies the signature of the method, namely the return type, the method name and the formal parameter as well as the body of the method.
+Notice that in our theoretical development we use unary methods, without loss of generality: tuples of arguments can be modelled by an object that instantiates a ``special''class containing as fields all the needed arguments.
+Expressions $e$ include variables, constants, field accesses, method invocations and object creations.
 
-\section{Guidance}
-Design should cover the abstract design in such a way that someone else might be able to do what you did, but with a different language or library or tool.
+%===================================================================================================
 
-%==================================================================================================================================
 \chapter{Implementation}
+
 What did you do to implement this idea, and what technical achievements did you make?
+
 \section{Guidance}
@@
 \subsection{Figures}
+
 \emph{Always} refer to figures included, like Figure \ref{fig:relu}, in the body of the text. Include full, explanatory captions and make sure the figures look good on the page.
-You may include multiple figures in one float, as in Figure \ref{fig:synthetic}, using \texttt{subcaption}, which is enabled in the template.
-
-% Figures are important. Use them well.
-\begin{figure}
-    \centering
-    \includegraphics[width=0.5\linewidth]{images/relu.pdf}
-
-    \caption{In figure captions, explain what the reader is looking at: ``A schematic of the rectifying linear unit, where $a$ is the output amplitude,
-        $d$ is a configurable dead-zone, and $Z_j$ is the input signal'', as well as why the reader is looking at this:
-        ``It is notable that there is no activation \emph{at all} below 0, which explains our initial results.''
-        \textbf{Use vector image formats (.pdf) where possible}. Size figures appropriately, and do not make them over-large or too small to read.
-    }
-
-    % use the notation fig:name to cross reference a figure
-    \label{fig:relu}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \begin{subfigure}[b]{0.45\textwidth}
-        \includegraphics[width=\textwidth]{images/synthetic.png}
-        \caption{Synthetic image, black on white.}
-        \label{fig:syn1}
-    \end{subfigure}
-    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc.
-    %(or a blank line to force the subfigure onto a new line)
-    \begin{subfigure}[b]{0.45\textwidth}
-        \includegraphics[width=\textwidth]{images/synthetic_2.png}
-        \caption{Synthetic image, white on black.}
-        \label{fig:syn2}
-    \end{subfigure}
-    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc.
-    %(or a blank line to force the subfigure onto a new line)
-    \caption{Synthetic test images for edge detection algorithms. \subref{fig:syn1} shows various gray levels that require an adaptive algorithm. \subref{fig:syn2}
-        shows more challenging edge detection tests that have crossing lines. Fusing these into full segments typically requires algorithms like the Hough transform.
-        This is an example of using subfigures, with \texttt{subref}s in the caption.
-    }\label{fig:synthetic}
-\end{figure}
-
-\clearpage
+You may include multiple figures in one float, as in Figure, using \texttt{subcaption}, which is enabled in the template.
 
@@
 \subsection{Algorithms}
-Algorithms can be set using \texttt{algorithm2e}, as in Algorithm \ref{alg:metropolis}.
 
-% NOTE: line ends are denoted by \; in algorithm2e
-\begin{algorithm}
-    \DontPrintSemicolon
-    \KwData{$f_X(x)$, a probability density function returing the density at $x$.\; $\sigma$ a standard deviation specifying the spread of the proposal distribution.\;
-        $x_0$, an initial starting condition.}
-    \KwResult{$s=[x_1, x_2, \dots, x_n]$, $n$ samples approximately drawn from a distribution with PDF $f_X(x)$.}
-    \Begin{
-        $s \longleftarrow []$\;
-        $p \longleftarrow f_X(x)$\;
-        $i \longleftarrow 0$\;
-        \While{$i < n$}
-        {
-            $x^\prime \longleftarrow \mathcal{N}(x, \sigma^2)$\;
-            $p^\prime \longleftarrow f_X(x^\prime)$\;
-            $a \longleftarrow \frac{p^\prime}{p}$\;
-            $r \longleftarrow U(0,1)$\;
-            \If{$r<a$}
-            {
-                $x \longleftarrow x^\prime$\;
-                $p \longleftarrow f_X(x)$\;
-                $i \longleftarrow i+1$\;
-                append $x$ to $s$\;
-            }
-        }
-    }
-
-    \caption{The Metropolis-Hastings MCMC algorithm for drawing samples from arbitrary probability distributions,
-        specialised for normal proposal distributions $q(x^\prime|x) = \mathcal{N}(x, \sigma^2)$.
-The symmetry of the normal distribution means the acceptance rule takes the simplified form.}\label{alg:metropolis}
-\end{algorithm}
+Algorithms can be set using \texttt{algorithm2e}, as in Algorithm.
 
@@
 
-If you need to include tables, like Table \ref{tab:operators}, use a tool like https://www.tablesgenerator.com/ to generate the table as it is
-extremely tedious otherwise.
+If you need to include tables, like Table , use a tool like https://www.tablesgenerator.com/ to generate the table as it is extremely tedious otherwise.
 
-\begin{table}[]
-    \caption{The standard table of operators in Python, along with their functional equivalents from the \texttt{operator} package. Note that table
-        captions go above the table, not below. Do not add additional rules/lines to tables. }\label{tab:operators}
-    %\tt
-    \rowcolors{2}{}{gray!3}
-    \begin{tabular}{@{}lll@{}}
-        %\toprule
-        \textbf{Operation}    & \textbf{Syntax}                         & \textbf{Function}                          \\ %\midrule % optional rule for header
-        Addition              & \texttt{a + b}                          & \texttt{add(a, b)}                         \\
-        Concatenation         & \texttt{seq1 + seq2}                    & \texttt{concat(seq1, seq2)}                \\
-        Containment Test      & \texttt{obj in seq}                     & \texttt{contains(seq, obj)}                \\
-        Division              & \texttt{a / b}                          & \texttt{div(a, b) }                        \\
-        Division              & \texttt{a / b}                          & \texttt{truediv(a, b) }                    \\
-        Division              & \texttt{a // b}                         & \texttt{floordiv(a, b)}                    \\
-        Bitwise And           & \texttt{a \& b}                         & \texttt{and\_(a, b)}                       \\
-        Bitwise Exclusive Or  & \texttt{a \textasciicircum b}           & \texttt{xor(a, b)}                         \\
-        Bitwise Inversion     & \texttt{$\sim$a}                        & \texttt{invert(a)}                         \\
-        Bitwise Or            & \texttt{a | b}                          & \texttt{or\_(a, b)}                        \\
-        Exponentiation        & \texttt{a ** b}                         & \texttt{pow(a, b)}                         \\
-        Identity              & \texttt{a is b}                         & \texttt{is\_(a, b)}                        \\
-        Identity              & \texttt{a is not b}                     & \texttt{is\_not(a, b)}                     \\
-        Indexed Assignment    & \texttt{obj{[}k{]} = v}                 & \texttt{setitem(obj, k, v)}                \\
-        Indexed Deletion      & \texttt{del obj{[}k{]}}                 & \texttt{delitem(obj, k)}                   \\
-        Indexing              & \texttt{obj{[}k{]}}                     & \texttt{getitem(obj, k)}                   \\
-        Left Shift            & \texttt{a \textless{}\textless b}       & \texttt{lshift(a, b)}                      \\
-        Modulo                & \texttt{a \% b}                         & \texttt{mod(a, b)}                         \\
-        Multiplication        & \texttt{a * b}                          & \texttt{mul(a, b)}                         \\
-        Negation (Arithmetic) & \texttt{- a}                            & \texttt{neg(a)}                            \\
-        Negation (Logical)    & \texttt{not a}                          & \texttt{not\_(a)}                          \\
-        Positive              & \texttt{+ a}                            & \texttt{pos(a)}                            \\
-        Right Shift           & \texttt{a \textgreater{}\textgreater b} & \texttt{rshift(a, b)}                      \\
-        Sequence Repetition   & \texttt{seq * i}                        & \texttt{repeat(seq, i)}                    \\
-        Slice Assignment      & \texttt{seq{[}i:j{]} = values}          & \texttt{setitem(seq, slice(i, j), values)} \\
-        Slice Deletion        & \texttt{del seq{[}i:j{]}}               & \texttt{delitem(seq, slice(i, j))}         \\
-        Slicing               & \texttt{seq{[}i:j{]}}                   & \texttt{getitem(seq, slice(i, j))}         \\
-        String Formatting     & \texttt{s \% obj}                       & \texttt{mod(s, obj)}                       \\
-        Subtraction           & \texttt{a - b}                          & \texttt{sub(a, b)}                         \\
-        Truth Test            & \texttt{obj}                            & \texttt{truth(obj)}                        \\
-        Ordering              & \texttt{a \textless b}                  & \texttt{lt(a, b)}                          \\
-        Ordering              & \texttt{a \textless{}= b}               & \texttt{le(a, b)}                          \\
-        % \bottomrule
-    \end{tabular}
-\end{table}
 \subsection{Code}
 
-Avoid putting large blocks of code in the report (more than a page in one block, for example). Use syntax highlighting if possible, as in Listing \ref{lst:callahan}.
+Avoid putting large blocks of code in the report (more than a page in one block, for example). Use syntax highlighting if possible, as in Listing.
 
-\begin{lstlisting}[language=python, float, caption={The algorithm for packing the $3\times 3$ outer-totalistic binary CA successor rule into a
-    $16\times 16\times 16\times 16$ 4 bit lookup table, running an equivalent, notionally 16-state $2\times 2$ CA.}, label=lst:callahan]
-    def create_callahan_table(rule="b3s23"):
-        """Generate the lookup table for the cells."""
-        s_table = np.zeros((16, 16, 16, 16), dtype=np.uint8)
-        birth, survive = parse_rule(rule)
+%===================================================================================================
 
-        # generate all 16 bit strings
-        for iv in range(65536):
-            bv = [(iv >> z) & 1 for z in range(16)]
-            a, b, c, d, e, f, g, h, i, j, k, l, m, n, o, p = bv
-
-            # compute next state of the inner 2x2
-            nw = apply_rule(f, a, b, c, e, g, i, j, k)
-            ne = apply_rule(g, b, c, d, f, h, j, k, l)
-            sw = apply_rule(j, e, f, g, i, k, m, n, o)
-            se = apply_rule(k, f, g, h, j, l, n, o, p)
-
-            # compute the index of this 4x4
-            nw_code = a | (b << 1) | (e << 2) | (f << 3)
-            ne_code = c | (d << 1) | (g << 2) | (h << 3)
-            sw_code = i | (j << 1) | (m << 2) | (n << 3)
-            se_code = k | (l << 1) | (o << 2) | (p << 3)
-
-            # compute the state for the 2x2
-            next_code = nw | (ne << 1) | (sw << 2) | (se << 3)
-
-            # get the 4x4 index, and write into the table
-            s_table[nw_code, ne_code, sw_code, se_code] = next_code
-
-        return s_table
-
-\end{lstlisting}
-
-%==================================================================================================================================
 \chapter{Evaluation}
@@
 \section{Guidance}
+
 \begin{itemize}
@@
 
-If you visualise, follow the basic rules, as illustrated in Figure \ref{fig:boxplot}:
+If you visualise, follow the basic rules, as illustrated in Figure:
 \begin{itemize}
@@
 
-\begin{figure}
-    \centering
-    \includegraphics[width=1.0\linewidth]{images/boxplot_finger_distance.pdf}
+%===================================================================================================
 
-    \caption{Average number of fingers detected by the touch sensor at different heights above the surface, averaged over all gestures. Dashed lines indicate
-        the true number of fingers present. The Box plots include bootstrapped uncertainty notches for the median. It is clear that the device is biased toward
-        undercounting fingers, particularly at higher $z$ distances.
-    }
+\chapter{Conclusion}
 
-    % use the notation fig:name to cross reference a figure
-    \label{fig:boxplot}
-\end{figure}
+Summarise the whole project for a lazy reader who didn't read the rest (e.g. a prize-awarding committee).
 
-%==================================================================================================================================
-\chapter{Conclusion}
-Summarise the whole project for a lazy reader who didn't read the rest (e.g. a prize-awarding committee).
 \section{Guidance}
@@
 
-%==================================================================================================================================
-%
-%
-%==================================================================================================================================
+%===================================================================================================
+
 %  APPENDICES
@@
 
-    \chapter{Appendices}
+    \chapter{SFJ ANTLR Grammar}
 
-    Typical inclusions in the appendices are:
+    \begin{lstlisting}
+        grammar jsf;
 
-    \begin{itemize}
-        \item
-              Copies of ethics approvals (required if obtained)
-        \item
-              Copies of questionnaires etc. used to gather data from subjects.
-        \item
-              Extensive tables or figures that are too bulky to fit in the main body of
-              the report, particularly ones that are repetitive and summarised in the body.
+        @header{
+        package jsf;
+        }
 
-        \item Outline of the source code (e.g. directory structure), or other architecture documentation like class diagrams.
+        program
+                :       classDecl* expression EOF
+                ;
 
-        \item User manuals, and any guides to starting/running the software.
+        classDecl
+                :   CLASS classlbl=ID (EXTEND extendlbl=ID)? LBRAC
+                        fieldDecl* constructorDecl methodDecl*
+                    RBRAC
+                ;
 
-    \end{itemize}
+        fieldDecl
+                :       type ID SEMI
+                ;
 
-    \textbf{Don't include your source code in the appendices}. It will be
-    submitted separately.
+        constructorDecl
+                :       constructorname=ID LPAR (type ID (COMMA type ID)*)? RPAR LBRAC
+                            superDecl
+                            fieldAssignment*
+                        RBRAC
+                ;
+
+        superDecl
+                :       SUPER LPAR (ID (COMMA ID)*)? RPAR SEMI
+                ;
+
+        fieldAssignment
+                :       THIS DOT field=ID EQ parameter=ID SEMI
+                ;
+
+        methodDecl
+                :       returntype=type name=ID LPAR (paramtype=methodType paramname=ID)? RPAR LBRAC
+                            RETURN expression SEMI
+                        RBRAC
+                ;
+
+        expression
+                :       e1=primExpression
+                            (op=(PLUS | MINUS | DIV | MULT) e2=primExpression)?
+                ;
+
+        primExpression
+                :       NUMBER
+                |       DECIMAL
+                |       TRUE
+                |       FALSE
+                |       ID
+                |       THIS
+                |       primExpression DOT ID
+                |       primExpression DOT ID LPAR (expression)? RPAR
+                |       NEW ID LPAR (expression (COMMA expression)*)? RPAR
+                ;
+
+        type
+                :       basicType
+                |       ID
+                |       NOT classlbl=type
+                |       LPAR type1=type bool=(AND | OR) type2=type RPAR
+                ;
+
+        basicType
+                :       BYTE | INT | LONG | FLOAT | DOUBLE | CHAR | BOOL
+                ;
+
+        methodType
+                : type
+                | LPAR (NOT)? ID COLON param=type ARROW returnType=type RPAR
+                ;
+
+        BYTE    :       'byte'                      ;
+        INT     :       'int'                       ;
+        LONG    :       'long'                      ;
+        FLOAT   :       'float'                     ;
+        DOUBLE  :       'double'                    ;
+        CHAR    :       'char'                      ;
+        BOOL    :       'bool'                      ;
+
+        TRUE    :       'true'                      ;
+        FALSE   :       'false'                     ;
+
+        AND     :       'and'                       ;
+        OR      :       'or'                        ;
+        NOT     :       'not'                       ;
+
+        CLASS   :       'class'                     ;
+        SUPER   :       'super'                     ;
+        EXTEND  :       'extends'                   ;
+        THIS    :       'this'                      ;
+        RETURN  :       'return'                    ;
+        NEW     :       'new'                       ;
+
+        LPAR    :       '('                         ;
+        RPAR    :       ')'                         ;
+        LBRAC   :       '{'                         ;
+        RBRAC   :       '}'                         ;
+        EQ      :       '='                         ;
+        PLUS    :       '+'                         ;
+        MINUS   :       '-'                         ;
+        DIV     :       '/'                         ;
+        MULT    :       '*'                         ;
+        COMMA   :       ','                         ;
+        DOT     :       '.'                         ;
+        SEMI    :       ';'                         ;
+        COLON   :       ':'                         ;
+        ARROW   :       '->'                        ;
+
+        ID      :       LETTER (LETTER | DIGIT)*    ;
+        NUMBER  :       DIGIT (DIGIT)*              ;
+        DECIMAL :       DIGIT DOT (DIGIT)*          ;
+        SPACE   :       (' ' | '\t')+   -> skip     ;
+        EOL     :       '\r'? '\n'      -> skip     ;
+        EMPTY   :       'EMPTY'                     ;
+
+        fragment LETTER :   'a'..'z' | 'A'..'Z'     ;
+        fragment DIGIT  :   '0'..'9'                ;
+    \end{lstlisting}
+
+    \chapter{TODO}
 
@@
 
-%==================================================================================================================================
-%   BIBLIOGRAPHY
-
-% The bibliography style is abbrvnat
-% The bibliography always appears last, after the appendices.
+%===================================================================================================
 

### 2020-04-02 dissertation/l4proj.tex
implementation

@@
 \usepackage{amsmath, amssymb}
+
+\SetKwRepeat{Do}{do}{while}
+\SetKwProg{Fn}{Function}{:}{end}
 
@@
 On the one hand, this is often because the original problem is not statically defined.
-If it is a solution for an external client, then their requirements and needs will change over time, or similarly the needs and wants of the users using this project can also change.
+If, for example, it is a solution for an external client, then their requirements and needs will change over time.
 This means the project needs to be adapted over time.
@@
 
-These are in general called \emph{bugs} in a program, and have become more and more common as projects have become more complex compared to the early days of computing and the underlying hardware we work on has also become more complex.
+In general these are called \emph{bugs} in a program, and have become more and more common as projects and the underlying hardware they run on have become more complex since the early days of computing.
 Therefore in both academia and industry there have been great efforts over the years to create tools that allow us to decrease the amount of bugs in our programs and increase our productivity.
@@
 Spark is a formally defined language based on the Ada \citep{Ada1979} language intended for the development of high integrity software such as flight control systems.
-Others exist as separate, well-known tools that exist alongside the language such as the Clang analysis tools for C and C++ \citep{kremenek2008}.
+Others exist as separate, well-known tools that exist alongside the language such as the Clang analysis tools for C and C++. \footnote{https://clang.llvm.org/docs/ClangTools.html}
 
@@
 
-\subsection{Into Static Analysis}
+\subsection{How static analysis works}
 
-Underlying all of the previously mentioned static code analysis tools, is their use of the \emph{type system} of the language.
+Underlying how all of the previously mentioned static code analysis tools work, is through their use of the \emph{type system} of the language.
 Type systems at their most basic work by assigning types to various constructs of a program so that incompatible types cannot be used.
@@
 For example, one common area of error in C code that the Clang tools warn about is implicit conversions, where the value of an expression is a different type from the one expected, such as the use of a floating point number where an integer is expected.
-The C language will implicitly convert the value to an integer, however this can cause unexpected behaviour, unbeknownst to the programmer.
-Whilst being inconvenient, this conversion needs to happen as the type system of the language dictates that you cannot use any other type other than an integer where an integer is expected.
+The C language will implicitly convert the value to an integer, however this happens unbeknownst to the programmer, which can cause unexpected behaviour later on in the program.
+The tools, using the same knowledge that you can only use an integer where an integer is expected, can therefore flag this as being a potential bug.
 
@@
 The effectiveness of static analysis tools and more generally the \emph{safety} and \emph{expressivity} of a programming language therefore depends on the type system that is used.
-Type systems have evolved from the relatively simple systems in languages like C to more complex ones such as in Java where there are custom object types and subtyping relations where types can be substituted for one another.
+Type systems have evolved from the relatively simple systems in languages like C to more complex ones such as in Java which have class types and subtyping relations where types can be substituted for one another.
 
@@
 
-The aim of this project is to implement a new programming language, which we further call Semantic Featherweight Java (SFJ), which uses a new proposed type system consisting of boolean types and a semantic subtyping algorithm.
-This project aims to transform the mathematical definition of such a type system into an actual software implementation, solving the issues encountered doing so.
-The language should be able to illustrate the improved solutions to the common example problems encountered in a language without these features such as Java.
+The aim of this project is to implement a new programming language, which we further call Semantic Featherweight Java (SFJ), which uses a new proposed type system consisting of a semantic subtyping algorithm, which is based on set-theoretic models, and boolean type connectives such as: \emph{and}, \emph{or} and \emph{not}.
+
+While there is some non-trivial mathematical machinery involved in implementing such a type system, once it is implemented, this more powerful type system can be used by programmers without knowledge of the implementation underneath.
+
+This project aims to transform the mathematical definition of such a type system by \citet{Dardha2017} into an actual software implementation, solving the issues encountered doing so.
+The language should be able to illustrate improved solutions to the example problems which are encountered in a language without these features such as Java.
 Furthermore, the language will be evaluated so see whether these additional language features are intuitive to use.
@@
           \textbf{Chapter 2} provides background about type systems and Featherweight Java, the language on top of which the new type system is implemented.
-          We also discuss some existing work on semantic type systems and boolean types.
+          We also discuss the tools which were used in this project and existing work on semantic type systems and boolean types.
     \item
-          \textbf{Chapter 3} TODO
+          \textbf{Chapter 3} introduces the design a    nd implementation of SFJ. This includes the syntax of the language, and the key algorithms which were used to implement the type system.
     \item
@@
 
+\section{Nominal vs Structural Subtyping}
+
+Orthogonally to the discussion above of syntactic versus semantic approaches, for object-oriented languages such as SFJ, there also exists two other opposing approaches to defining a subtyping relation: the \emph{nominal} and \emph{structural} approaches.
+
+Nominal subtyping is based on explicit declarations by the developer of the class-hierarchy and is the approach used in Java.
+$A$ is a subtype of $B$ if and only if it declared to be so by including an explicit declaration that $A$ extends $B$.
+In Java, every user-defined class either extends another user-defined class or extends the root $Object$ class, creating a tree-like subtyping hierarchy.
+
+Structural subtyping uses the structure of classes, namely its fields and methods, to define the subtyping relation.
+$A$ is a subtype of $B$ if and only if the fields and methods of $B$ are a subset of the fields and methods of $A$, and that field and method return types are covariant and the method argument types are contravariant.
+This is similar to the requirements in Liskov's substitution principle \citep{Liskov1994} without the behavioural conditions.
+
+$$
+    \begin{array}{ll}
+        \syndecl{Apple}{\mathit{Object}} {
+         &
+            \qquad
+            \syndecl{Steel}{\mathit{Object}} {
+                \\
+                \qquad\quad \mathbf{float}\ \mathit{setPrice}\ (\mathbf{int}\ a)\{\ \ldots\ \}\qquad
+         &
+                \qquad\qquad\quad \mathbf{int}\ \mathit{quality}\
+                \\
+            }
+            \qquad\quad\qquad
+         &
+            \qquad\qquad\quad \mathbf{int}\ \mathit{setPrice}\ (\mathbf{float}\ a)\{\ \ldots\ \}
+            \\
+         &
+            \qquad }
+    \end{array}
+$$
+
+We can see in the example above that $Steel$ is a structural subtype of $Apple$, as we can see that the use of any field or method of $Apple$ would also be valid for an object of type $Steel$.
+Similarly any field or method return types in $Steel$ are the same or less generic than in $Apple$, and any method argument types are the same or more generic in $Steel$ than in $Apple$.
+However, $Apple$ is not a subtype of $Steel$ as $Apple$ is missing the $quality$ field.
+However in some cases a pair of classes can both be subtypes of one another.
+
+Nominal subtyping is more popular than structural subtyping, which makes it perhaps unsurprising that nominal subtyping aligns well with the syntactic approach, and structural subtyping aligns well with the semantic approach.
+
+In our implementation of SFJ based on the work of \citet{Dardha2017, Dardha2013}, we implement a semantic type system however we include both nominal and structural subtyping as we will further show that this approach gives the programmer a choice in which approach he may wish to use, but also allows greater flexibility and compactness for the logic a programmer may wish to implement.
+
 \section{Featherweight Java}
@@
 
-\paragraph{On multimethods}
-Since FJ is a core language, some features of the full Java are removed: overloading methods is one of them.
-In our framework, leveraging the expressivity of boolean connectives and semantic subtyping, many of these features, and language constructs, e.g., lists, are restored: overloading methods is one of them.
-We can thus model \emph{multimethods} \citep{BC97}, which according to the authors is ``\emph{very clean and easy to understand [...] it would be the best solution for a brand new language}".
-As an example \citet{Dardha2013,Dardha2017}, consider the following class declarations:
+\paragraph{Multimethods}
+While FJ removes some feature of the full Java language, the introduction of boolean types and semantic subtyping restores some of these for free.
+Overloaded methods is one of these.
+As suggested by \citet{Dardha2017}, we can thus model \emph{multimethods} \citep{BC97}, which according to the authors is ``\emph{very clean and easy to understand [...] it would be the best solution for a brand new language}".
+As an example \citet{Dardha2013,Dardha2017} consider the following class declarations:
+
 $$
@@
 $$
+
 Method {\it length} has type $\mathbf{string} \rightarrow \mathbf{int}$ in $A$.
@@
 The syntax of terms is given by the following grammar and is based on the standard syntax of terms in FJ \cite{Igarashi1999,Dardha2013,Dardha2017}.
+
 We assume an infinite countable set of names, with some special names: $\mathit{Object}$ indicates the root class, $\synthis$  indicates the current and $\synsuper$ indicates the parent object.
 We let  $A, B, C, \ldots$ range over classes; $a, b, \ldots$ over fields; $m, n, \ldots$ over methods and $x, y, z, \ldots$ range over variables.
-Constants $c$ range over an infinite countable set $\mathcal{K}$.
 
@@
      & \mbox{\textit{Method declaration}} & M \; ::= \; & \alpha \ m\ (\alpha \ x)\ \{\ \synret e; \}                             \\
-     & \mbox{\textit{Expressions}}        & e \; ::=\;  & x\ |\  c\ |\ e.a\ |\ e.m(e) \ | \ \synnew C(\wt{e})
+     & \mbox{\textit{Expressions}}        & e \; ::=\;  & x\ |\ e.a\ |\ e.m(e) \ | \ \synnew C(\wt{e})
 \end{align*}
 
-A {program} $(\wt{L}, e)$ is a pair of a sequence of class declarations $\wt L$, giving rise to a class hierarchy as specified by the inheritance relation, and an expression $e$ to be evaluated.
-%
+A {program} $(\wt{L}, e)$ consists of a sequence of class declarations $\wt L$ and an expression $e$ to be evaluated.
+
 A class declaration $L$ specifies the name of the class, the name of the parent class it extends, its typed fields , the constructor $K$ and its method declarations $M$.
 The constructor $K$ initializes the fields of the object by assigning values to the fields inherited by the super-class and to the fields declared in the current \textbf{this} class.
-%
 A method declaration $M$ specifies the signature of the method, namely the return type, the method name and the formal parameter as well as the body of the method.
-Notice that in our theoretical development we use unary methods, without loss of generality: tuples of arguments can be modelled by an object that instantiates a ``special''class containing as fields all the needed arguments.
 Expressions $e$ include variables, constants, field accesses, method invocations and object creations.
+
+In the theoretical development by \citet{Dardha2017}, unary methods are used without loss of generality: tuples of arguments can be modelled by an object that instantiates a ``special''class containing as fields all the needed arguments.
+
+We include the translation of the syntax of term into the ANTLR grammar for SFJ in Appendix \ref{lst:sfjgram}.
+
+\section{Implementation}
+
+\subsection{Finite Types}
+
+Since we want to use types $\tau$ in practice in SFJ, we restrict them to finite trees whose leaves are basic types with no cycles.
+For example, a recursive type $A = [a : A]$ denotes an infinite program tree \textbf{new} ${A}$(\textbf{new} $A$($\cdots{}$)), hence we avoid it as it is uninhabitable.
+Similarly the types $A = [b: B]$, $B = [a = A]$ create a cycle in our program tree and would also be impossible to instantiate.
+
+These kind of type definitions are allowed but still inhabitable in Java since you can always instantiate a object of any type by assigning the value $null$ to it.
+However, since we restrict ourselves to the functional fragment of Java, we do not allow this.
+
+Given a SFJ program, we use the ANTLR grammar to create a parser for our language and run this parser on the program to give us an abstract syntax tree (AST).
+When processing the AST of a program in SFJ, we forbid fields to have the same type as the type of the class they are defined in.
+Secondly, we mark any classes containing fields typed with only basic types as \emph{resolved} otherwise, as \emph{unresolved}.
+This is then used by Algorithm \ref{alg:types}, which checks if the type definitions in the program are valid.
+
+\begin{algorithm}
+    \DontPrintSemicolon
+    \KwData{$classes$, a set of classes, each marked as resolved or unresolved depending on if their fields contain only basic types}
+    \KwResult{$True$ if classes are all valid type definitions, $False$ if not.}
+    \Begin{
+        \Do{$resolutionOccured = true$}{
+            $resolutionOccured \longleftarrow false$ \;
+            \For{class that is unresolved in classes}{
+                $resolved \longleftarrow true$ \;
+                \For{field in class that contains a class type}{
+                    \For{type in field that is a class type}{
+                        \If{type is unresolved}{
+                            $resolved \longleftarrow false$
+                        }
+                    }
+                }
+                \;
+                \If{$resolved = true$}{
+                    $class \longleftarrow resolved$ \;
+                    $resolutionOccured \longleftarrow true$ \;
+                }
+            }
+        }
+        \;
+        \uIf{all classes are not resolved}{
+            \KwRet $False$
+        }
+        \Else{
+            \KwRet $True$
+        }
+    }
+
+\caption{Algorithm for given a set of classes which are marked as resolved or unresolved, determines if all the classes can be resolved, i.e.
+if all the types are finite trees with no cycles and with basic types as leaves}
+\label{alg:types}
+\end{algorithm}
+
+If the types in the program are finite trees whose leaves are constants with no cycles, then at each iteration of the algorithm we are going to be able to resolve at least one type or all the types are going to be resolved.
+If we do not resolve at least one type and not all types are resolved, we know we have encountered a cycle in the type definition.
+
+\subsection{Defining the subtyping relation}
+
+Now, given that we know that the class type definitions in our program are valid, we can define the subtyping relation for this program.
+
+Building upon the interpretation of types as sets of values, we define the subtyping relation by defining a map from a type to the set of its subtypes, with the property that the set of values of a subtype is included in the set of values of the type.
+As a first step, we start with basic types.
+
+\begin{flalign}
+    \label{fig:initialSubtype}
+    &Double  = \left\{Double, Float, Int, Short, Byte\right\} &&Float = \left\{Float, Short, Byte\right\}\\
+    &Long = \left\{Long, Int, Short, Byte\right\}  &&Int = \left\{Int, Short, Byte\right\}\\
+    &Short = \left\{Short, Byte\right\} && Byte = \left\{Byte\right\}\\
+    &Boolean = \left\{Boolean \right\}
+\end{flalign}
+
+We would like to note that \emph{Int} is not a subtype of \emph{Float}, as by the definition of floating-point numbers in Java \footnote{See IEEE-754 for reference}, \emph{Float} only has 23 bits for the mantissa, meaning it cannot represent all possible values of a 32 bit \emph{Int} accurately and therefore \emph{Int} is not fully set-contained.
+However, this is not the case for \emph{Int} and \emph{Double} as the latter has a 52 bit mantissa.
+This is also why \emph{Long} is not a subtype of \emph{Double}.
+
+Finally, Algorithm \ref{alg:subtyping} defines the subtyping relation for all class types, which concludes the process.
+
+\begin{algorithm}
+    \DontPrintSemicolon
+    \KwData{
+        $classes$, \;
+
+        $relation$, a map of types to the set of its subtypes.\ref{fig:initialSubtype}
+    }
+
+    \Begin{
+        \Fn{generateRelation(classes: List<Class>)}{
+            $untyped: List<Class>\ \longleftarrow []$
+
+            \For{class in classes}{
+                \If{addClass(class) = $False$}{
+                    $untyped.add(class)$
+                }
+            }
+            \If{$untyped \neq []$}{
+                $generateRelation(untyped)$
+            }
+        }
+        \;
+        \Fn{addClass(class: Class) $\rightarrow$ boolean}{
+            \For{existing class type in relation}{
+                \If{checkSuperSet(class, existingClass) = false}{
+                    \KwRet $False$\;
+                }
+                $checkSuperSet(existingClass, class)$\;
+            }
+            $relation[class].add(class)$ \;
+            \KwRet $True$ \;
+        }
+        \;
+        \Fn{checkSuperSet(class: Class, other: Class) $\rightarrow$ boolean}{
+            $flag \longleftarrow True$ \;
+            \;
+            \For{field in class}{
+                \If{field contains type not in relation}{
+                    \KwRet $False$
+                }
+                \uIf{other does not contain field}{
+                    $flag \longleftarrow False$
+                }
+                \Else{
+                    \If{other.field.types does not fully contain field.types}{
+                        $flag \longleftarrow False$
+                    }
+                }
+            }
+            \;
+            \For{method in class}{
+                \If{method contains type not in relation}{
+                    \KwRet $False$
+                }
+                \uIf{other does not contain method}{
+                    $flag \longleftarrow False$
+                }
+                \Else{
+                    \If{other.method.types does not fully contain method.types}{
+                        $flag \longleftarrow False$
+                    }
+                }
+            }
+            \;
+            \If{$flag = True$}{
+                $relation[other].add$
+            }
+        }
+    }
+
+\caption{Recursive algorithm for creating a semantic subtyping relation, given a set of class types that are finite trees and the initial relation for basic types.
+Given that we know all classes are valid types, we are guaranteed it will terminate}
+\label{alg:subtyping}
+\end{algorithm}
 
@@
 
-    \begin{lstlisting}
-        grammar jsf;
-
-        @header{
-        package jsf;
-        }
+    \begin{lstlisting}[label={lst:sfjgram}, caption={The full ANTLR grammar for SFJ language}]
+        grammar sfj;
 

### 2020-04-06 src/main/java/jsf/App.java
allow to specify the input file from command line

@@ -25,7 +25,7 @@ public final class App {
     try {
       // TODO change this to get input from args
       final var inputStream = CharStreams.fromPath(
-          Paths.get(System.getProperty("user.dir"), "src/main/resources/example.jsf"));
+          Paths.get(args[0]));
 
       final var lexer = new jsfLexer(inputStream);
       final var commonTokenStream = new CommonTokenStream(lexer);

### 2020-04-06 dissertation/l4proj.tex
final dissertation

@@
 \usepackage{amsmath, amssymb}
+\usepackage{float}
+\usepackage{changepage}
 
@@
 
-\title{Boolean types and semantic subtyping for Featherweight Java}
+\title{SFJ - Boolean types and semantic subtyping for Featherweight Java}
 \author{Artem Usov}
-\date{January 18, 2020}
+\date{April 6, 2020}
 
@@
 \begin{abstract}
-    Every abstract follows a similar pattern. Motivate; set aims; describe work; explain results.
+    The type system of a programming language often dictates its safety and expressivity.
+    There are two distinct approaches to defining a type system: the \emph{syntactic} and \emph{semantic}.
+    In the semantic approach, one defines a model of the language and an interpretation of types as subsets of this model.
+    Subtyping is defined as inclusion of subsets denoting types.
     \vskip 0.5em
-    ``XYZ is bad. This project investigated ABC to determine if it was better.
-    ABC used XXX and YYY to implement ZZZ. This is particularly interesting as XXX and YYY have
-    never been used together. It was found that
-    ABC was 20\% better than XYZ, though it caused rabies in half of subjects.''
+    In this paper we present SFJ---Semantic Featherweight Java, an extension of Featherweight Java, which implements such a semantic type system defined by \cite{Dardha2013, Dardha2017}.
+    We further exploit it to integrate boolean types as well as nominal and structural subtyping.
+    This paper illustrates the benefits of such language constructs by allowing us to write programs that have higher guarantees of correctness by increasing the static type-checking power of the language whilst being more intuitive at the same time.
 \end{abstract}
@@
 
+\renewcommand{\abstractname}{Acknowledgements}
+
+\begin{abstract}
+    I would like to thank my supervisor Dr Ornela Dardha for her support and guidance throughout the year as well as her enthusiasm for lecturing which is made interested in programming languages and type systems.
+    \vskip 0.5em
+    I would also like to thank my parents who have always supported and believed in me.
+\end{abstract}
+
+%===================================================================================================
+
 \def\consentname {Artem Usov}
-\def\consentdate {19 January 2020}
+\def\consentdate {6 April 2020}
 \educationalconsent
@@
 
-\section{End goal of programming}
+\section{End Goal of Programming}
 
 The end goal in all programming projects is to create a solution to the task at hand that exactly solves the initially defined problem and performs exactly as expected by the programmer.
-However, this very rarely actually happens.
+However, this rarely actually happens.
 
@@
 
-In general these are called \emph{bugs} in a program, and have become more and more common as projects and the underlying hardware they run on have become more complex since the early days of computing.
+In general these are called \emph{bugs} in a program, and have become more and more common as projects and the underlying hardware they run on have become more complex.
 Therefore in both academia and industry there have been great efforts over the years to create tools that allow us to decrease the amount of bugs in our programs and increase our productivity.
@@
 Spark is a formally defined language based on the Ada \citep{Ada1979} language intended for the development of high integrity software such as flight control systems.
-Others exist as separate, well-known tools that exist alongside the language such as the Clang analysis tools for C and C++. \footnote{https://clang.llvm.org/docs/ClangTools.html}
-
-Whilst they are not necessarily needed to write correct programs, these tools have quickly become industry standards for maintaining a level of correctness within codebases and avoiding typical programming errors.
-
-\subsection{How static analysis works}
-
-Underlying how all of the previously mentioned static code analysis tools work, is through their use of the \emph{type system} of the language.
+Others exist as separate, well-known tools that exist alongside the language such as the Clang analysis tools for C and C++ \footnote{https://clang.llvm.org/docs/ClangTools.html}.
+
+Whilst they are not necessary to write correct programs, these tools have quickly become industry standards for maintaining a level of correctness within codebases and avoiding typical programming errors.
+
+\subsection{How Static Analysis Works}
+
+Underlying how all of the previously mentioned static code analysis tools work, is their use of the \emph{type system} of the language.
 Type systems at their most basic work by assigning types to various constructs of a program so that incompatible types cannot be used.
@@
 The C language will implicitly convert the value to an integer, however this happens unbeknownst to the programmer, which can cause unexpected behaviour later on in the program.
-The tools, using the same knowledge that you can only use an integer where an integer is expected, can therefore flag this as being a potential bug.
+The tools, using their knowledge of the type system that you can only use an integer where an integer is expected, can therefore flag this as being a potential bug.
 
@@
 
-The effectiveness of static analysis tools and more generally the \emph{safety} and \emph{expressivity} of a programming language therefore depends on the type system that is used.
+The effectiveness of static analysis tools and more generally the \emph{safety} and \emph{expressivity} of a programming language therefore depends on the type system that is uses.
 Type systems have evolved from the relatively simple systems in languages like C to more complex ones such as in Java which have class types and subtyping relations where types can be substituted for one another.
 
-More modern advances in type systems include the use of linear types in the Cyclone \citep{grossman2002} language that allow us to more naturally and logically define finite state machines.
-There also exist other new developments in type systems such as Boolean types which are explored in this paper.
+More modern advances in type systems include the use of linear types in the Cyclone \citep{grossman2002} language which allows us to more naturally and logically define finite state machines.
+There also exist other new developments in type systems such as boolean types which are explored in this paper.
 
@@
 The typical hierarchical subtyping definitions in Java quite often restrict and complicate the logic that the programmer wants to implement.
-These problems come up particularly often when working with legacy code that has incorrectly defined abstractions in its classes.
-This therefore leads to less understandable and less maintainable code that could cause problems in the future.
+These problems come up particularly often when working with legacy code which has incorrectly defined abstractions in its classes.
+This therefore leads to less understandable and less maintainable code which could cause problems in the future.
 
@@
 
-This project aims to transform the mathematical definition of such a type system by \citet{Dardha2017} into an actual software implementation, solving the issues encountered doing so.
+This project aims to transform the mathematical definition of such a type system by \citet{Dardha2013, Dardha2017} into an actual software implementation, solving the issues encountered doing so.
 The language should be able to illustrate improved solutions to the example problems which are encountered in a language without these features such as Java.
@@
     \item
-          \textbf{Chapter 3} introduces the design a    nd implementation of SFJ. This includes the syntax of the language, and the key algorithms which were used to implement the type system.
+          \textbf{Chapter 3} introduces the design and implementation of SFJ. This includes the syntax of the language, and the key algorithms which were used to implement the type system.
     \item
-          \textbf{Chapter 4} TODO
+          \textbf{Chapter 4} shows how using the semantic type system, we can now easily add language constructs that exploit the set-theoretic implementation of subtyping.
     \item
-          \textbf{Chapter 5} TODO
+          \textbf{Chapter 5} explains the process of code generation for a SFJ program so that it can be run on a computer.
     \item
@@
     \item
-          Finally \textbf{chapter 7} concludes the dissertation with a summary of the important results.
+          \textbf{Chapter 7} briefly describes the conference paper that was submitted detailing SFJ as a new innovative programming tool.
+    \item
+          Finally \textbf{chapter 8} concludes the dissertation with a summary of the important results.
 \end{itemize}
@@
 
-In this chapter, we define and explore type systems and the different ways they can be defined.
-We especially explore the semantic method of defining a type system and explore the benefits and challenges that it brings.
-We also discuss the tools that were used in this project and the related work that has already been done with boolean types or semantic type systems.
+In this chapter, we explore type systems and the different ways they can be defined.
+We especially explore the semantic approach of defining a type system and explore the benefits and challenges that it brings.
+We also discuss the tools that were used in this project and the related work that has already been done relating to boolean types or semantic type systems.
 
 \section{Type Systems}
+\label{sec:framework}
 
@@
 Most importantly, we want to emphasize from the first definition that in a type system, we aim to classify phrases in our language into a specific \emph{type}.
-Then, using the type system, we can check that the types of the arguments given to any operation are always correct, thus avoiding a type error.
+Then, using the type system, we can check that the types of the arguments given to any operation are always correct, thus avoiding type errors.
 
 For example, taking the addition operation which takes two arguments of the number type, then given the arguments $42$ and $2 + 2$, we know they both hold or compute a number type and so there is guaranteed to be no type error.
-Conversely, given the arguments $42$ and the boolean value $true$, they are not both a number type and so cause a type error.
+Conversely, given the arguments $42$ and the boolean value $true$, they are not both a number type and so would cause a type error.
 We therefore avoid unexpected bad run-time behaviour and instead find these problems before even running the program.
@@
 However what the first definition is missing is that it defines a type system to be solely a syntactic method, in which the system is defined as a list of formal deduction rules.
-The syntactic approach is certainly by far the most common approach for a type system, but there does exist an alternative method in the semantic approach.
+The syntactic approach is certainly by far the most common approach for a type system, but there does exist an alternative in the semantic approach.
 \citet{Frisch2002} describe the semantic approach as instead starting with a model of the language and all the possible values in the language and defining an interpretation of types as subsets of values the model.
@@
 
-\begin{equation*}
+\begin{equation}
     D = \left[ a + bi \mid a, b \in \mathbb{R}\right]
-\end{equation*}
+\end{equation}
 
@@
 Such as model comes with several advantages.
-For example, given two types $s$ and $t$, which represent subsets of values in the model, when $s \subseteq t$ does not hold, it is possible to exhibit an element of the model which is in the interpretation of type $s$ and not $t$.
+For example, given two types $s$ and $t$ which represent subsets of values in the model, when $s \subseteq t$ does not hold then it is possible to exhibit an element of the model which is in the interpretation of type $s$ and not $t$.
 This can then be used to show more informative error messages to the programmer as shown by \citet{Castagna2005}, such as showing which value in $s$ is causing the equality not to hold.
 Secondly the set based approach to the type system leads itself to able to incorporate boolean operators into its types easily.
-However due to being a more technical approach, such as the non-triviality of defining the interpretation of types as subsets of a model as discussed below, the semantic method has received less attention than syntactic methods.
-
-The first problem we encounter in the semantic approach is that in such a model, $t_{1}$ is a subtype of $t_{2}$ if all the $t_{1}$-values are also $t_{2}$-values, i.e.
+However due to being a more technical approach, such as the non-triviality of defining the interpretation of types as subsets of a model discussed below, the semantic method has received less attention than syntactic methods.
+
+The first problem we encounter in the semantic approach is that in such a model $t_{1}$ is a subtype of $t_{2}$ if all $t_{1}$-values are also $t_{2}$-values, i.e.
 the equality $t_{1} \subseteq t_{2}$ holds.
-However, in this way, subtyping is defined by relying on the notion of well-typed values; hence, we need the typing relation to able to type values; but the typing rules require a subtyping relation to be fully defined.
+However, in this way, subtyping is defined by relying on the notion of well-typed values; hence we need the typing relation to able to type values; but the typing rules require a subtyping relation to be fully defined.
 So, there is a circularity in our definition, as demonstrated in Figure \ref{fig:circ}.
@@
 The general idea of the framework is that we first extend the types in the language with \emph{Boolean Combinators}: union $\lor$, intersection $\land$ and negation $\neg$ to give a type algebra $\tau$.
-We can then define an abstract model $\beta$ with a interpretation function $[\![ ]\!]_{\beta} : \tau \to \rho(\beta)$ (where $\rho(\beta)$ is the power-set of $\beta$).
-This model must capture the meaning of each type but also interprets the combinators in a set-theoretic way, such as $[\![s \land t]\!]_{\beta} = [\![s]\!]_{\beta} \cap [\![t]\!]_{\beta}$ etc.
+We can then define an abstract model $B$ with a interpretation function $[\![ ]\!]_{B} : \tau \to \rho(B)$ (where $\rho(B)$ is the power-set of $B$).
+This model must capture the meaning of each type but also interprets the combinators in a set-theoretic way, such as $[\![s \land t]\!]_{B} = [\![s]\!]_{B} \cap [\![t]\!]_{B}$.
 This model then induces the following subtyping relation:
 
-\begin{equation*}
-    s \leq_{\beta} t \iff [\![s]\!]_{\beta} \subseteq [\![t]\!]_{\beta}
-\end{equation*}
+\begin{equation}
+    s \leq_{B} t \iff [\![s]\!]_{B} \subseteq [\![t]\!]_{B}
+\end{equation}
 
@@
 
-Separately, now that we have typing rules, we can define a new interpretation of types based on values $[\![t]\!]_{\mathrm{v}} = \left\{\mathrm{v} \in V \mid \: \vdash_{\beta} \mathrm{v} : t \right\}$ and then define a ``new''subtyping relation as above:
-
-\begin{equation*}
+Separately, now that we have typing rules, we can define a new interpretation of types based on values $[\![t]\!]_{\mathrm{v}} = \left\{\mathrm{v} \in V \mid \: \vdash_{B} \mathrm{v} : t \right\}$ and then define a ``new''subtyping relation as above:
+
+\begin{equation}
     s \leq_{\mathrm{v}} t \iff [\![s]\!]_{\mathrm{v}} \subseteq [\![t]\!]_{\mathrm{v}}
-\end{equation*}
+\end{equation}
 
 While these may be different relations, if the models are chosen carefully such that:
-\begin{equation*}
-    s \leq_{\beta} t \iff s \leq_{\nu} t,
-\end{equation*}
+
+\begin{equation}
+    s \leq_{B} t \iff s \leq_{\nu} t,
+\end{equation}
 
@@
     \includegraphics[width=0.4\linewidth]{images/circularity.PNG}
-    \caption{Circularity in the semantic subtyping definition \citep{Castagna2005}}
+    \caption{Circularity in trying to define a semantic subtyping relation \citep{Castagna2005}}
     \label{fig:circ}
@@
 \section{Nominal vs Structural Subtyping}
+\label{sec:struct}
 
@@
 $A$ is a subtype of $B$ if and only if it declared to be so by including an explicit declaration that $A$ extends $B$.
-In Java, every user-defined class either extends another user-defined class or extends the root $Object$ class, creating a tree-like subtyping hierarchy.
+In Java, every user-defined class either extends another class or the root $Object$ class, creating a tree-like subtyping hierarchy.
 
 Structural subtyping uses the structure of classes, namely its fields and methods, to define the subtyping relation.
-$A$ is a subtype of $B$ if and only if the fields and methods of $B$ are a subset of the fields and methods of $A$, and that field and method return types are covariant and the method argument types are contravariant.
-This is similar to the requirements in Liskov's substitution principle \citep{Liskov1994} without the behavioural conditions.
-
-$$
+$A$ is a subtype of $B$ if and only if the fields and methods of $B$ are a subset of the fields and methods of $A$, and if the field and method return types are covariant and the method argument types are contravariant.
+This is similar to the requirements in Liskov's Substitution Principle \citep{Liskov1994} without the behavioural conditions.
+
+\begin{equation}
+    \label{eq:substitutionprinciple}
     \begin{array}{ll}
@@
                 \\
-                \qquad\quad \mathbf{float}\ \mathit{setPrice}\ (\mathbf{int}\ a)\{\ \ldots\ \}\qquad
-         &
-                \qquad\qquad\quad \mathbf{int}\ \mathit{quality}\
+                \qquad\quad \ldots \qquad
+         &
+                \qquad\qquad\quad \ldots
+                \\
+                \qquad\quad \textbf{float}\ \mathit{setPrice}\ (\textbf{int}\ a)\{\ \ldots\ \}\qquad
+         &
+                \qquad\qquad\quad \textbf{int}\ \mathit{quality}
                 \\
@@
          &
-            \qquad\qquad\quad \mathbf{int}\ \mathit{setPrice}\ (\mathbf{float}\ a)\{\ \ldots\ \}
+            \qquad\qquad\quad \textbf{int}\ \mathit{setPrice}\ (\textbf{float}\ a)\{\ \ldots\ \}
             \\
@@
     \end{array}
-$$
-
-We can see in the example above that $Steel$ is a structural subtype of $Apple$, as we can see that the use of any field or method of $Apple$ would also be valid for an object of type $Steel$.
-Similarly any field or method return types in $Steel$ are the same or less generic than in $Apple$, and any method argument types are the same or more generic in $Steel$ than in $Apple$.
+\end{equation}
+
+We can see in Equation \ref{eq:substitutionprinciple} that $Steel$ is a structural subtype of $Apple$, as the use of any field or method of $Apple$ would also be defined in $Steel$.
+All field or method return types in $Steel$ are the same or more restrictive than those that also occur in $Apple$, and all method argument types are the same or less restrictive in $Steel$ than in $Apple$.
 However, $Apple$ is not a subtype of $Steel$ as $Apple$ is missing the $quality$ field.
-However in some cases a pair of classes can both be subtypes of one another.
+In other cases a pair of classes can both be subtypes of one another.
 
@@
 
-In our implementation of SFJ based on the work of \citet{Dardha2017, Dardha2013}, we implement a semantic type system however we include both nominal and structural subtyping as we will further show that this approach gives the programmer a choice in which approach he may wish to use, but also allows greater flexibility and compactness for the logic a programmer may wish to implement.
+In our implementation of SFJ based on the work of \citet{Dardha2017, Dardha2013}, we implement a semantic type system however we include both nominal and structural subtyping.
+This approach gives the programmer a choice in which approach he may wish to use, but also allows both to be used concurrently for greater flexibility and compactness in the logic a programmer may wish to implement.
 
@@
 The syntax of SFJ is exactly that of Featherweight Java (FJ) \citep{Igarashi1999}, with only the associated type system being different.
-FJ is closely related to Java, with a key simplification in the removal of assignment operators.
+FJ is closely related to Java, but with a key simplification in the removal of the assignment operation.
 All fields of an object cannot be changed after initialisation and all methods are pure functions.
 While this restricts FJ to what is essentially a functional fragment of Java, it is still fully computationally complete.
-SFJ was intentionally based on FJ instead of the full Java language because features such as concurrency and reflection are orthogonal to the purpose of demonstrating the implementation of our novel semantic type system.
+SFJ was intentionally based on FJ instead of the full Java language because features such as concurrency and reflection are orthogonal to the purpose of demonstrating of our novel type system.
 
 \paragraph{Multimethods}
-While FJ removes some feature of the full Java language, the introduction of boolean types and semantic subtyping restores some of these for free.
+
+While FJ removes some feature of the  Java language, the introduction of boolean types and semantic subtyping restores some of these.
 Overloaded methods is one of these.
-As suggested by \citet{Dardha2017}, we can thus model \emph{multimethods} \citep{BC97}, which according to the authors is ``\emph{very clean and easy to understand [...] it would be the best solution for a brand new language}".
+As suggested by \citet{Dardha2017}, we can model overloaded methods as \emph{multimethods} \citep{BC97}, which according to the authors is ``\emph{very clean and easy to understand [...] it would be the best solution for a brand new language}".
 As an example \citet{Dardha2013,Dardha2017} consider the following class declarations:
 
-$$
+\begin{equation}
+    \label{multi1}
     \begin{array}{ll}
@@
                 \\
-                \qquad\quad \mathbf{int}\ \mathit{length}\ (\mathbf{string}\ s)\{\ \ldots\ \}\qquad
-         &
-                \qquad\qquad\quad \mathbf{int}\ \mathit{length}\ (\mathbf{int}\ n)\{\ \ldots\ \}
+                \qquad\quad \ldots \qquad
+         &
+                \qquad\qquad\quad \ldots
+                \\
+                \qquad\quad \textbf{int}\ \mathit{length}\ (\textbf{string}\ s)\{\ \ldots\ \}\qquad
+         &
+                \qquad\qquad\quad \textbf{int}\ \mathit{length}\ (\textbf{int}\ n)\{\ \ldots\ \}
                 \\
@@
     \end{array}
-$$
-
-Method {\it length} has type $\mathbf{string} \rightarrow \mathbf{int}$ in $A$.
-However, in $B$ it has type $(\mathbf{string} \rightarrow \mathbf{int}) \synwedge (\mathbf{int} \rightarrow \mathbf{int})$,
-which can be simplified as $(\mathbf{string} \synvee \mathbf{int}) \rightarrow \mathbf{int}$.
+\end{equation}
+
+Method \textit{length} has type $\textbf{string} \rightarrow \textbf{int}$ in $A$.
+However, in $B$ it has type $(\textbf{string} \rightarrow \textbf{int}) \synwedge (\textbf{int} \rightarrow \textbf{int})$, which can be simplified as $(\textbf{string} \synvee \textbf{int}) \rightarrow \textbf{int}$.
 
@@
 The type system of FSJ was built using ANTLR \citep{parr2013} to define the grammar of the language and automatically create a parser for this grammar rather than having to create one by hand, which made the initial development process much quicker.
-It accepts a grammar using Extended Backus-Naur Form (EBNF) \footnote{See ISO/IEC 14977 for reference, although it is not the standard used in the examples} notation to create ANTLR rules which are a list of productions or alternatives.
+It accepts a grammar using Extended Backus-Naur Form (EBNF) \footnote{See ISO/IEC 14977 for reference, although this is not the standard used in the examples} notation to create ANTLR rules which are a list of productions or alternatives.
 A general form of a rule is as follow:
 
-$$
+\begin{equation}
     \begin{array}{llll}
@@
     \end{array}
-$$
-
-Each alternative production in a rule can itself be a list of elements, where an element can be another rule or just a terminating token.
-Since we are using EBNF notation, we can also use \emph{*} and \emph{?} to respectively signify repeated and conditional elements in a production.
+\end{equation}
+
+Each alternative production in a rule can itself be a list of elements, where an element can be another rule or a terminating token.
+Since we are using EBNF notation, we can also use $*$ and $?$ to respectively signify repeated and conditional elements in a production.
 We can also use $|$ to give several alternatives for a single element.
@@
 
-$$
+\begin{equation}
     \begin{array}{llll}
-        expression     & : & \mbox{primExpression ( (PLUS | MINUS | DIV | MULT) primExpression)?}
+        expression     & : & primExpression\ ( (PLUS | MINUS | DIV | MULT)\ primExpression)?
         \\
@@
         \\
-        primExpression & : & \mbox{NUMBER}
-        \\
-                       & | & \mbox{TRUE}
+        primExpression & : & NUMBER
+        \\
+                       & | & TRUE
         \\
@@
         \\
-        TRUE           & : & \mbox{'true'}
+        TRUE           & : & 'true'
         \\
@@
         \\
-        NUMBER         & : & \mbox{DIGIT (DIGIT)*}
+        NUMBER         & : & DIGIT\ (DIGIT)*
         \\
@@
     \end{array}
-$$
+\end{equation}
 
 \section{Related Work}
-The closest similar area of research would likely be the work on $\mathbb{C}$Duce
-\citep{Benzaken2003}, which is xalso a functional language with a semantic type system designed for
-working with XML documents and a continuation of the work on XDuce \citep{Hosoya2003}. The language
-extended XDuce by introducing less XML specific types such as records, boolean connectives and arrow
-types. This therefore makes it similar to (\textit{TODO reword}) our language in that a class-based
-semantic type system is a combination of the $\mathbb{C}$Duce record types with arrow types
-\textit{(fields and methods)}
-
-Our work and the work on $\mathbb{C}$Duce follow the functional style of $\lambda$-calculus,
-whereas the work by \citet{Castagna2008} extends $\pi$-calculus with semantic subtyping.
-(\textit{TODO reword}) Similar work to ours creating an implementation for this would result in a
-Golang-like \footnote{https://golang.org/} like language, creating a concurrency-focused language with more
-intuitive types. The paper found that it was required to be able to decide and resolve the atomicy,
-that is whether the only proper subtype is the empty type, of types in order to decide the subtyping
-relation, and observes that this same problem appears in $\lambda$-calculus and any other
-semantic-based system, for which we will show where we encountered this problem further in this
-paper.
+The closest area of research would likely be the work on $\mathbb{C}$Duce
+\citep{Benzaken2003}, which is also a functional language with a semantic type system designed for
+working with XML documents and a continuation of the work on XDuce \citep{Hosoya2003}.
+The language extended XDuce by introducing less XML specific types such as records, boolean connectives and arrow types.
+This therefore makes it similar to our language in that a class-based semantic type system is a combination of the $\mathbb{C}$Duce record types with arrow types.
+Muehlboeck and Tate \cite{Muehlboeck2018} define a syntactic framework with boolean connectives which has been implemented in the Ceylon programming language \citep{Ceylon2016}.
+
+Our work and the work on $\mathbb{C}$Duce follow the functional style of $\lambda$-calculus, whereas the work by \citet{Castagna2008} extends $\pi$-calculus with semantic subtyping.
+Similar work to ours creating an implementation for this would result in a Golang-like \footnote{https://golang.org/} language, creating a concurrency-focused language with more
+intuitive types.
+\citet{Castagna2008} found that it was required to be able to decide and resolve the atomicy, that is whether the only proper subtype is the empty type, of types in order to decide the subtyping relation, and observes that this same problem appears in $\lambda$-calculus and any other semantic-based system.
+This is exactly the problem we find and solve in that we need to validate whether our type definitions are finite trees with basic types as leaves and with not cycles.
 
@@
 
-% \chapter{Analysis/Requirements}
-% What is the problem that you want to solve, and how did you arrive at it?
-% \section{Guidance}
-% Make it clear how you derived the constrained form of your problem via a clear and logical process.
-
-%===================================================================================================
-\chapter{SFJ: Design and Implementation}
+\chapter{Syntax and Subtyping Implementation}
+
+This chapter provides the formal syntax of the language and gives a detailed explanation of the implementation of the first core novel feature in SFJ, the semantic type system and subtyping relation.
+We then consider the algorithms used in the implementation for their efficiency and any flaws the subtyping relation may have.
 
 \section{Syntax}
+\label{sec:syntax}
 
@@
 
-\label{sec:design}
 The syntax of types is given by the following grammar \citep{Dardha2013,Dardha2017}:
-$$
+
+\begin{equation}
+    \label{eq:syntaxtypes}
     \begin{array}{llll}
         \tau   & ::=                                          & \alpha \ |\ \mu
-               & \mbox{{\em Type term}}
+               & \mbox{\textit{Type term}}
         \\
         \alpha & ::=                                          & \zero \ |\ \btypes \ |\ [\wt{l:\tau}] \ |\ \alpha \ \bf{and}\ \alpha \ |\ \bf{not}\ \alpha
-        \qquad
-               & \mbox{{\em Object type} ($\alpha $-type)}
+               & \mbox{\textit{Object type} ($\alpha $-type)}
         \\
         \mu    & ::=                                          & \alpha \to \alpha \ |\ \mu \ \bf{and}\ \mu \ |\ \bf{not}\ \mu
-               & \mbox{{\em Method type} ($\mu $-type)}
-    \end{array}
-$$
+               & \mbox{\textit{Method type} ($\mu $-type)}
+    \end{array}
+\end{equation}
 
@@
 Type $\btypes$ denotes the \emph{basic} types, such as integers, booleans, etc.
-    {Record} types $[\wt{l:\tau}]$, where $\wt l$ is a sequence of disjoint labels are used to type objects.
+Record types $[\wt{l:\tau}]$, where $\wt l$ is a sequence of disjoint labels, are used to type objects.
 Arrow types $\alpha \to \alpha$ are used to type methods.
-The boolean types using ${and}$ and ${not}$ have their expected set-theoretic meanings, and ${or}$ is obtained by their combination.\\
+The boolean types using ${and}$ and ${not}$ have their expected set-theoretic meanings, and ${or}$ is obtained by their combination.
 
@@
 
-We assume an infinite countable set of names, with some special names: $\mathit{Object}$ indicates the root class, $\synthis$  indicates the current and $\synsuper$ indicates the parent object.
+We assume an infinite countable set of names, with some special names: $\mathit{Object}$ indicates the root class, $\mathit{this}$  indicates the current and $\mathit{super}$ indicates the parent object.
 We let  $A, B, C, \ldots$ range over classes; $a, b, \ldots$ over fields; $m, n, \ldots$ over methods and $x, y, z, \ldots$ range over variables.
-
-\begin{align*}
-     & \mbox{\textit{Class declaration}}  & L \; ::= \; & \syndecl{C}{C}{\wt{\alpha \ a};\ K; \ \wt{M}\ }                         \\
-     & \mbox{\textit{Constructor}}        & K \;::=\;   & C\ (\wt{\alpha\ x})\ \{\ \synsuper(\wt{x});\ \wt{\synthis.a}=\wt{x}; \} \\
-     & \mbox{\textit{Method declaration}} & M \; ::= \; & \alpha \ m\ (\alpha \ x)\ \{\ \synret e; \}                             \\
-     & \mbox{\textit{Expressions}}        & e \; ::=\;  & x\ |\ e.a\ |\ e.m(e) \ | \ \synnew C(\wt{e})
-\end{align*}
-
-A {program} $(\wt{L}, e)$ consists of a sequence of class declarations $\wt L$ and an expression $e$ to be evaluated.
-
-A class declaration $L$ specifies the name of the class, the name of the parent class it extends, its typed fields , the constructor $K$ and its method declarations $M$.
-The constructor $K$ initializes the fields of the object by assigning values to the fields inherited by the super-class and to the fields declared in the current \textbf{this} class.
+Constants $c$ range over an infinite countable set $\mathbb{K}$
+
+\begin{equation}
+    \label{eq:syntaxterms}
+    \begin{array}{llllll}
+         & \mbox{\textit{Class declaration}}  & L & \; ::= \; & \syndecl{C}{C}{\wt{\alpha \ a};\ K; \ \wt{M}\ }                         \\
+         & \mbox{\textit{Constructor}}        & K & \;::=\;   & C\ (\wt{\alpha\ x})\ \{\ \synsuper(\wt{x});\ \wt{\synthis.a}=\wt{x}; \} \\
+         & \mbox{\textit{Method declaration}} & M & \; ::= \; & \alpha | \mu \ m\ (\alpha | \mu \ x)\ \{\ \synret e; \}                 \\
+         & \mbox{\textit{Expressions}}        & e & \; ::=\;  & x\ |\  c\ |\ e.a\ |\ e.m(e) \ | \ \synnew C(\wt{e})
+    \end{array}
+\end{equation}
+
+A program $(\wt{L}, e)$ consists of a sequence of class declarations $\wt L$ and an expression $e$ to be evaluated.
+
+A class declaration $L$ specifies the name of the class, the name of the parent class it extends, its typed fields, the constructor $K$ and its method declarations $M$.
+The constructor $K$ initializes the fields of the object by assigning values to the fields inherited by the \textit{super} class and to the fields declared in the current \textit{this} class.
 A method declaration $M$ specifies the signature of the method, namely the return type, the method name and the formal parameter as well as the body of the method.
@@
 
-In the theoretical development by \citet{Dardha2017}, unary methods are used without loss of generality: tuples of arguments can be modelled by an object that instantiates a ``special''class containing as fields all the needed arguments.
-
-We include the translation of the syntax of term into the ANTLR grammar for SFJ in Appendix \ref{lst:sfjgram}.
-
-\section{Implementation}
+We do not allow programs that would cause $\wt L$ to be ill-defined, such as declaring a constructor called $B$ in class $A$, multiple field or methods with the same name or declaring class $A$ which extends a non-existing type.
+All these checks are also the same ones used in FJ \citep{Igarashi1999}.
+
+In the theoretical development by \citet{Dardha2017}, unary methods are used without loss of generality: tuples of arguments can be modelled by an object that instantiates a \emph{special} class containing as fields all the needed arguments.
+
+We include the translation of the syntax of terms into the ANTLR grammar for SFJ in Appendix \ref{lst:sfjgrammar}.
+
+\section{Implementation of subtyping algorithm}
 
@@
 Since we want to use types $\tau$ in practice in SFJ, we restrict them to finite trees whose leaves are basic types with no cycles.
-For example, a recursive type $A = [a : A]$ denotes an infinite program tree \textbf{new} ${A}$(\textbf{new} $A$($\cdots{}$)), hence we avoid it as it is uninhabitable.
-Similarly the types $A = [b: B]$, $B = [a = A]$ create a cycle in our program tree and would also be impossible to instantiate.
-
-These kind of type definitions are allowed but still inhabitable in Java since you can always instantiate a object of any type by assigning the value $null$ to it.
-However, since we restrict ourselves to the functional fragment of Java, we do not allow this.
-
-Given a SFJ program, we use the ANTLR grammar to create a parser for our language and run this parser on the program to give us an abstract syntax tree (AST).
-When processing the AST of a program in SFJ, we forbid fields to have the same type as the type of the class they are defined in.
-Secondly, we mark any classes containing fields typed with only basic types as \emph{resolved} otherwise, as \emph{unresolved}.
-This is then used by Algorithm \ref{alg:types}, which checks if the type definitions in the program are valid.
+For example, the recursive type $A = [a : A]$ denotes an infinite program tree $\textbf{new}\ A(\textbf{new}\ A(\ \cdots\ ))$, hence we avoid it as it is uninhabitable.
+Similarly the types $A = [b: B]$, $B = [a = A]$ create a cycle in our program tree and would also be impossible to inhabit.
+
+The above type definitions are allowed but still uninhabitable in Java since you can always instantiate a object of any type by assigning the value $null$ to it.
+However, since we restrict SFJ to the functional fragment of Java, we cannot do this.
+
+Given a SFJ program, we use the ANTLR grammar, given in Appendix \ref{lst:sfjgrammar}, to create a parser and run this parser to give us an abstract syntax tree (AST) of the program.
+The AST can be visited to make sure that our program is well-defined.
+We mark any classes containing fields typed with only basic types as \emph{resolved} otherwise, as \emph{unresolved}.
+This is then used by Algorithm \ref{alg:types}, which checks if the type definitions in the program are valid, i.e. they are finite trees whose leaves are basic types with no cycles.
 
 \begin{algorithm}
+    \caption{Algorithm which given a set of classes which are marked as resolved or unresolved, determines if all the classes can be resolved, i.e.
+if all the types are finite trees with no cycles and with basic types as leaves}
+    \label{alg:types}
+
     \DontPrintSemicolon
-    \KwData{$classes$, a set of classes, each marked as resolved or unresolved depending on if their fields contain only basic types}
-    \KwResult{$True$ if classes are all valid type definitions, $False$ if not.}
+    \KwData{$classes$, a set of classes, each marked as resolved or unresolved depending on if their fields contain only basic types.}
+    \KwResult{$True$ if all classes are valid type definitions, $False$ otherwise.}
     \Begin{
@@
                 \For{field in class that contains a class type}{
-                    \For{type in field that is a class type}{
-                        \If{type is unresolved}{
+                    \If{type of field is unresolved}{
                         $resolved \longleftarrow false$
                     }
-                    }
+
                 }
@@
         \;
-        \uIf{all classes are not resolved}{
+        \uIf{not all classes are resolved}{
             \KwRet $False$
@@
     }
-
-\caption{Algorithm for given a set of classes which are marked as resolved or unresolved, determines if all the classes can be resolved, i.e.
-if all the types are finite trees with no cycles and with basic types as leaves}
-\label{alg:types}
 \end{algorithm}
 
-If the types in the program are finite trees whose leaves are constants with no cycles, then at each iteration of the algorithm we are going to be able to resolve at least one type or all the types are going to be resolved.
+If the types in the program are finite trees whose leaves are constants with no cycles, then at each iteration of the algorithm we are going to be able to resolve at least one type or all the types are resolved.
 If we do not resolve at least one type and not all types are resolved, we know we have encountered a cycle in the type definition.
 
-\subsection{Defining the subtyping relation}
-
-Now, given that we know that the class type definitions in our program are valid, we can define the subtyping relation for this program.
+\subsection{Defining the Subtyping Relation}
+\label{sec:sub}
+
+Now, given that we know that the type definitions in our program are valid, we can define the subtyping relation for this program.
 
 Building upon the interpretation of types as sets of values, we define the subtyping relation by defining a map from a type to the set of its subtypes, with the property that the set of values of a subtype is included in the set of values of the type.
-As a first step, we start with basic types.
-
-\begin{flalign}
-    \label{fig:initialSubtype}
-    &Double  = \left\{Double, Float, Int, Short, Byte\right\} &&Float = \left\{Float, Short, Byte\right\}\\
-    &Long = \left\{Long, Int, Short, Byte\right\}  &&Int = \left\{Int, Short, Byte\right\}\\
-    &Short = \left\{Short, Byte\right\} && Byte = \left\{Byte\right\}\\
-    &Boolean = \left\{Boolean \right\}
-\end{flalign}
-
-We would like to note that \emph{Int} is not a subtype of \emph{Float}, as by the definition of floating-point numbers in Java \footnote{See IEEE-754 for reference}, \emph{Float} only has 23 bits for the mantissa, meaning it cannot represent all possible values of a 32 bit \emph{Int} accurately and therefore \emph{Int} is not fully set-contained.
-However, this is not the case for \emph{Int} and \emph{Double} as the latter has a 52 bit mantissa.
-This is also why \emph{Long} is not a subtype of \emph{Double}.
-
-Finally, Algorithm \ref{alg:subtyping} defines the subtyping relation for all class types, which concludes the process.
+As a first step, in Equation \ref{eq:initialSubtype} we define the relation for the basic types as these will be the same for all SFJ programs and will not change.
+We also define the \emph{Universe} type that holds the set of all types.
+
+\begin{equation}
+    \label{eq:initialSubtype}
+    \begin{array}{llllll}
+         & Double   & = \left\{Double, Float, Int, Short, Byte\right\} &  & Float & = \left\{Float, Short, Byte\right\} \\
+         & Long     & =    \left\{Long, Int, Short, Byte\right\}       &  & Int   & = \left\{Int, Short, Byte\right\}   \\
+         & Short    & =    \left\{Short, Byte\right\}                  &  & Byte  & = \left\{Byte\right\}               \\
+         & Boolean  & =    \left\{Boolean\right\}                      &  & Void  & = \left\{Void\right\}               \\
+        \\
+         & Universe & = \{Double, Float, Long, Int,                                                                     \\
+         &          & \quad Short, Byte, Boolean, Void\}
+    \end{array}
+\end{equation}
+
+It can be seen from the initial definitions that the subtyping relation is reflexive and transitive, but not necessarily symmetric.
+
+We would also like to note that $Int$ is not a subtype of $Float$, as in the definition of  floating-point numbers that is used in Java \footnote{See IEEE-754 for reference}, $Float$ only has 23 bits for the mantissa, meaning it cannot represent all possible values of a 32 bit $Int$ accurately and therefore $Int$ is not fully set-contained.
+However, this is not the case for $Int$ and $Double$ as the latter has a 52 bit mantissa. This is also why $Long$ is not a subtype of $Double$.
+
+\subsubsection{Final Relation}
+
+Finally, Algorithm \ref{alg:subtyping} defines the subtyping relation for all class types.
+The order in which the class types are added to the relation does not matter, as the algorithm guarantees that all subtypes for all types will be found, no matter the order.
+We give a description of Algorithm \ref{alg:subtyping} to aid comprehension of the algorithm.
+
+We iterate over the keys of the relation mapping, further called the existing type.
+However we can skip all those that are a basic type as no class type will have or be a subtype of a basic type.
+We first check if the the class we are currently processing, further called the new type, has types in its fields or methods that are currently not in the subtyping relation.
+If so, we stop and add the new type to the list of unprocessed classes, as we are not be able to correctly reason if it is a subtype due to the missing type.
+This list of unprocessed classes will get recursively called to be processed again, and this process will is guaranteed to terminate because we have already checked using Algorithm \ref{alg:types} that all types are valid.
+
+We check if the new type is a subtype of the existing type. The fields and methods of the new type must be subsets of the existing type.
+We also check the structural subtyping rules for field, method return and method argument types as given in Section \ref{sec:struct}.
+If these all hold, then the new type is a subtype of the existing type and is added to its relation mapping.
+
+We then repeat the above process but instead check if the existing type is a subtype of the new type.
+However we do not have to check if the call to $checkSuperSet(existingClass, class)$ return $False$ as we are guaranteed that since the existing type is already in the subtyping relation, all of its field and methods contain types that are also in the subtyping relation.
+We then add the new type to its own relation mapping for reflexivity and to the \emph{Universe} type.
 
 \begin{algorithm}
+    \caption{Recursive algorithm for creating a semantic subtyping relation by calling the function \textit{generateRelation} with the set of class types that are finite trees.
+        Given that we know all classes are valid types, we are guaranteed it will terminate}
+    \label{alg:subtyping}
+
     \DontPrintSemicolon
     \KwData{
-        $classes$, \;
-
-        $relation$, a map of types to the set of its subtypes.\ref{fig:initialSubtype}
-    }
-
+        $classes$, the set of classes in the program for which we have not yet defined a subtyping relation. \;
+        $relation$, a map of types to the set of its subtypes.
+        Initially the map only contains the entries defined in Equation \ref{eq:initialSubtype}. \;
+    }
     \Begin{
         \Fn{generateRelation(classes: List<Class>)}{
-            $untyped: List<Class>\ \longleftarrow []$
+            $unprocessed: List<Class>\ \longleftarrow []$
 
@@
                 \If{addClass(class) = $False$}{
-                    $untyped.add(class)$
+                    $unprocessed.add(class)$
                 }
@@
             \If{$untyped \neq []$}{
-                $generateRelation(untyped)$
+                $generateRelation(unprocessed)$
             }
@@
             $relation[class].add(class)$ \;
+            $relation[Universe].add(class)$ \;
             \KwRet $True$ \;
@@
     }
-
-\caption{Recursive algorithm for creating a semantic subtyping relation, given a set of class types that are finite trees and the initial relation for basic types.
-Given that we know all classes are valid types, we are guaranteed it will terminate}
-\label{alg:subtyping}
 \end{algorithm}
 
+The subtyping algorithm finds all nominal and structural subtypes as it examines all pairs of types.
+It finds all nominal subtypes as child types inherit all fields and methods of their parent class, so when considering the parent-child pair, the child is always guaranteed to be a superset.
+Similarly, for all other pairs of types, the algorithm checks if one is a superset of the other, and if so, they are related by structural subtyping.
+For example, in the extreme situation, the type $empty = []$, will have all classes as structural subtypes as any other class can be safely structurally substituted for it.
+Therefore the same algorithm finds both nominal and structural subtypes without needing to examine the nominal subtyping hierarchy as field and method inheritance guarantees they will be found through structural subtyping.
+
+\subsection{Typing Values and Closing the Circularity}
+\label{numtypes}
+
+Following the framework in Section \ref{sec:framework} to define a sematic type system, we next need typing rules for values, i.e.
+we should be able to type any expression that is able to be constructed in the syntax of terms, given in Equation\ref{eq:syntaxterms}.
+
+\begin{equation}
+    \mbox{\textit{Expressions}}        \qquad e  \; ::=\;   x\ |\  c\ |\ e.a\ |\ e.m(e) \ | \ \synnew C(\wt{e})
+\end{equation}
+
+For the expressions $e.a$ and $e.m(e)$, we can type these values simply by looking at the type of the field or return type of the method.
+For object constructors $new\ C(\wt{e})$, the type of this value will be the class type $C$.
+For variables $x$, we can look at what the type is of the method argument, as this is the only possible situation a variable is allowed to be used.
+
+For the constants $c$, these are the literal values in our language such as $42$, $3.14$ or $True$.
+It is not as simple as it may appear to type these correctly so that boolean type operators in our language behave correctly.
+
+In most languages such as Java, literal values are cast to the type that is expected in the context they are used.
+For example, a method $\textbf{long}\ getNum()\{return\ 42\}$ automatically casts the value to the type $long$.
+In other situation such as variable type inference in Java $\textbf{var}\ x = 42$, the compiler treats the value as an $int$ too as this is the most efficient number type in hardware operations.
+
+However, in our system, types represent sets of values which must interpret boolean connectives in a set-theoretic way.
+Given the type with boolean connectives $(int\ \textbf{and not}\ byte)$, this type represents the values $-2^31$ to $2^31 - 1$ from which are removed the values $-128$ to $127$.
+If we were to type the value $42$ as an integer, this would be accepted by the boolean type, but would not be in the set of values of the type, thus breaking our semantic type system.
+
+Therefore, for all constants $c$, we use a strong typing to assign values the most restrictive types possible, i.e. the type with the smallest amount of values which still contains this value.
+
+With this typing rule, the value $42$ is assigned the type $Byte$, so would correctly not be accepted as a value of the type $(int\ \textbf{and not}\ byte)$.
+
+We have now defined how to type any expression that can be formed by the syntax and so have defined our typing rules.
+The relation defined by the typing rules and the abstract model defined in Section \ref{sec:sub} coincide, so we have closed the circularity and defined our semantic type system.
+
+\section{Analysis of the Subtyping Relation}
+
+\subsection{Complexity and Efficiency of the Subtyping Algorithm}
+
+The complexity of Algorithm \ref{alg:subtyping} is $\mathcal{O}(n)$, however the complexity of Algorithm \ref{alg:subtyping} is $\mathcal{O}(n^{2})$.
+For the small example programs that SFJ has currently compiled for this has not been an issue, however if the language was extended with a standard library with several hundred if not thousand of classes, then each additional class will incur larger penalties in compiler performance.
+However, the real-life impact of such a situation has not yet been explored.
+
+Due to the fact that our subtyping relation can possibly be symmetric because of our inclusion of structural subtyping, this $n^{2}$ behaviour is needed to ensure we fully find all subtypes for each type.
+In a system with solely nominal subtyping, we only need to traverse the subtype hierarchy to find all subtypes, which is a linear operation.
+
+\subsection{Flaws of Structural Subtyping}
+\label{sec:flaws}
+
+Given the observed flaws in complexity in Algorithm \ref{alg:subtyping} due to structural subtyping, we would like to point out another flaw the reader might see in the structural approach but we argue for why we disagree with this flaw.
+
+Consider two structurally equivalent class types $coordinate = [x:int, y:int, z:int]$ and $colour = [x:int, y:int, z:int]$.
+They are structural subtypes so they can be used interchangeably in our type system, but their \emph{meaning} is completely different.
+Therefore, we lose some safety in our language as it is unlikely we would like to use a colour where we expect a coordinate, but in the eyes of the type-checker, it is correct.
+
+\citet{Dardha2017} suggest a possible extension to the language by including a $nominal$ keyword that can be used in the class signature to indicate that this class should only be considered by the nominal component of the type system.
+However, being optional, we argue that this would create more bugs than problems it would solve due to forgetting the keyword.
+A similar issue occurs for example with destructors in C++ classes. Leaving out the virtual keyword of the parent class destructor in certain situations can lead to memory leaks.
+We therefore argue instead that an additional \emph{structural} keyword could be introduced as an extension to the language which specifies that this class can have structural subtypes.
+However, a class with this keyword is not added to the subtyping relation of other structurally similar classes which are not marked \emph{structural} as this would weaken the invariant conditions of this class.
+
+In languages with solely nominal subtyping such as Java, one can define an overridden method to perform the opposite logic to what the super class is expecting, such as:
+
+\begin{equation}
+    \begin{array}{ll}
+        \syndecl{A}{\mathit{Object}} {
+         &
+            \qquad
+            \syndecl{B}{A} {
+                \\
+                \qquad\quad \ldots \qquad
+         &
+                \qquad\qquad\quad \ldots
+                \\
+                \qquad\quad \textbf{int}\ length\qquad
+         &
+                \qquad\qquad\quad \textbf{int}\ getLength()\{\ \synret{-length}\ \}
+                \\
+                \qquad\quad \textbf{int}\ getLength()\{\ \synret{length}\ \}\qquad
+         &
+                \qquad}
+            \\   }
+    \end{array}
+\end{equation}
+
+Therefore, even languages without structural subtyping leave an expectation on the developer to check what they are doing is correct.
+Hence we argue that the integration of both in SFJ is a valid idea as either subtyping system is no more or no less than the other, so having both available gives the user the choice to use either to where they are best suited, which should lead to more correctly defined abstractions in the program, which should lead to less bugs in the program.
+
+For example, when structural subtyping is used correctly, it makes certain recurring challenging programming tasks effortless.
+Take the task of dependency injection in testing for mocking class behaviour.
+In languages like C++, we may wish to not make the methods in our classes virtual (making them not overridable) to perhaps not incur the performance overhead of virtual methods.
+In these situations, mocking such a class is near impossible without changing the type hierarchy or class definition.
+With structural subtyping, we simply define a class with the same fields and method signatures, and can provide our own implementations in the methods.
+Because the type will be structurally similar, it is effortless to use our new class as an argument where the old class is expected.
+
+\chapter{Exploiting Semantic Subtyping}
+
+In this chapter we demonstrate that having defined the mathematical machinery behind the semantic type system, it is easy to add features into the language which exploit the set-theoretic subtyping system and make writing SFJ programs more concise and intuitive.
+This introduces the second core novel addition in SFJ of boolean types, along with method types and multi-methods.
+
+\section{Implementation of Boolean Types}
+\label{sec:bool}
+
+Now that we have define our semantic type system, it is easy to implement boolean types in our language.
+
+\subsection{Illustrative example of Boolean Types}
+
+We demonstrate the implementation of boolean types by using our main motivational example given by \citet{Dardha2013} which illustrates the benefits of boolean types.
+
+\subsubsection{The Polygons}
+
+Our example considers a set of polygons, such as triangles, squares and rhombuses.
+We want to define a method \emph{diagonal} that takes a polygon and returns the length of its longest diagonal, however such a method only is defined for polygons that have at least four sides
+Therefore we need to find a way of implementing the type hierarchy in such a way that triangles are excluded.
+In Java this could be done in the following way:
+
+\begin{equation}
+    \begin{array}{l}
+        \synclass Polygon \ \{ \ldots\}
+        \\
+        \syndecl{Triangle}{Polygon} {\ldots}
+        \\
+        \syndecl{Other\_Polygons}{Polygon} {
+            \\
+            \qquad \ldots
+            \\
+            \qquad \textbf{double}\ diagonal(Other\_Polygons \ shape)\ \{\ldots \}
+            \\
+        }
+        \\
+        \syndecl{Square}{Other\_Polygons}{\ldots}
+        \\
+        \syndecl{Rhombus}{Other\_Polygons} {\ldots}
+    \end{array}
+\end{equation}
+
+Or by means of an interface \emph{Diagonal}:
+
+\begin{equation}
+    \label{eq:interface}
+    \begin{array}{l}
+        public \ \syninterface Diagonal\ \{
+        \\
+        \qquad \textbf{double}\ diagonal(Polygon \ shape);
+        \\
+        \}
+        \\
+        \synclass Polygon \ \{\ldots\}
+        \\
+        \syndecl{Triangle}{Polygon} {\ldots}
+        \\
+        \synidecl{Square}{Polygon} \synimpl Diagonal \ \{\ldots\}
+        \\
+        \synidecl{Rhombus}{Polygon} \synimpl Diagonal \ \{\ldots\}
+        \\
+        \hspace{4.5cm} \ldots
+    \end{array}
+\end{equation}
+
+However, the first is error-prone and could be said to be not self-documenting \citep{schach2007}, i.e. the type hierarchy does not explain why this extra class $Other_Polygons$ has been created.
+The second solution is more informative, however given many other methods we may want wish implement, it would quickly become cumbersome to have a long list of interfaces for each polygon we add.
+
+If we were not able to write our own abstractions and rather were given a legacy system, the most likely and natural abstraction would be for \emph{Polygon} to be the parent type and all more specific shapes extend this class.
+Now the only way we would be able to define such a method \emph{diagonal} is to use \emph{instanceof} to see the type of the argument $Polygon$ at run-time and throw an error if we encounter a triangle.
+However now we no longer check at compile-time whether the arguments we are passing are correct, leading to possible bugs at run-time.
+
+In SFJ, we instead propose a more elegant solution with the use of boolean types where the check for excluding triangles happens at compile-time.
+We can simply define a method \emph{diagonal} which takes an argument of type \emph{Polygon} \textbf{and not} \emph{Triangle}.
+
+\begin{equation}
+    \label{eq:polygons}
+    \begin{array}{l}
+        \synclass Polygon \ \{\ldots\}
+        \\
+        \syndecl{Triangle}{Polygon} {\ldots}
+        \\
+        \syndecl{Square}{Polygon} {\ldots}
+        \\
+        \syndecl{Rhombus}{Polygon} {\ldots}
+        \\
+        \synclass Diagonal \ \{
+        \\
+        \qquad \ldots \\
+        \qquad \textbf{double}\ diagonal((Polygon\ \textbf{and not}\ Triangle) \ shape)\{\ldots\}
+        \\
+        \}
+    \end{array}
+\end{equation}
+
+\subsection{Type-Checking Boolean Types}
+\label{typecheck}
+
+\begin{equation}
+    \label{eq:initialSubtype}
+    \begin{array}{llllll}
+         & Double   & = \left\{Double, Float, Int, Short, Byte\right\}     &  & Float    & = \left\{Float, Short, Byte\right\} \\
+         & Long     & =    \left\{Long, Int, Short, Byte\right\}           &  & Int      & = \left\{Int, Short, Byte\right\}   \\
+         & Short    & =    \left\{Short, Byte\right\}                      &  & Byte     & = \left\{Byte\right\}               \\
+         & Boolean  & =    \left\{Boolean\right\}                          &  & Void     & = \left\{Void\right\}               \\
+         & Polygon  & = \left\{Polygon, Triangle, Square, Rhombus \right\} &  & Triangle & = \left\{Triangle\right\}           \\
+         & Square   & = \left\{Square\right\}                              &  & Rhombus  & = \left\{Rhombus\right\}            \\
+         & Diagonal & = \left\{Diagonal\right\}
+        \\
+        \\
+         & Universe & = \{Double, Float, Long, Int, Short, Byte                                                                \\
+         &          & \quad Boolean, Void, Polygon, Square                                                                     \\
+         &          & \quad Square, Rhombus, Diagonal\}
+    \end{array}
+\end{equation}
+
+We first generate the subtyping relation for our solution in Equation \ref{eq:polygons} to give the relation in Equation \ref{eq:initialSubtype}.
+
+Recall the method \emph{diagonal} in class \emph{Diagonal}, we can see that the result of the set operation on its parameter type gives the following set of polygons:
+
+\begin{equation}
+    \label{eq:polygonsandnottriangle}
+    Polygon\ \textbf{and not}\ Triangle = \left\{Polygon, Square, Rhombus \right\}
+\end{equation}
+
+The \emph{\textbf{not} Triangle} operation is achieved by $Universe \setminus Triangle$.
+Therefore if we type-check the following SFJ expression:
+
+\begin{equation}
+    (\textbf{new}\ Diagonal()).diagonal(\textbf{new}\ Square())
+\end{equation}
+
+the argument $new\ Square()$ is typed using our typing rules given in Section \ref{numtypes} to be of type $Square$, and the subtypes of Square is $\{ Square \}$ which is fully contained in the set of the parameter type $\left\{Polygon, Square, Rhombus \right\}$, so this expression successfully type-checks.
+
+On the other hand, given the expression:
+
+\begin{equation}
+    (\textbf{new}\ Diagonal()).diagonal(\textbf{new}\ Triangle())
+\end{equation}
+
+the subtypes of the argument is $\{ Triangle \}$ which is not contained in the set of subtypes for the parameter so therefore this does not type check.
+
+We can also demonstrate the example from Section \ref{numtypes}. Recall the type $(int\ \textbf{and not}\ byte)$. The result of the set operations would be:
+
+\begin{equation}
+    \label{eq:intnotbyte}
+    int\ \textbf{and not}\ byte = \left\{Int,\ Byte\right\}
+\end{equation}
+
+Then given the classes:
+
+\begin{equation}
+    \begin{array}{l}
+        \synclass{Integer} \ \{
+        \\
+        \qquad \ldots
+        \\
+        \qquad \textbf{int}\ num
+        \\
+        \qquad \textbf{int}\ getNum()\{return\ this.num\}
+        \\
+        \}
+        \\
+        \synclass{Calculator}\ \{
+        \\
+        \qquad \ldots \\
+        \qquad \textbf{double}\ acceptLargeNum((int \textbf{ and not } byte) \ n)\{\ldots\}
+        \\
+        \}
+    \end{array}
+\end{equation}
+
+with the following expression to evaluate:
+
+\begin{equation}
+    (\textbf{new}\ Calculator()).acceptLargeNum((\textbf{new}\ Integer(2)).getNum())
+\end{equation}
+
+we can show that this expression does not type-check.
+The value $2$ is given the type $Byte$, but is valid for the field \emph{num} as the subtypes of $Byte$, namely just itself, is fully contained in the set of subtypes of $Int$.
+However, the argument to the function $acceptLargeNum$ is of type $Int$, which has subtypes $\{Int, Short, Byte\}$.
+We can see that this is not fully set-contained in the set of subtypes of Equation \ref{eq:intnotbyte}, so the expression does not type-check.
+
+\subsection{How Boolean Types are Represented in SFJ}
+
+We next show how we ensure that compound boolean statements are evaluated consistently and how we represent the boolean types in our implementation.
+
+If we refer to the SFJ grammar in Appendix \ref{lst:sfjgrammar}, we can see in the \emph{type} production rule we enforce that every single boolean connective is surrounded by brackets to enforce consistency in the definitions.
+Take the boolean type $type_{1} \textbf{ and } type_{2} \textbf{ or } type_{3}$.
+Without brackets, we do not know how to consistently evaluate this type as they could mean different things:
+
+\begin{equation}
+    \begin{array}{llll}
+        type_{1} \textbf{ and } (type_{2} \textbf{ or } type_{3}) & =  & (type_{1} \textbf{ and } type_{2}) \textbf{ or } (type_{1}\ \textbf{and} \ type_{3})
+        \\
+                                                                  & or &
+        \\
+        (type_{1} \textbf{ and } type_{2}) \textbf{ or } type_{3} & =  & (type_{1} \textbf{ or } type_{2}) \textbf{ and } (type_{1}\ \textbf{or} \ type_{3})
+    \end{array}
+\end{equation}
+
+Therefore such a type is not allowed and the user has to specify which of the two alternatives they mean to use.
+
+Secondly, once we have a boolean type which is consistent, we can evaluate the final result by storing it as a tree, such as shown in Figure \ref{fig:boolfig}.
+Starting from the root node, we check if it is a type or a boolean connective.
+If it's a type, we can look up the type in our subtyping relation and return the set of of its subtypes.
+However, if its a boolean connective, we know it will have two child nodes and we can inspect them  and apply the same process until they both return us a set of subtypes.
+We can then get the result of the boolean connective node by applying the corresponding set operation to the two sets of subtypes it received.
+
+\begin{figure}
+    \label{fig:boolfig}
+    \centering
+    \includegraphics[width=0.7\linewidth]{images/diag.pdf}
+    \caption{Representation of how the types $int$, $int \ \textbf{and not}\ byte$ and $type_{1}\ \textbf{and}\ (type_{2}\ \textbf{or}\ type_{3})$ are stored.}
+\end{figure}
+
+\section{Unusable Boolean Types}
+\label{unusable}
+
+Similar to structural subtyping, we would like to highlight a potential issue the reader might notice with the use of boolean types.
+
+The use of boolean types can restrict the set of subtypes we accept, such as with the example of $Polygon \textbf{ and not } Triangle$ in Equation \ref{eq:polygonsandnottriangle}.
+However, a problem arises in that we can widen the set of subtypes that we accept such that there is no guarantee about how the object can be used at run-time.
+
+Take the subtype relation from the Polygons example, Equation \ref{eq:initialSubtype}.
+If we create a boolean type $\textbf{not}\ Byte$, this corresponds to the set of accepted subtypes $\{Short, Int, Long, Float, Double, Boolean, \allowbreak Void, Polygon, Square, Rhombus, Triangle\}$.
+An argument of such a type accepts many values, but what can such an argument actually be used for?
+
+It is not guaranteed that it is a numerical value that can be used in a mathematical expression, nor is it guaranteed to be a polygon object whose field or method we can access.
+If we were to try and access a field, then at runtime this would sometimes be correct but in other times this would cause an error.
+So in which ways should such a type be allowed to be used in expressions?
+
+We follow the opinion we expressed in Section \ref{sec:flaws} and leave an expectation on the developer to use boolean types in such a way that they are useful.
+We implement the simple check that in use of a value that has a boolean type, then at least one of the subtype alternatives is valid, e.g.
+the field we are trying to access is defined for at least one subtype.
+However otherwise leave the developer to check that the type is used correctly.
+
+An alternative would be strict checking that the use of a value is defined for all of its subtypes.
+The \emph{diagonal} method from the Polygons example \ref{eq:polygons} would still be an accepted type, however almost any use of the $\textbf{not}\ Byte$ type would result in a compiler error.
+
+However we feel this would go against our aims of creating a language where the programmer is given the choice to use a more expressive type system.
+Additionally, a static type checking tool could be developed for SFJ which warns the user if a value is used in a way which is not defined for all of its subtypes, giving the user greater confidence in using boolean types.
+
+\section{Method Types}
+
+While our more elegant solution to the Polygons problem \ref{eq:polygons} is an improvement in terms of compactness and clarity, it is not perfect.
+One oversight of this solution is that we define a single method for all accepted alternatives, however the implementation of calculating the longest diagonal will be different for each of them.
+To solve this using the language constructs available to us would again involve us having to use \emph{instanceof} to check the type at run-time and executing the correct implementation for each type.
+Alternatively we could also have the following implementation:
+
+\begin{equation}
+    \begin{array}{l}
+        \synclass Polygon \ \{
+        \\
+        \qquad \ldots
+        \\ \qquad \textbf{double}\ diagonal()\{\ldots\} \\ \}
+        \\
+        \syndecl{Triangle}{Polygon} {\ldots}
+        \\
+        \syndecl{Square}{Polygon} {
+            \\
+            \qquad \ldots
+
+            \\ \qquad \textbf{double}\ diagonal()\{\ldots\} \\ }
+        \\
+        \syndecl{Rhombus}{Polygon} {
+            \\
+            \qquad \ldots
+
+            \\ \qquad \textbf{double}\ diagonal()\{\ldots\} \\ }
+        \\
+        \\
+        \synclass Diagonal \ \{
+        \\
+        \qquad \ldots \\
+        \qquad \textbf{double}\ diagonal((Polygon\ \textbf{and not}\ Triangle) \ shape)\{\synret\ shape.diagonal()\}
+        \\
+        \}
+    \end{array}
+\end{equation}
+
+However what if the class \emph{Polygon} also did not implement the method diagonal, seeing how it is a generic class and we do not how we would be able to calculate a diagonal for it.
+The type declaration that is currently used in the method \emph{diagonal} would no longer be valid and we instead would have to specify each alternative manually $(Square\ \textbf{or}\ Rhombus\ \textbf{or}\ \ldots)$ which is tedious and error-prone.
+
+This is where the arrow type (method type) from our syntax of types \ref{eq:syntaxtypes} can be used. We can instead define the class \emph{Diagonal} as follows:
+
+\begin{equation}
+    \begin{array}{l}
+        \synclass Diagonal \ \{
+        \\
+        \qquad \ldots \\
+        \qquad \textbf{double}\ diagonal((diagonal: \textbf{Void} \rightarrow \textbf{Double})\ shape)\{\synret\ shape.diagonal()\}
+        \\
+        \}
+    \end{array}
+\end{equation}
+
+We define the type as accepting any type and its subtypes which implement a method called \emph{diagonal} which is a method from the type \emph{Void} to a \emph{Double}.
+We also finally see that the type \emph{Void} is used in SFJ to indicate a method that has no arguments, as due to being a functional language, we do not have any fields or method return types which can be of type \emph{Void}.
+
+In order to type-check the value passed to such an argument, we can at compile-time build a collection of types $\{type_{1},\ type_{2},\ \ldots\}$ which have this method.
+We can do this the simply iterating over the same list of classes in the program as when we were creating our subtyping relation and checking for the presence of the required method.
+The set of accepted types would then be the union of all their corresponding subtypes $([\![type_{1}]\!]_{B} \cup [\![type_{2}]\!]_{B} \cup\ \ldots)$.
+
+However, calculating this collection of types for each method of every class would be computationally unnecessary as only few of these would be used for method types.
+Therefore we only compute on demand during type-checking them when we come across such a type.
+This makes compilation times exceedingly quicker the more classes are in our program, especially if we cache the method types we have already encountered, as it is likely to come up again if it has already been used once.
+However, this caching behaviour has currently not been implemented in SFJ as compilation speed has so far been orthogonal to the aims of this project.
+
+We can therefore use method types to statically include or exclude a portion of our type hierarchy.
+However unlike the use of interfaces such as in the suggested Java solution \ref{eq:interface} to the Polygons problem, the values that can be accepted by a method type do not have to be related to each other in any way in the class hierarchy.
+This becomes increasingly useful if we have a legacy system as we can still accept all the classes that have defined \emph{Diagonal} methods without having to go back and add interface implementations.
+
+\section{Multi-methods}
+
+One key feature that we restore in SFJ that we lost from Java is that of overloaded methods, however we implement these as multi-methods.
+Methods within a class must still have unique names, however when using nominal subtyping, we can extend the inherited method.
+
+For example, looking back at Equation \ref{multi1}, the method \emph{length} has type $\mathbf{string} \rightarrow \mathbf{int}$ in $A$.
+However, in $B$ it has type $(\textbf{string} \rightarrow \textbf{int}) \synwedge (\textbf{int} \rightarrow \textbf{int})$,
+which can be simplified as $(\textbf{string} \synvee \textbf{int}) \rightarrow \textbf{int}$.
+
+However, what happens if the extended input types are not disjoint, unlike the example in Equation \ref{multi1}:
+
+\begin{equation}
+    \label{multi}
+    \begin{array}{ll}
+        \syndecl{A}{\mathit{Object}} {
+         &
+            \qquad
+            \syndecl{B}{A} {
+                \\
+                \qquad\quad \ldots \qquad
+         &
+                \qquad\qquad\quad \ldots \
+                \\
+                \qquad\quad \textbf{float}\ \mathit{length}\ (\textbf{float}\ n)\{\ \ldots\ \}\qquad
+         &
+                \qquad\qquad\quad \textbf{int}\ \mathit{length}\ (\textbf{int}\ n)\{\ \ldots\ \}
+                \\
+            }
+         &
+            \qquad }
+    \end{array}
+\end{equation}
+
+Considering the example in Equation \ref{multi}, if we called method \emph{length} on an object of class \emph{B} with the value $2$, should it perform the expression in its own method or the method of $A$ considering they both accept it as an argument.
+No matter which one we choose, the return type of the value is still going to be the same $\textbf{float}\ \synvee\ \textbf{int}$ which simplifies to $\textbf{float}$, however the expressions in the methods could be entirely different.
+
+We decided to follow common object-oriented conventions and we perform the expression of the multi-method that is closest to the class of the object we are calling it on, i.e.
+in the case of our example, the method \emph{length} in $B$ would be called.
+
+We do this storing the type of \emph{length} in $B$ without regard of multi-methods, as we can always generate the return type by walking up the tree of parent types, but it allows us to decide which implementation to choose.
+
+\section{Additional Features}
+
+One of the features that we implemented to increase the usability of SFJ as a language is removing the need for forward declarations of classes.
+This means that classes can be can be used before they are defined later on in the file of the program, such as in Java.
+
+For example for our polygons example \ref{eq:polygons} this is not necessarily obvious, but it could have been also defined as follows:
+
+\begin{equation}
+    \begin{array}{l}
+        \synclass Diagonal \ \{
+        \\
+        \qquad \ldots \\
+        \qquad \textbf{double}\ diagonal((\textit{Polygon} \textbf{ and not }  \textit{Triangle}) \ shape)\{\ldots\}
+        \\
+        \}
+        \\
+        \synclass Polygon \ \{\ldots\}
+        \\
+        \syndecl{Triangle}{Polygon} {\ldots}
+        \\
+        \vdots
+    \end{array}
+\end{equation}
+
+In a language like C++, this would have required defining the structure of all the classes in a header file, however by incorporating it into the language, we make the language easier to use.
+
+Implementing this requires that we do multiple passes over the AST of our program, which will cause an impact on compilation speed.
+On our first pass of the AST we gather all the signatures of the classes and their field and methods and store it in a program tree, which is a condensed version of the AST with all unneeded information removed.
+At this point we have all the information needed to create the subtyping relation by running Algorithm \ref{alg:subtyping} and Algorithm\ref{alg:subtyping} on the all the class nodes at the top of our program tree.
+
+We can now do a second pass over our program, this time on our program tree.
+During this pass we check what we were unable on our first pass such as checking that the arguments given to the \emph{super} call inside class constructors type-check with respect to the subtyping relation.
+Now that we have our typing rules and subtyping relation, we can now also make sure that the expressions in methods type-check according to their expected return type.
+
+\chapter{Code Generation}
+
+In this chapter, we define for code generation for a SFJ program as just a type system which we can use to type-check programs is without much use if we cannot run these programs on a computer.
+
+\section{Implementation}
+
+Due to extenuating circumstances around the COVID-19 pandemic, the code generation implementation of SFJ was not able to be completed in this project, however we still provide the theoretical implementation for the use of any future work.
+
+Given the similarity of SFJ to Java, we felt that the usage of the Java Virtual Machine (JVM) \footnote{See https://docs.oracle.com/javase/specs/jvms/se7/html/ for reference} would be most suitable for code generation by transpiling SFJ programs into Java bytecode.
+This is the approach used by object-oriented languages such as Kotlin \footnote{https://kotlinlang.org/} and many others.
+
+The key challenge in implementing code generation compared to other languages which transpile their programs into bytecode is that for example, a single field in SFJ with type $int\ or\ bool$ actually represents two Java fields, one of type $int$ and another of type $bool$, of which only one is inhabited with a value.
+
+Therefore to reduce the amount of alternatives for which we would need to generate code, we first analyse our program and reduce the boolean types by keeping only the alternatives which actually get used in the program.
+For example, if the field of type $int\ or\ bool$ only ever gets initialised with a boolean value, we can reduce it and make it a field of a single type $bool$.
+Similarly the numerical basic types can also be reduced to their largest type without consequence, such as $int\ \textbf{and not}\ byte$ can be reduced to just $int$ as we have already type checked all literal values passed to such a type.
+
+After reducing the boolean types in our program, we have to expand all our still existing boolean types and consider each alternative in the result of the set operation.
+Staying with the example of a field named $f1$ of type $int\ or\ bool$, we would define in our class two fields $int\_f1$ and $bool\_f1$ with types $int$ and $boolean$ respectively.
+In order to initialise these field, we use the constructor overloading capabilities of Java to generate an overloaded version for all combinations of constructor parameters.
+In each constructor, each expanded parameter alternative that is not used is initialised to $null$ and only the expanded parameter that matches the type of the parameter in this constructor is initialised with a value.
+
+For methods, we also use the overloading capabilities of Java to define a method for each type in the expanded method parameter, all with the same method body.
+Depending on which alternative the argument inhabits at run-time, a different method will be dispatched to.
+Then depending on if the actual argument is valid for the method body, the method will return a value or we will encounter an exception from the JVM.
+
+For the access of any field of an object, we generate code that checks for each alternative of the field if it is non-null and if it is non-null, then we perform the generated code for the remaining expression.
+However at runtime, only one branch of will be true, and this is the branch of generated code which will be executed.
+
+For structural subtypes, we create an actual nominal subtype with the same implementation for each structural subtype a type has, so that it is accepted as a proper subtype in Java.
+
+Like methods with boolean types, we implement methods with method types the same way by defining an overloaded alternative for each type that implements the method of the method type.
+
+We have now defined the code generation for each expression in our syntax of terms and for field, constructor and method declarations, so therefore have defined how to generate the bytecode for any SFJ program.
+
+\section{Problems with Code Generation}
+
+One of the problems that can easily be seen is that when the boolean types are used incorrectly such as the example $\textbf{not}\ Byte$ from Section \ref{unusable}, then we will generate lots of bytecode as we have to consider each alternative.
+While this should theoretically not affect the time for the bytecode to run on the JVM, the time for the compiler to generate bytecode for each alternative could take a significant amount of time, especially when more and more classes are added to the program.
+However, it would have to be evaluated if compilation times would be affected so much that compilation is infeasible for large programs.
+Yet, considering that there exist project with multi-million lines of Java code \footnote{https://www.visualcapitalist.com/millions-lines-of-code/, accessed 31/03/2020} which compile in an acceptable amount of time, we do not think this should be a problem for most programs.
+
 %===================================================================================================
 
-\chapter{Implementation}
-
-What did you do to implement this idea, and what technical achievements did you make?
-
-\section{Guidance}
-You can't talk about everything. Cover the high level first, then cover important, relevant or impressive details.
-
-\section{General points}
-
-These points apply to the whole dissertation, not just this chapter.
-
-\subsection{Figures}
-
-\emph{Always} refer to figures included, like Figure \ref{fig:relu}, in the body of the text. Include full, explanatory captions and make sure the figures look good on the page.
-You may include multiple figures in one float, as in Figure, using \texttt{subcaption}, which is enabled in the template.
-
-\subsection{Equations}
-
-Equations should be typeset correctly and precisely. Make sure you get parenthesis sizing correct, and punctuate equations correctly
-(the comma is important and goes \textit{inside} the equation block). Explain any symbols used clearly if not defined earlier.
-
-For example, we might define:
-\begin{equation}
-    \hat{f}(\xi) = \frac{1}{2}\left[ \int_{-\infty}^{\infty} f(x) e^{2\pi i x \xi} \right],
-\end{equation}
-where $\hat{f}(\xi)$ is the Fourier transform of the time domain signal $f(x)$.
-
-\subsection{Algorithms}
-
-Algorithms can be set using \texttt{algorithm2e}, as in Algorithm.
-
-\subsection{Tables}
-
-If you need to include tables, like Table , use a tool like https://www.tablesgenerator.com/ to generate the table as it is extremely tedious otherwise.
-
-\subsection{Code}
-
-Avoid putting large blocks of code in the report (more than a page in one block, for example). Use syntax highlighting if possible, as in Listing.
+\chapter{Evaluation}
+
+This chapter describes the evaluation methods for this project.
+The quality of the language as a software product was tested.
+The quality if the language in terms of ease of use of new users was also evaluated.
+Finally, the language was evaluated if it met all the requirements that we set of the type of programming problems it should have made easier.
+
+\section{Testing}
+
+One of the major limitations we found of testing a programming language which was built as non-commercial, prototype product is that code ends up very tightly coupled, which makes unit testing difficult.
+Therefore it was decided that the code was mainly tested using regression testing of a single constantly evolving test program which aimed to cover all possible edge cases of class definitions.
+However, another difficulty which was found is that due to the infinite possibilities of programs, it is much more difficult to find edge cases.
+When your input is simpler such as integers, there are well known edge cases such as zero, one, negative one, and the max values.
+
+The regression test was used to create a code coverage report, pictured in Figure \ref{fig:coverage}.
+The testing achieves a line coverage of 71\%, which is close to the 80\% figure that is aimed for in software project by companies such as Microsoft \footnote{https://docs.microsoft.com/en-us/visualstudio/test/using-code-coverage-to-determine-how-much-code-is-being-tested}
+
+\begin{figure}[H]
+    \label{fig:coverage}
+    \centering
+    \includegraphics[width=0.7\linewidth]{images/coverage.png}
+    \caption{Coverage Report indicating a line coverage of 71\%}
+\end{figure}
+
+\section{User Evaluation}
+
+The second evaluation metric of the language was to evaluate whether users found the semantic type system and boolean types obvious to use, however due to extenuating circumstances this could not be conducted.
+We include our evaluation plan as an area of future work.
+
+\subsection{Evaluation Strategy}
+
+Users are to be given the following scenario:
+
+``
+Our problem considers a set of polygons, such as triangles, squares and rhombuses.
+We want to define a method \emph{diagonal} in a class \emph{Diagonal} that takes a polygon and returns the length of its longest diagonal, however such a method only is defined for polygons that have at least four sides.
+Therefore we need to find a way of implementing this method in such a way that triangles are excluded.
+
+You are required to define at least the following four classes: Triangle, Square and Rhombus and Diagonal.
+The class Diagonal must have a single method \emph{diagonal} that upon being passed an object of type Square or Rhombus, returns the integer two.
+Upon being passed an object of type Triangle, it raises any sort of exception, whether it be at compile-time or run-time.
+The classes Triangle, Square, Rhombus and Diagonal are not required to have any fields.
+You do not have to run the program to complete the task but can do so if wanted.
+''
+
+and are given one minute to read it.
+They are required to write a solution to the scenario in Java on a Windows machine using the program Notepad without any assistance of other people or programs.
+They are timed and monitored for how long it takes them to create a solution.
+
+The same participants are then repeat this but implement the solution in SFJ.
+All are given Section \ref{sec:syntax} as documentation to the language.
+Half of the participants are also given the additional documentation as follows:
+
+``
+The interpretation of types is of sets of values, so we define the subtyping relation by defining a map from a type to the set of its subtypes, with the property that the set of values of a subtype is included in the set of values of the type.
+
+\begin{equation*}
+    \begin{array}{l}
+        Int = \{Int, Short, Byte\}
+        \\
+        Byte = \{Byte\}
+        \\
+        int \textbf{ and not } byte = \left\{Int,\ Short\right\}
+    \end{array}
+\end{equation*}
+''
+
+The participants shall all be students who are enrolled on a single or joint honours Computer Science degree of any year of study.
+The time to solve both problems shall be anonymised, but their year of study and programming language of choice shall be recorded for analysis purposes.
+
+The aim of the experiment is that being given the problem and being allowed to think it through in Java beforehand, then when given the second problem, the participants who are not given the additional documentation perform the same task in the same or longer time, whereas the participants who are given the additional documentation perform the task quicker than their first time.
+It shows that when given an unrelated example of the semantic type system and boolean types, users can apply it to the problem given to them, showing that the language is intuitive to use.
+
+\section{Summary}
+
+One of the main requirements of SFJ was that it made programming more intuitive and therefore reduced the number of lines needed for the same problem.
+We can see from Listings \ref{lst:polygonsjava} and \ref{lst:polygonsfj} that the solution in SFJ took only thirteen lines whilst the equivalent solution in Java took sixteen lines of code.
+While we do not have such a comparison for a program of larger length and complexity, it is already visible that if we restrict Java to an equivalent functional subset as SFJ, then we are able to express all functional Java programs in SFJ, meaning that we can only make our programs more concise.
+
+\begin{lstlisting}[language=Java, caption={Polygons example solution in Java}, label=lst:polygonsjava]
+    interface Diagonal {
+        double diagonal(Polygon shape);
+    }
+    class Polygon {
+        Polygon(){}
+    }
+    class Square extends Polygon implements Diagonal {
+        Square(){}
+        double diagonal(){return 2;}
+    }
+    class Triangle extends Polygon {
+        Triangle(){}
+    }
+    class DiagonalCalculator {
+        double getDiagonal(Diagonal shape){return shape.diagonal();}
+    }
+\end{lstlisting}
+
+\begin{lstlisting}[language=Java, caption={Polygons example solution in SFJ}, label=lst:polygonsfj]
+    class Polygon {
+        Polygon(){}
+    }
+    class Square extends Polygon {
+        Square(){}
+        double diagonal(){return 2;}
+    }
+    class Triangle extends Polygon {
+        Triangle(){}
+    }
+    class DiagonalCalculator {
+        double diagonal((Polygon and not Triangle) shape){return shape.diagonal();}
+    }
+\end{lstlisting}
+
+The language satisfies our minimum requirements of a semantic type system which represents type as sets of values, and complements this type system with boolean types, as shown in Section \ref{sec:bool}.
+
+We can also see how expressive the end language is by seeing if we are able to reintroduce the typical programming constructs that were removed in FJ given by \citet{Dardha2017} Section 8.3.
+
+Our current implementation of multi-methods is able to reintroduce the \emph{instanceof} construct, exactly as defined.
+However for the implementation of the \emph{if-then-else}, we are missing the ability to use singleton values from our types of sets of values as parameter types, such as the $[\textbf{true}]$ type.
+If we did have this construct, we would also be able to add the \emph{try-catch} construct back into our language.
+It is visible however how the mathematical model is still more powerful than our language due to the much greater flexibility it has, but our language is close and has implemented other additional features such as method types.
+
+As a software project, the language could be improved by the introduction of a formal testing suite with targeted unit tests.
+However what would be better is to expand the current regression testing even more.
+Given that the regression tests are testing a compiler, they also automatically show which feature in our language has regressed by the error message of the compiler.
 
@@
 
-\chapter{Evaluation}
-How good is your solution? How well did you solve the general problem, and what evidence do you have to support that?
-
-\section{Guidance}
-
-\begin{itemize}
-    \item
-          Ask specific questions that address the general problem.
-    \item
-          Answer them with precise evidence (graphs, numbers, statistical
-          analysis, qualitative analysis).
-    \item
-          Be fair and be scientific.
-    \item
-          The key thing is to show that you know how to evaluate your work, not
-          that your work is the most amazing product ever.
-\end{itemize}
-
-\section{Evidence}
-Make sure you present your evidence well. Use appropriate visualisations, reporting techniques and statistical analysis, as appropriate.
-
-If you visualise, follow the basic rules, as illustrated in Figure:
-\begin{itemize}
-    \item Label everything correctly (axis, title, units).
-    \item Caption thoroughly.
-    \item Reference in text.
-    \item \textbf{Include appropriate display of uncertainty (e.g. error bars, Box plot)}
-    \item Minimize clutter.
-\end{itemize}
-
-See the file \texttt{guide\_to\_visualising.pdf} for further information and guidance.
+\chapter{Demonstration Paper}
+
+This chapter briefly describes the tool paper about SFJ which was submitted to COORDINATION 2020.
+We include a reference to the paper and demonstration video that were submitted \citep{UD20}.
+
+\section{COORDINATION 2020}
+
+The 22nd International Conference on Coordination Models and Languages (COORDINATION) is a conference which aims to promote research into new models, architectures, languages and verification techniques in order to cope with the complexity induced by the demands of modern software development.
+A paper was submitted as a tool paper to demonstrate our innovative prototype language which through semantic subtyping and boolean types, allows users to define more intuitive types which should lead to fewer logical bugs in programs.
+The paper is titled \emph{SFJ: An implementation of Semantic Featherweight Java} and discusses a subset of this dissertation in less detail.
+As of the time of writing, the acceptance of the paper is still pending.
+The abstract is as follows:
+
+\begin{adjustwidth}{50pt}{50pt}
+    Subtyping is a key notion in programming languages, as it allows more flexibility in coding.
+    There are two approaches to defining subtyping relations: the \emph{syntactic} and the \emph{semantic} approach.
+    In semantic subtyping, one defines a model of the language and an interpretation of types as subsets of this model.
+    Subtyping is defined as inclusion of subsets denoting types.
+    An orthogonal subtyping question, typical of object-oriented languages, is the \emph{nominal} vs. \emph{structural} subtyping.
+    Dardha \emph{et al.} \cite{Dardha2013,Dardha2017} defined boolean types and semantic subtyping for Featherweight Java (FJ) and integrated both structural and nominal subtyping, thus exploiting the benefits of both approaches.
+    However, these benefits were illustrated only at a theoretical level, but not exploited practically.
+
+    In this paper, we present SFJ---Semantic Featherweight Java, an implementation of FJ which features boolean types and structural subtyping as well as nominal subtyping.
+    The benefits of SFJ, illustrated in the paper and the video (with audio/subtitles) \cite{UD20}, show how static typechecking of boolean types and semantic subtyping gives higher guarantees of program correctness, more flexibility and compactness of program writing.
+\end{adjustwidth}
 
@@
 
-Summarise the whole project for a lazy reader who didn't read the rest (e.g. a prize-awarding committee).
-
-\section{Guidance}
-\begin{itemize}
-    \item
-          Summarise briefly and fairly.
-    \item
-          You should be addressing the general problem you introduced in the
-          Introduction.
-    \item
-          Include summary of concrete results (``the new compiler ran 2x
-          faster'')
-    \item
-          Indicate what future work could be done, but remember: \textbf{you
-              won't get credit for things you haven't done}.
-\end{itemize}
+This paper has discussed the reason why type systems exist, and the motivation for why we should research more expressive type systems.
+We incorporate a semantic type system as an extension to Featherweight Java and present this as Semantic Featherweight Java.
+The syntax and implementation of SFJ and its semantic subtyping relation were explained in depth, before we exploited it to easily implement boolean types.
+This chapter provides a full summary of the paper and project and looks at what future work remains to be done on this project and on type systems.
+
+\section{Summary}
+
+In summary, a semantic type system is one where types represent subsets of values in the model of our language.
+The subtyping relation is then defined as inclusion of sets denoting types.
+We extend Featherweight Java with such a type system to present Semantic Featherweight Java.
+SFJ has an extended type algebra with boolean connectives: \emph{and}, \emph{or} and \emph{not} which behave according to their expected set-theoretic interpretation.
+SFJ also gives the programmer the choice to use either nominal or structural subtyping or both.
+We describe the syntax of the language and then give the two main algorithms of the project, the first of which decides whether the class definitions in our program are finite trees with no cycles.
+The second defines the subtyping relation as a map from a type to the set of its subtypes, with the property that the set of values of a subtype is included in the set of values of the type.
+This subtyping relation allows us to type all the expression in our language, therefore fully defining the semantic type system.
+We show how the type system is easily extended with boolean connectives and give examples how this allows us to write more concise and intuitive programs through the \emph{Polygons} example.
+We discuss how both types with boolean connectives and structural subtyping can also be used in ways that guarantee less program correctness but argue that they are no more unsafe than other language features in commonly used languages.
+We show another feature, method types, that allows us to select parts of our class hierarchy which is also easily implemented by having a semantic type system.
+
+Throughout the paper its been clear that our novel type system allows us to more easily choose and prune parts of class hierarchy which we want to accept or reject when we have a clear idea of the system we want to build.
+However it has also been clear when the resulting system we want to build is not as clear, it is possible to decrease rather than increase the correctness of the program, which indicates that more research is required into this topic.
+
+\section{Future Work}
+
+We have mentioned several areas of future research in relation to this project already in the paper.
+The first of these is another extension of the language with a \emph{nominal} or \emph{structural} keyword that allows us finer control over when the nominal and structural type systems apply.
+The second and perhaps largest area of future research is the implementation of code generation for SFJ.
+We provide a theoretical implementation of how it could be implemented but just as our work for the actual implementation of SFJ found unexpected issues to solve, so could this.
+The third area of research is research into the efficiency issues we have mentioned for both the semantic subtype relation generation and code generation and how these perform computationally as we increase the size of the program we are compiling.
+
+Separately areas of interesting future work could be in guaranteeing that boolean types and their use are logically coherent and correct, perhaps by only allowing boolean types which only allow us to restrict the tree of accepted types.
+If such types were to be always coherent, it could be analysed whether these additional guarantees in the correctness of our program allows us to further optimise the generated machine code for these programs.
 
@@
 
-%  APPENDICES
-
 \begin{appendices}
@@
 
-    \begin{lstlisting}[label={lst:sfjgram}, caption={The full ANTLR grammar for SFJ language}]
+    \begin{lstlisting}[label={lst:sfjgrammar}, caption={The full ANTLR grammar for SFJ language}]
         grammar sfj;
@@
 
-    \chapter{TODO}
-
 \end{appendices}

### 2020-04-17 conference/short/background.tex
restructuring for camera ready

\section{Background}
\label{sec:background}
The technical developments behind semantic subtyping and its properties are not easy, however, they are completely transparent to the programmer.
In this section we will illustrate the types and typing rules behind SFJ, however, the technical details behind defining the semantic subtyping are beyond the scope of the paper and we refer the reader to the relevant work \cite{Dardha2013,Dardha2017}.

\paragraph{Types} are defined by following grammar:
$$
\begin{array}{llll}
\tau & ::= & \alpha \ |\ \mu
& \mbox{{\em Type term}}
\\
\alpha & ::=  & \zero \ |\ \btypes \ |\ [\wt{l:\tau}] \ |\ \alpha \synwedge \alpha \ |\ \synneg\alpha
\qquad
& \mbox{{\em Object type} ($\alpha $-type)}
\\
\mu & ::=  & \alpha \to \alpha \ |\ \mu \synwedge \mu \ |\ \synneg\mu
& \mbox{{\em Method type} ($\mu $-type)}
\end{array}
$$
$\alpha$-types are used to type fields and $\mu$-types are used to type methods.
Type $\zero$ is the empty type.
Type $\btypes$ denotes the \emph{basic} types, such as integers, booleans, etc.
We will use $\rho$ range over \emph{record} types $[\wt{l:\tau}]$, where $\wt l$ is a sequence of disjoint labels ranging over an infinite countable set ${\cal L}$.
Given a type $\rho = [\wt{a:\alpha},\wt{m:\mu}]$, $\rho(a_i)$ is the type of the field $a_i$ and $\rho(m_j)$ is the type of the method $m_j$.
The boolean types using $\synwedge$ and $\synneg$ connectives in both $\alpha$ and $\mu$ syntactic category have the expected meaning.
We will use $\one$ to denote the type $\synneg \zero$, corresponding to the universal type. We let $\alpha \syndiff \alpha'$  denote $\alpha \synwedge \synneg \alpha'$ and $\alpha \synvee \alpha'$  denote $\synneg(\synneg \alpha \synwedge \synneg \alpha')$. The same holds for the $\mu$-types.
Arrow types $\alpha \to \alpha$ are used to type methods.

\paragraph{Terms}
are based on FJ \cite{featherweight}.
We assume an infinite countable set of names, with some special names: $\mathit{Object}$ indicates the root class, $\synthis$  indicates the current and $\synsuper$ indicates the parent object.
We let  $A, B, C, \ldots$ range over classes; $a, b, \ldots$ over fields; $m, n, \ldots$ over methods and $x, y, z, \ldots$ range over variables.
Constants $c$ range over an infinite countable set $\mathcal{K}$.
The syntax of terms is given by the following grammar:
\begin{align*}
& \mbox{\textit{Class declaration}} & L \; ::= \; & \syndecl{C}{C}{\wt{\alpha \ a};\ K; \ \wt{M}\ }\\
& \mbox{\textit{Constructor}} & K \;::=\; & C\ (\wt{\alpha\ x})\ \{\ \synsuper(\wt{x});\ \wt{\synthis.a}=\wt{x}; \}\\
& \mbox{\textit{Method declaration}} & M \; ::= \; & \alpha \ m\ (\alpha \ x)\ \{\ \synret e; \} \\
& \mbox{\textit{Expressions}} & e \; ::=\; & x\ |\  c\ |\ e.a\ |\ e.m(e) \ | \ \synnew C(\wt{e})
\end{align*}
A {program} $(\wt{L}, e)$ is a pair of a sequence of class declarations $\wt L$, giving rise to a class hierarchy as specified by the inheritance relation, and an expression $e$ to be evaluated. 
%
A class declaration $L$ specifies the name of the class, the name of the parent class it extends, its typed fields , the constructor $K$ and its method declarations $M$.
The constructor $K$ initializes the fields of the object by assigning values to the fields inherited by the super-class and to the fields declared in the current \textbf{this} class.
%
A method declaration $M$ specifies the signature of the method, namely the return type, the method name and the formal parameter as well as the body of the method.
Notice that in our theoretical development we use unary methods, without loss of generality: tuples of arguments can be modelled by an object that instantiates a ``special''class containing as fields all the needed arguments.
Expressions $e$ include variables, constants, field accesses, method invocations and object creations.

\paragraph{On multimethods}
Since FJ is a core language, some features of the full Java are removed: overloading methods is one of them.
In our framework, leveraging the expressivity of boolean connectives and semantic subtyping, many of these features, and language constructs, e.g., lists, are restored: overloading methods is one of them. We can thus model \emph{multimethods}, \cite{BC97}, which according to the authors is ``\emph{very clean and easy to understand [...] it would be the best solution for a brand new language}".
As an example Dardha \emph{et al.}  \cite{Dardha2013,Dardha2017}, consider the following class declarations:
$$
\begin{array}{ll}
\syndecl{A}{\mathit{Object}} {
&
\qquad
\syndecl{B}{A} {
\\
	%\qquad\quad \ldots
%&
	%\qquad\qquad\quad \ldots
%\\
	\qquad\quad \mathbf{int}\ \mathit{length}\ (\mathbf{string}\ s)\{\ \ldots\ \}\qquad
&
	\qquad\qquad\quad \mathbf{int}\ \mathit{length}\ (\mathbf{int}\ n)\{\ \ldots\ \}
\\
}
&
\qquad }
\end{array}
$$
Method {\it length} has type $\mathbf{string} \rightarrow \mathbf{int}$ in $A$. However, in $B$ it has type $(\mathbf{string} \rightarrow \mathbf{int}) \synwedge (\mathbf{int} \rightarrow \mathbf{int})$,
which can be simplified as $(\mathbf{string} \synvee \mathbf{int}) \rightarrow \mathbf{int}$.

\paragraph{Typing Rules} use the subtyping relation $\leq_{\mathcal B}$ defined semantically as set inclusion of types which are interpreted as subsets of a model, we call it $\mathcal B$ of FJ. Details of model $\mathcal B$ and the subtyping relation $\leq_{\mathcal B}$ are fully presented in \cite{Dardha2013,Dardha2017}.

Typing judgements for expression are of the form $\Gamma \vdash_{\mathcal B} e : \tau$ and for method declarations are $\Gamma \vdash_{\mathcal B} M$. The same for class declarations and programs.

Let us assume a sequence of class declarations $\wt{L}$.
We first determine the (structural) type of every class $C$ in $\wt{L}$.
To this aim, we have to take into account the inheritance relation specified in the class declarations in $\wt{L}$.
We write ``$a \in C$'' to mean that there is a field declaration of name $a$ in class $C$ within the hierarchy $\wt L$.
Similarly, we write ``$a \in C$ with type $\alpha$'' to also specify the declared type $\alpha$. Similar notations also hold for method names $m$.

### 2020-04-17 conference/camera ready/paper.tex
a first pass...

--- 2020-03-24 conference/paper.tex
@@
 \title{SFJ: An implementation of Semantic Featherweight Java
-    %\thanks{Research supported by the UK EPSRC grant EP/K034413/1, ``From Data Types to Session Types: A Basis for Concurrency and Distribution (ABCD)'', and by the EU HORIZON 2020 MSCA RISE project 778233 ``BehAPI: Behavioural Application Program Interfaces''.}
+    \thanks{Research supported by the UK EPSRC grant EP/K034413/1, ``From Data Types to Session Types: A Basis for Concurrency and Distribution (ABCD)'', and by the EU HORIZON 2020 MSCA RISE project 778233 ``BehAPI: Behavioural Application Program Interfaces''.}
 }
@@
 
-\institute{School of Computing Science, University of Glasgow, United Kingdom \\
-    \email{\{artem.usov, ornela.dardha\}@glasgow.ac.uk}}
+\institute{School of Computing Science, University of Glasgow, Glasgow, United Kingdom \\
+\email{2296905U@student.gla.ac.uk, ornela.dardha@glasgow.ac.uk}}
 
@@
 \begin{abstract}    
-    {Subtyping} is a key notion in programming languages, as it allows more flexibility in coding.
     There are two approaches to defining subtyping relations: the \emph{syntactic} and the \emph{semantic} approach.
-    %
     In semantic subtyping, one defines a model of the language and an interpretation of types as subsets of this model.
     Subtyping is defined as inclusion of subsets denoting types.
-    %
-    An orthogonal subtyping question, typical of object-oriented languages, is the \emph{nominal} vs. \emph{structural} subtyping.
-    Dardha \emph{et al.} \cite{Dardha2013,Dardha2017} defined boolean types and semantic subtyping for Featherweight Java (FJ) and integrated both structural and nominal subtyping, thus exploiting the benefits of both approaches.
+
+    An orthogonal subtyping question, typical of object-oriented languages, is the \emph{nominal} and the \emph{structural} subtyping.
+    Dardha \emph{et al.} \cite{Dardha2013,Dardha2017} defined boolean types and semantic subtyping for Featherweight Java (FJ) and integrated both nominal and structural subtyping, thus exploiting the benefits of both approaches.
     However, these benefits were illustrated only at a theoretical level, but not exploited practically.
 
-    In this paper, we present SFJ---Semantic Featherweight Java, an implementation of FJ which features boolean types and structural subtyping as well as nominal subtyping.
-    The benefits of SFJ, illustrated in the paper and the video (with audio/subtitles) \cite{UD20}, show how static typechecking of boolean types and semantic subtyping gives higher guarantees of program correctness, more flexibility and compactness of program writing.
-    %\keywords{Nominal subtyping \and Structural subtyping  \and Semantic Featherweight Java \and Object-oriented languages \and Boolean types \and Type theory.}
+    We present SFJ---Semantic Featherweight Java, an implementation of FJ which features boolean types and semantic subtyping and integrates nominal as well as structural subtyping.
+    The benefits of SFJ, illustrated in the paper and the accompanying video (with audio/subtitles) \cite{UD20}, show how static typechecking of boolean types and semantic subtyping gives higher guarantees of program correctness, more flexibility and compactness of program writing.
+    \keywords{Nominal subtyping \and Structural subtyping  \and Semantic Featherweight Java \and Object-oriented languages \and Boolean types \and Type theory.}
 \end{abstract}

### 2020-04-20 conference/camera ready/background.tex
working on background

\section{Background}
\label{sec:background}
The technical developments behind semantic subtyping and its properties are not easy, however, they are completely transparent to the programmer. Moreover, they are beyond the scope of this paper and we refer the reader to the relevant work \cite{Dardha2013,Dardha2017}.
%
In this section we will briefly detail the types and terms of SFJ.

\subsection{Types}
The syntax of types $\tau$ is defined by the following grammar:
$$
\begin{array}{llll}
\tau & ::= & \alpha \ |\ \mu
& \mbox{{\em Types}}
\\
\alpha & ::=  & \zero \ |\ \btypes \ |\ [\wt{l:\tau}] \ |\ \alpha\ \synwedge\ \alpha \ |\ \synneg\ \alpha
\qquad
& \mbox{{\em Object types} ($\alpha $-types)}
\\
\mu & ::=  & \alpha \to \alpha \ |\ \mu\ \synwedge\ \mu \ |\ \synneg\ \mu
& \mbox{{\em Method types} ($\mu $-types)}
\end{array}
$$

The $\alpha$-types are used to type fields and the $\mu$-types are used to type methods.
Type $\zero$ is the empty type.
Type $\btypes$ denotes the \emph{basic} types, such as integers, booleans, etc.
{Record} types $[\wt{l:\tau}]$, where $\wt l$ is a sequence of disjoint labels, are used to type objects.
Arrow types $\alpha \to \alpha$ are used to type methods.

The boolean connectives $\bf{and}$ and $\bf{not}$ in the $\alpha$-types and $\mu$-types have their expected set-theoretic meanings.
We let $\alpha\ \syndiff\ \alpha'$  denote $\alpha\ \synwedge\ (\synneg\ \alpha')$, and $\alpha\ \synvee\ \alpha'$  denote $\synneg(\synneg\ \alpha\ \synwedge\ (\synneg\ \alpha'))$.

\subsection{Terms}
The syntax of terms is defined by the following grammar and is based on the standard syntax of terms in FJ \cite{featherweight}:
\begin{align*}
     & \mbox{\textit{Class declaration}}  & L \; ::= \; & \syndecl{C}{C}{\wt{\alpha \ a};\ K; \ \wt{M}\ }                         \\
     & \mbox{\textit{Constructor}}        & K \;::=\;   & C\ (\wt{\alpha\ x})\ \{\ \synsuper(\wt{x});\ \wt{\synthis.a}=\wt{x}; \} \\
     & \mbox{\textit{Method declaration}} & M \; ::= \; & \alpha \ m\ (\alpha \ x)\ \{\ \synret e; \}                             \\
     & \mbox{\textit{Expressions}}        & e \; ::=\;  & x\ |\  c\ |\ e.a\ |\ e.m(e) \ | \ \synnew C(\wt{e})
\end{align*}

We assume an infinite set of names, with some special names:
$\mathit{Object}$ denotes the root class, $\synthis$ denotes the current object and $\synsuper$ denotes the parent object.
We let  $A, B,\ldots$ range over classes; $a, b,\ldots$ over fields; $m, n,\ldots$ over methods and $x, y, z, \ldots$ over variables.
A {program} $(\wt{L}, e)$ is a pair of a sequence of class declarations $\wt L$, giving rise to a class hierarchy as specified by the inheritance relation, and an expression $e$ to be evaluated. 
%
A class declaration $L$ specifies the name of the class, the name of the parent class it extends, its typed fields, the constructor $K$ and its method declarations $M$.
The constructor $K$ initializes the fields of the object by assigning values to the fields inherited by the super-class and to the fields declared in the current \textbf{this} class.
%
A method declaration $M$ specifies the signature of the method, namely the return type, the method name and the formal parameter as well as the body of the method.
Notice that in our theoretical development we use unary methods, without loss of generality: tuples of arguments can be modelled by an object that instantiates a ``special''class containing as fields all the needed arguments.
Expressions $e$ include variables, constants, field accesses, method invocations and object creations.

\paragraph{On multimethods}
Since FJ is a core language, some features of the full Java are removed, such as overloading methods.
In our framework, by leveraging the expressivity of boolean connectives and semantic subtyping, we are able to restore overloading, among other features \cite[\S 8.4]{Dardha2017}.
We can thus model \emph{multimethods}, \cite{BC97}, which according to the authors is ``\emph{very clean and easy to understand [...] it would be the best solution for a brand new language}".
As an example, taken from Dardha \emph{et al.}  \cite{Dardha2013,Dardha2017}, consider the following class declarations:
$$
\begin{array}{ll}
\syndecl{A}{\mathit{Object}} {
&
\qquad
\syndecl{B}{A} {
\\
	%\qquad\quad \ldots
%&
	%\qquad\qquad\quad \ldots
%\\
	\qquad\quad \mathbf{int}\ \mathit{length}\ (\mathbf{string}\ s)\{\ \ldots\ \}\qquad
&
	\qquad\qquad\quad \mathbf{int}\ \mathit{length}\ (\mathbf{int}\ n)\{\ \ldots\ \}
\\
}
&
\qquad }
\end{array}
$$
Method {\it length} has type $\mathbf{string} \rightarrow \mathbf{int}$ in $A$. However, in $B$ it has type $(\mathbf{string} \rightarrow \mathbf{int}) \synwedge (\mathbf{int} \rightarrow \mathbf{int})$,
which can be simplified as $(\mathbf{string} \synvee \mathbf{int}) \rightarrow \mathbf{int}$.

% \subsection{Typing Rules} use the subtyping relation $\leq_{\mathcal B}$ defined semantically as set inclusion of types which are interpreted as subsets of a model, we call it $\mathcal B$ of FJ. Details of model $\mathcal B$ and the subtyping relation $\leq_{\mathcal B}$ are fully presented in \cite{Dardha2013,Dardha2017}.

% Typing judgements for expression are of the form $\Gamma \vdash_{\mathcal B} e : \tau$ and for method declarations are $\Gamma \vdash_{\mathcal B} M$. The same for class declarations and programs.

% Let us assume a sequence of class declarations $\wt{L}$.
% We first determine the (structural) type of every class $C$ in $\wt{L}$.
% To this aim, we have to take into account the inheritance relation specified in the class declarations in $\wt{L}$.
% We write ``$a \in C$'' to mean that there is a field declaration of name $a$ in class $C$ within the hierarchy $\wt L$.
% Similarly, we write ``$a \in C$ with type $\alpha$'' to also specify the declared type $\alpha$. Similar notations also hold for method names $m$.

### 2020-04-20 conference/camera ready/paper.tex
working on background

@@
 \title{SFJ: An implementation of Semantic Featherweight Java
-    \thanks{Research supported by the UK EPSRC grant EP/K034413/1, ``From Data Types to Session Types: A Basis for Concurrency and Distribution (ABCD)'', and by the EU HORIZON 2020 MSCA RISE project 778233 ``BehAPI: Behavioural Application Program Interfaces''.}
+    \thanks{Supported by the UK EPSRC grant EP/K034413/1, ``From Data Types to Session Types: A Basis for Concurrency and Distribution'' (ABCD), and by the EU HORIZON 2020 MSCA RISE project 778233
+    ``Behavioural Application Program Interfaces'' (BehAPI).}
 }
@@
 
-    An orthogonal subtyping question, typical of object-oriented languages, is the \emph{nominal} and the \emph{structural} subtyping.
+    An orthogonal subtyping question, typical of object-oriented languages, is the \emph{nominal} versus the \emph{structural} subtyping.
     Dardha \emph{et al.} \cite{Dardha2013,Dardha2017} defined boolean types and semantic subtyping for Featherweight Java (FJ) and integrated both nominal and structural subtyping, thus exploiting the benefits of both approaches.
@@
 There are two approaches to defining subtyping relations: the \emph{syntactic} and the \emph{semantic} approach.
-Syntactic subtyping is more mainstream in programming languages and is defined by means of a set of formal deductive subtyping rules.
-Semantic subtyping is less know:  one defines a formal model of the  language and an interpretation of types as subsets of this model, ultimately, these will be subsets of values in the language.
-Then, subtyping is defined as set inclusion of these subsets denoting types.
-%
-Orthogonally, for object-oriented languages there are two approaches to defining subtyping relations: the \emph{nominal} and the \emph{structural} approach.
-Nominal subtyping is based on {declarations} by the developer and is \emph{name}-based: $A$ is a subtype of $B$ if and only if it is declared to be so, that is if the class $A$ extends (or implements) the class (or interface) $B$.
-Structural subtyping instead is based on the \emph{structure} of a class, its fields and methods: a class $A$ is a subtype of a class $B$ if and only if the fields and methods of $A$ are a superset of the fields and methods of $B$, and their types in $A$ are subtypes of the types in $B$.
-It is not surprising that the nominal subtyping aligns well with the syntactic approach, and in fact it is more well-known than the structural subtyping, which in turn aligns well with the semantic approach.
+Syntactic subtyping \cite{Liskov1994} is more mainstream in programming languages and is defined by means of a set of formal deductive subtyping rules.
+Semantic subtyping \cite{Castagna2005,Cas05,Frisch2008} is more recent and less known:  one defines a formal model of the language and an interpretation of types as subsets of this model.
+Then, subtyping is defined as set inclusion of subsets denoting types.
 
-Dardha \emph{et al.} \cite{Dardha2013,Dardha2017} define {boolean types} and semantic subtyping for \emph{Featherweight Java} (FJ) \cite{featherweight}.
-This approach allows for the integration of both nominal and structural subtyping in FJ, bringing in higher guarantees of program correctness, flexibility and compactness in programming.
+Orthogonally, for object-oriented languages there are two approaches to defining subtyping relations: the \emph{nominal} and the \emph{structural} approach \cite{Integrating,Useful}.
+Nominal subtyping is based on {declarations} by the developer and is \emph{name}-based: ``$A$ is a subtype of $B$ if and only if it is declared to be so, that is if the class $A$ extends (or implements) the class (or interface) $B$''.
+Structural subtyping instead is based on the \emph{structure} of a class, its fields and methods: ``a class $A$ is a subtype of a class $B$ if and only if the fields and methods of $A$ are a superset of the fields and methods of $B$, and their types in $A$ are subtypes of the types in $B$''.
+Intutively, extending a class with more fields and methods results in a smaller set of inhabitants for that class.
+For example, the set of objects of a class \emph{Student} is smaller than the set of objects of a class \emph{Person} as each Student is a Person, but not the other way around.
+Hence, \emph{Student} is a structural subtype of \emph{Person}, even if it is not declared to be so.
+
+Dardha \emph{et al.} \cite{Dardha2013,Dardha2017} define boolean types---based on set-theoretic operations such as \textbf{and}, \textbf{not}, \textbf{or}---and semantic subtyping for \emph{Featherweight Java} (FJ) \cite{featherweight}.
+This approach allows for the integration of both nominal and structural subtyping in FJ, bringing in higher guarantees of program correctness, flexibility and compactness in program writing.
 Unfortunately, these benefits were only presented at a theoretical level and not exploited practically, due to the lack of an implementation of the language and its types and type system.
@@
 In SFJ the developer has a larger and more expressive set of types, by using boolean connectives \textbf{and}, \textbf{not}, \textbf{or}, with the expected set-theoretic interpretation.
-On the other hand, this added expressivity does not introduce added complexity, but rather the opposite is true, as the developer has an easier, more compact and more elegant way of programming.
+On the other hand, this added expressivity does not add complexity.
+Rather the opposite is true, as the developer has an easier, more compact and elegant way of programming.
 SFJ integrates both structural and nominal subtyping, and the developer can choose which one to use.
-Finally, as discussed in Dardha \emph{et al.} \cite[\S 8.4]{Dardha2017}, thanks to semantic subtyping, we can easily encode in SFJ standard programming constructs and features of Java, such as lists, or overloading classes via multimethods \cite{BC97}, which are missing in FJ, thus making SFJ a more complete language closer to Java.
+Finally, as discussed in Dardha \emph{et al.} \cite[\S 8.4]{Dardha2017}, thanks to semantic subtyping, we can easily encode in SFJ standard programming constructs and features of the full Java language, such as lists, or overloading classes via multimethods \cite{BC97}, which are missing in FJ, thus making SFJ a more complete language closer to Java.
 
@@
 
-    Now, suppose that our class hierarchy is such that \emph{Polygon} is the parent class and all other geometric figure classes extend \emph{Polygon}, which is how one would naturally define the set of polygons.
-    If we think of the class hierarchy as given and part of legacy code, which cannot be changed, then again a natural way to implement this in Java is by defining the method \emph{diagonal} in the class \emph{Polygon} and using an \textbf{instanceof}, for e.g., inside a \textbf{try-catch} which throws an exception at runtime, if the argument passed to the method is a triangle.
+    Now, suppose our class hierarchy is such that \emph{Polygon} is the parent class and all other geometric figures extend \emph{Polygon}, which is how one would naturally define the set of polygons.
+    Suppose the class hierarchy is given and is part of legacy code, which cannot be changed.
+Then again, a natural way to implement this in Java is by defining the method \emph{diagonal} in the class \emph{Polygon} and using an \textbf{instanceof}, for example, inside a \textbf{try-catch} construct.
+Then, an exception would be thrown at runtime, if the argument passed to the method is a triangle.
 
-    We propose a more elegant solution, by combining boolean types and semantic subtyping, where only static typechecking is required and implement this in SFJ \cite{UD20}: it is enough to define a method \emph{diagonal} that has an argument of type \emph{Polygon} \textbf{and not} \emph{Triangle}, thus allowing the typechecker to check at compile time the restrictions on types:
+    We propose a more elegant solution, by combining boolean types and semantic subtyping, where only static typechecking is required and we implement this in SFJ \cite{UD20}: it is enough to define a method \emph{diagonal} that has an argument of type \emph{Polygon} \textbf{and not} \emph{Triangle}, thus allowing the typechecker to check at compile time the restrictions on types:
     $$
@@
 
-The technical developments behind semantic subtyping and its properties are complex, however, they are completely transparent to the developer.
-Due to space limit we will remove most theoretical details from this submission, as they are not necessary to understand the design and implementation of SFJ, which is the focus of this paper.
+%The technical developments behind semantic subtyping and its properties are complex, however, they are completely transparent to the developer.
+%Due to space limit we will remove most theoretical details from this submission, as they are not necessary to understand the design and implementation of SFJ, which is the focus of this paper.
 We refer the reader to the relevant work \cite{Dardha2013,Dardha2017} and we will include a more complete background on semantic subtyping in a possible full paper.
+
+The rest of the paper is organised as follows:
+\autoref{sec:background} details the technical developments behind semantic subtyping and its properties.
+\autoref{sec:design} details the design and implementation of SFJ.
+\autoref{sec:conclusion} concludes the paper.
+
+\input{background.tex}
 
@@
 \label{sec:design}
-The syntax of types is given by the following grammar \cite{Dardha2013,Dardha2017}:
-
-$$
-    \begin{array}{llll}
-        \tau   & ::=                                       & \alpha \ |\ \mu
-               & \mbox{{\em Type term}}
-        \\
-        \alpha & ::=                                       & \zero \ |\ \btypes \ |\ [\wt{l:\tau}] \ |\ \alpha \ \bf{and}\ \alpha \ |\ \bf{not}\ \alpha
-        \qquad
-               & \mbox{{\em Object type} ($\alpha $-type)}
-        \\
-        \mu    & ::=                                       & \alpha \to \alpha \ |\ \mu \ \bf{and}\ \mu \ |\ \bf{not}\ \mu
-               & \mbox{{\em Method type} ($\mu $-type)}
-    \end{array}
-$$
-
-$\alpha$-types are used to type fields and $\mu$-types are used to type methods.
-Type $\zero$ is the empty type.
-Type $\btypes$ denotes the \emph{basic} types, such as integers, booleans, etc.
-    {Record} types $[\wt{l:\tau}]$, where $\wt l$ is a sequence of disjoint labels are used to type objects.
-Arrow types $\alpha \to \alpha$ are used to type methods.
-The boolean types using $\bf{and}$ and $\bf{not}$ have their expected set-theoretic meanings, and $\bf{or}$ is obtained by their combination.\\
-The syntax of terms is given by the following grammar and is based on the standard syntax of terms in FJ \cite{featherweight,Dardha2013,Dardha2017}:
-\begin{align*}
-     & \mbox{\textit{Class declaration}}  & L \; ::= \; & \syndecl{C}{C}{\wt{\alpha \ a};\ K; \ \wt{M}\ }                         \\
-     & \mbox{\textit{Constructor}}        & K \;::=\;   & C\ (\wt{\alpha\ x})\ \{\ \synsuper(\wt{x});\ \wt{\synthis.a}=\wt{x}; \} \\
-     & \mbox{\textit{Method declaration}} & M \; ::= \; & \alpha \ m\ (\alpha \ x)\ \{\ \synret e; \}                             \\
-     & \mbox{\textit{Expressions}}        & e \; ::=\;  & x\ |\  c\ |\ e.a\ |\ e.m(e) \ | \ \synnew C(\wt{e})
-\end{align*}
 Since we want to use types $\tau$ in practice in SFJ, we restrict them to finite trees whose leaves are constants with no cycles.
-For e.g., a recursive type $\alpha = [a : \alpha]$ denotes an infinite program tree \textbf{new} $C$(\textbf{new} $C$($\cdots{}$)), hence we avoid it as it is inhabitable.
+For example, a recursive type $\alpha = [a : \alpha]$ denotes an infinite program tree \textbf{new} $C$(\textbf{new} $C$($\cdots{}$)), hence we avoid it as it is inhabitable.
 %Similarly the types $\alpha = [b: \beta]$, $\beta = [a = \alpha]$ would also be impossible to instantiate.
@@
 Similarly, given all pairs of types, the algorithm checks if one is a superset of the other, i.e.,\ they share common fields and methods, and if so, they are related by structural subtyping.
-For e.g., type $empty = []$, will have all classes as structural subtypes.
+For example, type $empty = []$, will have all classes as structural subtypes.
 %The above algorithm will without differentiating, find all nominal and structural subtypes as we always consider all other classes in the program to be potential subtype candidates.
@@
 
-To conclude, in this paper we presented the design and implementation of SFJ, an extension of Featherweight Java with boolean types and semantic subtyping, which allow for both structural and nominal subtyping in FJ as well as restore standard Java constructs for e.g., lists and features for e.g., overloading which were not present in FJ, making SFJ a more complete language.
+To conclude, in this paper we presented the design and implementation of SFJ, an extension of Featherweight Java with boolean types and semantic subtyping, which allow for both structural and nominal subtyping in FJ as well as restore standard Java constructs for example, lists and features for example, overloading which were not present in FJ, making SFJ a more complete language.
 
@@
 \bibliography{paper}
-
 \end{document}

### 2020-04-21 conference/camera ready/background.tex
more on background

@@
 \label{sec:background}
-The technical developments behind semantic subtyping and its properties are not easy, however, they are completely transparent to the programmer.
-Moreover, they are beyond the scope of this paper and we refer the reader to the relevant work \cite{Dardha2013,Dardha2017}.
-%
+The technical developments behind semantic subtyping and its properties are complex, however, they are completely transparent to the programmer.
+The framework is detailed and proved correct in the relevant work by Dardha \emph{et al.} \cite{Dardha2013,Dardha2017}, and SFJ builds on that framework.
+
 In this section we will briefly detail the types and terms of SFJ.
@@
 \qquad
-& \mbox{{\em Object types} ($\alpha $-types)}
+& \mbox{{\em Field types} ($\alpha $-types)}
 \\
@@
 $$
-Method {\it length} has type $\mathbf{string} \rightarrow \mathbf{int}$ in $A$.
-However, in $B$ it has type $(\mathbf{string} \rightarrow \mathbf{int}) \synwedge (\mathbf{int} \rightarrow \mathbf{int})$,
-which can be simplified as $(\mathbf{string} \synvee \mathbf{int}) \rightarrow \mathbf{int}$.
+Method {\it length} has type $\mathbf{string} \rightarrow \mathbf{int}$ in class $A$.
+However, because class $B$ extends class $A$,  {\it length} has type $(\mathbf{string} \rightarrow \mathbf{int})\ \synwedge\ (\mathbf{int} \rightarrow \mathbf{int})$ in class $B$,
+which can be simplified to $(\mathbf{string}\ \synvee\ \mathbf{int}) \rightarrow \mathbf{int}$.
 

### 2020-04-21 conference/camera ready/background.tex
first algorithm presented

@@
 \subsection{Types}
+\label{sec:types}
 The syntax of types $\tau$ is defined by the following grammar:
@@
 \subsection{Terms}
+\label{sec:terms}
 The syntax of terms is defined by the following grammar and is based on the standard syntax of terms in FJ \cite{featherweight}:
@@
 We let  $A, B,\ldots$ range over classes; $a, b,\ldots$ over fields; $m, n,\ldots$ over methods and $x, y, z, \ldots$ over variables.
+
 A {program} $(\wt{L}, e)$ is a pair of a sequence of class declarations $\wt L$, giving rise to a class hierarchy as specified by the inheritance relation, and an expression $e$ to be evaluated. 
@@
 
+Following FJ \cite{featherweight}, we rule out ill-formed programs, such as declaring a constructor named $B$ within a class named $A$; or multiple fields or methods having the same name; or fields having the same type as the type of the class they are defined in.
+
 % \subsection{Typing Rules} use the subtyping relation $\leq_{\mathcal B}$ defined semantically as set inclusion of types which are interpreted as subsets of a model, we call it $\mathcal B$ of FJ.

### 2020-04-21 conference/camera ready/paper.tex
more on background
first algorithm presented

@@
 
+% Algorithm typesetting package
+\usepackage[plain, shortend, linesnumberedhidden, ruled]{algorithm2e}
+\SetAlFnt{\sffamily \small}
+\SetKwRepeat{Do}{do}{while}
+\SetKwProg{Fn}{Function}{:}{end}
+\SetKwInOut{Input}{input}\SetKwInOut{Output}{output}
+
+%% For section renaming in autoref
 \addto\extrasenglish{%
   \renewcommand{\sectionautorefname}{\S}%
+  \renewcommand{\subsectionautorefname}{\S}%
+  \renewcommand{\algorithmautorefname}{Algo.}%
 }
@@
 
-The rest of the paper is organised as follows:
+\paragraph{Structure of the paper:}
 \autoref{sec:background} details the technical developments behind semantic subtyping and its properties.
@@
 \label{sec:design}
-Since we want to use types $\tau$ in practice in SFJ, we restrict them to finite trees whose leaves are constants with no cycles.
-For example, a recursive type $\alpha = [a : \alpha]$ denotes an infinite program tree \textbf{new} $C$(\textbf{new} $C$($\cdots{}$)), hence we avoid it as it is inhabitable.
-%Similarly the types $\alpha = [b: \beta]$, $\beta = [a = \alpha]$ would also be impossible to instantiate.
-When processing the abstract syntax tree (AST) of a program in SFJ, we forbid fields to have the same type as the type of the class they are defined in.
-Secondly, we mark any classes containing fields typed with only basic types as \emph{resolved} otherwise, as \emph{unresolved}.
-This is captured by the following algorithm, which checks if the type definitions in the program are valid.
-\begin{verbatim}
-do
-    boolean resolutionOccured = false
-    for class that is unresolved:
-        boolean resolved = true
-        for field in class that contains a class type:
-            if class type is not resolved
-                resolved = false
-        if resolved is true:
-            set class to be resolved
-            resolutionOccured = true
-while resolutionOccured is true
+Since we want to use types $\tau$ in practice in SFJ, we restrict them to \emph{finite trees} whose leaves are basic types, $\btypes$ \autoref{sec:terms} with no cycles.
+For example, a recursive type $\alpha = [a : \alpha]$ denotes an infinite program tree \textbf{new} $C$(\textbf{new} $C$($\cdots{}$)), hence we rule it out as it is not inhabitable.
+Similarly, the types $\alpha = [b: \beta]$, $\beta = [a = \alpha]$ create a cycle and thus would not be inhabitable.
+Notice that the above types can be defined in Java but as well are not inhabitable.
 
-if all classes are not resolved:
-    program contains invalid type definition
-\end{verbatim}
-%If the types in the program are finite trees whose leaves are constants with no cycles, then at each iteration of the algorithm, either all the types are going to be resolved, or we are going to be able to resolve at least one type.
-%If we do not resolve at least one type and not all types are resolved, we know we have encountered a cycle in the type definition.
+The type system of SFJ builds upon the type system by Dardha \emph{et al.} \cite{Dardha2013,Dardha2017}.
+It is implemented using ANTLR \cite{parr2013}.
+We start by defining the grammar of the language in Extended Backus-Naur Form (EBNF), following \autoref{sec:types} and by using ANTLR, we can automatically generate a parser for SFJ and extend it in order to implement the required checks for our type system.
+Running the parser returns an abstract syntax tree (AST) of a program in SFJ.
+
+When visiting the AST of a program, we check if it is well-formed, following the intuition at the end of \autoref{sec:background}.
+We mark any classes containing fields typed with only basic types as \emph{resolved} otherwise, as \emph{unresolved}.
+Using this information, \autoref{alg:types} checks if the type definitions in a program are valid, namely, if they are finite trees whose leaves are basic types with no cycles.
+%
+At each iteration of the algorithm we resolve at least one type, until all the types are resolved, and return $True$.
+However, if there is at least one type unresolved, that means there is a cycle in the type definition and we return $False$.
+
+\begin{algorithm}[t]
+    \SetAlgoLined
+    \caption{ValidityCheck}
+    \label{alg:types}
+
+    \DontPrintSemicolon
+    \Input{classes marked $resolved/unresolved$ depending if their fields contain only basic types.}
+    \Output{$True$ if all classes are valid type definitions, $False$ otherwise.}
+    \BlankLine
+    \Begin{
+        \Do{$resolutionOccured = true$}{ %\tcc {comment}
+            $resolutionOccured \longleftarrow false$ \;
+            \For{class that is unresolved in classes}{
+                $resolved \longleftarrow true$ \;
+                \For{field in class that contains a class type}{
+                    \If{type of field is unresolved}{
+                        $resolved \longleftarrow false$
+                    }
+
+                }
+                \;
+                \If{$resolved = true$}{
+                    $class \longleftarrow resolved$ \;
+                    $resolutionOccured \longleftarrow true$ \;
+                }
+            }
+        }
+        \;
+        \uIf{not all classes are resolved}{
+            \KwRet $False$
+        }
+        \Else{
+            \KwRet $True$
+        }
+    }
+\end{algorithm}
+
 %

### 2020-04-21 conference/camera ready/paper.tex
second algorithm entered

@@
 %%%%%%%%%%%%%%%%%%%%%%%%
-\usepackage[utf8]{inputenc}
+%\usepackage[utf8]{inputenc}
+\usepackage[T1]{fontenc}
 \usepackage[english]{babel}
@@
 
+\renewcommand{\theequation}{\arabic{section}.\arabic{equation}}
+
 % Algorithm typesetting package
@@
 \SetKwProg{Fn}{Function}{:}{end}
-\SetKwInOut{Input}{input}\SetKwInOut{Output}{output}
+\SetKwInOut{Input}{Input}\SetKwInOut{Output}{Output}
 
@@
   \renewcommand{\subsectionautorefname}{\S}%
-  \renewcommand{\algorithmautorefname}{Algo.}%
+  \renewcommand{\algorithmautorefname}{Algorithm}%
+  %\renewcommand{\exampleautorefname}{Ex.}%
 }
@@
 \label{sec:design}
-Since we want to use types $\tau$ in practice in SFJ, we restrict them to \emph{finite trees} whose leaves are basic types, $\btypes$ \autoref{sec:terms} with no cycles.
+\subsection{Valid type definitions}
+Since we want to use types $\tau$ in practice in SFJ, we restrict them to \emph{finite trees} whose leaves are basic types $\btypes$ \autoref{sec:terms} with no cycles.
 For example, a recursive type $\alpha = [a : \alpha]$ denotes an infinite program tree \textbf{new} $C$(\textbf{new} $C$($\cdots{}$)), hence we rule it out as it is not inhabitable.
 Similarly, the types $\alpha = [b: \beta]$, $\beta = [a = \alpha]$ create a cycle and thus would not be inhabitable.
-Notice that the above types can be defined in Java but as well are not inhabitable.
+Notice that the above types can be defined in Java, but as well are not inhabitable.
 
-The type system of SFJ builds upon the type system by Dardha \emph{et al.} \cite{Dardha2013,Dardha2017}.
-It is implemented using ANTLR \cite{parr2013}.
-We start by defining the grammar of the language in Extended Backus-Naur Form (EBNF), following \autoref{sec:types} and by using ANTLR, we can automatically generate a parser for SFJ and extend it in order to implement the required checks for our type system.
-Running the parser returns an abstract syntax tree (AST) of a program in SFJ.
-
-When visiting the AST of a program, we check if it is well-formed, following the intuition at the end of \autoref{sec:background}.
+SFJ is implemented using ANTLR \cite{parr2013}.
+We start by defining the grammar of the language in Extended Backus-Naur Form (EBNF), following \autoref{sec:types} and by using ANTLR, we can automatically generate a parser for SFJ and extend it in order to implement the required checks for our types and type system.
+Running the parser on an SFJ program returns an abstract syntax tree (AST) of that program.
+%
+When visiting the AST, we check if the program is well-formed, following the intuition at the end of \autoref{sec:background}.
 We mark any classes containing fields typed with only basic types as \emph{resolved} otherwise, as \emph{unresolved}.
@@
 %
-At each iteration of the algorithm we resolve at least one type, until all the types are resolved, and return $True$.
-However, if there is at least one type unresolved, that means there is a cycle in the type definition and we return $False$.
+At each iteration of the algorithm we can resolve at least one type, until all the types are resolved, and return $True$.
+However, if there is at least one type unresolved, that means there is a cycle in the type definition, and we return $False$.
 
@@
     \SetAlgoLined
-    \caption{ValidityCheck}
+    \caption{Validity Check for Type Definitions}
     \label{alg:types}
@@
     \DontPrintSemicolon
-    \Input{classes marked $resolved/unresolved$ depending if their fields contain only basic types.}
+    \Input{$classes$, the set of classes marked $resolved/unresolved$ depending if their fields contain only basic types.}
     \Output{$True$ if all classes are valid type definitions, $False$ otherwise.}
@@
 
-%
-Now, we can define the subtyping relation.
-Building upon the interpretation of types as sets of values to define the semantic subtyping for FJ \cite{Dardha2013,Dardha2017}, in our implementation of SFJ we keep track of the subtyping relation by defining a map from a type to the set of its subtypes, with the property that the set of values of a subtype is included in the set of values of the type.
-As a first step, we start with basic types.
-\setlength{\abovedisplayskip}{5pt}
-\setlength{\belowdisplayskip}{5pt}
-\begin{flalign*}
-    &Double  = \left\{Double, Float, Int, Short, Byte\right\} && Float = \left\{Float, Short, Byte\right\}\\
-    &Long = \left\{Long, Int, Short, Byte\right\}  && Int = \left\{Int, Short, Byte\right\}\\
-    &Short = \left\{Short, Byte\right\} \qquad Byte = \left\{Byte\right\} && Boolean = \left\{Boolean \right\}
-\end{flalign*}
+\subsection{Building semantic subtyping for SFJ}
+If \autoref{alg:types} returns $True$, we can proceed to building semantic subtyping for SFJ.
+Leveraging the interpretation of types as sets of values to define the semantic subtyping for FJ \cite{Dardha2013,Dardha2017}, in our implementation of SFJ we keep track of the semantic subtyping relation by defining a map from a type to the set of its subtypes, satisfying the property that the set of values of a subtype is included in the set of values of the type.
+We start with basic types and define $Universe$ to be the set of all types.
+The mapping for basic types is defined as follows.
+\begin{equation}
+    \label{eq:initialSubtype}
+    \begin{array}{llllll}
+        & Double   & = \left\{Double, Float, Int, Short, Byte\right\} &  & Float & = \left\{Float, Short, Byte\right\} \\
+        & Long     & =    \left\{Long, Int, Short, Byte\right\}       &  & Int   & = \left\{Int, Short, Byte\right\}   \\
+        & Short    & =    \left\{Short, Byte\right\}                  &  & Byte  & = \left\{Byte\right\}               \\
+        & Boolean  & =    \left\{Boolean\right\}                      &  & Void  & = \left\{Void\right\}               \\
+        & Universe & = \{Double, Float, Long, Int,                                                                     \\
+        &          & \quad Short, Byte, Boolean, Void\}
+   \end{array}
+\end{equation}
 
-Note that \emph{Int} is not a subtype of \emph{Float} as a 32-bit \emph{float} cannot represent the whole set of 32-bit \emph{integer} values accurately and therefore \emph{Int} is not fully set-contained, however this is not the case for \emph{Int} and \emph{Double}.
+Note that \emph{Int} is not a subtype of \emph{Float} as a 32-bit \emph{float} cannot represent the whole set of 32-bit \emph{integer} values accurately and therefore \emph{Int} is not fully set-contained in \emph{Float}, however this is not the case for \emph{Int} and \emph{Double}.
 Similarly, \emph{Long} is not a subtype of \emph{Double}.
 
-Finally, the following algorithm defines the subtyping relation for all class types, which concludes the process.
-\begin{verbatim}
-function generateRelation(classes):
-    List<class> untyped = []
-    for class in classes:
-        if addClass(class) is false:
-            untyped.add(class)
-    if untyped is not []:
-        generateRelation(untyped)
+\autoref{alg:subtyping} builds the semantic subtyping relation for all class types of an SFJ program by calling the function \textit{generateRelation}.
+Given that classes are valid type definitions by \autoref{alg:types}, we are guaranteed that \autoref{alg:subtyping} will terminate.
 
-function addClass(class):
-    for existing class type in relation:
-        if checkSuperSet(class, existingClass) is false:
-            return false
-        checkSuperSet(existingClass, class)
-        add class to its own subtype relation
+%
+\begin{algorithm}
+    \caption{Semantic Subtyping for SFJ}
+    \label{alg:subtyping}
 
-function checkSuperSet(class, other):
-    boolean flag = true
-    for field in class:
-        if field contains type not in relation:
-            return false
-        if other does not have field:
-            flag = false
-        else:
-            if other.field.types not fully contains field.types:
-                flag = false
-    for method in class:
-        if method contains type not in relation:
-            return false
-        if other does not have method:
-            flag = false
-        else:
-            if other.method.types not fully contains method.types:
-                flag = false
-    if flag == true:
-        add class to other subtype relation
-\end{verbatim}
+    \DontPrintSemicolon
+    \Input{$classes$, the set of classes in an SFJ program for which we have not yet defined a subtyping relation.\\
+    $relation$, mapping of types to the set of subtypes. Initially this relation\\ only contains the entries defined in Equation \ref{eq:initialSubtype}. \;
+    }
+    \Begin{
+        \Fn{generateRelation(classes: List<Class>)}{
+            $unprocessed: List<Class>\ \longleftarrow []$
+
+            \For{class in classes}{
+                \If{addClass(class) = $False$}{
+                    $unprocessed.add(class)$
+                }
+            }
+            \If{$untyped \neq []$}{
+                $generateRelation(unprocessed)$
+            }
+        }
+        \;
+        \Fn{addClass(class: Class) $\rightarrow$ boolean}{
+            \For{existing class type in relation}{
+                \If{checkSuperSet(class, existingClass) = false}{
+                    \KwRet $False$\;
+                }
+                $checkSuperSet(existingClass, class)$\;
+            }
+            $relation[class].add(class)$ \;
+            $relation[Universe].add(class)$ \;
+            \KwRet $True$ \;
+        }
+        \;
+        \Fn{checkSuperSet(class: Class, other: Class) $\rightarrow$ boolean}{
+            $flag \longleftarrow True$ \;
+            \For{field in class}{
+                \If{field contains type not in relation}{
+                    \KwRet $False$
+                }
+                \uIf{other does not contain field}{
+                    $flag \longleftarrow False$
+                }
+                \Else{
+                    \If{other.field.types does not fully contain field.types}{
+                        $flag \longleftarrow False$
+                    }
+                }
+            }
+
+            \For{method in class}{
+                \If{method contains type not in relation}{
+                    \KwRet $False$
+                }
+                \uIf{other does not contain method}{
+                    $flag \longleftarrow False$
+                }
+                \Else{
+                    \If{other.method.types does not fully contain method.types}{
+                        $flag \longleftarrow False$
+                    }
+                }
+            }
+
+            \If{$flag = True$}{
+                $relation[other].add$
+            }
+        }
+    }
+\end{algorithm}
+%
+\subsection{Polygons: Continued}
 Let us illustrate the subtyping algorithm with our \emph{Polygons} \autoref{polygons}.

### 2020-04-22 conference/camera ready/background.tex
Add details of method types and structural/nominal discussion. Fix typos

@@
 A class declaration $L$ specifies the name of the class, the name of the parent class it extends, its typed fields, the constructor $K$ and its method declarations $M$.
-The constructor $K$ initializes the fields of the object by assigning values to the fields inherited by the super-class and to the fields declared in the current \textbf{this} class.
+The constructor $K$ initialises the fields of the object by assigning values to the fields inherited by the super-class and to the fields declared in the current \textbf{this} class.
 %
 A method declaration $M$ specifies the signature of the method, namely the return type, the method name and the formal parameter as well as the body of the method.
-Notice that in our theoretical development we use unary methods, without loss of generality: tuples of arguments can be modelled by an object that instantiates a ``special''class containing as fields all the needed arguments.
+Notice that in our theoretical development we use unary methods, without loss of generality: tuples of arguments can be modelled by an object that instantiates a ``special'' class containing as fields all the needed arguments.
 Expressions $e$ include variables, constants, field accesses, method invocations and object creations.

### 2020-04-22 conference/camera ready/paper.tex
Add details of method types and structural/nominal discussion. Fix typos

@@
 Structural subtyping instead is based on the \emph{structure} of a class, its fields and methods: ``a class $A$ is a subtype of a class $B$ if and only if the fields and methods of $A$ are a superset of the fields and methods of $B$, and their types in $A$ are subtypes of the types in $B$''.
-Intutively, extending a class with more fields and methods results in a smaller set of inhabitants for that class.
+Intuitively, extending a class with more fields and methods results in a smaller set of inhabitants for that class.
 For example, the set of objects of a class \emph{Student} is smaller than the set of objects of a class \emph{Person} as each Student is a Person, but not the other way around.
@@
 Similarly, the types $\alpha = [b: \beta]$, $\beta = [a = \alpha]$ create a cycle and thus would not be inhabitable.
-Notice that the above types can be defined in Java, but as well are not inhabitable.
+Notice that the above types can be defined in Java. These types can be inhabited by assigning \emph{null} to all fields, but are not useful in practise.
 
@@
 While structurally they can be used interchangeably in our type system, their ``meaning'' is completely different.
+Therefore, we lose some safety in our language as it is unlikely we would like to use a colour where we expect a coordinate, but in the eyes of the type-checker, it is correct.
+
+Dardha \emph{et al.}\cite{Dardha2017} suggest a possible extension to the language by including a $nominal$ keyword that can be used in the class signature to indicate that this class should only be considered by the nominal component of the type system.
+This would allow us finer control over the type system and eliminate the previously mentioned problems.
+However, being optional, we argue this would create more bugs than problems it would solve, due to easily forgetting the keyword.
+For example with destructors in C++ classes,leaving out the virtual keyword of the parent class destructor in certain situations can lead to memory leaks.
+Instead a \emph{structural} keyword could be introduced as an extension to the language which specifies that this class is allowed to have structural subtypes.
+A class with this keyword is not added to the subtyping relation of other structurally similar classes which are not marked \emph{structural} as this would weaken the invariant conditions of that class.
+
 On the other hand, with nominal subtyping in Java one can define an overridden method to perform the opposite logic to what the super class is expecting.
-Therefore, both approaches leave an expectation on the developer to check what they are doing is correct, hence the integration of both in SFJ is the key to overcoming these drawbacks, as one can choose which approach to use for a given task.
+Therefore, both approaches leave an expectation on the developer to ensure what they are doing is correct, hence the integration of both in SFJ is the key to overcoming these drawbacks, as one can choose which approach to use for a given task.
 
-Furthermore, we have extended FJ with method types, demonstrated below to show another possible implementation of the \emph{Polygons} \autoref{polygons}.
+Furthermore, we have extended FJ with \emph{method types}, demonstrated below to show another possible implementation of the \emph{Polygons} \autoref{polygons}.
 $$
@@
 
-We instead only implement the \emph{getDiagonal} method on polygons that support it.
-We can then specify to only accept classes that implement the method \emph{getDiagonal}.
-If we pass an argument of type \emph{Triangle}, it is not in the computed set of subtypes, so we get a typechecking error at compile time.
-We do not have to generate a relation for all methods in all classes as this is computationally unnecessary and instead
-only generate the set of subtypes on encountering a method type.
+We define the method type as accepting any type and its subtypes which implement a method called \emph{diagonal} which is a method from the type \emph{Void} to a \emph{Double}.
+
+In order to type-check the value passed to such an argument, we can at compile-time build a collection of types $\{type_{1},\ type_{2},\ \ldots\}$ which have this method.
+We can do this the simply iterating over the same list of classes in the program as when we were creating our subtyping relation and checking for the presence of the required method.
+The set of accepted types would then be the union of all their corresponding subtypes $([\![type_{1}]\!]_{B} \cup [\![type_{2}]\!]_{B} \cup\ \ldots)$.
+
+However, calculating this collection of types for each method of every class would be computationally unnecessary as only few of these would be used for method types.
+Therefore we only compute them on demand during type-checking when we come across such a type.
+
+We can therefore use method types to statically include or exclude a portion of our type hierarchy.
+However unlike the use of interfaces such as in the suggested Java solution to the Polygons problem \autoref{polygons}, the values that can be accepted by a method type do not have to be related to each other in any way in the class hierarchy.
+This becomes increasingly useful if we have a legacy system as we can still accept all the classes that have defined \emph{Diagonal} methods without having to go back and add interface implementations.
 

### 2020-04-22 conference/camera ready/paper.tex
more on subtyping

--- 2020-04-21 conference/camera ready/paper.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%
-%\usepackage[utf8]{inputenc}
+\usepackage[utf8]{inputenc}
 \usepackage[T1]{fontenc}
@@
 % Algorithm typesetting package
-\usepackage[plain, shortend, linesnumberedhidden, ruled]{algorithm2e}
+\usepackage[plain, shortend, linesnumbered]{algorithm2e}
 \SetAlFnt{\sffamily \small}
@@
   %\renewcommand{\exampleautorefname}{Ex.}%
+  \renewcommand{\equationautorefname}{Mapping}%
 }
@@
 \section{Introduction}
+\label{sec:intro}
 There are two approaches to defining subtyping relations: the \emph{syntactic} and the \emph{semantic} approach.
@@
 SFJ is implemented using ANTLR \cite{parr2013}.
-We start by defining the grammar of the language in Extended Backus-Naur Form (EBNF), following \autoref{sec:types} and by using ANTLR, we can automatically generate a parser for SFJ and extend it in order to implement the required checks for our types and type system.
+We start by defining the grammar of the language in Extended Backus-Naur Form (EBNF), following \autoref{sec:types} and by running ANTLR, we can automatically generate a parser for SFJ and extend it in order to implement the required checks for our types and type system.
 Running the parser on an SFJ program returns an abstract syntax tree (AST) of that program.
-%
+
 When visiting the AST, we check if the program is well-formed, following the intuition at the end of \autoref{sec:background}.
@@
 %
-At each iteration of the algorithm we can resolve at least one type, until all the types are resolved, and return $True$.
-However, if there is at least one type unresolved, that means there is a cycle in the type definition, and we return $False$.
+At each iteration of the algorithm we can resolve at least one type, until all the types in the SFJ program are resolved, and return $True$.
+However, if there is at least one type unresolved, meaning there is a cycle in the type definition, we return $False$.
 
@@
     \DontPrintSemicolon
-    \Input{$classes$, the set of classes marked $resolved/unresolved$ depending if their fields contain only basic types.}
+    \Input{$classes$, the set of classes marked $resolved$ if their fields contain only basic types, $unresolved$ otherwise.}
     \Output{$True$ if all classes are valid type definitions, $False$ otherwise.}
@@
     \Begin{
-        \Do{$resolutionOccured = true$}{ %\tcc {comment}
-            $resolutionOccured \longleftarrow false$ \;
+        \Do{$resolutionOccured = True$}{ %\tcc {comment}
+            $resolutionOccured \longleftarrow False$ \;
             \For{class that is unresolved in classes}{
-                $resolved \longleftarrow true$ \;
+                $resolved \longleftarrow True$ \;
                 \For{field in class that contains a class type}{
                     \If{type of field is unresolved}{
-                        $resolved \longleftarrow false$
+                        $resolved \longleftarrow False$
                     }
@@
                 \;
-                \If{$resolved = true$}{
+                \If{$resolved = True$}{
                     $class \longleftarrow resolved$ \;
-                    $resolutionOccured \longleftarrow true$ \;
+                    $resolutionOccured \longleftarrow True$ \;
                 }
@@
 
-\subsection{Building semantic subtyping for SFJ}
-If \autoref{alg:types} returns $True$, we can proceed to building semantic subtyping for SFJ.
+\subsection{Building the semantic subtyping for SFJ}
+If \autoref{alg:types} returns $True$, meaning all type definitions in the program are valid, we can proceed to building the semantic subtyping for SFJ.
 Leveraging the interpretation of types as sets of values to define the semantic subtyping for FJ \cite{Dardha2013,Dardha2017}, in our implementation of SFJ we keep track of the semantic subtyping relation by defining a map from a type to the set of its subtypes, satisfying the property that the set of values of a subtype is included in the set of values of the type.
@@
 Given that classes are valid type definitions by \autoref{alg:types}, we are guaranteed that \autoref{alg:subtyping} will terminate.
+The semantic subtyping generated by \autoref{alg:subtyping} is a preorder: it is reflexive and transitive. This is also illustrated by \autoref{eq:initialSubtype}.
+
+Some comments on \autoref{alg:subtyping} will follow.
+In function \textit{generateRelation} we iterate over the set of classes in an SFJ program.
+If the class currently being processed contains fields or methods not present in the subtyping relation (line 5), then we add the current class to the list of \emph{unprocessed} classes (line 6), so we can process its fields and methods first.
+The set of unprocessed classes will then be inspected again in a recursive call (line 10).
+The next two functions of the algorithm, \emph{addClass} and \emph{checkSuperSet}, check subtyping for the current $class$ being processed and update $relation$, which is a mapping from a type to its subtypes and originally only consists of entries from \autoref{eq:initialSubtype}.
+In function $addClass(class)$ we check if the type $class$ is a subtype of an existing type in $relation$ (lines 15-18), as well as the opposite, meaning if $class$ is a supertype of an existing type in $relation$ (line 19).
+In order to do so $checkSuperSet$ checks all fields (lines 28-39) and all methods (lines 40-51) in $class$ and compares them with an \emph{existingClass} in $relation$.
+If a subtyping relation is established, then it is added to $relation$ (line 53).
+Finally, upon returning from $checkSuperSet$, we also add $class$ to its own relation (line 21) to satisfy reflexivity and to $Universe$ (line 22), which is a supertype of all types.
+
+It is worth noticing that the subtyping algorithm finds all nominal and structural subtypes of a given type. This is due to the fact that all pairs of types are inspected.
+Recall from \autoref{sec:intro} that nominal subtyping is name-based and given by the class hierarchy defined by the programmer, whether structural subtyping is structure-based and given by the set-inclusion of fields and methods.
+In particular, it is contra-variant with respect to this set-inclusion.
+\autoref{alg:subtyping} finds all structural subtypes of a given class because it checks that its fields and methods are a superset of existing types in $relation$.
+On the other hand, it also finds all nominal subtypes because a class inherits all fields and methods of its superclass and as such its fields and methods are a superset of its superclass.
+This means that checking for structural subtyping is enough because nominal subtyping will be captured due to inheritance of fields and methods.
 
@@
 \begin{algorithm}
-    \caption{Semantic Subtyping for SFJ}
+    \caption{Semantic Subtyping for SFJ Classes}
     \label{alg:subtyping}
@@
     \DontPrintSemicolon
-    \Input{$classes$, the set of classes in an SFJ program for which we have not yet defined a subtyping relation.\\
-    $relation$, mapping of types to the set of subtypes. Initially this relation\\ only contains the entries defined in Equation \ref{eq:initialSubtype}. \;
+    \Input{$classes$, the set of classes in an SFJ program for which we have not yet defined the subtyping relation.\;
+    $relation$, the mapping of types to the set of subtypes, initially being Mapping \ref{eq:initialSubtype}. \;
     }
@@
             \For{existing class type in relation}{
-                \If{checkSuperSet(class, existingClass) = false}{
+                \If{checkSuperSet(class, existingClass) = False}{
                     \KwRet $False$\;

### 2020-04-22 conference/camera ready/paper.tex

@@
 Similarly, the types $\alpha = [b: \beta]$, $\beta = [a = \alpha]$ create a cycle and thus would not be inhabitable.
-Notice that the above types can be defined in Java, but as well are not inhabitable.
+Notice that the above types can be defined in Java. These types can be inhabited by assigning \emph{null} to all fields, but are not useful in practice.
 
 SFJ is implemented using ANTLR \cite{parr2013}.
-We start by defining the grammar of the language in Extended Backus-Naur Form (EBNF), following \autoref{sec:types} and by running ANTLR, we can automatically generate a parser for SFJ and extend it in order to implement the required checks for our types and type system.
+We start by defining the grammar of the language in Extended Backus-Naur Form (EBNF), following \autoref{sec:types} and by using ANTLR, we can automatically generate a parser for SFJ and extend it in order to implement the required checks for our types and type system.
 Running the parser on an SFJ program returns an abstract syntax tree (AST) of that program.

### 2020-04-22 conference/camera ready/paper.tex
minor

@@
 Structural subtyping instead is based on the \emph{structure} of a class, its fields and methods: ``a class $A$ is a subtype of a class $B$ if and only if the fields and methods of $A$ are a superset of the fields and methods of $B$, and their types in $A$ are subtypes of the types in $B$''.
-Intuitively, extending a class with more fields and methods results in a smaller set of inhabitants for that class.
-For example, the set of objects of a class \emph{Student} is smaller than the set of objects of a class \emph{Person} as each Student is a Person, but not the other way around.
-Hence, \emph{Student} is a structural subtype of \emph{Person}, even if it is not declared to be so.
+For example, the set of inhabitants of a class \emph{Student} is smaller than the set of inhabitants of a class \emph{Person}, as each Student is a Person, but not the other way around.
+However, the set of fields and methods of \emph{Student} is a superset of that of \emph{Person}.
+Hence, \emph{Student} is a structural subtype of \emph{Person}, even if it is not declared so.
 
@@
 SFJ is implemented using ANTLR \cite{parr2013}.
-We start by defining the grammar of the language in Extended Backus-Naur Form (EBNF), following \autoref{sec:types} and by using ANTLR, we can automatically generate a parser for SFJ and extend it in order to implement the required checks for our types and type system.
+We start by defining the grammar of the language in Extended Backus-Naur Form (EBNF), following \autoref{sec:types} and by running ANTLR, we can automatically generate a parser for SFJ and extend it in order to implement the required checks for our types and type system.
 Running the parser on an SFJ program returns an abstract syntax tree (AST) of that program.
@@
 
-\subsection{Building the semantic subtyping for SFJ}
-If \autoref{alg:types} returns $True$, meaning all type definitions in the program are valid, we can proceed to building the semantic subtyping for SFJ.
+\subsection{Building semantic subtyping for SFJ}
+If \autoref{alg:types} returns $True$, meaning all type definitions in an SFJ program are valid, we can proceed with building the semantic subtyping.
 Leveraging the interpretation of types as sets of values to define the semantic subtyping for FJ \cite{Dardha2013,Dardha2017}, in our implementation of SFJ we keep track of the semantic subtyping relation by defining a map from a type to the set of its subtypes, satisfying the property that the set of values of a subtype is included in the set of values of the type.

### 2020-04-22 conference/camera ready/paper.tex
on complexity

@@
 
+Finally, a note on complexity.
+The complexity of \autoref{alg:types} is $\mathcal{O}(n)$, and the complexity of \autoref{alg:subtyping} is $\mathcal{O}(n^{2})$, with $n$ being the size of the input.
+The reason for an exponential complexity \autoref{alg:subtyping} is due to the symmetric check of structural subtyping between a $class$ and an $existingClass$ in $relation$.
+Notice that if we were to only work with nominal subtyping, then we only require traversing the class hierarchy once,  which gives an $\mathcal{O}(n)$ complexity.
+
+\subsection{Type system for SFJ}
+The subtyping relation generated by \autoref{alg:subtyping} is used in the type system for SFJ terms, given in \autoref{sec:terms}.
+The complete type system and its properties are presented in relevant work \cite{Dardha2013,Dardha2017} and are beyond the scope of this paper.
+Let us briefly comment on typing expressions in SFJ.
+Field access $e.a$, method call $e.m(e)$ and object creation $\synnew C(\wt{e})$ are typed in the same way as in Java.
+Let us comment on constants...
+
 %

### 2020-04-23 conference/camera ready/paper.tex
new place for subtyping
add typing for constants, fix typos

@@
 
-\makeatletter
-\def\verbatim@font{\linespread{0.75}\normalfont\ttfamily}
-\makeatother
-
 % Non standard
 %\usepackage{varacca}
-\input semmacros
+\input{semmacros}
 
@@
 
-    We present SFJ---Semantic Featherweight Java, an implementation of FJ which features boolean types and semantic subtyping and integrates nominal as well as structural subtyping.
+    We present SFJ---Semantic Featherweight Java, an implementation of FJ which features boolean types, semantic subtyping and integrates nominal as well as structural subtyping.
     The benefits of SFJ, illustrated in the paper and the accompanying video (with audio/subtitles) \cite{UD20}, show how static typechecking of boolean types and semantic subtyping gives higher guarantees of program correctness, more flexibility and compactness of program writing.
@@
 This approach allows for the integration of both nominal and structural subtyping in FJ, bringing in higher guarantees of program correctness, flexibility and compactness in program writing.
-Unfortunately, these benefits were only presented at a theoretical level and not exploited practically, due to the lack of an implementation of the language and its types and type system.
+Unfortunately, these benefits were only presented at a theoretical level and not exploited practically, due to the lack of an implementation of the language, its types and type system.
 
@@
     In Java this could be implemented in the following ways:
-    $$
+
+    \begin{equation*}
         \begin{array}{l}
@@
         \end{array}
-    $$
+    \end{equation*}
+
     Or by means of an interface \emph{Diagonal}:
-    $$
+
+    \begin{equation*}
         \begin{array}{l}
@@
         \end{array}
-    $$
+    \end{equation*}
 
@@
     We propose a more elegant solution, by combining boolean types and semantic subtyping, where only static typechecking is required and we implement this in SFJ \cite{UD20}: it is enough to define a method \emph{diagonal} that has an argument of type \emph{Polygon} \textbf{and not} \emph{Triangle}, thus allowing the typechecker to check at compile time the restrictions on types:
-    $$
+
+    \begin{equation*}
         \begin{array}{l}
@@
         \end{array}
-    $$
+    \end{equation*}
+
     We can now call \emph{diagonal} on an argument of type \emph{Polygon}: if the polygon is \textbf{not} a \emph{Triangle}, then the method returns the length of its longest diagonal; otherwise, if the polygon is a triangle, then there will be a type error at compile time.
@@
 Similarly, the types $\alpha = [b: \beta]$, $\beta = [a = \alpha]$ create a cycle and thus would not be inhabitable.
-Notice that the above types can be defined in Java. These types can be inhabited by assigning \emph{null} to all fields, but are not useful in practice.
+Notice that the above types can be defined in Java as these types can be inhabited by assigning \emph{null} to all fields, but are not useful in practice.
 
@@
 \subsection{Building semantic subtyping for SFJ}
+
 If \autoref{alg:types} returns $True$, meaning all type definitions in an SFJ program are valid, we can proceed with building the semantic subtyping.
@@
 The mapping for basic types is defined as follows.
+
 \begin{equation}
@@
 Some comments on \autoref{alg:subtyping} will follow.
-In function \textit{generateRelation} we iterate over the set of classes in an SFJ program.
-If the class currently being processed contains fields or methods not present in the subtyping relation (line 5), then we add the current class to the list of \emph{unprocessed} classes (line 6), so we can process its fields and methods first.
+In function $generateRelation$ we iterate over the set of classes in an SFJ program.
+If the class currently being processed contains types in its fields or methods not present in the subtyping relation (line 5, 30, 42), then we add the current class to the list of \emph{unprocessed} classes (line 6) so we can process its later as it is currently undecidable.
 The set of unprocessed classes will then be inspected again in a recursive call (line 10).
@@
 
-It is worth noticing that the subtyping algorithm finds all nominal and structural subtypes of a given type. This is due to the fact that all pairs of types are inspected.
-Recall from \autoref{sec:intro} that nominal subtyping is name-based and given by the class hierarchy defined by the programmer, whether structural subtyping is structure-based and given by the set-inclusion of fields and methods.
-In particular, it is contra-variant with respect to this set-inclusion.
-\autoref{alg:subtyping} finds all structural subtypes of a given class because it checks that its fields and methods are a superset of existing types in $relation$.
-On the other hand, it also finds all nominal subtypes because a class inherits all fields and methods of its superclass and as such its fields and methods are a superset of its superclass.
-This means that checking for structural subtyping is enough because nominal subtyping will be captured due to inheritance of fields and methods.
-
-Finally, a note on complexity.
-The complexity of \autoref{alg:types} is $\mathcal{O}(n)$, and the complexity of \autoref{alg:subtyping} is $\mathcal{O}(n^{2})$, with $n$ being the size of the input.
-The reason for an exponential complexity \autoref{alg:subtyping} is due to the symmetric check of structural subtyping between a $class$ and an $existingClass$ in $relation$.
-Notice that if we were to only work with nominal subtyping, then we only require traversing the class hierarchy once,  which gives an $\mathcal{O}(n)$ complexity.
-
-\subsection{Type system for SFJ}
-The subtyping relation generated by \autoref{alg:subtyping} is used in the type system for SFJ terms, given in \autoref{sec:terms}.
-The complete type system and its properties are presented in relevant work \cite{Dardha2013,Dardha2017} and are beyond the scope of this paper.
-Let us briefly comment on typing expressions in SFJ.
-Field access $e.a$, method call $e.m(e)$ and object creation $\synnew C(\wt{e})$ are typed in the same way as in Java.
-Let us comment on constants...
-
-%
 \begin{algorithm}
@@
 \end{algorithm}
-%
+
+It is worth noticing that the subtyping algorithm finds all nominal and structural subtypes of a given type.
+This is due to the fact that all pairs of types are inspected.
+Recall from \autoref{sec:intro} that nominal subtyping is name-based and given by the class hierarchy defined by the programmer, whereas structural subtyping is structure-based and given by the set-inclusion of fields and methods.
+In particular, it is contra-variant with respect to this set-inclusion.
+\autoref{alg:subtyping} finds all structural subtypes of a given class because it checks that its fields and methods are a superset of existing types in $relation$.
+On the other hand, it also finds all nominal subtypes because a class inherits all fields and methods of its superclass and as such its fields and methods are a superset of its superclass.
+This means that checking for structural subtyping is enough because nominal subtyping will be captured due to inheritance of fields and methods.
+
+Finally, a note on complexity.
+The complexity of \autoref{alg:types} is $\mathcal{O}(n)$, and the complexity of \autoref{alg:subtyping} is $\mathcal{O}(n^{2})$, with $n$ being the size of the input.
+The reason for a quadratic complexity \autoref{alg:subtyping} is due to the symmetric check of structural subtyping between a $class$ and an $existingClass$ in $relation$.
+Notice that if we were to only work with nominal subtyping, then we only require traversing the class hierarchy once,  which gives an $\mathcal{O}(n)$ complexity.
+
+\subsection{Type system for SFJ}
+
+The subtyping relation generated by \autoref{alg:subtyping} is used in the type system for SFJ terms, given in \autoref{sec:terms}.
+The complete type system and its properties are presented in relevant work \cite{Dardha2013,Dardha2017} and are beyond the scope of this paper.
+Let us briefly comment on typing expressions in SFJ.
+Field access $e.a$, method call $e.m(e)$, variables $x$ and object creation $\synnew C(\wt{e})$ are typed in the same way as in Java.
+
+For the constants $c$, these are the literal values in our language such as $42$, $3.14$ or $True$.
+In most languages such as Java, literal values are cast to the type that is expected in the context they are used.
+For example, a method $\textbf{int}\ getNum()\{return\ 42\}$ automatically casts the value to the type $int$.
+
+However, in our system, types represent sets of values which must interpret boolean connectives in a set-theoretic way.
+Given the boolean type $int\ \textbf{and not}\ byte$, this type represents the values $-2^31$ to $2^31 - 1$ from which are removed the values $-128$ to $127$.
+If we were to type the value $42$ as an integer, this would be accepted by the boolean type, but would not be in the set of values of the type, thus breaking our semantic type system.
+Therefore, for all constants $c$, we use a strong typing to assign values the most restrictive types possible, i.e. the type with the smallest amount of values which still contains this value.
+With this typing rule, the value $42$ is assigned the type $Byte$, so would correctly not be accepted as a value of the type $(int\ \textbf{and not}\ byte)$.
+
+We have now defined how to type any expression that can be formed by the syntax and so have defined our typing rules.
+
 \subsection{Polygons: Continued}
+
 Let us illustrate the subtyping algorithm with our \emph{Polygons} \autoref{polygons}.
 The algorithm generates the following subtyping relation, which also includes the subtyping relation for basic types defined in the first step.
+
 \begin{flalign*}
@@
 \end{flalign*}
+
 Recall method \emph{diagonal} in class \emph{Diagonal}, we can see that the result of the set operation on its parameter type gives the following set of polygons:
+
 \begin{flalign*}
@@
 \end{flalign*}
+
 If we pass to \emph{diagonal} an argument of type \emph{Triangle}, as it is not contained in the computed set of subtypes, we get a typechecking error at compile time \cite{UD20}.
 
-The subtyping algorithm finds all nominal and structural subtypes.
-It finds all nominal subtypes as they inherit all fields and methods of their super class, so are always guaranteed to be a superset.
-Similarly, given all pairs of types, the algorithm checks if one is a superset of the other, i.e.,\ they share common fields and methods, and if so, they are related by structural subtyping.
-For example, type $empty = []$, will have all classes as structural subtypes.
+%It finds all nominal subtypes as they inherit all fields and methods of their super class, so are always guaranteed to be a superset.
+%Similarly, given all pairs of types, the algorithm checks if one is a superset of the other, i.e.,\ they share common fields and methods, and if so, they are related by structural subtyping.
+%For example, type $empty = []$, will have all classes as structural subtypes.
 %The above algorithm will without differentiating, find all nominal and structural subtypes as we always consider all other classes in the program to be potential subtype candidates.
+
+\subsection{Nominal vs Structural}
 
@@
 Therefore, both approaches leave an expectation on the developer to ensure what they are doing is correct, hence the integration of both in SFJ is the key to overcoming these drawbacks, as one can choose which approach to use for a given task.
+
+\subsection{Method Types}
 
@@
 \label{sec:conclusion}
+
 Semantic subtyping approach goes back to more than two decades ago \cite{Aiken,Damm}.
@@
 Hosoya and Pierce \cite{XML1,XML2,XML3} who define XDuce, an XML-oriented language designed specifically to transform XML documents in other XML documents satisfying certain properties.
-Castagna \emph{et al.} \cite{gentle,Cas05,FCB08} extend XDuce with first-class functions and arrow types and implement it as $\mathbb{C}$Duce.
+Castagna \emph{et al.} \cite{Castagna2005,Cas05,FCB08} extend XDuce with first-class functions and arrow types and implement it as $\mathbb{C}$Duce.
 The starting point of their framework is a higher-order $\lambda-$calculus with pairs and projections.

### 2020-04-23 conference/camera ready/paper.tex
no need for typing

@@
 
-It is worth noticing that the subtyping algorithm finds all nominal and structural subtypes of a given type.
-This is due to the fact that all pairs of types are inspected.
-Recall from \autoref{sec:intro} that nominal subtyping is name-based and given by the class hierarchy defined by the programmer, whereas structural subtyping is structure-based and given by the set-inclusion of fields and methods.
+It is worth noticing that the subtyping algorithm finds all nominal and structural subtypes of a given type. This is due to the fact that all pairs of types are inspected.
+Recall from \autoref{sec:intro} that nominal subtyping is name-based and given by the class hierarchy defined by the programmer, whether structural subtyping is structure-based and given by the set-inclusion of fields and methods.
 In particular, it is contra-variant with respect to this set-inclusion.
 \autoref{alg:subtyping} finds all structural subtypes of a given class because it checks that its fields and methods are a superset of existing types in $relation$.
+For example, all classes are structural subtypes of type $empty = []$.
 On the other hand, it also finds all nominal subtypes because a class inherits all fields and methods of its superclass and as such its fields and methods are a superset of its superclass.
@@
 
-\subsection{Type system for SFJ}
+% \subsection{Type system for SFJ}
+% The subtyping relation generated by \autoref{alg:subtyping} is used in the type system for SFJ terms, given in \autoref{sec:terms}.
+% The complete type system and its soundness properties are presented in relevant work \cite{Dardha2013,Dardha2017} and are beyond the scope of this paper.
+% However, let us briefly comment on typing terms in SFJ.
+% A program $(\tilde L,e)$ is well typed if both $\tilde{L}$ and $e$ are well typed.
+% Class declaration $L$ and method declaration $M$ are well typed if all their components are well typed.
+% Let us move onto expressions.
+% Field access $e.a$, method call $e.m(e)$ and object creation $\synnew C(\wt{e})$ are typed in the same way as in Java: we inspect the type of the field and the type of the method and its arguments to determine the type of the field access and method call, respectively.
+The type of an object creation is determined by the type of its class.
+% Constants are typed by their basic type.
 
-The subtyping relation generated by \autoref{alg:subtyping} is used in the type system for SFJ terms, given in \autoref{sec:terms}.
-The complete type system and its properties are presented in relevant work \cite{Dardha2013,Dardha2017} and are beyond the scope of this paper.
-Let us briefly comment on typing expressions in SFJ.
-Field access $e.a$, method call $e.m(e)$, variables $x$ and object creation $\synnew C(\wt{e})$ are typed in the same way as in Java.
-
-For the constants $c$, these are the literal values in our language such as $42$, $3.14$ or $True$.
-In most languages such as Java, literal values are cast to the type that is expected in the context they are used.
-For example, a method $\textbf{int}\ getNum()\{return\ 42\}$ automatically casts the value to the type $int$.
+% Subtyping generated in \autoref{alg:subtyping} is used in a subsumssion typing rule:
+% $$
+% \frac{\Gamma\vdash e:\alpha_1 \qquad \alpha_1 \leq \alpha_2}{\Gamma \vdash e:\alpha_2}
+% $$
+% We read this typing rule as follows:
+% If an expression $e$ is of type $\alpha_1$ under a typing context $\Gamma$ (details of a typing context are irrelevant here) and type $\alpha$ is a subtype $\leq$ of $\alpha_2$, then expression $e$ can be typed with $\alpha_2$.
 

### 2020-04-23 conference/camera ready/paper.tex
on type system for SFJ

--- 2020-04-23 conference/camera ready/paper.tex
@@
 
-    We can now call \emph{diagonal} on an argument of type \emph{Polygon}: if the polygon is \textbf{not} a \emph{Triangle}, then the method returns the length of its longest diagonal; otherwise, if the polygon is a triangle, then there will be a type error at compile time.
+    We can now call \emph{diagonal} on an argument of type \emph{Polygon}: if the polygon is \textbf{not} a \emph{Triangle}, then the method computes and returns the length of its longest diagonal; otherwise, there will be a type error at compile time.
 \end{example}
@@
 Similarly, the types $\alpha = [b: \beta]$, $\beta = [a = \alpha]$ create a cycle and thus would not be inhabitable.
-Notice that the above types can be defined in Java as these types can be inhabited by assigning \emph{null} to all fields, but are not useful in practice.
+Notice that these types can be defined and inhabited in Java by assigning \emph{null} to all fields in a class, however they are not useful in practice.
 
@@
 
-If \autoref{alg:types} returns $True$, meaning all type definitions in an SFJ program are valid, we can proceed with building the semantic subtyping.
-Leveraging the interpretation of types as sets of values to define the semantic subtyping for FJ \cite{Dardha2013,Dardha2017}, in our implementation of SFJ we keep track of the semantic subtyping relation by defining a map from a type to the set of its subtypes, satisfying the property that the set of values of a subtype is included in the set of values of the type.
-We start with basic types and define $Universe$ to be the set of all types.
-The mapping for basic types is defined as follows.
+If \autoref{alg:types} returns $True$, meaning all type definitions in a program are valid, we can then build the semantic subtyping.
+Leveraging the interpretation of types as sets of values to define semantic subtyping for FJ \cite{Dardha2013,Dardha2017}, in SFJ we keep track of the semantic subtyping relation by defining a map from a type to the set of its subtypes, satisfying the property that the set of values of a subtype is included in the set of values of the type.
+We start with basic types and map $Universe$ to the set of all types.
+The full mapping for basic types is defined as follows:
 
@@
 
-Some comments on \autoref{alg:subtyping} will follow.
+Some comments on \autoref{alg:subtyping} follow.
 In function $generateRelation$ we iterate over the set of classes in an SFJ program.
-If the class currently being processed contains types in its fields or methods not present in the subtyping relation (line 5, 30, 42), then we add the current class to the list of \emph{unprocessed} classes (line 6) so we can process its later as it is currently undecidable.
+If the class currently being processed contains types in its fields or methods not present in the subtyping relation (lines 5, 30, 42), then we add the current class to the list of \emph{unprocessed} classes (line 6) so we can process its fields and methods first and the class itself later after having all required type information.
 The set of unprocessed classes will then be inspected again in a recursive call (line 10).
@@
 The complexity of \autoref{alg:types} is $\mathcal{O}(n)$, and the complexity of \autoref{alg:subtyping} is $\mathcal{O}(n^{2})$, with $n$ being the size of the input.
-The reason for a quadratic complexity \autoref{alg:subtyping} is due to the symmetric check of structural subtyping between a $class$ and an $existingClass$ in $relation$.
+The reason for a quadratic complexity of \autoref{alg:subtyping} is due to the symmetric check of structural subtyping between a $class$ and an $existingClass$ in $relation$.
 Notice that if we were to only work with nominal subtyping, then we only require traversing the class hierarchy once,  which gives an $\mathcal{O}(n)$ complexity.
@@
 
-The subtyping relation generated by \autoref{alg:subtyping} is used in the type system for SFJ terms, given in \autoref{sec:terms}.
-The complete type system and its properties are presented in relevant work \cite{Dardha2013,Dardha2017} and are beyond the scope of this paper.
-Let us briefly comment on typing expressions in SFJ.
-Field access $e.a$, method call $e.m(e)$, variables $x$ and object creation $\synnew C(\wt{e})$ are typed in the same way as in Java.
+The type system for the SFJ language, given in \autoref{sec:terms}, is based on the type system by Dardha \emph{et al.} \cite{Dardha2013,Dardha2017} where the formal typing rules and soundness properties are detailed.
+As these formal developments are beyond the scope of this paper, we discuss typing for SFJ only informally.
 
-For the constants $c$, these are the literal values in our language such as $42$, $3.14$ or $True$.
-In most languages such as Java, literal values are cast to the type that is expected in the context they are used.
-For example, a method $\textbf{int}\ getNum()\{return\ 42\}$ automatically casts the value to the type $int$.
+A program $(\tilde L,e)$ is well typed if both $\tilde{L}$ and $e$ are well typed.
+Class declaration $L$ and method declaration $M$ are well typed if all their components are well typed.
+Let us move onto expressions $E$.
+Field access $e.a$, method call $e.m(e)$ and object creation $\synnew C(\wt{e})$ are typed in the same way as in Java: we inspect the type of the field and the type of the method and its arguments to determine the type of the field access and method call, respectively.
+The type of an object creation is determined by the type of its class.
+Regarding constants, in order to respect the set-theoretic interpretation of types as sets of values, we type constants with the most restrictive type, i.e., the type representing the smallest set of values containing the value itself.
+For example, the type system would assign to the value $42$ the type \textbf{byte}, which is the smallest in the sequence \textbf{byte}, \textbf{short}, \textbf{int} (see \autoref{eq:initialSubtype} for details).
 
-However, in our system, types represent sets of values which must interpret boolean connectives in a set-theoretic way.
-Given the boolean type $int\ \textbf{and not}\ byte$, this type represents the values $-2^31$ to $2^31 - 1$ from which are removed the values $-128$ to $127$.
-If we were to type the value $42$ as an integer, this would be accepted by the boolean type, but would not be in the set of values of the type, thus breaking our semantic type system.
-Therefore, for all constants $c$, we use a strong typing to assign values the most restrictive types possible, i.e. the type with the smallest amount of values which still contains this value.
-With this typing rule, the value $42$ is assigned the type $Byte$, so would correctly not be accepted as a value of the type $(int\ \textbf{and not}\ byte)$.
+% Let us move onto constants, where we can also illustrate the boolean connectives. 
+% Typing constants in Java uses type casting where the type of a constant changes to be adapted to the context it is being used in.
+% %For example, a method $\textbf{long}\ getNum()\{return\ 42;\}$ casts $42$ to \textbf{long}.
+% %Instead, in variable type inference $\textbf{var}\ x = 42$, the compiler treats $42$ as having type \textbf{int}, as this is the most efficient type for hardware operations.
 
-We have now defined how to type any expression that can be formed by the syntax and so have defined our typing rules.
+% Since in our framework types represent sets of values and the boolean connectives \textbf{and}, \textbf{not}, \textbf{or} are interpreted in a set-theoretic way, typing constants changes from Java as follows.
+% Consider the boolean type \textbf{int and} (\textbf{not byte}), this type is interpreted as the set of values $-2^31$ to $2^31 - 1$ removing the values $-128$ to $127$.
+% In SFJ, if we were to type $42$ as \textbf{int}, this would also mean that $42$ can be typed by \textbf{int and} (\textbf{not byte}), however, the value itself would not be in the set of values of the type, thus breaking our semantic type system.
+% Therefore, our design choice is to type constants with the most restrictive type, i.e., the type representing the smallest set of values containing the value itself.
+% With this typing rule, the value $42$ has type \textbf{byte}, and as a consequence it cannot have type \textbf{int and} (\textbf{not byte}).
 
-\subsection{Polygons: Continued}
+Finally, the subtyping relation generated by \autoref{alg:subtyping} is used in the type system for the SFJ language via a subsumption typing rule:
+ $$
+ \frac{\Gamma\vdash e:\alpha_1 \qquad \alpha_1 \leq \alpha_2}{\Gamma \vdash e:\alpha_2}
+ $$
+We read this typing rule as follows:
+if an expression $e$ is of type $\alpha_1$ under a typing context $\Gamma$ (details of a typing context are irrelevant here) and type $\alpha$ is a subtype $\leq$ of $\alpha_2$, then expression $e$ can be typed with $\alpha_2$.
+
+\subsection{Polygons: continued}
 

### 2020-04-24 conference/camera ready/background.tex
no need for typing
on polygons

@@
 $$
-
 The $\alpha$-types are used to type fields and the $\mu$-types are used to type methods.
@@
 \end{align*}
-
 We assume an infinite set of names, with some special names:
@@
 Notice that in our theoretical development we use unary methods, without loss of generality: tuples of arguments can be modelled by an object that instantiates a ``special'' class containing as fields all the needed arguments.
-Expressions $e$ include variables, constants, field accesses, method invocations and object creations.
+Expressions $e$ include variables, constants, field access, method call and object creation.
 

### 2020-04-24 conference/camera ready/paper.tex
on polygons

@@
     In Java this could be implemented in the following ways:
-
     \begin{equation*}
@@
     \end{equation*}
-
     Or by means of an interface \emph{Diagonal}:
-
     \begin{equation*}
@@
             \\
-            \hspace{4.5cm} \ldots
+            \hspace{2.5cm} // \text{ other geometric figures} \ldots
         \end{array}
     \end{equation*}
-
     Now, suppose our class hierarchy is such that \emph{Polygon} is the parent class and all other geometric figures extend \emph{Polygon}, which is how one would naturally define the set of polygons.
@@
     We propose a more elegant solution, by combining boolean types and semantic subtyping, where only static typechecking is required and we implement this in SFJ \cite{UD20}: it is enough to define a method \emph{diagonal} that has an argument of type \emph{Polygon} \textbf{and not} \emph{Triangle}, thus allowing the typechecker to check at compile time the restrictions on types:
-
     \begin{equation*}
@@
     \end{equation*}
-
     We can now call \emph{diagonal} on an argument of type \emph{Polygon}: if the polygon is \textbf{not} a \emph{Triangle}, then the method computes and returns the length of its longest diagonal; otherwise, there will be a type error at compile time.
@@
 \label{sec:design}
-\subsection{Valid type definitions}
+\subsection{On Valid Type Definitions}
 Since we want to use types $\tau$ in practice in SFJ, we restrict them to \emph{finite trees} whose leaves are basic types $\btypes$ \autoref{sec:terms} with no cycles.
@@
 
-\subsection{Building semantic subtyping for SFJ}
+\subsection{Building Semantic Subtyping for SFJ}
 
@@
 Leveraging the interpretation of types as sets of values to define semantic subtyping for FJ \cite{Dardha2013,Dardha2017}, in SFJ we keep track of the semantic subtyping relation by defining a map from a type to the set of its subtypes, satisfying the property that the set of values of a subtype is included in the set of values of the type.
-We start with basic types and map $Universe$ to the set of all types.
-The full mapping for basic types is defined as follows:
-
+We start with basic types and let $Universe$ be a supertype of all types.
+The full mapping for basic types is defined in \autoref{eq:initialSubtype}.
 \begin{equation}
@@
 \end{equation}
-
 Note that \emph{Int} is not a subtype of \emph{Float} as a 32-bit \emph{float} cannot represent the whole set of 32-bit \emph{integer} values accurately and therefore \emph{Int} is not fully set-contained in \emph{Float}, however this is not the case for \emph{Int} and \emph{Double}.
@@
 
-\subsection{Type system for SFJ}
+\subsection{Type System for SFJ}
 
@@
 
-\subsection{Polygons: continued}
+\subsection{Polygons: Continued}
 
-Let us illustrate the subtyping algorithm with our \emph{Polygons} \autoref{polygons}.
-The algorithm generates the following subtyping relation, which also includes the subtyping relation for basic types defined in the first step.
+Let us illustrate the semantic subtyping algorithm on our \emph{Polygons} given in \autoref{polygons}.
+\autoref{alg:subtyping} generates the subtyping relation given in \autoref{eq:sugtypePolygons}, together with the subtyping relation for basic types, omitted here and defined in \autoref{eq:initialSubtype}.
+Notice that the mapping for $Universe$ is extended with the new types for polygons.
+\begin{equation}
+    \label{eq:sugtypePolygons}
+    \begin{array}{llllll}
+         & Polygon  & = \left\{Polygon, Triangle, Square, Rhombus \right\} &  & Triangle & = \left\{Triangle\right\}           \\
+         & Square   & = \left\{Square\right\}                              &  & Rhombus  & = \left\{Rhombus\right\}            \\
+         & Universe & = \{Double, Float, Long, Int, Short, Byte                                                              &  & Diagonal & = \left\{Diagonal\right\}          \\
+         &          & \quad Boolean, Void, Polygon, Square                                                                     \\
+         &          & \quad Square, Rhombus, Diagonal\}
+    \end{array}
+\end{equation}
+Recall the method \emph{diagonal} in class \emph{Diagonal}, with signature
+$$
+\mathbf{double}\ diagonal((\textit{Polygon} \textbf{ and not }  \textit{Triangle}) \ shape)
+$$
+The result of the set operation on its parameter type gives the following set of polygons:
+$$
+Polygon \textbf{ and not } Triangle = \left\{Polygon, Square, Rhombus \right\}
+$$
+In order to define the \textbf{not} $Triangle$ type we need the $Universe$ type so that we can define it as $Universe\ \mathbf{\setminus}\ Triangle$.
+Then, the \textbf{and} connective is the intersection of sets of $Polygon$ with \textbf{not} $Triangle$.
 
-\begin{flalign*}
-    &Polygon = \left\{Polygon, Triangle, Square, Rhombus \right\} \qquad Triangle = \left\{Triangle\right\}\\
-    &Square = \left\{Square\right\} \qquad  Rhombus  = \left\{Rhombus\right\} \qquad
-    Diagonal = \left\{Diagonal\right\}
-\end{flalign*}
+Thus, if we pass to \emph{diagonal} an argument of type \emph{Triangle}, we get a typechecking error at compile time, as illustrated in the accompanying video \cite{UD20}.
 
-Recall method \emph{diagonal} in class \emph{Diagonal}, we can see that the result of the set operation on its parameter type gives the following set of polygons:
-
-\begin{flalign*}
-    Polygon \textbf{ and not } Triangle = \left\{Polygon, Square, Rhombus \right\}
-\end{flalign*}
-
-If we pass to \emph{diagonal} an argument of type \emph{Triangle}, as it is not contained in the computed set of subtypes, we get a typechecking error at compile time \cite{UD20}.
-
-%It finds all nominal subtypes as they inherit all fields and methods of their super class, so are always guaranteed to be a superset.
-%Similarly, given all pairs of types, the algorithm checks if one is a superset of the other, i.e.,\ they share common fields and methods, and if so, they are related by structural subtyping.
-%For example, type $empty = []$, will have all classes as structural subtypes.
-%The above algorithm will without differentiating, find all nominal and structural subtypes as we always consider all other classes in the program to be potential subtype candidates.
-
-\subsection{Nominal vs Structural}
+\subsection{Nominal vs. Structural Subtyping}
 

### 2020-04-24 conference/camera ready/paper.tex
more on polygons

@@
 \subsection{Type System for SFJ}
+\label{sec:typing}
 
@@
 
-Thus, if we pass to \emph{diagonal} an argument of type \emph{Triangle}, we get a typechecking error at compile time, as illustrated in the accompanying video \cite{UD20}.
+If we write in our SFJ program the following expression:
+$$
+    (\textbf{new}\ Diagonal()).diagonal(\textbf{new}\ Square())
+$$
+the argument \textbf{new} $Square()$ of the $diagonal$ method is of type $Square$, by the type system in \autoref{sec:typing} and $Square$ is contained in the set of the parameter type of the method, so this expression will successfully type checks.
+
+However, if we write the following SFJ expression:
+$$
+    (\textbf{new}\ Diagonal()).diagonal(\textbf{new}\ Triangle())
+$$
+Type $Triangle$ is not contained in $\left\{Polygon, Square, Rhombus \right\}$, therefore this expression will not type check and will return a type error at compile time.
+
+This is further illustrated in the accompanying video of this paper \cite{UD20}.
 

### 2020-04-24 conference/camera ready/paper.tex
nominal vs structural

@@
 
-Note that both approaches to subtyping have their drawbacks.
-%However, it can be argued that structural subtyping is counterproductive in writing logically correct programs.
-Consider two structurally equivalent class/record types $coordinate = [x:int, y:int, z:int]$ and $colour = [x:int, y:int, z:int]$.
-While structurally they can be used interchangeably in our type system, their ``meaning'' is completely different.
-Therefore, we lose some safety in our language as it is unlikely we would like to use a colour where we expect a coordinate, but in the eyes of the type-checker, it is correct.
+In this section we will comment on pros and cons of nominal vs. structural subtyping.
 
-Dardha \emph{et al.}\cite{Dardha2017} suggest a possible extension to the language by including a $nominal$ keyword that can be used in the class signature to indicate that this class should only be considered by the nominal component of the type system.
-This would allow us finer control over the type system and eliminate the previously mentioned problems.
-However, being optional, we argue this would create more bugs than problems it would solve, due to easily forgetting the keyword.
-For example with destructors in C++ classes,leaving out the virtual keyword of the parent class destructor in certain situations can lead to memory leaks.
-Instead a \emph{structural} keyword could be introduced as an extension to the language which specifies that this class is allowed to have structural subtypes.
-A class with this keyword is not added to the subtyping relation of other structurally similar classes which are not marked \emph{structural} as this would weaken the invariant conditions of that class.
+Structural subtyping allows for more flexibility in defining this relation and the user does not need to explicitly definite it, as would do with nominal subtyping.
+However, for this flexibility we might need to pay in meaning.
+For example, consider the following two structurally equivalent classes, hence record types
+$coordinate = [x:int, y:int, z:int]$ and $colour = [x:int, y:int, z:int]$.
+While they can be used interchangeably in a type system using structural subtyping, their ``meaning'' is different and we might want to prohibit it, because intuitively speaking we do not want to use a $colour$ where a $coordinate$ is expected.
 
-On the other hand, with nominal subtyping in Java one can define an overridden method to perform the opposite logic to what the super class is expecting.
-Therefore, both approaches leave an expectation on the developer to ensure what they are doing is correct, hence the integration of both in SFJ is the key to overcoming these drawbacks, as one can choose which approach to use for a given task.
+On the other hand, while nominal subtyping can avoid the above problem, it can introduce others and in particular, a developer can define an overridden method to perform the opposite logic to what the super class is expecting, as illustrated by the following classes in Java:
+\begin{equation*}
+    \begin{array}{ll}
+        \syndecl{A}{\mathit{Object}} {
+         &
+            \qquad
+            \syndecl{B}{A} {
+                \\
+               \qquad \ldots \qquad
+         &
+                \qquad\qquad \ldots
+                \\
+                \quad \textbf{int}\ n;\qquad
+         &
+                \qquad\quad \textbf{int}\ length()\{\ \synret{-n};\ \}
+                \\
+                \quad \textbf{int}\ length()\{\ \synret{n};\ \}\qquad
+         &
+                \qquad}
+                \\   }
+    \end{array}
+\end{equation*}
+Both approaches have their pros and cons, and they leave an expectation on the developer to use the logic behind subtyping correctly when writing code.
+Hence, the integration of both approaches in SFJ makes it possible to  overcome these drawbacks, as one can choose on which subtyping to focus for a given task.
+
+% Dardha \emph{et al.}\cite{Dardha2017} suggest a possible extension to the language by including a $nominal$ keyword that can be used in the class signature to indicate that this class should only be considered by the nominal component of the type system.
+% This would allow us finer control over the type system and eliminate the previously mentioned problems.
+% However, being optional, we argue this would create more bugs than problems it would solve, due to easily forgetting the keyword.
+% For example with destructors in C++ classes,leaving out the virtual keyword of the parent class destructor in certain situations can lead to memory leaks.
+% Instead a \emph{structural} keyword could be introduced as an extension to the language which specifies that this class is allowed to have structural subtypes.
+% A class with this keyword is not added to the subtyping relation of other structurally similar classes which are not marked \emph{structural} as this would weaken the invariant conditions of that class.
 

### 2020-04-24 conference/camera ready/background.tex
methods and method types

@@
 
-\paragraph{On multimethods}
-Since FJ is a core language, some features of the full Java are removed, such as overloading methods.
-In our framework, by leveraging the expressivity of boolean connectives and semantic subtyping, we are able to restore overloading, among other features \cite[\S 8.4]{Dardha2017}.
-We can thus model \emph{multimethods}, \cite{BC97}, which according to the authors is ``\emph{very clean and easy to understand [...] it would be the best solution for a brand new language}".
-As an example, taken from Dardha \emph{et al.}  \cite{Dardha2013,Dardha2017}, consider the following class declarations:
-$$
-\begin{array}{ll}
-\syndecl{A}{\mathit{Object}} {
-&
-\qquad
-\syndecl{B}{A} {
-\\
-	%\qquad\quad \ldots
-%&
-	%\qquad\qquad\quad \ldots
-%\\
-	\qquad\quad \mathbf{int}\ \mathit{length}\ (\mathbf{string}\ s)\{\ \ldots\ \}\qquad
-&
-	\qquad\qquad\quad \mathbf{int}\ \mathit{length}\ (\mathbf{int}\ n)\{\ \ldots\ \}
-\\
-}
-&
-\qquad }
-\end{array}
-$$
-Method {\it length} has type $\mathbf{string} \rightarrow \mathbf{int}$ in class $A$.
-However, because class $B$ extends class $A$,  {\it length} has type $(\mathbf{string} \rightarrow \mathbf{int})\ \synwedge\ (\mathbf{int} \rightarrow \mathbf{int})$ in class $B$,
-which can be simplified to $(\mathbf{string}\ \synvee\ \mathbf{int}) \rightarrow \mathbf{int}$.
-
 Following FJ \cite{featherweight}, we rule out ill-formed programs, such as declaring a constructor named $B$ within a class named $A$; or multiple fields or methods having the same name; or fields having the same type as the type of the class they are defined in.

### 2020-04-24 conference/camera ready/paper.tex
methods and method types

@@
             \\
-            \hspace{2.5cm} // \text{ other geometric figures} \ldots
+            \hspace{2.5cm} // \text{ other polygons } \ldots
         \end{array}
@@
 \paragraph{Structure of the paper:}
-\autoref{sec:background} details the technical developments behind semantic subtyping and its properties.
-\autoref{sec:design} details the design and implementation of SFJ.
-\autoref{sec:conclusion} concludes the paper.
+In \autoref{sec:background} we present the types and terms of the SFJ language.
+In \autoref{sec:design} we present the design and implementation of SFJ; we discuss our two main algorithms, \autoref{alg:types} in \autoref{sec:valid} which checks the validity of type definitions, and \autoref{alg:subtyping} in \autoref{sec:subtypingAlgo} which generates the semantic subtyping relation.
+Further, we discuss typing in SFJ in \autoref{sec:typing}; nominal vs.
+structural subtyping in \autoref{sec:nominal}; and method types in \autoref{sec:methods}.
+We discuss future work, related work and conclude the paper in \autoref{sec:conclusion}.
 
@@
 
-\section{SFJ: Design and Implementation}
+\section{The SFJ Language}
 \label{sec:design}
 \subsection{On Valid Type Definitions}
+\label{sec:valid}
 Since we want to use types $\tau$ in practice in SFJ, we restrict them to \emph{finite trees} whose leaves are basic types $\btypes$ \autoref{sec:terms} with no cycles.
@@
 \subsection{Building Semantic Subtyping for SFJ}
-
+\label{sec:subtypingAlgo}
 If \autoref{alg:types} returns $True$, meaning all type definitions in a program are valid, we can then build the semantic subtyping.
@@
 \subsection{Nominal vs. Structural Subtyping}
-
+\label{sec:nominal}
 In this section we will comment on pros and cons of nominal vs. structural subtyping.
@@
                 \\
-               \qquad \ldots \qquad
+               \qquad \qquad\ldots \qquad
          &
-                \qquad\qquad \ldots
+                \qquad\qquad\qquad \ldots
                 \\
@@
 
-\subsection{Method Types}
+\subsection{Methods in SFJ}
+\label{sec:methods}
+\paragraph{On multimethods}
+Since FJ is a core language, some features of the full Java are removed, such as overloading methods.
+In our framework, by leveraging the expressivity of boolean connectives and semantic subtyping, we are able to restore overloading, among other features \cite[\S 8.4]{Dardha2017}.
+We can thus model \emph{multimethods}, \cite{BC97}, which according to the authors is ``\emph{very clean and easy to understand [...] it would be the best solution for a brand new language}".
+As an example, taken from Dardha \emph{et al.}  \cite{Dardha2013,Dardha2017}, consider the following class declarations:
+$$
+\begin{array}{ll}
+\syndecl{A}{\mathit{Object}} {
+&
+\qquad
+\syndecl{B}{A} {
+\\
+	%\qquad\quad \ldots
+%&
+	%\qquad\qquad\quad \ldots
+%\\
+	\qquad\quad \mathbf{int}\ \mathit{length}\ (\mathbf{string}\ s)\{\ \ldots\ \}\qquad
+&
+	\qquad\qquad\quad \mathbf{int}\ \mathit{length}\ (\mathbf{int}\ n)\{\ \ldots\ \}
+\\
+}
+&
+\qquad }
+\end{array}
+$$
+Method {\it length} has type $\mathbf{string} \rightarrow \mathbf{int}$ in class $A$.
+However, because class $B$ extends class $A$,  {\it length} has type $(\mathbf{string} \rightarrow \mathbf{int})\ \synwedge\ (\mathbf{int} \rightarrow \mathbf{int})$ in class $B$,
+which can be simplified to $(\mathbf{string}\ \synvee\ \mathbf{int}) \rightarrow \mathbf{int}$.
 
-Furthermore, we have extended FJ with \emph{method types}, demonstrated below to show another possible implementation of the \emph{Polygons} \autoref{polygons}.
+\paragraph{Method types}
+Let us illustrate the method types given in \autoref{sec:types} via an alternative implementation of the class $Diagonal$ at the end of \autoref{polygons}.
 $$
+\begin{array}{l}
     \synclass Diagonal \ \{
-    \mathbf{int}\ diagonal(\textit{(getDiagonal: void$\rightarrow$int)}  \ shape)\{\ldots\}
+    \\
+    \qquad\qquad\qquad \ldots \\
+    \qquad \textbf{double}\ diagonal((diagonal: \textbf{void} \rightarrow \textbf{double})\ shape)\\
+    \hfill
+    \{\ \synret\ shape.diagonal();\ \}
+    \\
     \}
+\end{array}
 $$
+We define the type of the (outside) $diagonal$ method as accepting any type and its subtypes implementing the (inside) \emph{diagonal} method with type signature \textbf{void} to \textbf{double}.
 
-We define the method type as accepting any type and its subtypes which implement a method called \emph{diagonal} which is a method from the type \emph{Void} to a \emph{Double}.
+In order to type check an argument passed to the (outside) $diagonal$ method, at compile time we build a collection of types $\{type_{1},\ type_{2},\ldots\}$ which are class types where the (inside) $diagonal$ method is defined.
+As such, we iterate over the list of \emph{classes} in an SFJ program (as we did in \autoref{alg:subtyping})) to check for the required method.
+The resulting collection of types is the union of all classes where $diagonal$ is defined together with their subtypes $([\![type_{1}]\!] \cup [\![type_{2}]\!] \cup \ldots)$, where each $[\![type_{i}]\!]$ denotes a mapping of $type_i$ to the set of its subtypes, similar to \autoref{eq:sugtypePolygons}.
 
-In order to type-check the value passed to such an argument, we can at compile-time build a collection of types $\{type_{1},\ type_{2},\ \ldots\}$ which have this method.
-We can do this the simply iterating over the same list of classes in the program as when we were creating our subtyping relation and checking for the presence of the required method.
-The set of accepted types would then be the union of all their corresponding subtypes $([\![type_{1}]\!]_{B} \cup [\![type_{2}]\!]_{B} \cup\ \ldots)$.
+However, calculating this collection of types for each method of every class would be computationally inefficient and most importantly unnecessary as only few methods would in turn be used as method types.
+Therefore we only compute them on demand during typechecking when we come across such a type.
 
-However, calculating this collection of types for each method of every class would be computationally unnecessary as only few of these would be used for method types.
-Therefore we only compute them on demand during type-checking when we come across such a type.
-
-We can therefore use method types to statically include or exclude a portion of our type hierarchy.
-However unlike the use of interfaces such as in the suggested Java solution to the Polygons problem \autoref{polygons}, the values that can be accepted by a method type do not have to be related to each other in any way in the class hierarchy.
-This becomes increasingly useful if we have a legacy system as we can still accept all the classes that have defined \emph{Diagonal} methods without having to go back and add interface implementations.
+We can therefore use method types to statically include or exclude a portion of our class hierarchy.
+However, unlike the use of interfaces, as in one of the proposed Java approaches in \autoref{polygons}, the values that can be accepted by a method type do not have to be related to each other in any way in the class hierarchy.
+This indeed is useful if we are dealing with legacy code as we can still accept all classes where \emph{diagonal} is defined, without having to go back and add interface implementations.
 
@@
 %Improper use of this allows for run-time exceptions however it more closely follows the idea of giving the programmer tools to create more expressive programs at the cost of checking program correctness such as with the use of structural subtypes.
-
 \section{Related Work and Conclusion}

### 2020-04-24 conference/camera ready/paper.tex
type-checking

@@
     We present SFJ---Semantic Featherweight Java, an implementation of FJ which features boolean types, semantic subtyping and integrates nominal as well as structural subtyping.
-    The benefits of SFJ, illustrated in the paper and the accompanying video (with audio/subtitles) \cite{UD20}, show how static typechecking of boolean types and semantic subtyping gives higher guarantees of program correctness, more flexibility and compactness of program writing.
+    The benefits of SFJ, illustrated in the paper and the accompanying video (with audio/subtitles) \cite{UD20}, show how static type-checking of boolean types and semantic subtyping gives higher guarantees of program correctness, more flexibility and compactness of program writing.
     \keywords{Nominal subtyping \and Structural subtyping  \and Semantic Featherweight Java \and Object-oriented languages \and Boolean types \and Type theory.}
@@
 
-    We propose a more elegant solution, by combining boolean types and semantic subtyping, where only static typechecking is required and we implement this in SFJ \cite{UD20}: it is enough to define a method \emph{diagonal} that has an argument of type \emph{Polygon} \textbf{and not} \emph{Triangle}, thus allowing the typechecker to check at compile time the restrictions on types:
+    We propose a more elegant solution, by combining boolean types and semantic subtyping, where only static type-checking is required and we implement this in SFJ \cite{UD20}: it is enough to define a method \emph{diagonal} that has an argument of type \emph{Polygon} \textbf{and not} \emph{Triangle}, thus allowing the type-checker to check at compile time the restrictions on types:
     \begin{equation*}
@@
 $$
-the argument \textbf{new} $Square()$ of the $diagonal$ method is of type $Square$, by the type system in \autoref{sec:typing} and $Square$ is contained in the set of the parameter type of the method, so this expression will successfully type checks.
+the argument \textbf{new} $Square()$ of the $diagonal$ method is of type $Square$, by the type system in \autoref{sec:typing} and $Square$ is contained in the set of the parameter type of the method, so this expression will successfully type-checks.
 
@@
 $$
-Type $Triangle$ is not contained in $\left\{Polygon, Square, Rhombus \right\}$, therefore this expression will not type check and will return a type error at compile time.
+Type $Triangle$ is not contained in $\left\{Polygon, Square, Rhombus \right\}$, therefore this expression will not type-check and will return a type error at compile time.
 
@@
 However, calculating this collection of types for each method of every class would be computationally inefficient and most importantly unnecessary as only few methods would in turn be used as method types.
-Therefore we only compute them on demand during typechecking when we come across such a type.
+Therefore we only compute them on demand during type-checking when we come across such a type.
 
@@
 %Improper use of this allows for run-time exceptions however it more closely follows the idea of giving the programmer tools to create more expressive programs at the cost of checking program correctness such as with the use of structural subtypes.
-\section{Related Work and Conclusion}
+\section{Conclusion, Related Work and Future Work}
 \label{sec:conclusion}

### 2020-04-24 conference/camera ready/paper.tex
add code generation section

@@
 %Improper use of this allows for run-time exceptions however it more closely follows the idea of giving the programmer tools to create more expressive programs at the cost of checking program correctness such as with the use of structural subtypes.
+
+\section{Code generation}
+
+So far, SFJ only includes the typechecking component of the language. We provide a ``proof'' of the code generation algorithm for areas of future work.
+
+Given the similarity of SFJ to Java, we felt that the usage of the Java Virtual Machine (JVM) \cite{lindholm2013} would be most suitable for code generation by transpiling SFJ programs into JVM bytecode.
+This is the approach used by object-oriented languages such as Kotlin amongst others.
+
+The key challenge in implementing code generation compared to other languages is that a single field in SFJ with type $(int\ \textbf{or}\ bool)$ actually represents two JVM fields, one of type $int$ and another of type $bool$, of which only one is inhabited with a value.
+
+Therefore to reduce the amount of alternatives for which we would need to generate code, we first analyse our program and reduce the boolean types by keeping only the alternatives which actually get used in the program.
+For example, if the field of type $int\ or\ bool$ only ever gets initialised with a boolean value, we can reduce it and make it a field of a single type $bool$.
+Similarly the numerical basic types can also be reduced to their largest type without consequence, such as $int\ \textbf{and not}\ byte$ can be reduced to just $int$ as we have already type checked all literal values passed to such a type.
+
+After reducing the boolean types, we have to expand all boolean types that still remain by considering the result of their set operation.
+Staying with the example of a field named $f1$ of type $int\ or\ bool$, we would define in our class two fields $int\_f1$ and $bool\_f1$ with types $int$ and $boolean$ respectively.
+In order to initialise these field, we use the constructor overloading capabilities of the JVM to generate an overloaded version for each alternative type.
+In each constructor, only the alternative field that matches the type of the parameter in this constructor is initialised with a value and all other alternatives are set to $null$.
+
+For methods, we also use the overloading capabilities of the JVM to define an  overloaded method for each type in the expanded method parameter, all with the same method body.
+Depending on which alternative the argument inhabits at run-time, a different method will be dispatched to.
+
+To access the field of an object, we generate code that checks each alternative of the field if it is non-null and includes the rest of the code generation for the expression for each branch.
+At runtime, only one branch will be true, and this is the branch of generated code which will be executed.
+
+For structural subtyping, we create an new nominal subtype with the same implementation for each structural subtype a type has, so that it is accepted as a proper subtype in the JVM.
+
+Like methods with boolean types, we similarly implement methods with method types by defining an overloaded alternative for each type that implements the specified method.
+
+We have now defined the code generation for each expression in our syntax of terms and for field, constructor and method declarations, so therefore have defined how
+
 \section{Conclusion, Related Work and Future Work}

### 2020-04-24 conference/camera ready/paper.tex
more on related work

--- 2020-04-24 conference/camera ready/paper.tex
@@
 Syntactic subtyping \cite{Liskov1994} is more mainstream in programming languages and is defined by means of a set of formal deductive subtyping rules.
-Semantic subtyping \cite{Castagna2005,Cas05,Frisch2008} is more recent and less known:  one defines a formal model of the language and an interpretation of types as subsets of this model.
+Semantic subtyping \cite{Aiken,Damm} is more recent and less known:  one defines a formal model of the language and an interpretation of types as subsets of this model.
 Then, subtyping is defined as set inclusion of subsets denoting types.
@@
 
-\paragraph{Structure of the paper:}
+\subsubsection{Structure of the paper:}
 In \autoref{sec:background} we present the types and terms of the SFJ language.
@@
 \label{sec:methods}
-\paragraph{On multimethods}
+\subsubsection{On multimethods}
 Since FJ is a core language, some features of the full Java are removed, such as overloading methods.
@@
 
-\paragraph{Method types}
+\subsubsection{Method types}
 Let us illustrate the method types given in \autoref{sec:types} via an alternative implementation of the class $Diagonal$ at the end of \autoref{polygons}.
@@
 %Improper use of this allows for run-time exceptions however it more closely follows the idea of giving the programmer tools to create more expressive programs at the cost of checking program correctness such as with the use of structural subtypes.
-\section{Conclusion, Related Work and Future Work}
+
+\section{Conclusion, Related and Future Work}
 \label{sec:conclusion}
 
-Semantic subtyping approach goes back to more than two decades ago \cite{Aiken,Damm}.
-Notable lines of work include:
-Hosoya and Pierce \cite{XML1,XML2,XML3} who define XDuce, an XML-oriented language designed specifically to transform XML documents in other XML documents satisfying certain properties.
-Castagna \emph{et al.} \cite{Castagna2005,Cas05,FCB08} extend XDuce with first-class functions and arrow types and implement it as $\mathbb{C}$Duce.
-The starting point of their framework is a higher-order $\lambda-$calculus with pairs and projections.
-Muehlboeck and Tate \cite{Muehlboeck2018} define a syntactic framework with boolean connectives which has been implemented in the Ceylon programming language \cite{Ceylon2016}.
-Castagna \emph{et al.} \cite{Cpi} define $\mathbb{C}\pi$, a variant of the asynchronous $\pi-$calculus \cite{sangiorgi}, where channel types are augmented with boolean connectives.
-Ancona and Lagorio~\cite{AL10} define subtyping for infinite types coinductively by using union and object type constructors, where types are interpreted as sets of value of the language.
+In this paper we presented the design and implementation of SFJ---Semantic Featherweight Java, an extension of Featherweight Java featuring boolean types, semantic subtyping, and integrating both nominal and structural subtyping.
+Due to the expressivity of semantic subtyping, in SFJ we are able to restore standard Java constructs and features for example, lists and overloading meathods, which were not present in FJ, thus making SFJ a more complete language.
+We presented \autoref{alg:types} on validity of type definitions and
+\autoref{alg:subtyping} on semantic subtyping, which finds all nominal and structural subtypes for all types in an SFJ program.
+We also described typing of terms in SFJ, which follows that of Java and builds upon relevant work \cite{Dardha2013,Dardha2017}.
+
+When it comes to executing an SFJ program, our approach is to translate an SFJ program into Java bytecode and then run it on the Java Virtual Machine (JVM).
+This is a standard approach used also by other object-oriented languages, for example, Kotlin\footnote{https://kotlinlang.org/}.
+The main challenge in translating SFJ to bytecode is in translating boolean connectives.
+For example, a field typed by \textbf{string or int}, will be translated as two Java fields, one of type \textbf{string} and one of type \textbf{int}, and at any given time, only one of the two types will be inhabited by a value.
+We have fully detailed the strategy for the code generation phase of SFJ using ANTLR \cite{parr2013}, however this is work in progress.
+
+Semantic subtyping goes back to more than two decades ago \cite{Aiken,Damm}.
+%
+Hosoya and Pierce \cite{XML1,XML2,XML3} define XDuce, an XML-oriented language designed specifically to transform XML documents in other XML documents satisfying certain properties.
+%
+Frisch \emph{et al.} \cite{FCB08} extend XDuce by introducing less XML specific types such as records, boolean connectives and arrow types, and implement it as $\mathbb{C}$Duce.
+Their work is similar to ours in that our class-based semantic type system is a combination of the $\mathbb{C}$Duce record types with arrow types.
+%
+Castagna \emph{et al.} define $\mathbb{C}\pi$ \cite{Cpi}, a variant of the asynchronous $\pi$-calculus, where channel types are augmented with boolean connectives;
+semantic subtyping for ML-like languages \cite{CastagnaP016} and semantic subtyping in a gradual typing framework \cite{CastagnaL17}.
+%
+Ancona and Lagorio~\cite{AL10} define subtyping for infinite types by using union and object type constructors, where types are interpreted as sets of values of the language.
+%
 Bonsangue \emph{et al.}~\cite{BRABR14} study a coalgebraic approach to coinductive types and define a set-theoretic interpretation of coinductive types with union types.
+%
 Pearce \cite{Pearce19} defines semantic subtyping for rewriting rules in the Whiley Rewrite Language and for a flow-typing calculus  \cite{Pearce13}.
 
-To conclude, in this paper we presented the design and implementation of SFJ, an extension of Featherweight Java with boolean types and semantic subtyping, which allow for both structural and nominal subtyping in FJ as well as restore standard Java constructs for example, lists and features for example, overloading which were not present in FJ, making SFJ a more complete language.
+On the implementation side, in addition to $\mathbb{C}$Duce, to the best of our knowledge there are few work who have developed tools.
+%
+Muehlboeck and Tate \cite{Muehlboeck2018} define a syntactic framework with boolean connectives which has been implemented in the Ceylon programming language \cite{Ceylon2016}.
+%
+Ancona and Corradi \cite{AnconaC16} define semantic subtyping for an imperative object-oriented language with mutable fields.
+In our framework we are considering only the functional fragment of Java, which is FJ, and as a result the semantic subtyping framework is simpler.
+The authors also propose a prototype implementation of their subtyping algorithm.
+%
+Chaudhuri \emph{et al.} \cite{ChaudhuriVGRL17} present the design and implementation of \textsc{Flow}, which is a type checker for JavaScript.
+They use boolean connectives \textbf{and}, \textbf{not}, \textbf{or} for their predicates, however they do not define semantic subtyping for their language.
 
 %% References
-\newpage
 \bibliographystyle{splncs04}

### 2020-04-24 conference/camera ready/paper.tex

@@
 
-\section{Conclusion, Related and Future Work}
+\section{Code generation}
+
+So far, SFJ only includes the typechecking component of the language. We provide a ``proof'' of the code generation algorithm for areas of future work.
+
+Given the similarity of SFJ to Java, we felt that the usage of the Java Virtual Machine (JVM) \cite{lindholm2013} would be most suitable for code generation by transpiling SFJ programs into JVM bytecode.
+This is the approach used by object-oriented languages such as Kotlin amongst others.
+
+The key challenge in implementing code generation compared to other languages is that a single field in SFJ with type $(int\ \textbf{or}\ bool)$ actually represents two JVM fields, one of type $int$ and another of type $bool$, of which only one is inhabited with a value.
+
+Therefore to reduce the amount of alternatives for which we would need to generate code, we first analyse our program and reduce the boolean types by keeping only the alternatives which actually get used in the program.
+For example, if the field of type $int\ or\ bool$ only ever gets initialised with a boolean value, we can reduce it and make it a field of a single type $bool$.
+Similarly the numerical basic types can also be reduced to their largest type without consequence, such as $int\ \textbf{and not}\ byte$ can be reduced to just $int$ as we have already type checked all literal values passed to such a type.
+
+After reducing the boolean types, we have to expand all boolean types that still remain by considering the result of their set operation.
+Staying with the example of a field named $f1$ of type $int\ or\ bool$, we would define in our class two fields $int\_f1$ and $bool\_f1$ with types $int$ and $boolean$ respectively.
+In order to initialise these field, we use the constructor overloading capabilities of the JVM to generate an overloaded version for each alternative type.
+In each constructor, only the alternative field that matches the type of the parameter in this constructor is initialised with a value and all other alternatives are set to $null$.
+
+For methods, we also use the overloading capabilities of the JVM to define an  overloaded method for each type in the expanded method parameter, all with the same method body.
+Depending on which alternative the argument inhabits at run-time, a different method will be dispatched to.
+
+To access the field of an object, we generate code that checks each alternative of the field if it is non-null and includes the rest of the code generation for the expression for each branch.
+At runtime, only one branch will be true, and this is the branch of generated code which will be executed.
+
+For structural subtyping, we create an new nominal subtype with the same implementation for each structural subtype a type has, so that it is accepted as a proper subtype in the JVM.
+
+Like methods with boolean types, we similarly implement methods with method types by defining an overloaded alternative for each type that implements the specified method.
+
+We have now defined the code generation for each expression in our syntax of terms and for field, constructor and method declarations, so therefore have defined how
+
+\section{Conclusion, Related Work and Future Work}
 \label{sec:conclusion}

### 2020-04-24 conference/camera ready/paper.tex
polishing code generation

--- 2020-04-24 conference/camera ready/paper.tex
@@
 
-\section{Conclusion, Related and Future Work}
+\subsection{Code Generation}
+\label{sec:code_generation}
+%%
+SFJ only includes the typechecking component of the language. We provide a ``proof'' of the code generation algorithm, which is work in progress at this stage.
+
+Given the similarity of SFJ to Java, our approach is to translate an SFJ program into Java bytecode and then run it on the Java Virtual Machine (JVM) \cite{lindholm2013}.
+This is a standard approach used also by other object-oriented languages, for example, Kotlin\footnote{https://kotlinlang.org/}.
+
+The main challenge in translating SFJ into bytecode is in translating boolean connectives.
+For example, a field $f1$ of type \textbf{int or bool}, will be translated as two Java fields, one of type \textbf{int} and one of type \textbf{bool}, and at any given time, only one of the two types will be inhabited by a value.
+
+Therefore to reduce the amount of alternatives for which we would need to generate code, we first analyse our program and reduce the boolean types by keeping only the alternatives which actually get used in the program.
+For example, if the field of type \textbf{int or bool} only ever gets initialised with a boolean value, we can reduce it and make it a field of a single type \textbf{bool}.
+%Similarly, basic types can be reduced to their largest type without consequence, such as \textbf{int and not bool} can be reduced to just \textbf{int} as we have already type checked all literal values passed to such a type.
+%
+After this reduction phase of boolean types, we have to expand all other boolean types that still remain by considering the result of their set operation.
+On the example above, consider the field  $f1$ of type \textbf{int or bool}, when translating we need define in our class two fields \textbf{int}$_{f1}$ and \textbf{bool}$\_{f1}$ with the corresponding types.
+In order to initialise these fields, we use the constructor overloading capabilities of the JVM to generate an overloaded version for each alternative type.
+In each constructor, only the alternative field that matches the type of the parameter in this constructor is initialised with a value and all other alternatives are set to $null$.
+%
+For methods, we also use the overloading capabilities of the JVM to define an  overloaded method for each type in the expanded method parameter, all with the same method body.
+Depending on which alternative the argument inhabits at run-time, a different method will be dispatched to.
+%
+To access the field of an object, we generate code that checks each alternative of the field if it is non-null and includes the rest of the code generation for the expression for each branch.
+At runtime, only one branch will be true, and this is the branch of generated code which will be executed.
+
+For structural subtyping, we create an new nominal subtype with the same implementation for each structural subtype a type has, so that it is accepted as a proper subtype in the JVM.
+
+Like methods with boolean types, we similarly implement methods with method types by defining an overloaded alternative for each type that implements the specified method.
+
+This concludes the code generation phase for all expressions in SFJ.
+
+\section{Conclusion}
 \label{sec:conclusion}
@@
 We also described typing of terms in SFJ, which follows that of Java and builds upon relevant work \cite{Dardha2013,Dardha2017}.
-
-When it comes to executing an SFJ program, our approach is to translate an SFJ program into Java bytecode and then run it on the Java Virtual Machine (JVM).
-This is a standard approach used also by other object-oriented languages, for example, Kotlin\footnote{https://kotlinlang.org/}.
-The main challenge in translating SFJ to bytecode is in translating boolean connectives.
-For example, a field typed by \textbf{string or int}, will be translated as two Java fields, one of type \textbf{string} and one of type \textbf{int}, and at any given time, only one of the two types will be inhabited by a value.
-We have fully detailed the strategy for the code generation phase of SFJ using ANTLR \cite{parr2013}, however this is work in progress.
 

### 2020-04-24 conference/camera ready/paper.tex
final trimmings

@@
     Then again, a natural way to implement this in Java is by defining the method \emph{diagonal} in the class \emph{Polygon} and using an \textbf{instanceof}, for example, inside a \textbf{try-catch} construct.
-    Then, an exception would be thrown at runtime, if the argument passed to the method is a triangle.
+    Then, an exception would be thrown at run time, if the argument passed to the method is a triangle.
 
@@
 In \autoref{sec:design} we present the design and implementation of SFJ; we discuss our two main algorithms, \autoref{alg:types} in \autoref{sec:valid} which checks the validity of type definitions, and \autoref{alg:subtyping} in \autoref{sec:subtypingAlgo} which generates the semantic subtyping relation.
-Further, we discuss typing in SFJ in \autoref{sec:typing}; nominal vs.
-structural subtyping in \autoref{sec:nominal}; and method types in \autoref{sec:methods}.
-We discuss future work, related work and conclude the paper in \autoref{sec:conclusion}.
+Further, we discuss typing of SFJ in \autoref{sec:typing}; nominal vs.
+structural subtyping in \autoref{sec:nominal}; method types in \autoref{sec:methods}; and code generation in \autoref{sec:code_generation}.
+We discuss related work and conclude the paper in \autoref{sec:conclusion}.
 
@@
 
-The main challenge in translating SFJ into bytecode is in translating boolean connectives.
-For example, a field $f1$ of type \textbf{int or bool}, will be translated as two Java fields, one of type \textbf{int} and one of type \textbf{bool}, and at any given time, only one of the two types will be inhabited by a value.
+The main challenge in translating SFJ into bytecode is translating types using boolean connectives.
 
-Therefore to reduce the amount of alternatives for which we would need to generate code, we first analyse our program and reduce the boolean types by keeping only the alternatives which actually get used in the program.
+For example, a field $f_1$ of type \textbf{int or bool}, will be translated as two Java fields, one of type \textbf{int} and one of type \textbf{bool}, and only one of the two types will be inhabited by a value.
+%
+In order to achieve this, we first analyse our program and reduce the boolean types by keeping only the alternatives which actually get used in the program.
 For example, if the field of type \textbf{int or bool} only ever gets initialised with a boolean value, we can reduce it and make it a field of a single type \textbf{bool}.
@@
 %
-After this reduction phase of boolean types, we have to expand all other boolean types that still remain by considering the result of their set operation.
-On the example above, consider the field  $f1$ of type \textbf{int or bool}, when translating we need define in our class two fields \textbf{int}$_{f1}$ and \textbf{bool}$\_{f1}$ with the corresponding types.
+After this reduction phase, we then consider the remaining types which use boolean connectives and could not be reduced.
+On the example above, consider again field $f_1$ of type \textbf{int or bool}. This will be translated as two fields \textbf{int}$\_{f_1}$ and \textbf{bool}$\_{f_1}$ with the corresponding types.
 In order to initialise these fields, we use the constructor overloading capabilities of the JVM to generate an overloaded version for each alternative type.
-In each constructor, only the alternative field that matches the type of the parameter in this constructor is initialised with a value and all other alternatives are set to $null$.
-%
-For methods, we also use the overloading capabilities of the JVM to define an  overloaded method for each type in the expanded method parameter, all with the same method body.
-Depending on which alternative the argument inhabits at run-time, a different method will be dispatched to.
+In each constructor, only the field that matches the type of the parameter is initialised with all other fields set to $null$.
 %
 To access the field of an object, we generate code that checks each alternative of the field if it is non-null and includes the rest of the code generation for the expression for each branch.
-At runtime, only one branch will be true, and this is the branch of generated code which will be executed.
+At run time, only one branch will be true, and this is the branch of generated code which will be executed.
 
-For structural subtyping, we create an new nominal subtype with the same implementation for each structural subtype a type has, so that it is accepted as a proper subtype in the JVM.
-
-Like methods with boolean types, we similarly implement methods with method types by defining an overloaded alternative for each type that implements the specified method.
+For methods, we also use the overloading capabilities of the JVM to define an  overloaded method for each type in the expanded method parameter, all with the same method body.
+Depending on which alternative the argument inhabits at run time, a different method will be dispatched to.
+%
+%For structural subtyping, we create an new nominal subtype with the same implementation for each structural subtype a type has, so that it is accepted as a proper subtype in the JVM.
+Like methods with boolean types, we similarly implement methods with method types, as we discussed in \autoref{sec:methods}, by defining an overloaded alternative for each type that implements the specified method.
 

### 2020-04-27 conference/camera ready/background.tex
first pass on paper after revision

@@
 
-Following FJ \cite{featherweight}, we rule out ill-formed programs, such as declaring a constructor named $B$ within a class named $A$; or multiple fields or methods having the same name; or fields having the same type as the type of the class they are defined in.
+Following FJ \cite{featherweight}, we rule out ill-formed programs, such as declaring a constructor named $B$ within a class named $A$, or multiple fields or methods having the same name, or fields having the same type as the class they are defined in.
 

### 2020-04-27 conference/camera ready/paper.tex
first pass on paper after revision

@@
   \renewcommand{\subsectionautorefname}{\S}%
-  \renewcommand{\algorithmautorefname}{Algorithm}%
+  \renewcommand{\algorithmautorefname}{Algo.}%
   %\renewcommand{\exampleautorefname}{Ex.}%
-  \renewcommand{\equationautorefname}{Mapping}%
+  \renewcommand{\equationautorefname}{Map}%
 }
@@
 \label{sec:valid}
-Since we want to use types $\tau$ in practice in SFJ, we restrict them to \emph{finite trees} whose leaves are basic types $\btypes$ \autoref{sec:terms} with no cycles.
+Since we want to use types $\tau$ to program in SFJ, we restrict them to \emph{finite trees} whose leaves are basic types $\btypes$ \autoref{sec:terms} with no cycles.
 For example, a recursive type $\alpha = [a : \alpha]$ denotes an infinite program tree \textbf{new} $C$(\textbf{new} $C$($\cdots{}$)), hence we rule it out as it is not inhabitable.
@@
 We start by defining the grammar of the language in Extended Backus-Naur Form (EBNF), following \autoref{sec:types} and by running ANTLR, we can automatically generate a parser for SFJ and extend it in order to implement the required checks for our types and type system.
+
 Running the parser on an SFJ program returns an abstract syntax tree (AST) of that program.
-
 When visiting the AST, we check if the program is well-formed, following the intuition at the end of \autoref{sec:background}.
-We mark any classes containing fields typed with only basic types as \emph{resolved} otherwise, as \emph{unresolved}.
-Using this information, \autoref{alg:types} checks if the type definitions in a program are valid, namely, if they are finite trees whose leaves are basic types with no cycles.
+We mark any classes containing only fields typed with basic types as \emph{resolved}, otherwise we mark them as \emph{unresolved}.
+Using this information, \autoref{alg:types} checks if the type definitions in a program are valid, namely if they are finite trees whose leaves are basic types with no cycles.
 %
-At each iteration of the algorithm we can resolve at least one type, until all the types in the SFJ program are resolved, and return $True$.
-However, if there is at least one type unresolved, meaning there is a cycle in the type definition, we return $False$.
+The algorithm returns $True$ only if all the types in the program are resolved, otherwise it returns $False$, meaning there is at least one type definition which is invalid and contains a cycle.
 
@@
     \DontPrintSemicolon
-    \Input{$classes$, the set of classes marked $resolved$ if their fields contain only basic types, $unresolved$ otherwise.}
+    \Input{$classes$, the set of classes in an SFJ program marked $resolved$, if their fields contain only basic types, $unresolved$ otherwise.}
     \Output{$True$ if all classes are valid type definitions, $False$ otherwise.}
@@
     \DontPrintSemicolon
-    \Input{$classes$, the set of classes in an SFJ program for which we have not yet defined the subtyping relation.\;
-    $relation$, the mapping of types to the set of subtypes, initially being Mapping \ref{eq:initialSubtype}. \;
+    \Input{$classes$, the set of classes in an SFJ program.\\
+    $relation$, the mapping of types to the set of subtypes, initially Map \ref{eq:initialSubtype}. \;
     }
@@
 It is worth noticing that the subtyping algorithm finds all nominal and structural subtypes of a given type. This is due to the fact that all pairs of types are inspected.
-Recall from \autoref{sec:intro} that nominal subtyping is name-based and given by the class hierarchy defined by the programmer, whether structural subtyping is structure-based and given by the set-inclusion of fields and methods.
-In particular, it is contra-variant with respect to this set-inclusion.
+Recall from \autoref{sec:intro} that nominal subtyping is name-based and given by the class hierarchy defined by the programmer, whereas structural subtyping is structure-based and given by the set-inclusion of fields and methods.
+In particular, structural subtyping is contra-variant with respect to this set-inclusion.
 \autoref{alg:subtyping} finds all structural subtypes of a given class because it checks that its fields and methods are a superset of existing types in $relation$.
@@
 The reason for a quadratic complexity of \autoref{alg:subtyping} is due to the symmetric check of structural subtyping between a $class$ and an $existingClass$ in $relation$.
-Notice that if we were to only work with nominal subtyping, then we only require traversing the class hierarchy once,  which gives an $\mathcal{O}(n)$ complexity.
+Notice that if we were to only work with nominal subtyping, then we would only require traversing the class hierarchy once,  which gives an $\mathcal{O}(n)$ complexity.
 
@@
 Both approaches have their pros and cons, and they leave an expectation on the developer to use the logic behind subtyping correctly when writing code.
-Hence, the integration of both approaches in SFJ makes it possible to  overcome these drawbacks, as one can choose on which subtyping to focus for a given task.
+The integration of both subtyping approaches in SFJ gives the developer the freedom to choose the most suitable subtyping relation to use for a given task.
 
@@
 In order to type check an argument passed to the (outside) $diagonal$ method, at compile time we build a collection of types $\{type_{1},\ type_{2},\ldots\}$ which are class types where the (inside) $diagonal$ method is defined.
-As such, we iterate over the list of \emph{classes} in an SFJ program (as we did in \autoref{alg:subtyping})) to check for the required method.
+As such, we iterate over the list of \emph{classes} in an SFJ program (as we did in \autoref{alg:subtyping}) to check for the required method.
 The resulting collection of types is the union of all classes where $diagonal$ is defined together with their subtypes $([\![type_{1}]\!] \cup [\![type_{2}]\!] \cup \ldots)$, where each $[\![type_{i}]\!]$ denotes a mapping of $type_i$ to the set of its subtypes, similar to \autoref{eq:sugtypePolygons}.
@@
 We can therefore use method types to statically include or exclude a portion of our class hierarchy.
-However, unlike the use of interfaces, as in one of the proposed Java approaches in \autoref{polygons}, the values that can be accepted by a method type do not have to be related to each other in any way in the class hierarchy.
+However, unlike with interfaces as in \autoref{polygons}, the values that can be accepted by a method type do not have to be related to each other in any way in the class hierarchy.
 This indeed is useful if we are dealing with legacy code as we can still accept all classes where \emph{diagonal} is defined, without having to go back and add interface implementations.
@@
 %%
-SFJ only includes the typechecking component of the language. We provide a ``proof'' of the code generation algorithm, which is work in progress at this stage.
+SFJ only includes the typechecking component of the language. In this section, we provide a sketch of the code generation algorithm, which is work in progress.
 
 Given the similarity of SFJ to Java, our approach is to translate an SFJ program into Java bytecode and then run it on the Java Virtual Machine (JVM) \cite{lindholm2013}.
-This is a standard approach used also by other object-oriented languages, for example, Kotlin\footnote{https://kotlinlang.org/}.
+This is a standard approach also used by other object-oriented languages, for example, Kotlin\footnote{https://kotlinlang.org/}.
 
 The main challenge in translating SFJ into bytecode is translating types using boolean connectives.
-
+%
 For example, a field $f_1$ of type \textbf{int or bool}, will be translated as two Java fields, one of type \textbf{int} and one of type \textbf{bool}, and only one of the two types will be inhabited by a value.
@@
 Like methods with boolean types, we similarly implement methods with method types, as we discussed in \autoref{sec:methods}, by defining an overloaded alternative for each type that implements the specified method.
-
-This concludes the code generation phase for all expressions in SFJ.
+%
+This concludes the code generation phase for all expressions in SFJ given in \autoref{sec:terms}.
 
@@
 
-On the implementation side, in addition to $\mathbb{C}$Duce, to the best of our knowledge there are few work who have developed tools.
+Regarding implementations of semantic subtyping, to the best of our knowledge, there are only a few works in the literature.
 %

### 2020-04-27 conference/camera ready/paper.tex
splitting the algorithm

@@
 % Algorithm typesetting package
-\usepackage[plain, shortend, linesnumbered]{algorithm2e}
+\usepackage[plain, shortend, linesnumbered, ruled]{algorithm2e}
 \SetAlFnt{\sffamily \small}
@@
   \renewcommand{\algorithmautorefname}{Algo.}%
-  %\renewcommand{\exampleautorefname}{Ex.}%
+  %\newcommand{\exampleautorefname}{Ex.}%
+  \def\exampleautorefname{Ex.}% 
   \renewcommand{\equationautorefname}{Map}%
@@
 % Non standard
-%\usepackage{varacca}
 \input{semmacros}
@@
 
-Some comments on \autoref{alg:subtyping} follow.
-In function $generateRelation$ we iterate over the set of classes in an SFJ program.
-If the class currently being processed contains types in its fields or methods not present in the subtyping relation (lines 5, 30, 42), then we add the current class to the list of \emph{unprocessed} classes (line 6) so we can process its fields and methods first and the class itself later after having all required type information.
-The set of unprocessed classes will then be inspected again in a recursive call (line 10).
-The next two functions of the algorithm, \emph{addClass} and \emph{checkSuperSet}, check subtyping for the current $class$ being processed and update $relation$, which is a mapping from a type to its subtypes and originally only consists of entries from \autoref{eq:initialSubtype}.
-In function $addClass(class)$ we check if the type $class$ is a subtype of an existing type in $relation$ (lines 15-18), as well as the opposite, meaning if $class$ is a supertype of an existing type in $relation$ (line 19).
-In order to do so $checkSuperSet$ checks all fields (lines 28-39) and all methods (lines 40-51) in $class$ and compares them with an \emph{existingClass} in $relation$.
-If a subtyping relation is established, then it is added to $relation$ (line 53).
-Finally, upon returning from $checkSuperSet$, we also add $class$ to its own relation (line 21) to satisfy reflexivity and to $Universe$ (line 22), which is a supertype of all types.
-
-\begin{algorithm}
-    \caption{Semantic Subtyping for SFJ Classes}
+\begin{algorithm}[t]
+    \caption{Semantic Subtyping for SFJ---\emph{generateRelation}}
     \label{alg:subtyping}
-
+    \SetKwBlock{Begin}{begin}{}
     \DontPrintSemicolon
@@
         \;
+    }
+\end{algorithm}
+%%
+%%
+\setcounter{algocf}{1}
+\begin{algorithm}
+    \caption{Semantic Subtyping for SFJ---\emph{addClass} and \emph{checkSuperSet}}
+
+    \DontPrintSemicolon
+    \setcounter{AlgoLine}{12}
+    \SetKwBlock{Begin}{}{end}
+    \Begin{
         \Fn{addClass(class: Class) $\rightarrow$ boolean}{
@@
 \end{algorithm}
+Some comments on \autoref{alg:subtyping} follow.
+In function $generateRelation$ we iterate over the set of classes in an SFJ program.
+If the class currently being processed contains types in its fields or methods not present in the subtyping relation (lines 5 and 30, 42 in the continuation of the algorithm in the next page), then we add the current class to the list of \emph{unprocessed} classes (line 6) so we can process its fields and methods first and the class itself later after having all required type information.
+The set of unprocessed classes will then be inspected again in a recursive call (line 10).
+
+The next two functions of the algorithm, \emph{addClass} and \emph{checkSuperSet} given in the next page, check subtyping for the current $class$ being processed and update $relation$, which is a mapping from a type to its subtypes and originally only consists of entries from \autoref{eq:initialSubtype}.
+In function $addClass(class)$ we check if the type $class$ is a subtype of an existing type in $relation$ (lines 15-18), as well as the opposite, meaning if $class$ is a supertype of an existing type in $relation$ (line 19).
+In order to do so $checkSuperSet$ checks all fields (lines 28-39) and all methods (lines 40-51) in $class$ and compares them with an \emph{existingClass} in $relation$.
+If a subtyping relation is established, then it is added to $relation$ (line 53).
+Finally, upon returning from $checkSuperSet$, we also add $class$ to its own relation (line 21) to satisfy reflexivity and to $Universe$ (line 22), which is a supertype of all types.
 

### 2020-04-27 conference/camera ready/background.tex
giving space

@@
 $$
+
 The $\alpha$-types are used to type fields and the $\mu$-types are used to type methods.
@@
 \end{align*}
+
 We assume an infinite set of names, with some special names:

### 2020-04-27 conference/camera ready/paper.tex
giving space

@@
     \end{equation*}
+
     Now, suppose our class hierarchy is such that \emph{Polygon} is the parent class and all other geometric figures extend \emph{Polygon}, which is how one would naturally define the set of polygons.
@@
     \end{equation*}
+
     We can now call \emph{diagonal} on an argument of type \emph{Polygon}: if the polygon is \textbf{not} a \emph{Triangle}, then the method computes and returns the length of its longest diagonal; otherwise, there will be a type error at compile time.
@@
 %
-The algorithm returns $True$ only if all the types in the program are resolved, otherwise it returns $False$, meaning there is at least one type definition which is invalid and contains a cycle.
+The algorithm returns $True$ only if all the types in the SFJ program are resolved, otherwise it returns $False$, meaning there is at least one type definition which is invalid and contains a cycle.
 
@@
 Leveraging the interpretation of types as sets of values to define semantic subtyping for FJ \cite{Dardha2013,Dardha2017}, in SFJ we keep track of the semantic subtyping relation by defining a map from a type to the set of its subtypes, satisfying the property that the set of values of a subtype is included in the set of values of the type.
+
 We start with basic types and let $Universe$ be a supertype of all types.
@@
 Given that classes are valid type definitions by \autoref{alg:types}, we are guaranteed that \autoref{alg:subtyping} will terminate.
+
 The semantic subtyping generated by \autoref{alg:subtyping} is a preorder: it is reflexive and transitive. This is also illustrated by \autoref{eq:initialSubtype}.
@@
 \end{equation}
+
 Recall the method \emph{diagonal} in class \emph{Diagonal}, with signature
@@
 \end{equation*}
+
 Both approaches have their pros and cons, and they leave an expectation on the developer to use the logic behind subtyping correctly when writing code.
@@
 $$
+
 Method {\it length} has type $\mathbf{string} \rightarrow \mathbf{int}$ in class $A$.
@@
 $$
+
 We define the type of the (outside) $diagonal$ method as accepting any type and its subtypes implementing the (inside) \emph{diagonal} method with type signature \textbf{void} to \textbf{double}.

### 2020-04-27 conference/camera ready/paper.tex
on future work
commenting on algorithm

@@
     \DontPrintSemicolon
+    \SetNoFillComment
     \Input{$classes$, the set of classes in an SFJ program.\\
@@
     }
+    \tcc{algorithm continued on next page...}
 \end{algorithm}
@@
     \DontPrintSemicolon
+    \SetNoFillComment
+    \tcc{...algorithm continued from previous page}
+
     \setcounter{AlgoLine}{12}
@@
  $$
+
 We read this typing rule as follows:
@@
 \end{equation}
-
+%
 Recall the method \emph{diagonal} in class \emph{Diagonal}, with signature
@@
 
-\section{Conclusion}
+\section{Related Work and Conclusion}
 \label{sec:conclusion}
-
-In this paper we presented the design and implementation of SFJ---Semantic Featherweight Java, an extension of Featherweight Java featuring boolean types, semantic subtyping, and integrating both nominal and structural subtyping.
-Due to the expressivity of semantic subtyping, in SFJ we are able to restore standard Java constructs and features for example, lists and overloading meathods, which were not present in FJ, thus making SFJ a more complete language.
-We presented \autoref{alg:types} on validity of type definitions and
-\autoref{alg:subtyping} on semantic subtyping, which finds all nominal and structural subtypes for all types in an SFJ program.
-We also described typing of terms in SFJ, which follows that of Java and builds upon relevant work \cite{Dardha2013,Dardha2017}.
 
@@
 
+In this paper we presented the design and implementation of SFJ---Semantic Featherweight Java, an extension of Featherweight Java featuring boolean types, semantic subtyping, and integrating both nominal and structural subtyping.
+Due to the expressivity of semantic subtyping, in SFJ we are able to restore standard Java constructs and features for example, lists and overloading meathods, which were not present in FJ, thus making SFJ a more complete language.
+We presented \autoref{alg:types} on validity of type definitions and
+\autoref{alg:subtyping} on semantic subtyping, which finds all nominal and structural subtypes for all types in an SFJ program.
+We also described typing of terms in SFJ, which follows that of Java and builds upon relevant work \cite{Dardha2013,Dardha2017}.
+As future work, we aim to finalise the code generation phase, which is sketched in \autoref{sec:code_generation}.
+
 %% References
